### 2017-08-13 article.tex
report

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Journal Article
% LaTeX Template
% Version 1.4 (15/5/16)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% Original author:
% Frits Wenneker (http://www.howtotex.com) with extensive modifications by
% Vel (vel@LaTeXTemplates.com)
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------
\documentclass[a4paper,11pt, twoside,twocolumn]{article}
\usepackage{blindtext} % Package to generate dummy text throughout this template

\usepackage[sc]{mathpazo} % Use the Palatino font
\usepackage[T1]{fontenc} % Use 8-bit encoding that has 256 glyphs
\linespread{1.05} % Line spacing - Palatino needs more space between lines
\usepackage{microtype} % Slightly tweak font spacing for aesthetics

\usepackage[english]{babel} % Language hyphenation and typographical rules
\usepackage{multirow}
\usepackage{multicol}
\usepackage[left=2.5cm,right=2cm,top=2cm,bottom=2cm, hmarginratio=1:1,headheight=14pt,columnsep=15pt]{geometry} % Document margins
\usepackage[hang, small,labelfont=bf,up,textfont=it,up]{caption} % Custom captions under/above floats in tables or figures
\usepackage{booktabs} % Horizontal rules in tables
\usepackage{graphicx}
\usepackage{enumitem} % Customized lists
\setlist[itemize]{noitemsep} % Make itemize lists more compact

\usepackage{abstract} % Allows abstract customization
\renewcommand{\abstractnamefont}{\normalfont\bfseries} % Set the "Abstract" text to bold
\renewcommand{\abstracttextfont}{\normalfont\small} % Set the abstract itself to small italic text

\usepackage{fancyhdr} % Headers and footers
\pagestyle{fancy} % All pages have headers and footers
\fancyhead{} % Blank out the default header
\fancyfoot{} % Blank out the default footer
%\fancyhead[C]{Running title $\bullet$ May 2016 $\bullet$ Vol. XXI, No. 1} % Custom header text
\fancyfoot[RO,LE]{\thepage} % Custom footer text

\usepackage{titling} % Customizing the title section
\usepackage{titlesec}
\titleformat{name=\section}{\flushleft\bf\scshape\Large}{\thesection}{.5em}{}
\titleformat{name=\subsection}{\flushleft\bf\scshape\large}{\thesubsection}{.5em}{}
\titleformat{name=\subsubsection}{\flushleft\bf\scshape\normalsize}{\thesubsubsection}{.5em}{}

\usepackage{hyperref} % For hyperlinks in the PDF
\usepackage{amsfonts} % For more logic symbols
\usepackage{amssymb} % For more logic symbols
\usepackage[fleqn]{amsmath} % left aligned math
\usepackage{centernot} % to use not implies
\usepackage{array}
\usepackage{upgreek}
\setlength\parindent{0pt}

%----------------------------------------------------------------------------------------
%	Custom Logic Environment and  Bindings
%----------------------------------------------------------------------------------------
%eq numbering
\makeatletter
\newcommand{\eqnum}[1][]{\leavevmode\hfill\refstepcounter{equation}\label{#1}\textup{\tagform@{\theequation}}}
\makeatother
%environment formating
\newenvironment{logic}[1][null]
{\begin{flushleft} \small \label{#1}}
{\end{flushleft}}
%logic shortcuts
\newcommand*\rot{\rotatebox{90}}
\newcommand{\loin}{$\in$}
\newcommand{\loforall}{$\forall$}
\newcommand{\loexists}{$\exists$}
\newcommand{\loand}{$\land$}
\newcommand{\loor} {$\lor$}
\newcommand{\loneq} {$\neq$}
\newcommand{\losubseteq}{$\subseteq$}
\newcommand{\loimplies}{$\implies$}
\newcommand{\lonimplies}{$\centernot\implies$}
\newcommand{\loimpliedby}{$\Longleftarrow$}
\newcommand{\lonimpliedby}{$\centernot\Longleftarrow$}
\newcommand{\losigma}{$\upsigma$}
\newcommand{\loturns} {$\vDash$}
\newcommand{\lochi}{$\upchi$}
\newcommand{\lophi}{$\upvarphi$}
\newcommand{\lorarrow}{$\rightarrow$}
\newcommand{\lonturns} {$\nvDash$}
\newcommand{\loiff} {$\iff$}
\newcommand{\loleadsto} {$\leadsto$}
\newcommand{\lomapsto} {$\mapsto$}
\newcommand{\locup} {$\cup$}
\newcommand{\lobigcup} {$\bigcup$}
\newcommand{\lotimes} {$\times$}
\newcommand{\loneg}{$\neg$}
\newcommand{\loexec}[2] {$\lfloor$#1$\rfloor _{\text{#2}}$}
\newcommand{\loconj}[1] {$\bar{\text{#1}}$}
\newcommand{\ablock} {\null\qquad}
%-----------------------
%----------------------------------------------------------------------------------------
%	TITLE SECTION
%----------------------------------------------------------------------------------------
\setlength{\droptitle}{-4\baselineskip} % Move the title up

\pretitle{\begin{center}\Huge\bfseries} % Article title formatting
\posttitle{\end{center}} % Article title closing formatting
\title{Object Capability Patterns:\\Policies and Specifications} % Article title
\author{%
\textsc{Shu-Peng Loh} \\[1ex] % Your name
\normalsize Imperial College London \\ % Your institution
\normalsize \href{mailto:shu.loh16@imperial.ac.uk}{shu.loh16@imperial.ac.uk} % Your email address
\and % if 2 authors are required, duplicate these 4 lines if more
\textsc{Sophia Drossopoulou}\\[1ex] % Second author's name
\normalsize Imperial College London \\ % Second author's institution
\normalsize \href{mailto:s.drossopoulou@imperial.ac.uk}{s.drossopoulou@imperial.ac.uk} % Second author's email address
}
\date{} % Leave empty to omit a date

\renewcommand{\maketitlehookd}{%
\begin{abstract}
\noindent  We propose a set of first-order logic predicates to formally specify object-to-object interactions which can then be used to describe a set of policies in an object-oriented computational model. Using these predicates, we attempt to formally specify the policies of well-established Object-Capability (OCap) patterns within the OCap literature which we have implemented in the capability-safe language Pony. We also offer some preliminary insights on how such specifications can be used in the context of a non-Ocap model by describing the security properties of a pattern built on the Ethereum smart contract programming language Solidity, which we argue implements a form of stack-based access control.\end{abstract}
}
\begin{document}
\renewcommand{\ref}[1]{\textit{\autoref{#1}}}
% Print the title
\maketitle

%----------------------------------------------------------------------------------------
%	ARTICLE CONTENTS
%----------------------------------------------------------------------------------------

\section{Introduction}
Recent widespread adoption of distributed ledger technology (blockchain) has created multiple decentralized, distributed computational platforms where millions of dollars are transacted over codified constructs called smart contracts\footnote{For example, as of 10 Aug 2017, Ethereum is a US\$28 billion blockchain platform with an in-built Turing-complete programming language that can be used to create and deploy such contracts.}.
The power of distributed modern computing hence lies in facilitating cooperation between multiple agents, but it comes with risk as an agent is vulnerable to \textit{unexpected} outcomes\footnote{Representing in general any outcome arising from a piece of code execution that has deviated from an agent's original intention or objective independent from code.} from participating in these smart contracts. This might generally arise from two issues: 
\begin{itemize}
\item oversight or misconception of the outcomes of executing a piece of \textit{known} code
\item failure to defend against malicious execution of \textit{unknown} code components\end{itemize}
\noindent These two issue are often closely intertwined in any system of execution that has both trusted and untrusted code components (the second issue is often a result of the first).

In recent years the Object-Capability (OCap) model has received increasing attention as a compelling approach to building robust, distributed systems that promote what Miller\cite{miller2006} calls \textit{cooperation without vulnerability}. The OCap model attemps to address these two issues by alleviating security as a separate concern from the mind of the programmer, by leveraging the object-oriented programming paradigm and imposing certain prohibitions.
%----------------------------------------------------------------------------------------
%   OCap Model
%----------------------------------------------------------------------------------------
\section{OCap Model}
The OCap model uses the reference graph of the objects as the access graph, and strictly requires objects to interact with each other only by sending messages on object references\cite{miller2003b}.
\subsection{From Capability to Object-Capability}
\subsection{OCap Languages}
\begin{itemize}
\item Joe-E (inspired by Java)
\item Emily (inspired by OCaml)
\item Caja (inspired by JavaScript)
\item E
\item Pony
\end{itemize}
\subsubsection{Language Restrictions}
\subsection{OCap Patterns}
An OCap pattern is a concrete representation of the OCap model and comprises a set of objects connected to each other by capabilities. Objects interact with each other by sending messages on capabilities. An OCap pattern may be visualised as a directed graph---nodes represent objects, and each edge from an object \textit{o} to another \textit{o'} represents \textit{o} holding a capability that allows it to directly access \textit{o'}.
%----------------------------------------------------------------------------------------
%   Formal Specifications
%----------------------------------------------------------------------------------------
\section{Formal specifications}
\subsection{Definitions}\noindent
Here, we borrow liberally the definitions of runtime state, module and arising configurations from the appendix of \cite{drossopoulou2015b}.\\
\textbf{Runtime state:}
 \losigma\ consists of a stack frame \lophi, and a heap \lochi. A stack frame is a mapping from receiver (this) to its address, and from the local variables (VarId) and parameters (ParId) to their values. Values are integers, the booleans true or false, addresses, or null. The heap maps addresses to objects. Objects are tuples consisting of the class of the object, and a mapping from field identifiers onto values.
\begin{logic}[Runtime state]
\losigma\ \loin\ state = frame \lotimes\ heap \\
\lophi \loin\ frame = StackId $\rightarrow$ val \\
\lochi \loin\ heap = addr $\rightarrow$ object \\
v \loin\ val = \{\texttt{null}, \texttt{true}, \texttt{false}\} \locup\ \textit{addr} \locup\ $\mathbb{N}$ \\
StackId = \{\textit{this}\} \locup\ VarId \locup\ ParId \\
object = ClassId \lotimes (FieldId $\rightarrow$ val)
\end{logic}
\textbf{Module:}
\begin{logic}[Module]
M \loin\ Module = ClassId \locup\ SpecId\linebreak\ablock\qquad\qquad\quad $\rightarrow$ \linebreak
\ablock \qquad \qquad \quad (ClassDescr \locup\ Specification)
\end{logic}

\textbf{Reach and Execution:}\\
\begin{logic}[Reach and Execution]\end{logic}

\textbf{Arising Configurations}\\
\begin{logic}[Arising]\end{logic}

\textbf{Domination:}\\
\begin{logic}[Domination]\end{logic}

\subsubsection{MayAccess Definitions}
\begin{logic}[MayAccessDirNow]
\loforall M,\losigma,x,e. M,\losigma\ \loturns\ MayAccess$^{Dir,Now}$(x,e) \loiff \linebreak
\ablock \loexists f. \loexec{x.f}{\losigma} = \loexec{e}{\losigma}
\loor \linebreak
\ablock (\losigma(this) = \loexec{x}{\losigma} \loand\ \loexists y. \losigma(y) = \loexec{e}{\losigma})
\end{logic}
\begin{logic}[MayAccessDirEve]
\loforall M,\losigma,x,e. M,\losigma\ \loturns\ MayAccess$^{Dir,Eve}$(x,e) \loiff \linebreak
\ablock \loexists \losigma' \loin Arising(M,\losigma).\linebreak
\ablock M,\losigma'\ \loturns\ MayAccess$^{Dir,Now}$(x,e) 
\end{logic}
\begin{logic}[MayAccessIndNow]
\loforall M,\losigma,x,e. M,\losigma\ \loturns\ MayAccess$^{Ind,Now}$(x,e) \loiff \linebreak
\ablock \loexists \loconj{f}. \loexec{x.\loconj{f}}{\losigma} = \loexec{e}{\losigma}
\loor \linebreak
\ablock (\losigma(this) = \loexec{x}{\losigma} \loand\ \loexists y. \losigma(y.\loconj{f}) = \loexec{e}{\losigma})
\end{logic}
\begin{logic}[MayAccessIndEve]
\loforall M,\losigma,x,e. M,\losigma\ \loturns\ MayAccess$^{Ind,Eve}$(x,e) \loiff \linebreak
\ablock \loexists \losigma' \loin Arising(M,\losigma).\linebreak
\ablock M,\losigma'\ \loturns\ MayAccess$^{Ind,Now}$(x,e) 
\end{logic}
A note on \textit{f} and \textit{\loconj{f}}\ : While f can considered as an object field, it can also represent an object method that returns a val.
Similarly \loconj{f} can be considered a series of fields, or methods that return a val, or a combination of both.\\

Note that we define MayAccess$^{Ind,*}$(x,e) to mean a form of weak access---that a path exists from x to some well-formed e, but it does \textit{not} imply that such a path is navigateable by x. Take the following example where at a particular state, x1.next points to x2, and x2.next points to x3. MayAccess$^{Ind,*}$(x1,x3) holds, regardless of whether x2.next is navigateable by x1. If x2.next is a private method that can only be called internally by x2, the predicate still holds---we say that a path from x1 to x3 exists, but is not navigateable by x1.\\ \\
On the other hand, we define MayAccess$^{Dir,*}$(x,e) to mean a form of strong access---it implies that a patch exists from x0 to e and that such a path \textit{is} navigateable. This is because the reference to e exists within x0's state and can be used by x0. Notice how this differs from the previous example, where x1 cannot guarantee that it can execute x2.next, because it is possible that x2.next is protected by x2 through encapsulation and data-hiding.

\subsubsection{MayAffect Definitions}
\begin{logic}[MayAffectNow]
\loforall M,\losigma,x,e. M,\losigma\ \loturns\ MayAffect$^{Now}$(x,e) \loiff \linebreak
\ablock \loexists \loconj{m},\loconj{a},\losigma'. x.\loconj{m}(\loconj{a}) \loleadsto\ \losigma' \loand \loexec{e}{\losigma} \loneq \loexec{e}{\losigma'}
\end{logic}
\begin{logic}[MayAffectEve]
\loforall M,\losigma,x,e. M,\losigma\ \loturns\ MayAffect$^{Eve}$(x,e) \loiff \linebreak
\ablock \loexists \losigma \loin Arising(M,\losigma).\linebreak
\ablock \loexists \loconj{m},\loconj{a},\losigma'. x.\loconj{m}(\loconj{a}) \loleadsto\ \losigma' \loand \loexec{e}{\losigma} \loneq \loexec{e}{\losigma'}
\end{logic}
\begin{logic}[AffectObject]
\textit{If e is an object:} \\
\loforall\losigma,\losigma', e\loin Object. \loexec{e}{\losigma} \loneq \loexec{e}{\losigma'} \loiff \linebreak
\ablock \loexists f. \loexec{e.f}{\losigma} \loneq \loexec{e.f}{\losigma'}
\end{logic}

\subsection{Implications}

\subsubsection{Imposing OCap rules}
We assume \loforall \{o,o'\}\loin Object.\\
\begin{table}[hbtp]
\small
\caption{MayAccess(o,o') Relations in OCap}
\centering
\begin{tabular*}{\linewidth}{c|ccc}\toprule
& \bf Now & & \bf Eventually\\
\hline
\multirow{5}{*}{\rot{\bf Indirect \enspace Direct \:}} & \multirow{2}{*}{MayAccess$^{Dir,Now}$} & \loimplies & \multirow{2}{*}{MayAccess$^{Dir, Eve}$} \\
& & \lonimpliedby &  \\
& \rot{\loimpliedby} \rot{\lonimplies}& &\rot{\loimpliedby} \rot{\lonimplies} \\
& \multirow{2}{*}{MayAccess$^{Ind,Now}$} & \loimplies & \multirow{2}{*}{MayAccess$^{Ind, Eve}$} \\
& & \lonimpliedby &  \\
\bottomrule
\end{tabular*}
\end{table}

\begin{table}[hbtp]
\caption{MayAffect(o,o') Relations in OCap}
\centering
\begin{tabular*}{\linewidth}{ccc}\toprule
\multirow{2}{*}{MayAffect$^{Now}$} & \loimplies & \multirow{2}{*}{MayAffect$^{Eve}$} \\
& \lonimpliedby &  \\
\bottomrule
\end{tabular*}
\end{table}

\begin{logic}
\textbf{Results from Contraposition:}\\
\loneg MayAccess$^{Dir,Eve}$(o,o')\loimplies \linebreak
\ablock\loneg MayAccess$^{Dir,Now}$(o,o') \\

\loneg MayAccess$^{Ind,Eve}$(o,o')\loimplies \linebreak
\ablock\loneg MayAccess$^{Ind,Now}$(o,o') \\

\loneg MayAccess$^{Ind,Now}$(o,o')\loimplies \linebreak
\ablock\loneg MayAccess$^{Dir,Now}$(o,o') \\

\loneg MayAccess$^{Ind,Eve}$(o,o')\loimplies \linebreak
\ablock\loneg MayAccess$^{Dir,Eve}$(o,o')
\end{logic}

\textbf{Rule: Objects can only interact with each other through sending messages on capabilities.} Therefore, if an object o can affect o', then there must be a path from o to o':
\begin{logic}[Implication 1]
M,\losigma\ \loturns\ MayAffect$^{Eve}$(o,o') \loimplies\linebreak
	\ablock MayAccess$^{Ind,Eve}$(o,o') \eqnum[1]
\end{logic}
\begin{logic}[Implication 2]
\textit{By contraposition,}\linebreak
M,\losigma\ \loturns\ \loneg MayAccess$^{Ind,Eve}$(o,o') \loimplies\linebreak
	\ablock \loneg MayAffect$^{Eve}$(o,o') \eqnum[2]
\end{logic}
Here, we immediately see a first defensive outcome of the OCap model. Isolation of object o' from o guarantees that the state of object o' cannot be modified by object o.\\

\textbf{Rule: Objects cannot forge capabilities, and only connectivity begets connectivity.}
Therefore, if an object o' exists in a state \losigma\ and there are no possible paths from o to o' in that state, then it is not possible in any arising state \losigma' from \losigma\ where there is a path from o to o':
\begin{logic}[Implication 3]
M,\losigma\ \loturns\ \loneg MayAccess$^{Ind,Now}$(o,o') \loand\ Exists(o')$_{\text{\losigma}}$ \loimplies \linebreak
\ablock \loneg MayAccess$^{Ind,Eve}$(o,o')\eqnum[3]
\end{logic}
We require the enforcement of o' to exist in \losigma\ because o' can refer to an object that does not exist in \losigma\ but will be created and exist in some arising \losigma' from \losigma.
From \ref{3} and \ref{2}, to protect o' from o in all states arising from a state \losigma, we only need to ensure isolation of o' at \losigma\ :
\begin{logic}[Implication 4]
M,\losigma\ \loturns\ \loneg MayAccess$^{Ind,Now}$(o,o') \loand\ Exists(o')$_\text{\losigma}$ \loimplies \linebreak
	\ablock\loneg MayAffect$^{Eve}(o,o')$\eqnum[4]
\end{logic}

So far these results only serve as a base case, because the protection of an object's state requires the impractical configuration that objects are completely isolated from each other on the reference graph. Cooperation between objects imply that there needs to be some path established between the objects for interaction to take place and therefore the predicate \loneg MayAccess$^{Ind,Now}$(o,o') rarely holds in a system of cooperation.\\

The power of OCap patterns hence lies in providing concrete examples of a system of cooperation that allows the existence of paths between objects for cooperation while still dictating the degree of control of one object can have over another. For example, we can have variations of following configurations in an Ocap pattern:
\begin{itemize}\raggedright \small
\item M,\losigma\ \loturns\ MayAccess$^{Ind,Now}$(o,o') \loand\ Exists(o')$_\text{\losigma}$
\loimplies \loneg MayAffect$^{Eve}(o,o')$
\item M,\losigma\ \loturns\ MayAccess$^{Ind,Eve}$(o,o') \linebreak 
\loimplies \loneg MayAffect$^{Eve}(o,o')$
\end{itemize}

Notice how we have also also focused on the predicate MayAffect$^{Eve}$ rather than MayAffect$^{Now}$. Security concerns are often a question of whether what one object can do to another object's in any eventual state of a system. If we are concerned with the protection of o' from o, it is not very useful to have a policy where \loneg MayAffect$^{Now}$(o,o') holds while \loneg MayAffect$^{Eve}$(o,o') does not. Furthermore, by ensuring \loneg MayAffect$^{Eve}$ holds, we can also ensure \loneg MayAffect$^{Now}$ holds since, by contraposition, \loneg MayAffect$^{Eve}$ \loimplies \loneg MayAffect$^{Now}$.
\subsection{Pattern 1: The JavaScript DOM Tree}
We define the following variables throughout our pattern:
\begin{itemize}
\item o,o' \loin Object
\item Node,ReNode \losubseteq Object
\item n,n' \loin Node
\item rn,rn' \loin ReNode
\end{itemize}
\textbf{Vulnerability of a Node}:
The vulnerability of a node lies in the fact that its fields and methods are defined as public. Therefore, any object that has a path and can navigate a path to a node may change the state of a node.

\begin{logic}[DomP1]
\loforall o, n. MayAccess$^{Dir,Eve}$(o,n)\linebreak \ablock\ablock \loimplies \linebreak
\ablock \ablock MayAffect$^{Eve}$(o,n)\linebreak
\end{logic}
\textsc{Policy 1: Necessary Condition}
\begin{logic}[NodeP1]
\loforall n,o, RN\losubseteq ReNode.\linebreak
Dom(RN,n) \loand\ MayAffect$^{Eve}$(o,n)\linebreak \ablock\loimplies \linebreak
\ablock \loexists rn\loin RN. MayAccess$^{Ind,Eve}$(o,rn)\linebreak
\ablock \ablock \ablock\loand\ MayAccess$^{Dir,Eve}$(rn,n)
\end{logic}

\begin{logic}[DomP1]
\loforall o, n. MayAccess$^{Dir,Eve}$(o,n)\linebreak \ablock\ablock \loimplies \linebreak
\ablock \ablock MayAffect$^{Eve}$(o,n)\linebreak
\end{logic}
\textsc{Policy 1: Necessary Condition}
\begin{logic}[NodeP1]
\loforall n,o, RN\losubseteq ReNode.\linebreak
Dom(RN,n) \loand\ MayAffect$^{Eve}$(o,n)\linebreak \ablock\loimplies \linebreak
\ablock \loexists rn\loin RN. MayAccess$^{Ind,Eve}$(o,rn)\linebreak
\ablock \ablock \ablock\loand\ MayAccess$^{Dir,Eve}$(rn,n)
\end{logic}

\subsection{Pattern 2: Redell's Caretaker}
\subsection{Pattern 3: Membrane}

\textbf{Policy 1}:

%----------------------------------------------------------------------------------------
%	REFERENCE LIST
%----------------------------------------------------------------------------------------
\small
\bibliographystyle{plain}
\bibliography{ref}

%----------------------------------------------------------------------------------------

\end{document}

### 2017-08-14 article.tex
update

@@
 \usepackage{enumitem} % Customized lists
+
 \setlist[itemize]{noitemsep} % Make itemize lists more compact
@@
 \newcommand{\loneq} {$\neq$}
+\newcommand{\losimeq} {$\simeq$}
 \newcommand{\losubseteq}{$\subseteq$}
@@
 \newcommand{\ablock} {\null\qquad}
+
 %-----------------------
@@
 \begin{abstract}
-\noindent  We propose a set of first-order logic predicates to formally specify object-to-object interactions which can then be used to describe a set of policies in an object-oriented computational model.
+\noindent  We propose a set of higher-order predicate logic to formally specify object-to-object interactions which can then be used to describe reference dynamics in an object-oriented computational model.
 Using these predicates, we attempt to formally specify the policies of well-established Object-Capability (OCap) patterns within the OCap literature which we have implemented in the capability-safe language Pony.
@@
 \subsection{Definitions}\noindent
-Here, we borrow liberally the definitions of runtime state, module and arising configurations from the appendix of \cite{drossopoulou2015b}.\\
+We borrow liberally the definitions of runtime state, module and arising configurations from the appendix of \cite{drossopoulou2015b}.\\
 \textbf{Runtime state:}
@@
 \subsubsection{MayAccess Definitions}
+We define in total four flavours of MayAccess predicates that describe the relation between two entities in a system (of arity 2 that represent the identifiers of these entities).
+These four flavours represent a combination of space (distance) and time (state):
+\begin{itemize}
+\item distance: directly ($Dir$) or indirectly ($Ind$)
+\item state: now ($Now$) or eventually ($Eve$)
+\end{itemize}
+and are broad enough to describe both non-OCap and OCap models:
 \begin{logic}[MayAccessDirNow]
-\loforall M,\losigma,x,e. M,\losigma\ \loturns\ MayAccess$^{Dir,Now}$(x,e) \loiff \linebreak
+M,\losigma\ \loturns\ MayAccess$^{Dir,Now}$(x,e) \loiff \linebreak
 \ablock \loexists f. \loexec{x.f}{\losigma} = \loexec{e}{\losigma}
@@
 \begin{logic}[MayAccessDirEve]
-\loforall M,\losigma,x,e. M,\losigma\ \loturns\ MayAccess$^{Dir,Eve}$(x,e) \loiff \linebreak
+ M,\losigma\ \loturns\ MayAccess$^{Dir,Eve}$(x,e) \loiff \linebreak
 \ablock \loexists \losigma' \loin Arising(M,\losigma).\linebreak
@@
 \begin{logic}[MayAccessIndNow]
-\loforall M,\losigma,x,e. M,\losigma\ \loturns\ MayAccess$^{Ind,Now}$(x,e) \loiff \linebreak
+M,\losigma\ \loturns\ MayAccess$^{Ind,Now}$(x,e) \loiff \linebreak
 \ablock \loexists \loconj{f}. \loexec{x.\loconj{f}}{\losigma} = \loexec{e}{\losigma}
@@
 \begin{logic}[MayAccessIndEve]
-\loforall M,\losigma,x,e. M,\losigma\ \loturns\ MayAccess$^{Ind,Eve}$(x,e) \loiff \linebreak
+M,\losigma\ \loturns\ MayAccess$^{Ind,Eve}$(x,e) \loiff \linebreak
 \ablock \loexists \losigma' \loin Arising(M,\losigma).\linebreak
@@
 \end{logic}
-A note on \textit{f} and \textit{\loconj{f}}\ : While f can considered as an object field, it can also represent an object method that returns a val.
+
+A note on \textit{f} and \textit{\loconj{f}}\ : While f can considered as a field, it can also represent a method that returns a val.
 Similarly \loconj{f} can be considered a series of fields, or methods that return a val, or a combination of both.\\
 
-Note that we define MayAccess$^{Ind,*}$(x,e) to mean a form of weak access---that a path exists from x to some well-formed e, but it does \textit{not} imply that such a path is navigateable by x.
-Take the following example where at a particular state, x1.next points to x2, and x2.next points to x3.
-MayAccess$^{Ind,*}$(x1,x3) holds, regardless of whether x2.next is navigateable by x1.
-If x2.next is a private method that can only be called internally by x2, the predicate still holds---we say that a path from x1 to x3 exists, but is not navigateable by x1.\\ \\
-On the other hand, we define MayAccess$^{Dir,*}$(x,e) to mean a form of strong access---it implies that a patch exists from x0 to e and that such a path \textit{is} navigateable.
-This is because the reference to e exists within x0's state and can be used by x0.
-Notice how this differs from the previous example, where x1 cannot guarantee that it can execute x2.next, because it is possible that x2.next is protected by x2 through encapsulation and data-hiding.
-
-\subsubsection{MayAffect Definitions}
-\begin{logic}[MayAffectNow]
-\loforall M,\losigma,x,e. M,\losigma\ \loturns\ MayAffect$^{Now}$(x,e) \loiff \linebreak
-\ablock \loexists \loconj{m},\loconj{a},\losigma'. x.\loconj{m}(\loconj{a}) \loleadsto\ \losigma' \loand \loexec{e}{\losigma} \loneq \loexec{e}{\losigma'}
-\end{logic}
-\begin{logic}[MayAffectEve]
-\loforall M,\losigma,x,e. M,\losigma\ \loturns\ MayAffect$^{Eve}$(x,e) \loiff \linebreak
-\ablock \loexists \losigma \loin Arising(M,\losigma).\linebreak
-\ablock \loexists \loconj{m},\loconj{a},\losigma'. x.\loconj{m}(\loconj{a}) \loleadsto\ \losigma' \loand \loexec{e}{\losigma} \loneq \loexec{e}{\losigma'}
-\end{logic}
-\begin{logic}[AffectObject]
-\textit{If e is an object:} \\
-\loforall\losigma,\losigma', e\loin Object. \loexec{e}{\losigma} \loneq \loexec{e}{\losigma'} \loiff \linebreak
-\ablock \loexists f. \loexec{e.f}{\losigma} \loneq \loexec{e.f}{\losigma'}
-\end{logic}
-
-\subsection{Implications}
-
-\subsubsection{Imposing OCap rules}
-We assume \loforall \{o,o'\}\loin Object.\\
+We surmarise the relationships between the four flavours of MayAccess in \ref{MayAccessTable}:
 \begin{table}[hbtp]
 \small
-\caption{MayAccess(o,o') Relations in OCap}
+\caption{MayAccess$^{*,*}$(x,e) Relations}
 \centering
@@
 & \multirow{2}{*}{MayAccess$^{Ind,Now}$} & \loimplies & \multirow{2}{*}{MayAccess$^{Ind, Eve}$} \\
-& & \lonimpliedby &  \\
-\bottomrule
+& & \lonimpliedby &\\
+\label{MayAccessTable}
 \end{tabular*}
 \end{table}
+
+Let us assume that both x and e are well-defined.
+Note that without imposing any further assumptions (such as those from the OCap model), we have defined \textit{both} MayAccess$^{Dir,Now}$(x,e) and MayAccess$^{Ind,Now}$(x,e) to mean forms of very weak access---that a directed path exists from x to e, but we do \textit{not} imply that such a path is traverseable by x (it might or might not be traverseable).
+Indeed, these definitions by themselves represent mere \textit{possibilities} of interaction (or possible authorities), they do \textit{not} represent that interaction (or authority) would succeed.
+The difference between MayAccess$^{Dir,Now}$(x,e) and MayAccess$^{Ind,Now}$(x,e) is only the computational distance between x and e on the reference graph, where the latter involves possibly intermediate entities (or objects in an object-oriented model).\\
+
+What do the definitions mean then for non-OCap and OCap models in an object-oriented world?
+Again, let us assume both o and o' are well-defined, valid object references.
+In non-OCap models, the possible presence of a global ambient authority means the predicates above say \textit{nothing} about whether any interaction between an object o and o' would succeed.
+This is true even if o possesses directly the reference of o', where MayAccess$^{Dir,Now}$(o,o') holds.
+For all we know, we could easily have in a non-OCap language a feature to completely restrict access to o' using a global ambient authority, such that any object in the programming world which possesses references of o' cannot use them, and all paths leading to o' represent merely possibilities of interaction, but from which no authority can possibly arise.\\
+
+Could we say more about OCap systems?
+In OCap systems, there can be no global ambient authority so that an object reference by itself represents both the designation and the authority to use the object.
+This therefore leads us to be able to make a crucial distinction between MayAccess$^{Ind,Now}$(o,o') and MayAccess$^{Dir,Now}$(o,o') in the OCap model:
+\begin{itemize}
+\item MayAccess$^{Ind,Now}$(o,o') means---\textit{only}\\
+\ablock 1. a directed path from o to o' exists\\
+\ablock\ablock (possible authority)
+\item MayAccess$^{Dir,Now}$(o,o') means---\textit{both}\\
+\ablock 1. a directed path from o to o' exists\\
+\ablock\ablock (possible authority) \textit{and}\\
+\ablock 2. o's authority to use o' \textit{will} succeed
+\end{itemize}
+We elaborate the distinction with the following OCap example where there is a particular state \losigma\ of the system where, o$_1$.next points to o$_2$, and o$_2$.next points to o$_3$.
+o$_2$.next is a private method that can only be called internally by o$_2$.
+In this example, M,\losigma ~\loturns ~MayAccess$^{Ind,Now}$(o$_1$,o$_3$) holds, regardless of whether o$_2$.next is traverseable by o$_1$.
+We say here that a path from o$_1$ to o$_3$ exists, but is not traverseable by o$_1$.
+On the other hand, MayAccess$^{Dir,Now}$(o$_1$,o$_3$) does not hold true, because o$_3$ is not reachable from $o_1$ in a single step---o$_2$ sits between them on the reference graph as an intermediate object, and can possibly prevent or allow traversal from o$_1$ to o$_3$ (in this example, o$_2$ prevents such a traversal).\\
+
+What does it mean then for MayAccess$^{Dir,Now}$(*,*) to hold in an OCap model?
+From the same example, MayAccess$^{Dir,Now}$(o$_1$,o$_2$) holds and implies a stronger form of access---it means that a path exists from o$_1$ to o$_2$, and that such a path \textit{is} traverseable.
+This is because, by the definition of the predicate and configuration of the example, the object reference of o$_2$ exists within o$_1$'s state.
+Therefore, o$_2$ is guaranteed therefore to be accessible, and its authority exercisable, by o$_1$, without the interference of any ambient authority.
+Notice how this o$_1$-o$_2$ relationship differs from the o$_1$-o$_3$ relationship, where o$_1$ cannot guarantee that it can exercise the authority of o$_2$.next which points to o$_3$, since o$_2$.next does not exist within o$_1$'s state---it is possible that o$_2$.next is protected by x$_2$ through encapsulation and data-hiding.\\
+Within an OCap model, we can now be convinced that the stronger
+\begin{logic}
+MayAccess$^{Dir,Now}$(o$_1$,o$_2$) \losimeq \linebreak
+\ablock o$_1$ has the capability of o$_2$\end{logic}
+while the weaker MayAccess$^{Ind,Now}$(o$_1$,o$_3$) does not say anything about whether o$_1$ has the capability of o$_3$, but that only a directed path exists.
+It does however, represent a necessary condition for capability.
+\begin{logic}
+MayAccess$^{Ind,Now}$(o$_1$,o$_3$) \losimeq\ \linebreak
+\ablock there is a directed path from o$_1$ to o$_2$\linebreak
+\\
+o$_1$ has the capability of o$_3$ \loimplies \linebreak
+\ablock MayAccess$^{Ind,Now}$(o$_1$,o$_3$)\end{logic}
+\subsubsection{MayAffect Definitions}
+With our access predicates in place, we introduce a set of predicates that describe changes to the state of a system.
+Again, we highlight that these predicates are broad enough to describe both non-OCap and OCap models.
+\begin{logic}
+M,\losigma\ \loturns\ MayAffect$^{Now}$(x,e) \loiff \linebreak
+\ablock \loexists \loconj{m},\loconj{a},\losigma'. x.\loconj{m}(\loconj{a}) \loleadsto\ \losigma' \loand \loexec{e}{\losigma} \loneq \loexec{e}{\losigma'}
+\linebreak \\
+M,\losigma\ \loturns\ MayAffect$^{Eve}$(x,e) \loiff \linebreak
+\ablock \loexists \losigma \loin Arising(M,\losigma).\linebreak
+\ablock \loexists \loconj{m},\loconj{a},\losigma'. x.\loconj{m}(\loconj{a}) \loleadsto\ \losigma' \loand \loexec{e}{\losigma} \loneq \loexec{e}{\losigma'}
+\linebreak \\
+\textit{If e is an object:} \\
+\loforall e\loin Object. \loexec{e}{\losigma} \loneq \loexec{e}{\losigma'} \loiff \linebreak
+\ablock \loexists f. \loexec{e.f}{\losigma} \loneq \loexec{e.f}{\losigma'}
+\end{logic}
 
@@
 
+\subsection{OCap Security Implications}
+
+\subsubsection{What is protection?}
+In an object-oriented world\footnote{
+To simplify our discussion, we work with the variables \{o,o'\}\loin Object for this entire section.}
+, security concerns between objects are often a question of whether what one object can do to another object in \textit{any} eventual state of a system.
+Because an object encapsulates both internal state and behaviour, strictly speaking, security of an object should govern over both the integrity of the object's fields (internal state) and whether the objects' methods can be called (behaviour).
+Our predicates are broad enough to enable a discussion of both protection of state (MayAffect$^{Eve}$) and behaviour (MayAccess$^{Dir,Eve}$)\footnote{ Our MayAccess$^{Dir,Eve}$ is not weak enough to reason specifically which behaviours can be called.
+This is however not a big issue in \textit{pure} OCap systems, where often giving away the capability of an object typically means allowing \textit{all} public behaviours of the object to be called without distinction.}.
+We emphasize however that protecting either state or behaviour of an object, does \textit{not} necessarily imply the other.
+In fact, a common feature in OCap patterns is being able to protect a sensitive object's behaviours (they cannot be called directly by untrusted objects), but at the some allowing the same untrusted objects to modify the object's state in some controlled way.\\
+
+There are however, some flexibility in working with objects, that allows us to simplify our discussion and work with only a broad definition of state protection in terms of our MayAffect$^{Eve}$ predicate, \textit{without} thinking too much about specific fields of the object we want to protect or the protection of behaviour.
+Moving away from our broad definition of whether an object may be affected (we defined it as being able to change at least one field of the object), to more precise spcifications of which particular field(s) of the object may be affected, should be trivial.
+We can in theory also stay and reason within our framework by separating the particular concerned field(s) of the object into separately encapsulated objects.
+We merely have to be careful as to \textit{which} object's state we want to protect.
+Furthermore, in theory one can easily introduce a field within an object that behaves like a 'signal' which would be modified whenever a specified behaviour is called.
+Preventing a particular behaviour to be called by an untrusted object then becomes equivalent to denying the untrusted object the ability to modify the particular state of the signal field of the object.\\
+
+With these simplifications, protection for us then becomes solely a matter of whether we can allow or deny an object to modify the state of another.
+In the subsections below we build the neccessary conditions for MayAffect$^{Eve}$, where MayAffect$^{Eve}$ is placed in the antecedent, and we examine which of the family of MayAccess predicates is placed in the consequent.
+To help us achieve this, we also formalize our assumptions and the rules of OCap, using our predicates, to help guide us in constructing our necessary conditions.\\
+
+In building our necessary conditions for state protection in the next subsection notice how we have focused on finding the necessary conditions for the weaker predicate MayAffect$^{Eve}$ rather than MayAffect$^{Now}$ in the antecedent.
+This is because negation on both sides of the implication, would yield a stronger \loneg MayAffect$^{Eve}$ in the consequent.
+In practical terms, if we are concerned with the protection of o' from o, it is also often not very useful to have a policy where \loneg MayAffect$^{Now}$(o,o') holds but \loneg MayAffect$^{Eve}$(o,o') does not.
+Furthermore, by ensuring \loneg MayAffect$^{Eve}$(o,o') holds, we can also ensure \loneg MayAffect$^{Now}$(o,o') holds since by contraposition:
+
 \begin{logic}
-\textbf{Results from Contraposition:}\\
-\loneg MayAccess$^{Dir,Eve}$(o,o')\loimplies \linebreak
-\ablock\loneg MayAccess$^{Dir,Now}$(o,o') \\
-
-\loneg MayAccess$^{Ind,Eve}$(o,o')\loimplies \linebreak
-\ablock\loneg MayAccess$^{Ind,Now}$(o,o') \\
-
-\loneg MayAccess$^{Ind,Now}$(o,o')\loimplies \linebreak
-\ablock\loneg MayAccess$^{Dir,Now}$(o,o') \\
-
-\loneg MayAccess$^{Ind,Eve}$(o,o')\loimplies \linebreak
-\ablock\loneg MayAccess$^{Dir,Eve}$(o,o')
+$[$MayAffect$^{Now}$(o,o') \loimplies MayAffect$^{Eve}$(o,o')$]$\linebreak
+\ablock \loimplies\linebreak
+$[$\loneg MayAffect$^{Eve}$(o,o') \loimplies \loneg MayAffect$^{Now}$(o,o')$]$\\
 \end{logic}
 
-\textbf{Rule: Objects can only interact with each other through sending messages on capabilities.} Therefore, if an object o can affect o', then there must be a path from o to o':
+In building the necessary conditions for the predicate MayAffect$^{Eve}$, we are also more concerned with finding the configuration of MayAccess$^{*,Now}$, rather than MayAccess$^{*,Eve}$.
+This is because it is easier to implement or prove a configuration of relations that holds in \textit{one} specific state and hence much more useful in practice, than think about whether a configuration holds in \textit{all} possible eventual states.\\
+\subsection{Formal Implications}
+\subsubsection{Formalizing OCap rules}
+We begin our reasoning of protection in our OCap model, by introducing the assumption that all fields in our objects can only be declared private.
+Consequently, this implies the necessary condition that an object's state can only be modified if one of it's behaviour is called, and therefore require at least one other object in the system that holds its capability.
+We do not think this is restrictive in any case, the programmer merely has to write explicitly a method to return an object's field.
+
+\begin{logic}
+\hrule\null
+\textsc{\normalsize *Private Fields Assumption (PFA)}\\
+M,\losigma\ \loturns\ \loexists X*\loin Obj. MayAffect$^{Now}$(X*,o') \loimplies \linebreak
+	\ablock \ablock \loexists Y*\loin Obj. MayAccess$^{Dir,Now}$(Y*,o')
+\linebreak \\
+\textit{By contraposition,}\linebreak
+M,\losigma\ \loturns\ 	\loforall Y*\loin Obj. \loneg MayAccess$^{Dir,Now}$(Y*,o') \loimplies \linebreak
+	\ablock \ablock \loforall X*\loin Obj. \loneg MayAffect$^{Now}$(X*,o')
+\linebreak
+\hrule
+\end{logic}
+
+This means that in order for an object X*'s state to change, it must be done through some object Y* calling its behaviour (Y* can refer to the same object as X*).
+Equivalently, denying all objects in the system the ability to call an object's behaviour implies that no object can modify the object's state.\\
+
+\textbf{Rule 1: Objects can only interact with each other through sending messages on capabilities.}
+Therefore, if an object o can affect o', then there must be a path from o to o':
 \begin{logic}[Implication 1]
+\hrule\null
+\textsc{\normalsize *Necessary Path Condition 1 (NPC1)}\\
 M,\losigma\ \loturns\ MayAffect$^{Eve}$(o,o') \loimplies\linebreak
-	\ablock MayAccess$^{Ind,Eve}$(o,o') \eqnum[1]
-\end{logic}
-\begin{logic}[Implication 2]
+	\ablock MayAccess$^{Ind,Eve}$(o,o')
+\linebreak \\
 \textit{By contraposition,}\linebreak
 M,\losigma\ \loturns\ \loneg MayAccess$^{Ind,Eve}$(o,o') \loimplies\linebreak
-	\ablock \loneg MayAffect$^{Eve}$(o,o') \eqnum[2]
+	\ablock \loneg MayAffect$^{Eve}$(o,o')
+\linebreak
+\hrule
 \end{logic}
-Here, we immediately see a first defensive outcome of the OCap model. Isolation of object o' from o guarantees that the state of object o' cannot be modified by object o.\\
+Here, we immediately see a first defensive outcome of the OCap model. Having no paths from object o to o' guarantees that the state of object o' cannot be modified by object o.\\
 
-\textbf{Rule: Objects cannot forge capabilities, and only connectivity begets connectivity.}
-Therefore, if an object o' exists in a state \losigma\ and there are no possible paths from o to o' in that state, then it is not possible in any arising state \losigma' from \losigma\ where there is a path from o to o':
-\begin{logic}[Implication 3]
-M,\losigma\ \loturns\ \loneg MayAccess$^{Ind,Now}$(o,o') \loand\ Exists(o')$_{\text{\losigma}}$ \loimplies \linebreak
-\ablock \loneg MayAccess$^{Ind,Eve}$(o,o')\eqnum[3]
+\textbf{Rule 2: Objects cannot forge capabilities, and only connectivity begets connectivity.}
+
+\begin{logic}
+\hrule\null
+\textsc{\normalsize *Global Path Connectivity (GPC)}\
+$[$\loexists Y*\loin{Object}. MayAccess$^{Ind,Now}$(Y*,o')$]$ \loiff\linebreak
+\ablock $[$\loexists X*\loin{Object}. MayAccess$^{Dir,Now}$(X*,o')$]$
+\linebreak\\
+\textit{By contraposition,}\linebreak
+$[$\loforall X*\loin{Object}. \loneg MayAccess$^{Dir,Now}$(X*,o')$]$ \loiff\linebreak
+\ablock $[$\loforall Y*\loin{Object}. \loneg MayAccess$^{Ind,Now}$(Y*,o')$]$\linebreak
+\hrule
 \end{logic}
-We require the enforcement of o' to exist in \losigma\ because o' can refer to an object that does not exist in \losigma\ but will be created and exist in some arising \losigma' from \losigma.
-From \ref{3} and \ref{2}, to protect o' from o in all states arising from a state \losigma, we only need to ensure isolation of o' at \losigma\ :
-\begin{logic}[Implication 4]
-M,\losigma\ \loturns\ \loneg MayAccess$^{Ind,Now}$(o,o') \loand\ Exists(o')$_\text{\losigma}$ \loimplies \linebreak
-	\ablock\loneg MayAffect$^{Eve}(o,o')$\eqnum[4]
+In addition to path being a necessary condition for capability, \textsc{gpc} gives us an additional new relation between the two concepts over the entire system, and says that in a given system of objects, $iff$ there exists an object Y* which has a path to o', then there exists an object X* that has the capability of o'.
+Looking from the left to right direction, this is directly derived from Rule 2, since the path from Y* to o' must either be a direct path (Y*==X*) or if not at the very least an object X* must have a direct path to o', in order for the path from Y* to o' to be well-established, where X* is the last object in the path before o'.
+Moving from the right to left direction, \textsc{gpc} says that if an object X* has a capability then there exist an object Y* with a path to o'.
+We can make this assertion because we know by definition that it holds when Y* refers to the same object as X*.\\
+
+We now look for a relation between a path configuration at a state \losigma\ ($Now$) and eventual path configurations arising from \losigma\ ($Eve$).
+
+\begin{logic}
+\hrule\null
+\textsc{\normalsize *Eventual Path Connectivity (EPC)}\\
+M,\losigma\ \loturns\ 
+MayAccess$^{Ind,Eve}$(o,o')\linebreak
+\ablock \loimplies\linebreak
+\ablock MayAccess$^{Ind,Now}$(o,o')\linebreak
+\ablock \loor\linebreak
+\ablock $\{$\loexists X*\loin Obj.$[($MayAccess$^{Ind,Now}$(o,X*) \loor \linebreak
+\ablock \ablock \ablock \ablock MayAccess$^{Ind,Now}$(X*,o)$)$\linebreak
+\ablock \ablock \ablock \ablock \loand \linebreak
+\ablock \ablock \ablock \ablock MayAccess$^{Ind,Eve}$(X*,o')$]\}$
+\linebreak \\
+\textit{By contraposition,}\linebreak
+M,\losigma\ \loturns\ 
+\loneg MayAccess$^{Ind,Now}$(o,o')\linebreak
+\ablock \loand\linebreak
+\ablock \loforall X*\loin Obj.\loneg$[($MayAccess$^{Ind,Now}$(o,X*) \loor \linebreak
+\ablock \ablock \ablock \ablock MayAccess$^{Ind,Now}$(X*,o)$)$\linebreak
+\ablock \ablock \ablock \ablock \loand \linebreak
+\ablock \ablock \ablock \ablock MayAccess$^{Ind,Eve}$(X*,o')$]$\linebreak
+\ablock \loimplies\linebreak
+MayAccess$^{Ind,Eve}$(o,o')\
+\linebreak
+\hrule
+\end{logic}
+The contraposition result of \textsc{epc} says that o' is eventually path-isolated from o in all possible states arising from \losigma\ when o is does not have a path to o' in \losigma\ \textit{and} for all object X* that has eventual paths to o, it is not true in \losigma\ that either o is connected to X* or X* is connected to o by some path.\\
+
+Using \textsc{npc1} and \textsc{epc}, we derive \textsc{npc2} below which is a relation between state protection in eventual outcomes, and present path configurations.
+\textsc{epc} says that to protect o' from o in all states arising from a state \losigma, we only need to ensure isolation of o' in \losigma, provided o' exists in \losigma :
+
+\begin{logic}
+\hrule\null
+\textsc{\normalsize *Necessary Path Condition 2 (NPC2)}\\
+M,\losigma\ \loturns\ MayAffect$^{Eve}$(o,o')\linebreak
+\ablock \loimplies\linebreak
+\ablock MayAccess$^{Ind,Now}$(o,o')\linebreak
+\ablock \loor\linebreak
+\ablock $\{$\loexists X*\loin Obj.$[($MayAccess$^{Ind,Now}$(o,X*) \loor \linebreak
+\ablock \ablock \ablock \ablock MayAccess$^{Ind,Now}$(X*,o)$)$\linebreak
+\ablock \ablock \ablock \ablock \loand \linebreak
+\ablock \ablock \ablock \ablock MayAccess$^{Ind,Eve}$(X*,o')$]\}$
+\linebreak \\
+\textit{By contraposition,}\linebreak
+M,\losigma\ \loturns\ 
+\loneg MayAccess$^{Ind,Now}$(o,o')\linebreak
+\ablock \loand\linebreak
+\ablock \loforall X*\loin Obj.\loneg$[($MayAccess$^{Ind,Now}$(o,X*) \loor \linebreak
+\ablock \ablock \ablock \ablock MayAccess$^{Ind,Now}$(X*,o)$)$\linebreak
+\ablock \ablock \ablock \ablock \loand \linebreak
+\ablock \ablock \ablock \ablock MayAccess$^{Ind,Eve}$(X*,o')$]$\linebreak
+\ablock \loimplies\linebreak
+	\ablock\loneg MayAffect$^{Eve}(o,o')$\linebreak
+\hrule
+\end{logic}
+The interpretation of \textsc{npc2} is identical to \textsc{epc}, with protection of state being made an explicit consequent from the conditions of \textsc{epc}.
+So far these results serve as a good base to enforce \loneg MayAffect$^{Eve}(o,o')$ but we require \textit{stronger} necessary conditions for \loneg MayAffect$^{Eve}(o,o')$, because enforcing protection of an object's state with just these conditions require the impractical configuration that o is path-isolated from o' on the reference graph (no paths).
+Cooperation between objects imply that there needs to be some path established between the objects for interaction to take place.
+Therefore, we can expect neither \loneg MayAccess$^{Ind,Now}$(o,o') nor \loneg MayAccess$^{Ind,Eve}$(o,o') to hold in a system that allows cooperation between o and o'.\\
+
+So far we have only defined the relation between the existence of paths and an object's state.
+This is not the full picture, as we have not said anything about capabilities, or whether the paths are traverseable.
+We now examine whether it is possible to build a relation from MayAffect$^{Eve}$ to MayAccess$^{Dir,Eve}$.
+\begin{logic}
+\hrule\null
+\textsc{\normalsize *Necessary Execution Condition 1 (NEC1)}\\
+M,\losigma\ \loturns\ MayAffect$^{Eve}$(o,o') \loimplies\linebreak
+	\ablock \loexists X*\loin{Object}. MayAccess$^{Dir,Eve}$(X*,o') 
+\linebreak\\
+\textit{By contraposition and quantifier equivalence,}\linebreak
+M,\losigma\ \loturns\ $[$\loforall X*\loin{Object}. \loneg MayAccess$^{Dir,Eve}$(X*,o')$]$\loimplies\linebreak
+	\ablock \loneg MayAffect$^{Eve}$(o,o') \linebreak
+\hrule
+\end{logic}
+\textsc{nec1} is a stronger version of \textsc{npc1} and says that in order for an object o to modify the state of o', there must exist an object X* that has a direct path to o', and that X* \textit{can} traverse such a path.
+This is derived trivially from our \textsc{pfa} assumption in the beginning.
+Consequently, in our contraposition result, we can deny \textit{all} objects in our system capability of o' to enforce the protection of o' from o.
+\textsc{nec1} does not yield a very useful result, as this is just another way of implementing protection of o' from o through denying all paths to o' in the system reference graph.
+To see why denying all objects the capability of o' is equivalent to denying all paths to o', see the results from \textsc{gpc}.\\
+
+In \textsc{nec1}, if X* refers to the same object as o, then we have a straightforward configuration where object o has the capability of o' and can therefore affect o'.
+However, very crucially, o having the capability of o' is \textit{not} a sufficient condition---denying o the capability of o' does \textit{not} deny o the ability to affect o', since X* can refer to an object that is \textit{not} o.
+\textsc{nec2} below is an expansion of \textsc{nec1} to illustrate this crucial result in the OCap model.
+\begin{logic}
+\hrule\null
+\textsc{\normalsize *Necessary Execution Condition 2 (NEC2)}\\
+M,\losigma\ \loturns\ MayAffect$^{Eve}$(o,o') \loimplies\linebreak
+	\ablock $[$MayAccess$^{Dir,Eve}$(o,o') \loor\linebreak
+	\ablock \loexists X*\loin{Object}, X*\loneq o. MayAccess$^{Dir,Eve}$(X*,o')$]$
+\linebreak\\
+\textit{By contraposition,}\linebreak
+M,\losigma\ \loturns\ $[$\loneg MayAccess$^{Dir,Eve}$(o,o') \loand\linebreak
+	\ablock \loforall X*\loin{Object}, X*\loneq o. \loneg MayAccess$^{Dir,Eve}$(X*,o')$]$ \linebreak
+	\ablock \loimplies\loneg MayAffect$^{Eve}$(o,o')\linebreak
+\hrule
+\end{logic}
+We now examine the second half of the necessary condition in \textsc{nec2}:\\
+$[$\loexists X*\loin{Object}, X*\loneq o. MayAccess$^{Dir,Eve}$(X*,o')$]$
+\\ which means that there exists object X* that is \textit{not} object o, that must have the capability of o'.
+We now make use of our result in \textsc{npc1}, which states that a path must exist from o to o' in order for MayAffect$^{Eve}$(o,o') to hold.
+Since X* is not o, and X* has a direct path to o', we have to connect o to X* in order to connect o to o'.
+With this, we can now construct \textsc{nec3} from \textsc{nec2} and \textsc{npc1}.
+
+\begin{logic}
+\hrule\null
+\textsc{\normalsize *Necessary Execution Condition 3 (NEC3)}\\
+M,\losigma\ \loturns\ MayAffect$^{Eve}$(o,o') \loimplies\linebreak
+	\ablock MayAccess$^{Dir,Eve}$(o,o') \loor \linebreak
+	\ablock $[$\loexists X*\loin Object, X*\loneq o. MayAccess$^{Ind,Eve}$(o,X*) \loand\linebreak
+	\ablock \ablock MayAccess$^{Dir,Eve}$(X*,o')$]$
+	\linebreak\\
+\textit{By contraposition,}\linebreak
+M,\losigma\ \loturns\ $[$\loexists X*\loin Object, X*\loneq o. MayAccess$^{Ind,Eve}$(o,X*) \linebreak
+\ablock \loand MayAccess$^{Dir,Eve}$(X*,o')] \loimplies \linebreak
+		\ablock MayAccess$^{Dir,Eve}$(o,o') \loor \linebreak
+	\ablock 
+\hrule
 \end{logic}
 
-So far these results only serve as a base case, because the protection of an object's state requires the impractical configuration that objects are completely isolated from each other on the reference graph.
-Cooperation between objects imply that there needs to be some path established between the objects for interaction to take place and therefore the predicate \loneg MayAccess$^{Ind,Now}$(o,o') rarely holds in a system of cooperation.\\
+ShuPeng: Questions for myself...
+Can I can form necessary AND sufficient conditions for MayAffect?
 
 The power of OCap patterns hence lies in providing concrete examples of a system of cooperation that allows the existence of paths between objects for cooperation while still dictating the degree of control of one object can have over another.
-For example, we can have variations of following configurations in an Ocap pattern:
-\begin{itemize}\raggedright \small
-\item M,\losigma\ \loturns\ MayAccess$^{Ind,Now}$(o,o') \loand\ Exists(o')$_\text{\losigma}$
-\loimplies \loneg MayAffect$^{Eve}(o,o')$
-\item M,\losigma\ \loturns\ MayAccess$^{Ind,Eve}$(o,o') \linebreak 
-\loimplies \loneg MayAffect$^{Eve}(o,o')$
-\end{itemize}
+The logics we have developed so far illuminate the power of attenuating objects X* that can enable protection.
+Indeed attenuating objects feature prominently in the literature of OCap patterns which we shall see in the next section.
 
-Notice how we have also also focused on the predicate MayAffect$^{Eve}$ rather than MayAffect$^{Now}$.
-Security concerns are often a question of whether what one object can do to another object's in any eventual state of a system.
-If we are concerned with the protection of o' from o, it is not very useful to have a policy where \loneg MayAffect$^{Now}$(o,o') holds while \loneg MayAffect$^{Eve}$(o,o') does not.
-Furthermore, by ensuring \loneg MayAffect$^{Eve}$ holds, we can also ensure \loneg MayAffect$^{Now}$ holds since, by contraposition, \loneg MayAffect$^{Eve}$ \loimplies \loneg MayAffect$^{Now}$.
 \subsection{Pattern 1: The JavaScript DOM Tree}
+We use a JavaScript DOM Tree OCap pattern largely inspired by the example in Devriese et al.\cite{devriese2016} where they use a Kripke worlds framework to reason about the pattern.
 We define the following variables throughout our pattern:
@@
 \end{itemize}
-\textbf{Vulnerability of a Node}:
-The vulnerability of a node lies in the fact that its fields and methods are defined as public. Therefore, any object that has a path and can navigate a path to a node may change the state of a node.
 
-\begin{logic}[DomP1]
+\begin{logic}[NodeVul]
+\hrule\null
+\textsc{\normalsize *Node Vulnerability}\\
 \loforall o, n. MayAccess$^{Dir,Eve}$(o,n)\linebreak \ablock\ablock \loimplies \linebreak
-\ablock \ablock MayAffect$^{Eve}$(o,n)\linebreak
+\ablock \ablock MayAffect$^{Eve}$(o,n) \loand \linebreak
+\ablock \ablock \loforall n'. MayAccess$^{Dir,Eve}$(o,n')
 \end{logic}
-\textsc{Policy 1: Necessary Condition}
+The vulnerability of a node lies in the fact that it contains a public method \texttt{setProperty(key,value)} that will modify an internal mapping data structure.
+A node also has a public field \texttt{parent} that will return the capability of its parent node.
+Consequently, this allows an object which has the capability of any one node in the tree to navigate up to the root node (Document), and consequently navigate to all other nodes in the tree.
+
 \begin{logic}[NodeP1]
+\hrule\null
+\textsc{\normalsize *Policy 1: Necessary Condition}\\
 \loforall n,o, RN\losubseteq ReNode.\linebreak
-Dom(RN,n) \loand\ MayAffect$^{Eve}$(o,n)\linebreak \ablock\loimplies \linebreak
+\ablock Dom(RN,n) \loand\ MayAffect$^{Eve}$(o,n)\linebreak \ablock\loimplies \linebreak
 \ablock \loexists rn\loin RN. MayAccess$^{Ind,Eve}$(o,rn)\linebreak
-\ablock \ablock \ablock\loand\ MayAccess$^{Dir,Eve}$(rn,n)
+\ablock \ablock \:\:\:\: \loand\ MayAccess$^{Dir,Eve}$(rn,n)
+\end{logic}
+\hrule\null
+This policy states that if an Object o may affect the state of a Node n, and that n is dominated(protected) by a set of ReNodes RN, then it implies that a path exists from o to some rn in RN, and that rn has strong access to n.
+From Policy 1, using contraposition, we derive the relation:
+
+\begin{logic}[NodeP1a]
+\loforall n,o, RN\losubseteq ReNode.\linebreak
+\ablock [ \loforall rn\loin RN. \loneg MayAccess$^{Ind,Eve}$(o,rn) \loor \linebreak
+\ablock \loforall rn\loin RN. \loneg MayAccess$^{Dir,Eve}$(rn,n) ]\linebreak
+\ablock \loimplies
+\loneg MayAffect$^{Eve}$(o,n) \loor\ \loneg Dom(RN,n)
 \end{logic}
 
-\begin{logic}[DomP1]
-\loforall o, n. MayAccess$^{Dir,Eve}$(o,n)\linebreak \ablock\ablock \loimplies \linebreak
-\ablock \ablock MayAffect$^{Eve}$(o,n)\linebreak
-\end{logic}
-\textsc{Policy 1: Necessary Condition}
-\begin{logic}[NodeP1]
-\loforall n,o, RN\losubseteq ReNode.\linebreak
-Dom(RN,n) \loand\ MayAffect$^{Eve}$(o,n)\linebreak \ablock\loimplies \linebreak
-\ablock \loexists rn\loin RN. MayAccess$^{Ind,Eve}$(o,rn)\linebreak
-\ablock \ablock \ablock\loand\ MayAccess$^{Dir,Eve}$(rn,n)
+Note that without prescribing any rules to ReNode, we can deny object o the ability to affect node n by enforcing that no path exists from o to any rn in the set RN that dominates n, which by implication is another way of saying isolation from o to n, since:
+\begin{logic}[NodeP1b]
+\loforall n,o. Dom(RN,n) \loand \linebreak \loforall rn\loin RN. \loneg MayAccess$^{Ind,Eve}$(o,rn) \loimplies \linebreak
+\ablock \loneg MayAccess$^{Ind,Eve}$(o,n).
 \end{logic}
 
-\subsection{Pattern 2: Redell's Caretaker}
+Another implication is that we can also deny object o the ability to affect node n by enforcing \loforall rn\loin RN.
+\loneg MayAccess$^{Dir,Eve}$(rn,n).
+This is however, \textit{not} enforceable.
+This is because by definition a ReNode rn holds a field containing the capability of the node n it is meant to protect when the ReNode is constructed, and field that holds the capability of the node is private and constant, that cannot be subsequently changed or removed.
+Consequently, there must exist at least one ReNode rn in the dominating RN set that holds the capability of n:
+\begin{logic}[NodeP1c]
+\loexists rn \loin RN. Dom(RN,n) \loand MayAccess$^{Dir,Eve}$(rn,n)\linebreak
+\ablock \loimplies \linebreak
+\ablock \loneg( \loforall rn\loin RN. \loneg MayAccess$^{Dir,Eve}$(rn,n) )
+\end{logic}
+
+\textsc{Policy 2: Domination of Nodes}
+\begin{logic}[NodeP2]
+\end{logic}
+
+\subsection{Pattern 2: Caretaker}
 \subsection{Pattern 3: Membrane}

### 2017-08-16 article.tex
update

@@
 \newcommand{\lotimes} {$\times$}
-\newcommand{\loneg}{$\neg$}
+\newcommand{\loneg}{$\boldsymbol \neg$}
 \newcommand{\loexec}[2] {$\lfloor$#1$\rfloor _{\text{#2}}$}
@@
 Notice how this o$_1$-o$_2$ relationship differs from the o$_1$-o$_3$ relationship, where o$_1$ cannot guarantee that it can exercise the authority of o$_2$.next which points to o$_3$, since o$_2$.next does not exist within o$_1$'s state---it is possible that o$_2$.next is protected by x$_2$ through encapsulation and data-hiding.\\
+
 Within an OCap model, we can now be convinced that the stronger
@@
 With these simplifications, protection for us then becomes solely a matter of whether we can allow or deny an object to modify the state of another.
-In the subsections below we build the neccessary conditions for MayAffect$^{Eve}$, where MayAffect$^{Eve}$ is placed in the antecedent, and we examine which of the family of MayAccess predicates is placed in the consequent.
-To help us achieve this, we also formalize our assumptions and the rules of OCap, using our predicates, to help guide us in constructing our necessary conditions.\\
+In the following \ref{protection} below we build the neccessary conditions for MayAffect$^{Eve}$, where MayAffect$^{Eve}$ is placed in the antecedent, and we examine which of the family of MayAccess predicates is placed in the consequent.
+To help us achieve this, we also formalize our assumptions and the rules of OCap, using our predicates, to help guide us in constructing our necessary conditions in \ref{ocaprules}.
 
-In building our necessary conditions for state protection in the next subsection notice how we have focused on finding the necessary conditions for the weaker predicate MayAffect$^{Eve}$ rather than MayAffect$^{Now}$ in the antecedent.
-This is because negation on both sides of the implication, would yield a stronger \loneg MayAffect$^{Eve}$ in the consequent.
+\subsection{Formal specifications of OCap Rules}\label{ocaprules}
+
+\textbf{Rule 1: Objects can only interact with each other through sending messages on capabilities.}
+We begin our reasoning of protection in our OCap model, by first re-asserting the necessary but not sufficient path condition of capability, and calling it \textsc{pec}.
+By definition of our predicates, an object having a capability of o' implies object o having a path to o', but not vice-versa:
+\begin{logic}
+\hrule\null
+\textsc{\normalsize *Path-Execution Connectivity (PEC)}\\
+M,\losigma\ \loturns\ MayAccess$^{Dir,Now}$(o,o') \loimplies \linebreak
+\ablock \ablock  MayAccess$^{Ind,Now}$(o,o')
+\linebreak \\
+\textit{By contraposition,}\linebreak
+M,\losigma\ \loturns\ \loneg MayAccess$^{Indir Now}$(o,o')\loimplies \linebreak
+\ablock \ablock \loneg MayAccess$^{Dir,Now}$(o,o')
+\linebreak
+\hrule
+\end{logic}
+Note also that \textsc{pec} is a direct intepretation from Rule 1, since rule 1 says owning a capability is the only way of sending messages to another object, and therefore implies a path between the objects.\\
+
+\textbf{Rule 2: Objects cannot forge capabilities, and only connectivity begets connectivity.}
+\begin{logic}
+\hrule\null
+\textsc{\normalsize *Global Path Existence (GPE)}\\
+M,\losigma\ \loturns\ $[$\loexists Y*\loin{Obj}. MayAccess$^{Ind,Now}$(Y*,o')$]$\linebreak
+\ablock \loiff\linebreak
+\ablock $[$\loexists X*\loin{Obj}. MayAccess$^{Dir,Now}$(X*,o')$]$
+\linebreak\\
+\textit{By contraposition,}\linebreak
+M,\losigma\ \loturns\ $[$\loforall X*\loin{Obj}. \loneg MayAccess$^{Dir,Now}$(X*,o')$]$\linebreak
+\ablock \loiff\linebreak
+\ablock $[$\loforall Y*\loin{Obj}. \loneg MayAccess$^{Ind,Now}$(Y*,o')$]$\linebreak
+\hrule
+\end{logic}
+In addition to path being a necessary condition for capability, \textsc{gpe} gives us an additional new relation between the two concepts over the entire system.
+\textsc{gpe} says that in a given system of objects, $iff$ there exists an object Y* which has a path to o', then there exists an object X* that has the capability of o'.
+Looking from the left to right direction, since the path from Y* to o' exists, then there must exist an object X* with a capability of o' so that the path from Y* to o' is well-established.
+Moving from the right to left direction, \textsc{gpe} says that if an object X* has a capability then there exist an object Y* with a path to o', and this can be proven from \textsc{pec} or by definition when (X*==Y*).\\
+
+However, \textsc{gpe} only postulates the existence of some object with some capability iff there exists some object with some path, and says nothing about how the objects are connected.
+From the connectivity begets connectivity rule, \textsc{pec}, and \textsc{gpe}, we construct global path-execution and execution-execution relationships over the entire system:
+\begin{logic}
+\hrule\null
+\textsc{\normalsize *Global Path-Execution Connect. (GPEC)}\\
+M,\losigma\ \loturns\ MayAccess$^{Ind,Now}$(o,o')\linebreak
+\ablock \loiff\linebreak
+\ablock \loexists X*\loin{Obj}. $[$MayAccess$^{Ind,Now}$(o, X*) \loand\linebreak
+\ablock \ablock \ablock \quad MayAccess$^{Dir,Now}$(X*,o')$]$
+\linebreak\\
+\textit{By contraposition,} \\
+M,\losigma\ \loforall X*\loin{Obj}. \loneg $[$MayAccess$^{Ind,Now}$(o, X*) \loand\linebreak
+\ablock \ablock \ablock \ablock MayAccess$^{Dir,Now}$(X*,o')$]$\linebreak
+\ablock \loiff\linebreak
+\ablock \loneg MayAccess$^{Ind,Now}$(o,o')\linebreak
+\hrule
+\end{logic}
+\textsc{gpec} introduces an intermediate object X* between o and o' that makes explicit the missing connection in \textsc{gpe}.\\
+The power in \textsc{gpec} lies in the recursive predicate MayAccess$^{Ind,Now}$(o, X*) that can eventually be expanded into a set of MayAccess$^{Dir,Now}$ predicates that finishes with terminating predicate MayAccess$^{Ind,Now}$(o, o), where the terminating predicate will always be true.
+We now look for a relation between a path configuration at a state \losigma\ ($Now$) and eventual path configurations arising from \losigma\ ($Eve$).
+\begin{logic}
+\hrule\null
+\textsc{\normalsize *Eventual Path Connectivity 1 (EPC1)}\\
+M,\losigma\ \loturns\ MayAccess$^{Ind,Eve}$(o,o')\linebreak
+\ablock \loimplies\linebreak
+\ablock \loexists X*\loin Obj.$[($MayAccess$^{Ind,Now}$(o,X*) \loor \linebreak
+\ablock \ablock \ablock \ablock MayAccess$^{Dir,Now}$(X*,o)$)$\linebreak
+\ablock \ablock \ablock \ablock \loand \linebreak
+\ablock \ablock \ablock \ablock MayAccess$^{Ind,Eve}$(X*,o')$]$
+\linebreak \\
+\textit{By contraposition,}\linebreak
+M,\losigma\ \loturns\ 
+\loforall X*\loin Obj. $[$\loneg MayAccess$^{Ind,Now}$(o,X*) \loand \linebreak
+\ablock \ablock \ablock \ablock \loneg MayAccess$^{Dir,Now}$(X*,o)$]$\linebreak
+\ablock \ablock \ablock \ablock \loor \linebreak
+\ablock \ablock \ablock \ablock \loneg $[$MayAccess$^{Ind,Eve}$(X*,o')$]$\linebreak
+\ablock \loimplies\linebreak
+	\ablock\loneg MayAccess$^{Ind,Eve}(o,o')$\linebreak
+\hrule
+\end{logic}
+\textsc{epc1} has the meaning that if an object o has an eventual path to o' in some arising state \losigma', then there must exist an object X* that has an eventual path to o', and there must exist a way for o to have a path to X* in \losigma'.
+o either must already have a path to X* in \losigma, or that it will eventually receive the capability of X* in \textsc{epc1} is a representation of connectivity begets connectivity across time:
+
+\begin{itemize}
+\item \textbf{Initial Conditions or Endowment:} o has an existing path to o' in \losigma, therefore X* refers to o'
+\item \textbf{Parenthood:} o will create o' in some arising \losigma', therefore X* refers to o
+\item \textbf{Introduction:} o will only obtain the path to o' in some arising \losigma', therefore X* refers to an object that is not the same object as o (X*\loneq o). 
+\end{itemize}
+Note that for o to have an eventual path to o', we require only o to have an eventual path to X* since we have stated that X* will have an eventual path to o'.
+If o already has a path to X* in \losigma\ then we know o can reach X* in \losigma'.
+If not, the capability of X* must be introduced to o.
+For X* to introduce itself, X* must have the capability of o in \losigma'.
+Well, what if the capability of X* is introduced by some \textit{other} object \~X*?
+Note that in such a case, \~X* will have the capability of X*, and will therefore also be able to eventually have a path to o'.
+Also \~X* must also be able to introduce itself to o.
+There is hence no logical difference between \~X* and X* in our formal description and \~X* might simply be referred to as X*.\\
+
+The contraposition result of \textsc{epc1} says there can be no eventual paths from o to o'. This can be achieved from:
+\begin{itemize}
+\item Global two-way path isolation of o in present state \losigma, where there are no paths leading to o \textit{and} no paths originating from o in \losigma
+\item Global one-way path isolation to o' in all possible states \losigma'\end{itemize}
+There is a final critical result from \textsc{epc1}.
+Notice how, there is a 'recursive' MayAccess$^{Ind,Eve}$(X*,o') in our condition for MayAccess$^{Ind,Eve}$(o,o').
+This allows us to recursively 'expand' the condition to incorporate \textit{all} X* intermediate objects in the path leading to o' that gives us both MayAccess$^{Dir,Now}$ and MayAccess$^{Ind,Now}$ predicates, where the terminating MayAccess$^{Ind,Eve}$(o',o') always returns true.
+This result allows us to define MayAccess$^{Ind,Eve}$(X*,o') in a configuration completely based in \losigma.\\
+What about capabilities?
+Using \textsc{gpec}, we can always expand MayAccess$^{Ind, Now}$(o,X*) into a set of MayAccess$^{Dir, Now}$ predicates that terminate with MayAccess$^{Ind, Now}$(o,o) which is always true.
+The final result we will get is a configuration based purely on MayAccess$^{Dir, Now}$ predicates!!!
+
+\begin{logic}
+\hrule\null
+\textsc{\normalsize *Eventual Path Connectivity 2 (EPC2)}\\
+M,\losigma\ \loturns\ 
+MayAccess$^{Ind,Eve}$(o,o')\linebreak
+\ablock \loimplies \linebreak
+\ablock \loexists X*\loin Obj. $\{[$MayAccess$^{Ind,Now}$(o,X*) \loor \linebreak
+\ablock \ablock \ablock  \ablock MayAccess$^{Dir,Now}$(X*,o)$]$\linebreak
+\ablock \ablock \ablock \ablock \loor \linebreak
+\ablock \ablock \quad \loexists Y*\loin Obj. $[$MayAccess$^{Ind,Eve}$(X*,Y*) \loand\linebreak
+\ablock \ablock \ablock \ablock \ablock MayAccess$^{Dir,Eve}$(Y*,o')$]$\linebreak
+\ablock \ablock \ablock \ablock  $\}$\linebreak
+\linebreak \\
+\textit{By contraposition and De Morgan rules}\linebreak
+M,\losigma\ \loturns\ 
+\loforall X*\loin Obj. $\{[$\loneg MayAccess$^{Ind,Now}$(o,X*) \loand \linebreak
+\ablock \ablock \ablock \ablock \quad \loneg MayAccess$^{Dir,Now}$(X*,o)$]$\linebreak
+\ablock \ablock \ablock \ablock \loor \linebreak
+\ablock \ablock \loneg \loexists Y*\loin Obj. $[$MayAccess$^{Ind,Eve}$(X*,Y*) \loand\linebreak
+\ablock \ablock \ablock \ablock \ablock MayAccess$^{Dir,Eve}$(Y*,o')$]$\linebreak
+\ablock \ablock \ablock \ablock  $\}$\linebreak
+\ablock \loimplies\linebreak
+\ablock \loneg MayAccess$^{Ind,Eve}$(o,o')\
+\linebreak
+\hrule
+\end{logic}
+The results of \textsc{epc2} are identical to \textsc{epc1}, with the second condition of \textsc{epc1} expanded in \textsc{epc2} to say global one-way path isolation to o in all possible states \losigma' to mean there does not exist any eventual path from any object to an object Y* that has the eventual capability of o'.\\
+
+So far we have developed results for paths, but what about execution?
+We construct \textsc{eec} from \textsc{epc1}, the only difference being we now require all those paths in \textsc{epc1} to be traversable.
+
+\begin{logic}
+\hrule\null
+\textsc{\normalsize *Eventual Execution Connectivity (EEC)}\\
+M,\losigma\ \loturns\ MayAccess$^{Dir,Eve}$(o,o')\linebreak
+\ablock \loimplies\linebreak
+\ablock \loexists X*\loin Obj.$[($MayAccess$^{Dir,Now}$(o,X*) \loor \linebreak
+\ablock \ablock \ablock \ablock MayAccess$^{Dir,Now}$(X*,o)$)$\linebreak
+\ablock \ablock \ablock \ablock \loand \linebreak
+\ablock \ablock \ablock \ablock MayAccess$^{Dir,Eve}$(X*,o')$]$
+\linebreak \\
+\textit{By contraposition,}\linebreak
+M,\losigma\ \loturns\ 
+\loforall X*\loin Obj. $\{[$\loneg MayAccess$^{Dir,Now}$(o,X*) \loand \linebreak
+\ablock \ablock \ablock \ablock \quad \loneg MayAccess$^{Dir,Now}$(X*,o)$]$\linebreak
+\ablock \ablock \ablock \ablock \loor \linebreak
+\ablock \ablock \ablock \ablock \loneg $[$MayAccess$^{Dir,Eve}$(X*,o')$]\}$\linebreak
+\ablock \loimplies\linebreak
+	\ablock\loneg MayAccess$^{Dir,Eve}(o,o')$\linebreak
+\hrule
+\end{logic}
+The power of \textsc{eec} is similar to \textsc{epc}\ in that it allows an arbitrary recursive expansion of the MayAccess$^{Dir,Eve}$(X*,o') till the terminating MayAccess$^{Dir,Eve}$(o',o') that is always true.
+
+We now turn our attention formalzing Ocap state changes rules in the system.\\
+
+\textbf{Rule 3: Objects may have private encapsulation of state and behaviour}
+\begin{logic}
+\hrule\null
+\textsc{\normalsize *Global State Change Existence (GSCE)}\\
+M,\losigma\ \loturns\ $[$\loexists X*\loin{Obj}. MayAffect$^{Now}$(X*,o')$]$\linebreak
+\ablock \loimplies \linebreak
+\ablock $[$\loexists Y*\loin{Obj}. MayAccess$^{Dir,Now}$(Y*,o')$]$\linebreak
+\ablock \loimplies \linebreak
+\ablock $[$\loexists Z*\loin{Obj}. MayAccess$^{Ind,Now}$(Z*,o')$]$
+\linebreak\\
+\textit{By contraposition,}\linebreak
+M,\losigma\ \loturns\ $[$\loforall Z*\loin{Obj}. \loneg MayAccess$^{Ind,Now}$(Z*,o')$]$\linebreak
+\ablock \loimplies\linebreak
+\ablock $[$\loforall Y*\loin{Obj}. \loneg MayAccess$^{Dir,Now}$(Y*,o')$]$\linebreak
+\ablock \loimplies\linebreak
+\ablock $[$\loforall X*\loin{Obj}. \loneg MayAffect$^{Dir,Now}$(X*,o')$]$\linebreak
+\hrule
+\end{logic}
+\textsc{gsce} can be derived from our predicate definitions.
+The first implication states that if an object can change the state of o', then there must exist an object in the system that holds the capability of o'.
+The second implication is that there must exist an object in the system that holds a path to o'.
+We view reasoning about state change as reasoning about protection, which we elaborate in the next subsection.
+\subsection{Protection}\label{protection}
+To reason about enforcing protection, we focus on finding the necessary conditions for the weaker predicate MayAffect$^{Eve}$ rather than MayAffect$^{Now}$ in the antecedent.
+This is because negation on both sides of the implication, would yield a stronger \loneg MayAffect$^{Eve}$ in the consequent, which is our area of interest (since it represents state protection).
 In practical terms, if we are concerned with the protection of o' from o, it is also often not very useful to have a policy where \loneg MayAffect$^{Now}$(o,o') holds but \loneg MayAffect$^{Eve}$(o,o') does not.
@@
 
-In building the necessary conditions for the predicate MayAffect$^{Eve}$, we are also more concerned with finding the configuration of MayAccess$^{*,Now}$, rather than MayAccess$^{*,Eve}$.
-This is because it is easier to implement or prove a configuration of relations that holds in \textit{one} specific state and hence much more useful in practice, than think about whether a configuration holds in \textit{all} possible eventual states.\\
-\subsection{Formal Implications}
-\subsubsection{Formalizing OCap rules}
-We begin our reasoning of protection in our OCap model, by introducing the assumption that all fields in our objects can only be declared private.
+Furthermore, in building the necessary conditions for the predicate MayAffect$^{Eve}$, we are also more concerned with finding some structure that contains the configuration of MayAccess$^{*,Now}$, rather than MayAccess$^{*,Eve}$.
+This is because it is much easier to prove a configuration of relations that holds in \textit{one} specific state than think about whether a configuration holds in \textit{all} possible eventual states, which makes the former much easier to implement.\\
+
+To begin, we simplify our reasoning, we introduce the assumption that all fields in our objects can only be declared private.
 Consequently, this implies the necessary condition that an object's state can only be modified if one of it's behaviour is called, and therefore require at least one other object in the system that holds its capability.
@@
 
-This means that in order for an object X*'s state to change, it must be done through some object Y* calling its behaviour (Y* can refer to the same object as X*).
+\textsc{pfa} means that in order for an object X*'s state to change, it must be done through some object Y* calling its behaviour (Y* can refer to the same object as X*).
 Equivalently, denying all objects in the system the ability to call an object's behaviour implies that no object can modify the object's state.\\
 
-\textbf{Rule 1: Objects can only interact with each other through sending messages on capabilities.}
-Therefore, if an object o can affect o', then there must be a path from o to o':
+\subsubsection{Protection through possible paths}
+We now divert our attention to reasoning about the relationship between protection of state (MayAffect) and paths (MayAccess{Ind,*}).
+The question we ask is what is the relation between state protection in eventual outcomes, and present path configurations?
+We build this relationship from the basics by progressively finding stronger conditions of MayAffect.
+We begin with the first condition, which says that for an object to affect another, an \textit{eventual} path has to exist.
+This can be derived from the definitions of our predicates.
 \begin{logic}[Implication 1]
@@
 \end{logic}
-Here, we immediately see a first defensive outcome of the OCap model. Having no paths from object o to o' guarantees that the state of object o' cannot be modified by object o.\\
+Here, we immediately see a first defensive outcome of the OCap model. Having no eventual paths from object o to o' guarantees that the state of object o' cannot be modified by object o.\\
 
-\textbf{Rule 2: Objects cannot forge capabilities, and only connectivity begets connectivity.}
+So far our results serve as a good base to enforce \loneg MayAffect$^{Eve}(o,o')$ but we require \textit{stronger} necessary conditions, because enforcing protection of an object's state with just these conditions require the impractical configuration that o is eventually path-isolated from o' on the reference graph.
+Cooperation between objects imply that there needs to be some path established between the objects for interaction to take place.
+Therefore, we can expect neither \loneg MayAccess$^{Ind,Now}$(o,o') nor \loneg MayAccess$^{Ind,Eve}$(o,o') to hold in a system that allows cooperation between o and o'.\\
 
-\begin{logic}
-\hrule\null
-\textsc{\normalsize *Global Path Connectivity (GPC)}\
-$[$\loexists Y*\loin{Object}. MayAccess$^{Ind,Now}$(Y*,o')$]$ \loiff\linebreak
-\ablock $[$\loexists X*\loin{Object}. MayAccess$^{Dir,Now}$(X*,o')$]$
-\linebreak\\
-\textit{By contraposition,}\linebreak
-$[$\loforall X*\loin{Object}. \loneg MayAccess$^{Dir,Now}$(X*,o')$]$ \loiff\linebreak
-\ablock $[$\loforall Y*\loin{Object}. \loneg MayAccess$^{Ind,Now}$(Y*,o')$]$\linebreak
-\hrule
-\end{logic}
-In addition to path being a necessary condition for capability, \textsc{gpc} gives us an additional new relation between the two concepts over the entire system, and says that in a given system of objects, $iff$ there exists an object Y* which has a path to o', then there exists an object X* that has the capability of o'.
-Looking from the left to right direction, this is directly derived from Rule 2, since the path from Y* to o' must either be a direct path (Y*==X*) or if not at the very least an object X* must have a direct path to o', in order for the path from Y* to o' to be well-established, where X* is the last object in the path before o'.
-Moving from the right to left direction, \textsc{gpc} says that if an object X* has a capability then there exist an object Y* with a path to o'.
-We can make this assertion because we know by definition that it holds when Y* refers to the same object as X*.\\
-
-We now look for a relation between a path configuration at a state \losigma\ ($Now$) and eventual path configurations arising from \losigma\ ($Eve$).
-
-\begin{logic}
-\hrule\null
-\textsc{\normalsize *Eventual Path Connectivity (EPC)}\\
-M,\losigma\ \loturns\ 
-MayAccess$^{Ind,Eve}$(o,o')\linebreak
-\ablock \loimplies\linebreak
-\ablock MayAccess$^{Ind,Now}$(o,o')\linebreak
-\ablock \loor\linebreak
-\ablock $\{$\loexists X*\loin Obj.$[($MayAccess$^{Ind,Now}$(o,X*) \loor \linebreak
-\ablock \ablock \ablock \ablock MayAccess$^{Ind,Now}$(X*,o)$)$\linebreak
-\ablock \ablock \ablock \ablock \loand \linebreak
-\ablock \ablock \ablock \ablock MayAccess$^{Ind,Eve}$(X*,o')$]\}$
-\linebreak \\
-\textit{By contraposition,}\linebreak
-M,\losigma\ \loturns\ 
-\loneg MayAccess$^{Ind,Now}$(o,o')\linebreak
-\ablock \loand\linebreak
-\ablock \loforall X*\loin Obj.\loneg$[($MayAccess$^{Ind,Now}$(o,X*) \loor \linebreak
-\ablock \ablock \ablock \ablock MayAccess$^{Ind,Now}$(X*,o)$)$\linebreak
-\ablock \ablock \ablock \ablock \loand \linebreak
-\ablock \ablock \ablock \ablock MayAccess$^{Ind,Eve}$(X*,o')$]$\linebreak
-\ablock \loimplies\linebreak
-MayAccess$^{Ind,Eve}$(o,o')\
-\linebreak
-\hrule
-\end{logic}
-The contraposition result of \textsc{epc} says that o' is eventually path-isolated from o in all possible states arising from \losigma\ when o is does not have a path to o' in \losigma\ \textit{and} for all object X* that has eventual paths to o, it is not true in \losigma\ that either o is connected to X* or X* is connected to o by some path.\\
-
-Using \textsc{npc1} and \textsc{epc}, we derive \textsc{npc2} below which is a relation between state protection in eventual outcomes, and present path configurations.
-\textsc{epc} says that to protect o' from o in all states arising from a state \losigma, we only need to ensure isolation of o' in \losigma, provided o' exists in \losigma :
-
+What about present path configurations?
+Using \textsc{npc1} and the results of \textsc{epc1}, we derive \textsc{npc2} below which is a relation between protection of o' in eventual states, and \textit{present} path configurations between o and o'.
 \begin{logic}
@@
 \ablock \loimplies\linebreak
-\ablock MayAccess$^{Ind,Now}$(o,o')\linebreak
-\ablock \loor\linebreak
-\ablock $\{$\loexists X*\loin Obj.$[($MayAccess$^{Ind,Now}$(o,X*) \loor \linebreak
-\ablock \ablock \ablock \ablock MayAccess$^{Ind,Now}$(X*,o)$)$\linebreak
+\ablock \loexists X*\loin Obj.$[($MayAccess$^{Ind,Now}$(o,X*) \loor \linebreak
+\ablock \ablock \ablock \ablock MayAccess$^{Dir,Now}$(X*,o)$)$\linebreak
 \ablock \ablock \ablock \ablock \loand \linebreak
-\ablock \ablock \ablock \ablock MayAccess$^{Ind,Eve}$(X*,o')$]\}$
+\ablock \ablock \ablock \ablock MayAccess$^{Ind,Eve}$(X*,o')$]$
 \linebreak \\
-\textit{By contraposition,}\linebreak
+\textit{By contraposition and De Morgan rules,}\linebreak
 M,\losigma\ \loturns\ 
-\loneg MayAccess$^{Ind,Now}$(o,o')\linebreak
-\ablock \loand\linebreak
-\ablock \loforall X*\loin Obj.\loneg$[($MayAccess$^{Ind,Now}$(o,X*) \loor \linebreak
-\ablock \ablock \ablock \ablock MayAccess$^{Ind,Now}$(X*,o)$)$\linebreak
-\ablock \ablock \ablock \ablock \loand \linebreak
-\ablock \ablock \ablock \ablock MayAccess$^{Ind,Eve}$(X*,o')$]$\linebreak
-\ablock \loimplies\linebreak
+\loforall X*\loin Obj. $[$\loneg MayAccess$^{Ind,Now}$(o,X*) \loand \linebreak
+\ablock \ablock \ablock \ablock \loneg MayAccess$^{Dir,Now}$(X*,o)$]$\linebreak
+\ablock \ablock \ablock \ablock \loor \linebreak
+\ablock \ablock \ablock \ablock \loneg MayAccess$^{Ind,Eve}$(X*,o')$]$\linebreak
+\ablock \loimplies \linebreak
 	\ablock\loneg MayAffect$^{Eve}(o,o')$\linebreak
@@
 \end{logic}
-The interpretation of \textsc{npc2} is identical to \textsc{epc}, with protection of state being made an explicit consequent from the conditions of \textsc{epc}.
-So far these results serve as a good base to enforce \loneg MayAffect$^{Eve}(o,o')$ but we require \textit{stronger} necessary conditions for \loneg MayAffect$^{Eve}(o,o')$, because enforcing protection of an object's state with just these conditions require the impractical configuration that o is path-isolated from o' on the reference graph (no paths).
-Cooperation between objects imply that there needs to be some path established between the objects for interaction to take place.
-Therefore, we can expect neither \loneg MayAccess$^{Ind,Now}$(o,o') nor \loneg MayAccess$^{Ind,Eve}$(o,o') to hold in a system that allows cooperation between o and o'.\\
-
-So far we have only defined the relation between the existence of paths and an object's state.
-This is not the full picture, as we have not said anything about capabilities, or whether the paths are traverseable.
-We now examine whether it is possible to build a relation from MayAffect$^{Eve}$ to MayAccess$^{Dir,Eve}$.
+The interpretation of \textsc{npc2} is identical to \textsc{epc1}, with protection of state being made an explicit consequent from the conditions of \textsc{epc1}.\\
+We now examine whether it is possible to build a relation between MayAffect$^{Eve}$ and MayAccess$^{Dir,Eve}$.
+Similarly, \textsc{npc2} can be augmented by \textsc{epc2} which results in \textsc{npc3}:
+\begin{logic}
+\hrule\null
+\textsc{\normalsize *Necessary Path Condition 3 (NPC3)}\\
+M,\losigma\ \loturns\ 
+MayAffect$^{Eve}$(o,o')\linebreak
+\ablock \loimplies \linebreak
+\ablock \loexists X*\loin Obj. $\{[$MayAccess$^{Ind,Now}$(o,X*) \loor \linebreak
+\ablock \ablock \ablock  \ablock MayAccess$^{Dir,Now}$(X*,o)$]$\linebreak
+\ablock \ablock \ablock \ablock \loor \linebreak
+\ablock \ablock \quad \loexists Y*\loin Obj. $[$MayAccess$^{Ind,Eve}$(X*,Y*) \loand\linebreak
+\ablock \ablock \ablock \ablock \ablock MayAccess$^{Dir,Eve}$(Y*,o')$]$\linebreak
+\ablock \ablock \ablock \ablock  $\}$\linebreak
+\linebreak \\
+\textit{By contraposition and De Morgan rules}\linebreak
+M,\losigma\ \loturns\ 
+\loforall X*\loin Obj. $\{[$\loneg MayAccess$^{Ind,Now}$(o,X*) \loand \linebreak
+\ablock \ablock \ablock \ablock \quad \loneg MayAccess$^{Dir,Now}$(X*,o)$]$\linebreak
+\ablock \ablock \ablock \ablock \loor \linebreak
+\ablock \ablock \loneg \loexists Y*\loin Obj. $[$MayAccess$^{Ind,Eve}$(X*,Y*) \loand\linebreak
+\ablock \ablock \ablock \ablock \ablock MayAccess$^{Dir,Eve}$(Y*,o')$]$\linebreak
+\ablock \ablock \ablock \ablock  $\}$\linebreak
+\ablock \loimplies\linebreak
+\ablock \loneg MayAffect$^{Eve}$(o,o')\
+\linebreak
+\hrule
+\end{logic}
+We are at the finishing line. The Ind,Now predicates can be expanded using GPEC to obtain direct capabilities. Ind,Eve can be expanded using EPC2. Dir,Eve can be expanded using EEC.
+\subsubsection{A note on execution and protection}
+\begin{logic}
+M,\losigma\ \loturns\ MayAffect$^{Eve}$(o,o') \lonimplies 
+MayAccess$^{Dir,Eve}$(o,o')\linebreak
+\\
+M,\losigma\ \loturns\ MayAffect$^{Eve}$(o,o') \loimplies 
+MayAccess$^{Ind,Eve}$(o,o')
+\end{logic}
 \begin{logic}
@@
 M,\losigma\ \loturns\ MayAffect$^{Eve}$(o,o') \loimplies\linebreak
-	\ablock \loexists X*\loin{Object}. MayAccess$^{Dir,Eve}$(X*,o') 
+	\ablock \loexists X*\loin{Obj}. MayAccess$^{Dir,Eve}$(X*,o') 
 \linebreak\\
 \textit{By contraposition and quantifier equivalence,}\linebreak
-M,\losigma\ \loturns\ $[$\loforall X*\loin{Object}. \loneg MayAccess$^{Dir,Eve}$(X*,o')$]$\loimplies\linebreak
+M,\losigma\ \loturns\ $[$\loforall X*\loin{Obj}. \loneg MayAccess$^{Dir,Eve}$(X*,o')$]$\loimplies\linebreak
 	\ablock \loneg MayAffect$^{Eve}$(o,o') \linebreak
@@
 \textsc{nec1} does not yield a very useful result, as this is just another way of implementing protection of o' from o through denying all paths to o' in the system reference graph.
-To see why denying all objects the capability of o' is equivalent to denying all paths to o', see the results from \textsc{gpc}.\\
+To see why denying all objects the capability of o' is equivalent to denying all paths to o', see the contraposition result from \textsc{gpe}.\\
 
@@
 However, very crucially, o having the capability of o' is \textit{not} a sufficient condition---denying o the capability of o' does \textit{not} deny o the ability to affect o', since X* can refer to an object that is \textit{not} o.
-\textsc{nec2} below is an expansion of \textsc{nec1} to illustrate this crucial result in the OCap model.
+\textsc{nec2} below is an expansion of \textsc{nec1} to illustrate this crucial point.
 \begin{logic}
@@
 	\ablock $[$MayAccess$^{Dir,Eve}$(o,o') \loor\linebreak
-	\ablock \loexists X*\loin{Object}, X*\loneq o. MayAccess$^{Dir,Eve}$(X*,o')$]$
+	\ablock \loexists X*\loin{Obj}, X*\loneq o. MayAccess$^{Dir,Eve}$(X*,o')$]$
 \linebreak\\
@@
 M,\losigma\ \loturns\ $[$\loneg MayAccess$^{Dir,Eve}$(o,o') \loand\linebreak
-	\ablock \loforall X*\loin{Object}, X*\loneq o. \loneg MayAccess$^{Dir,Eve}$(X*,o')$]$ \linebreak
+	\ablock \loforall X*\loin{Obj}, X*\loneq o. \loneg MayAccess$^{Dir,Eve}$(X*,o')$]$ \linebreak
 	\ablock \loimplies\loneg MayAffect$^{Eve}$(o,o')\linebreak
@@
 \\ which means that there exists object X* that is \textit{not} object o, that must have the capability of o'.
-We now make use of our result in \textsc{npc1}, which states that a path must exist from o to o' in order for MayAffect$^{Eve}$(o,o') to hold.
-Since X* is not o, and X* has a direct path to o', we have to connect o to X* in order to connect o to o'.
-With this, we can now construct \textsc{nec3} from \textsc{nec2} and \textsc{npc1}.
+But this does not say anything about o and X*.
+We need to connect o to X*, because \textsc{npc1} says MayAffect$^{Eve}$(o,o') implies that o must have some indirect path to o'.
 
@@
 \textsc{\normalsize *Necessary Execution Condition 3 (NEC3)}\\
-M,\losigma\ \loturns\ MayAffect$^{Eve}$(o,o') \loimplies\linebreak
+M,\losigma\ \loturns\ MayAffect$^{Eve}$(o,o')\linebreak
+	\ablock \loimplies\linebreak
 	\ablock MayAccess$^{Dir,Eve}$(o,o') \loor \linebreak
-	\ablock $[$\loexists X*\loin Object, X*\loneq o. MayAccess$^{Ind,Eve}$(o,X*) \loand\linebreak
+	\ablock $[$\loexists X*\loin Obj, X*\loneq o. MayAccess$^{Dir,Eve}$(o,X*) \loand\linebreak
 	\ablock \ablock MayAccess$^{Dir,Eve}$(X*,o')$]$
@@
 \textit{By contraposition,}\linebreak
-M,\losigma\ \loturns\ $[$\loexists X*\loin Object, X*\loneq o. MayAccess$^{Ind,Eve}$(o,X*) \linebreak
-\ablock \loand MayAccess$^{Dir,Eve}$(X*,o')] \loimplies \linebreak
-		\ablock MayAccess$^{Dir,Eve}$(o,o') \loor \linebreak
-	\ablock 
+M,\losigma\ \loturns\ \loneg MayAccess$^{Dir,Eve}$(o,o') \loand \linebreak
+	\ablock \loneg$[$\loexists X*\loin Obj, X*\loneq o. MayAccess$^{Dir,Eve}$(o,X*) \loand\linebreak
+	\ablock \ablock MayAccess$^{Dir,Eve}$(X*,o')$]$\linebreak
+	\ablock \loimplies\linebreak
+	\ablock \loneg MayAffect$^{Eve}$(o,o')
+	\linebreak
+\hrule
+\end{logic}
+\textsc{nec3} is not very useful because we do not have any knowledge of how the system's path or capability configuration should look like at \losigma\ for us to enable protection of o' in all possible states arising from \losigma.
+We now incorporate our results of \textsc{eec} to expand MayAccess$^{Dir,Eve}$ that will give us the present configuration of paths and capabilities.\\
+
+\begin{logic}
+\hrule\null
+\textsc{\normalsize *Necessary Execution Condition 4 (NEC4)}\\
+M,\losigma\ \loturns\ MayAffect$^{Eve}$(o,o') \loimplies\linebreak
+\ablock \loexists X*\loin Obj.$[($MayAccess$^{Dir,Now}$(o,Y*) \loor \linebreak
+\ablock \ablock \ablock \ablock MayAccess$^{Dir,Now}$(Y*,o)$)$\linebreak
+\ablock \ablock \ablock \ablock \loand \linebreak
+\ablock \ablock \ablock \ablock MayAccess$^{Dir,Eve}$(Y*,o')$]$\linebreak
+\ablock \loor \linebreak
+\ablock $\{$\loexists X*\loin Obj, X*\loneq o.\linebreak
+\ablock \quad \loexists X$^1$\loin Obj.$[($MayAccess$^{Dir,Now}$(o,X$^1$) \loor \linebreak
+\ablock \ablock \ablock \ablock MayAccess$^{Dir,Now}$(X$^1$,o)$)$\linebreak
+\ablock \ablock \ablock \ablock \loand \linebreak
+\ablock \ablock \ablock \ablock MayAccess$^{Dir,Eve}$(X$^1$,X*)$]$\linebreak
+\ablock \quad \loand \linebreak
+\ablock \quad \loexists X$^2$\loin Obj.$[($MayAccess$^{Dir,Now}$(X*,X$^2$) \loor \linebreak
+\ablock \ablock \ablock \ablock MayAccess$^{Dir,Now}$(X$^2$,X*)$)$\linebreak
+\ablock \ablock \ablock \ablock \loand \linebreak
+\ablock \ablock \ablock \ablock MayAccess$^{Dir,Eve}$(X$^2$,o')$]$\linebreak
+\ablock $\}$
+	\linebreak\\
+\textit{By contraposition,}\linebreak
+...\linebreak
+\\
 \hrule
@@
 
-ShuPeng: Questions for myself...
-Can I can form necessary AND sufficient conditions for MayAffect?
-
-The power of OCap patterns hence lies in providing concrete examples of a system of cooperation that allows the existence of paths between objects for cooperation while still dictating the degree of control of one object can have over another.
+From \textsc{nec4} we see the power of OCap patterns hence lies in providing concrete examples of a system of cooperation that allows the existence of paths between objects for cooperation while still dictating the degree of control of one object can have over another.
 The logics we have developed so far illuminate the power of attenuating objects X* that can enable protection.
@@
 \textsc{\normalsize *Node Vulnerability}\\
-\loforall o, n. MayAccess$^{Dir,Eve}$(o,n)\linebreak \ablock\ablock \loimplies \linebreak
-\ablock \ablock MayAffect$^{Eve}$(o,n) \loand \linebreak
-\ablock \ablock \loforall n'. MayAccess$^{Dir,Eve}$(o,n')
+\loforall o, n. MayAccess$^{Dir,Now}$(o,n)\linebreak \ablock\ablock \loimplies \linebreak
+\ablock \ablock MayAffect$^{Now}$(o,n) \loand \linebreak
+\ablock \ablock \loforall n'. MayAccess$^{Dir,Now}$(o,n')
 \end{logic}
@@
 \ablock Dom(RN,n) \loand\ MayAffect$^{Eve}$(o,n)\linebreak \ablock\loimplies \linebreak
-\ablock \loexists rn\loin RN. MayAccess$^{Ind,Eve}$(o,rn)\linebreak
-\ablock \ablock \:\:\:\: \loand\ MayAccess$^{Dir,Eve}$(rn,n)
+\ablock \loexists rn\loin RN. MayAccess$^{Dir,Now}$(o,rn)\linebreak
+\ablock \ablock \:\:\:\: \loand\ MayAccess$^{Dir,Now}$(rn,n)
 \end{logic}

### 2017-08-16 article.tex
update

@@
 \usepackage{enumitem} % Customized lists
-
+\usepackage{color}
 \setlist[itemize]{noitemsep} % Make itemize lists more compact
@@
 %environment formating
-\newenvironment{logic}[1][null]
-{\begin{flushleft} \small \label{#1}}
+\newenvironment{logic}[1][]
+{\begin{flushleft} \small }
 {\end{flushleft}}
@@
 With these simplifications, protection for us then becomes solely a matter of whether we can allow or deny an object to modify the state of another.
+To help us reason about protection, we first formalize our assumptions and the rules of OCap, using our predicates, to help guide us in constructing our necessary conditions in \ref{ocaprules}.
 In the following \ref{protection} below we build the neccessary conditions for MayAffect$^{Eve}$, where MayAffect$^{Eve}$ is placed in the antecedent, and we examine which of the family of MayAccess predicates is placed in the consequent.
-To help us achieve this, we also formalize our assumptions and the rules of OCap, using our predicates, to help guide us in constructing our necessary conditions in \ref{ocaprules}.
 
@@
 \textit{By contraposition,}\linebreak
-M,\losigma\ \loturns\ \loneg MayAccess$^{Indir Now}$(o,o')\loimplies \linebreak
+M,\losigma\ \loturns\ \loneg MayAccess$^{Ind,Now}$(o,o')\loimplies \linebreak
 \ablock \ablock \loneg MayAccess$^{Dir,Now}$(o,o')
@@
 Moving from the right to left direction, \textsc{gpe} says that if an object X* has a capability then there exist an object Y* with a path to o', and this can be proven from \textsc{pec} or by definition when (X*==Y*).\\
+The contraposition result of \textsc{gpe} says that we can prevent all paths to o' through one of these conditions:
+\begin{itemize}
+\item Deny all direct paths leading into o' 
+\end{itemize}
 
@@
 \textit{By contraposition,} \\
-M,\losigma\ \loforall X*\loin{Obj}. \loneg $[$MayAccess$^{Ind,Now}$(o, X*) \loand\linebreak
-\ablock \ablock \ablock \ablock MayAccess$^{Dir,Now}$(X*,o')$]$\linebreak
+M,\losigma\ \loforall X*\loin{Obj}. $[$\loneg MayAccess$^{Ind,Now}$(o, X*) \loor\linebreak
+\ablock \ablock \ablock \quad \loneg MayAccess$^{Dir,Now}$(X*,o')$]$\linebreak
 \ablock \loiff\linebreak
@@
 \textsc{gpec} introduces an intermediate object X* between o and o' that makes explicit the missing connection in \textsc{gpe}.\\
-The power in \textsc{gpec} lies in the recursive predicate MayAccess$^{Ind,Now}$(o, X*) that can eventually be expanded into a set of MayAccess$^{Dir,Now}$ predicates that finishes with terminating predicate MayAccess$^{Ind,Now}$(o, o), where the terminating predicate will always be true.
-We now look for a relation between a path configuration at a state \losigma\ ($Now$) and eventual path configurations arising from \losigma\ ($Eve$).
+The contraposition result of \textsc{gpec} says that we can prevent all paths from o to o' through one of these conditions:
+
+The contraposition result of \textsc{gpec} says that for there to be no paths from o to o' these configurations must hold:
+\begin{itemize}\setlength\itemsep{0.5em}
+\item if there is some object x that o has a path to in \losigma, then x cannot have a path to o' in \losigma.
+\item if there is some object x that has a path to o' in \losigma, then o cannot have a path to x in \losigma.
+\end{itemize}
+
+The power in \textsc{gpec} lies in the recursive predicate MayAccess$^{Ind,Now}$(o, X*) that can eventually be expanded into a set of MayAccess$^{Dir,Now}$ predicates that finishes with terminating predicate MayAccess$^{Ind,Now}$(o, o), where the terminating predicate will always be true, unless the object o does not exist in \losigma.
+We now look for a relation between a path configuration at a state \losigma\ ($Now$) and eventual path configurations in states \losigma' arising from \losigma\ ($Eve$).\vfill
+\begin{minipage}[b]{\linewidth}
 \begin{logic}
@@
 \ablock \loimplies\linebreak
-\ablock \loexists X*\loin Obj.$[($MayAccess$^{Ind,Now}$(o,X*) \loor \linebreak
-\ablock \ablock \ablock \ablock MayAccess$^{Dir,Now}$(X*,o)$)$\linebreak
-\ablock \ablock \ablock \ablock \loand \linebreak
-\ablock \ablock \ablock \ablock MayAccess$^{Ind,Eve}$(X*,o')$]$
+\ablock \loexists X*\loin Obj. $\{[$MayAccess$^{Ind,Now}$(o,X*)\linebreak
+\ablock \ablock \ablock \ablock\loor \linebreak
+\ablock \ablock \ablock \quad\: MayAccess$^{Dir,Now}$(X*,o)$]$\linebreak
+\ablock \ablock \ablock \quad \loand \linebreak
+\ablock \ablock \ablock \quad MayAccess$^{Ind,Eve}$(X*,o')$\}$
 \linebreak \\
 \textit{By contraposition,}\linebreak
-M,\losigma\ \loturns\ 
-\loforall X*\loin Obj. $[$\loneg MayAccess$^{Ind,Now}$(o,X*) \loand \linebreak
-\ablock \ablock \ablock \ablock \loneg MayAccess$^{Dir,Now}$(X*,o)$]$\linebreak
-\ablock \ablock \ablock \ablock \loor \linebreak
-\ablock \ablock \ablock \ablock \loneg $[$MayAccess$^{Ind,Eve}$(X*,o')$]$\linebreak
+M,\losigma\ \loturns\ \linebreak
+\ablock \loforall X*\loin Obj. $\{[$\loneg MayAccess$^{Ind,Now}$(o,X*)\linebreak
+\ablock \ablock \ablock \ablock \quad \loand \linebreak
+\ablock \ablock \ablock \quad\: \loneg MayAccess$^{Dir,Now}$(X*,o)$]$\linebreak
+\ablock \ablock \ablock \quad \loor \linebreak
+\ablock \ablock \ablock \quad \loneg MayAccess$^{Ind,Eve}$(X*,o')$\}$\linebreak
 \ablock \loimplies\linebreak
@@
 \end{logic}
+\end{minipage}
 \textsc{epc1} has the meaning that if an object o has an eventual path to o' in some arising state \losigma', then there must exist an object X* that has an eventual path to o', and there must exist a way for o to have a path to X* in \losigma'.
-o either must already have a path to X* in \losigma, or that it will eventually receive the capability of X* in \textsc{epc1} is a representation of connectivity begets connectivity across time:
+o either must already have a path to X* in \losigma, or that it will eventually receive the capability of X* in \losigma'.
+\textsc{epc1} is actually a formal representation of connectivity begets connectivity across time:
 
@@
 \end{itemize}
+
 Note that for o to have an eventual path to o', we require only o to have an eventual path to X* since we have stated that X* will have an eventual path to o'.
@@
 If not, the capability of X* must be introduced to o.
-For X* to introduce itself, X* must have the capability of o in \losigma'.
+For X* to introduce itself, X* must have the capability of o in \losigma', and the only way we can guarantee is to have X* possess the capability of o in \losigma.
 Well, what if the capability of X* is introduced by some \textit{other} object \~X*?
@@
 
-The contraposition result of \textsc{epc1} says there can be no eventual paths from o to o'. This can be achieved from:
-\begin{itemize}
-\item Global two-way path isolation of o in present state \losigma, where there are no paths leading to o \textit{and} no paths originating from o in \losigma
-\item Global one-way path isolation to o' in all possible states \losigma'\end{itemize}
-There is a final critical result from \textsc{epc1}.
+The contraposition result of \textsc{epc1} says that for there to be no eventual paths from o to o' these configurations must hold:
+\begin{itemize}\setlength\itemsep{0.5em}
+\item if there is some object x that o has a path to in \losigma, then x cannot have an eventual path to o'
+\item if there is some object x that possesses the capability of o in \losigma, then x cannot have an eventual path to o'
+\end{itemize}
+Note that object x can refer to the same object as o.
+There is one final critical result from \textsc{epc1}.
 Notice how, there is a 'recursive' MayAccess$^{Ind,Eve}$(X*,o') in our condition for MayAccess$^{Ind,Eve}$(o,o').
-This allows us to recursively 'expand' the condition to incorporate \textit{all} X* intermediate objects in the path leading to o' that gives us both MayAccess$^{Dir,Now}$ and MayAccess$^{Ind,Now}$ predicates, where the terminating MayAccess$^{Ind,Eve}$(o',o') always returns true.
-This result allows us to define MayAccess$^{Ind,Eve}$(X*,o') in a configuration completely based in \losigma.\\
+This allows us to recursively expand the condition to incorporate \textit{all} X* intermediate objects in the path leading to o'.
+This recursive expansion gives us both MayAccess$^{Dir,Now}$ and MayAccess$^{Ind,Now}$ predicates, where the terminating MayAccess$^{Ind,Eve}$(o',o') always returns true if o' exists.
+This result allows us to define MayAccess$^{Ind,Eve}$(X*,o') in a configuration of paths completely based in \losigma.\\
+
 What about capabilities?
-Using \textsc{gpec}, we can always expand MayAccess$^{Ind, Now}$(o,X*) into a set of MayAccess$^{Dir, Now}$ predicates that terminate with MayAccess$^{Ind, Now}$(o,o) which is always true.
-The final result we will get is a configuration based purely on MayAccess$^{Dir, Now}$ predicates!!!
+Using \textsc{gpec}, we can always expand MayAccess$^{Ind, Now}$(o,X*) into a set of MayAccess$^{Dir, Now}$ predicates that terminate with MayAccess$^{Ind, Now}$(o,o) which is always true if o exists.
+The final result we will get is a configuration based purely on MayAccess$^{Dir, Now}$ predicates.
+We expand \textsc{epc1} into \textsc{epc2} to illustrate:
 
@@
 \textsc{\normalsize *Eventual Path Connectivity 2 (EPC2)}\\
-M,\losigma\ \loturns\ 
-MayAccess$^{Ind,Eve}$(o,o')\linebreak
-\ablock \loimplies \linebreak
-\ablock \loexists X*\loin Obj. $\{[$MayAccess$^{Ind,Now}$(o,X*) \loor \linebreak
-\ablock \ablock \ablock  \ablock MayAccess$^{Dir,Now}$(X*,o)$]$\linebreak
-\ablock \ablock \ablock \ablock \loor \linebreak
-\ablock \ablock \quad \loexists Y*\loin Obj. $[$MayAccess$^{Ind,Eve}$(X*,Y*) \loand\linebreak
-\ablock \ablock \ablock \ablock \ablock MayAccess$^{Dir,Eve}$(Y*,o')$]$\linebreak
-\ablock \ablock \ablock \ablock  $\}$\linebreak
+M,\losigma\ \loturns\ MayAccess$^{Ind,Eve}$(o,o')\linebreak
+\ablock \loimplies\linebreak
+\ablock \loexists X*\loin Obj.$\{$\linebreak
+\ablock \ablock $[$\null
+\textcolor{red}{\loexists X$^1$\loin{Obj}. MayAccess$^{Ind,Now}$(o, X$^1$) \loand\linebreak
+\ablock \ablock \ablock \ablock \quad MayAccess$^{Dir,Now}$(X$^1$,X*)\linebreak
+}
+\ablock \ablock \quad \loor \linebreak
+\ablock \ablock MayAccess$^{Dir,Now}$(X*,o)$ ]$\linebreak
+\ablock \ablock \loand \linebreak
+\textcolor{blue}{
+\ablock \ablock \loexists Y*\loin Obj.$\{[$MayAccess$^{Ind,Now}$(X*,Y*)\linebreak
+\ablock \ablock \ablock \ablock \quad \quad \loor \linebreak
+\ablock \ablock \ablock \ablock \quad MayAccess$^{Dir,Now}$(Y*,X*)$]$\linebreak
+\ablock \ablock \ablock \ablock \quad \loand \linebreak
+\ablock \ablock \ablock \ablock \quad MayAccess$^{Ind,Eve}$(Y*,o')$\}$
+}\linebreak
+\ablock \ablock $\}$
 \linebreak \\
-\textit{By contraposition and De Morgan rules}\linebreak
-M,\losigma\ \loturns\ 
-\loforall X*\loin Obj. $\{[$\loneg MayAccess$^{Ind,Now}$(o,X*) \loand \linebreak
-\ablock \ablock \ablock \ablock \quad \loneg MayAccess$^{Dir,Now}$(X*,o)$]$\linebreak
-\ablock \ablock \ablock \ablock \loor \linebreak
-\ablock \ablock \loneg \loexists Y*\loin Obj. $[$MayAccess$^{Ind,Eve}$(X*,Y*) \loand\linebreak
-\ablock \ablock \ablock \ablock \ablock MayAccess$^{Dir,Eve}$(Y*,o')$]$\linebreak
-\ablock \ablock \ablock \ablock  $\}$\linebreak
+\textit{By contraposition,}\linebreak
+M,\losigma\ \loturns\ \linebreak
+\ablock \loforall X*\loin Obj.$\{$\linebreak
+\ablock \quad $[$\null
+\textcolor{red}{\loforall X$^1$\loin{Obj}. \loneg MayAccess$^{Ind,Now}$(o,X$^1$) \loor\linebreak
+\ablock \ablock \ablock \ablock \loneg MayAccess$^{Dir,Now}$(X$^1$,X*)\linebreak
+}
+\ablock \ablock \quad \loand \linebreak
+\ablock \quad \loneg MayAccess$^{Dir,Now}$(X*,o)$ ]$\linebreak
+\ablock \ablock \loor \linebreak
+\textcolor{blue}{
+\ablock \loforall Y*\loin Obj. $\{[$\loneg MayAccess$^{Ind,Now}$(X*,Y*)\linebreak
+\ablock \ablock \ablock \ablock \quad \loand \linebreak
+\ablock \ablock \ablock \quad\: \loneg MayAccess$^{Dir,Now}$(Y*,X*)$]$\linebreak
+\ablock \ablock \ablock \quad \loor \linebreak
+\ablock \ablock \ablock \quad \loneg MayAccess$^{Ind,Eve}$(Y*,o')$\}$
+}\linebreak
 \ablock \loimplies\linebreak
-\ablock \loneg MayAccess$^{Ind,Eve}$(o,o')\
-\linebreak
+	\ablock\loneg MayAccess$^{Ind,Eve}(o,o')$\linebreak
 \hrule
 \end{logic}
-The results of \textsc{epc2} are identical to \textsc{epc1}, with the second condition of \textsc{epc1} expanded in \textsc{epc2} to say global one-way path isolation to o in all possible states \losigma' to mean there does not exist any eventual path from any object to an object Y* that has the eventual capability of o'.\\
 
-So far we have developed results for paths, but what about execution?
-We construct \textsc{eec} from \textsc{epc1}, the only difference being we now require all those paths in \textsc{epc1} to be traversable.
+To give a concrete example, let us assume a simple system with only three objects, o, x, o'.
+\begin{logic}
+\hrule\null
+\textsc{\normalsize *Eventual Path Example}\\
+M,\losigma\ \loturns\ MayAccess$^{Ind,Eve}$(o,o')\linebreak
+\ablock \loimplies\linebreak
+\ablock \ablock $[$\null
+\textcolor{red}{MayAccess$^{Ind,Now}$(o,o) \loand\linebreak
+\ablock \ablock \quad MayAccess$^{Dir,Now}$(o,x)\linebreak
+}
+\ablock \ablock \quad \loor \linebreak
+\ablock \ablock MayAccess$^{Dir,Now}$(x,o)$ ]$\linebreak
+\ablock \ablock \loand \linebreak
+\textcolor{blue}{
+\ablock \ablock $\{[$
+\textcolor{red}{MayAccess$^{Ind,Now}$(x,x) \loand\linebreak
+\ablock \ablock \quad MayAccess$^{Dir,Now}$(x,o')\linebreak}
+\ablock \ablock \quad \quad \loor \linebreak
+\ablock \ablock \quad MayAccess$^{Dir,Now}$(o',x)$]$\linebreak
+\ablock \ablock \quad \loand \linebreak
+\ablock \ablock \quad MayAccess$^{Ind,Eve}$(o',o')$\}$
+}\linebreak
+\ablock \ablock $\}$
+\linebreak \\
+\hrule
+\end{logic}
+In the example, we make concrete \textsc{epc2} using X* = x and Y* = o', where the expansion terminates at the three predicates, MayAccess$^{Ind,Now}$(o,o), MayAccess$^{Ind,Now}$(x,x) and MayAccess$^{Ind,Eve}$(o',o').
+Let us assume o, x, and o' always exists, such that these predicates would return true.
+This example then illustrates an eventual path from o to o' can only exist if one of these four conditions in \losigma\ hold:
+\begin{itemize}\setlength\itemsep{0.5em}
+\item o has the capability of x, and x has the capability of o'
+\item o has the capability of x, and o' has the capability of x
+\item x has the capability of o, and x has the capability of o'
+\item o has the capability of x, and x has the capability of o'
+\end{itemize}
+Note how, these four configurations are present state configurations at \losigma, but allows us to reason about whether a potential path from o to o' can exist in all arising states \losigma' from \losigma.\\
+
+So far we have developed results for paths from o to o', but what about a direct path from o to o'?
+Luckily, we only need construct \textsc{eec} from \textsc{epc1}, the only difference being we now require the path from o to X* in \textsc{epc1} to be traversable.
 
@@
 \end{logic}
-The power of \textsc{eec} is similar to \textsc{epc}\ in that it allows an arbitrary recursive expansion of the MayAccess$^{Dir,Eve}$(X*,o') till the terminating MayAccess$^{Dir,Eve}$(o',o') that is always true.
+The power of \textsc{eec} is similar to \textsc{epc}\ in that it allows an arbitrary recursive expansion of the MayAccess$^{Dir,Eve}$(X*,o') till the terminating MayAccess$^{Dir,Eve}$(o',o') that is always true.\\
 
-We now turn our attention formalzing Ocap state changes rules in the system.\\
+We can always use an automated theorem prover to prove whether a path can arise between two objects in a present configuration system.
+We now turn our attention formalizing Ocap state changes rules in the system.\\
 
-\textbf{Rule 3: Objects may have private encapsulation of state and behaviour}
-\begin{logic}
-\hrule\null
-\textsc{\normalsize *Global State Change Existence (GSCE)}\\
-M,\losigma\ \loturns\ $[$\loexists X*\loin{Obj}. MayAffect$^{Now}$(X*,o')$]$\linebreak
-\ablock \loimplies \linebreak
-\ablock $[$\loexists Y*\loin{Obj}. MayAccess$^{Dir,Now}$(Y*,o')$]$\linebreak
-\ablock \loimplies \linebreak
-\ablock $[$\loexists Z*\loin{Obj}. MayAccess$^{Ind,Now}$(Z*,o')$]$
-\linebreak\\
-\textit{By contraposition,}\linebreak
-M,\losigma\ \loturns\ $[$\loforall Z*\loin{Obj}. \loneg MayAccess$^{Ind,Now}$(Z*,o')$]$\linebreak
-\ablock \loimplies\linebreak
-\ablock $[$\loforall Y*\loin{Obj}. \loneg MayAccess$^{Dir,Now}$(Y*,o')$]$\linebreak
-\ablock \loimplies\linebreak
-\ablock $[$\loforall X*\loin{Obj}. \loneg MayAffect$^{Dir,Now}$(X*,o')$]$\linebreak
-\hrule
-\end{logic}
-\textsc{gsce} can be derived from our predicate definitions.
-The first implication states that if an object can change the state of o', then there must exist an object in the system that holds the capability of o'.
-The second implication is that there must exist an object in the system that holds a path to o'.
-We view reasoning about state change as reasoning about protection, which we elaborate in the next subsection.
 \subsection{Protection}\label{protection}
-To reason about enforcing protection, we focus on finding the necessary conditions for the weaker predicate MayAffect$^{Eve}$ rather than MayAffect$^{Now}$ in the antecedent.
-This is because negation on both sides of the implication, would yield a stronger \loneg MayAffect$^{Eve}$ in the consequent, which is our area of interest (since it represents state protection).
+With our formalizations of OCap rules in place, we can now reason about enforcing protection.
+In this subsection we focus on finding the necessary conditions for the weaker predicate MayAffect$^{Eve}$ rather than MayAffect$^{Now}$ in the antecedent.
+This is because negation on both sides of the implication, would yield a stronger \loneg MayAffect$^{Eve}$ in the consequent.
 In practical terms, if we are concerned with the protection of o' from o, it is also often not very useful to have a policy where \loneg MayAffect$^{Now}$(o,o') holds but \loneg MayAffect$^{Eve}$(o,o') does not.
@@
 
-To begin, we simplify our reasoning, we introduce the assumption that all fields in our objects can only be declared private.
-Consequently, this implies the necessary condition that an object's state can only be modified if one of it's behaviour is called, and therefore require at least one other object in the system that holds its capability.
-We do not think this is restrictive in any case, the programmer merely has to write explicitly a method to return an object's field.
-
+\textbf{Rule 3: Objects may have private encapsulation of state and behaviour}\\
+To begin, we make clean in our reasoning that all fields in our objects can only be declared private.
+Consequently, this implies the necessary condition that an object's state can only be modified if one of it's behaviour is called, either by itself or one other object in the system that holds its capability.
+This is also implied by the OCap rule that there is no ambient authority that can interact with any object's behaviour.
 \begin{logic}
@@
 
-\textsc{pfa} means that in order for an object X*'s state to change, it must be done through some object Y* calling its behaviour (Y* can refer to the same object as X*).
+\textsc{pfa} means that in order for an object o' to change, it must be done through some object Y* calling its behaviour (Y* can refer to the same object as X*).
 Equivalently, denying all objects in the system the ability to call an object's behaviour implies that no object can modify the object's state.\\
 
-\subsubsection{Protection through possible paths}
-We now divert our attention to reasoning about the relationship between protection of state (MayAffect) and paths (MayAccess{Ind,*}).
-The question we ask is what is the relation between state protection in eventual outcomes, and present path configurations?
-We build this relationship from the basics by progressively finding stronger conditions of MayAffect.
-We begin with the first condition, which says that for an object to affect another, an \textit{eventual} path has to exist.
-This can be derived from the definitions of our predicates.
-\begin{logic}[Implication 1]
+\begin{logic}
 \hrule\null
-\textsc{\normalsize *Necessary Path Condition 1 (NPC1)}\\
-M,\losigma\ \loturns\ MayAffect$^{Eve}$(o,o') \loimplies\linebreak
-	\ablock MayAccess$^{Ind,Eve}$(o,o')
-\linebreak \\
-\textit{By contraposition,}\linebreak
-M,\losigma\ \loturns\ \loneg MayAccess$^{Ind,Eve}$(o,o') \loimplies\linebreak
-	\ablock \loneg MayAffect$^{Eve}$(o,o')
-\linebreak
+\textsc{\normalsize *Global State Change Existence (GSCE)}\\
+M,\losigma\ \loturns\ $[$\loexists X*\loin{Obj}. MayAffect$^{Now}$(X*,o')$]$\linebreak
+\ablock \loimplies \linebreak
+\ablock $[$\loexists Y*\loin{Obj}. MayAccess$^{Dir,Now}$(Y*,o')$]$\linebreak
+\ablock \loimplies \linebreak
+\ablock $[$\loexists Z*\loin{Obj}. MayAccess$^{Ind,Now}$(Z*,o')$]$
+\linebreak\\
 \hrule
 \end{logic}
-Here, we immediately see a first defensive outcome of the OCap model. Having no eventual paths from object o to o' guarantees that the state of object o' cannot be modified by object o.\\
+\textsc{gsce} can be derived from our predicate definitions, and makes explicit the connection betwen state change, paths, and capabilities.
+The first implication states that if an object can change the state of o', then there must exist an object in the system that holds the capability of o'.
+The second implication is that there must exist an object in the system that holds a path to o'.
+We view reasoning about state change as reasoning about protection, which we elaborate in the next subsection.
 
-So far our results serve as a good base to enforce \loneg MayAffect$^{Eve}(o,o')$ but we require \textit{stronger} necessary conditions, because enforcing protection of an object's state with just these conditions require the impractical configuration that o is eventually path-isolated from o' on the reference graph.
-Cooperation between objects imply that there needs to be some path established between the objects for interaction to take place.
-Therefore, we can expect neither \loneg MayAccess$^{Ind,Now}$(o,o') nor \loneg MayAccess$^{Ind,Eve}$(o,o') to hold in a system that allows cooperation between o and o'.\\
-
-What about present path configurations?
-Using \textsc{npc1} and the results of \textsc{epc1}, we derive \textsc{npc2} below which is a relation between protection of o' in eventual states, and \textit{present} path configurations between o and o'.
-\begin{logic}
-\hrule\null
-\textsc{\normalsize *Necessary Path Condition 2 (NPC2)}\\
-M,\losigma\ \loturns\ MayAffect$^{Eve}$(o,o')\linebreak
-\ablock \loimplies\linebreak
-\ablock \loexists X*\loin Obj.$[($MayAccess$^{Ind,Now}$(o,X*) \loor \linebreak
-\ablock \ablock \ablock \ablock MayAccess$^{Dir,Now}$(X*,o)$)$\linebreak
-\ablock \ablock \ablock \ablock \loand \linebreak
-\ablock \ablock \ablock \ablock MayAccess$^{Ind,Eve}$(X*,o')$]$
-\linebreak \\
-\textit{By contraposition and De Morgan rules,}\linebreak
-M,\losigma\ \loturns\ 
-\loforall X*\loin Obj. $[$\loneg MayAccess$^{Ind,Now}$(o,X*) \loand \linebreak
-\ablock \ablock \ablock \ablock \loneg MayAccess$^{Dir,Now}$(X*,o)$]$\linebreak
-\ablock \ablock \ablock \ablock \loor \linebreak
-\ablock \ablock \ablock \ablock \loneg MayAccess$^{Ind,Eve}$(X*,o')$]$\linebreak
-\ablock \loimplies \linebreak
-	\ablock\loneg MayAffect$^{Eve}(o,o')$\linebreak
-\hrule
-\end{logic}
-The interpretation of \textsc{npc2} is identical to \textsc{epc1}, with protection of state being made an explicit consequent from the conditions of \textsc{epc1}.\\
-We now examine whether it is possible to build a relation between MayAffect$^{Eve}$ and MayAccess$^{Dir,Eve}$.
-Similarly, \textsc{npc2} can be augmented by \textsc{epc2} which results in \textsc{npc3}:
-\begin{logic}
-\hrule\null
-\textsc{\normalsize *Necessary Path Condition 3 (NPC3)}\\
-M,\losigma\ \loturns\ 
-MayAffect$^{Eve}$(o,o')\linebreak
-\ablock \loimplies \linebreak
-\ablock \loexists X*\loin Obj. $\{[$MayAccess$^{Ind,Now}$(o,X*) \loor \linebreak
-\ablock \ablock \ablock  \ablock MayAccess$^{Dir,Now}$(X*,o)$]$\linebreak
-\ablock \ablock \ablock \ablock \loor \linebreak
-\ablock \ablock \quad \loexists Y*\loin Obj. $[$MayAccess$^{Ind,Eve}$(X*,Y*) \loand\linebreak
-\ablock \ablock \ablock \ablock \ablock MayAccess$^{Dir,Eve}$(Y*,o')$]$\linebreak
-\ablock \ablock \ablock \ablock  $\}$\linebreak
-\linebreak \\
-\textit{By contraposition and De Morgan rules}\linebreak
-M,\losigma\ \loturns\ 
-\loforall X*\loin Obj. $\{[$\loneg MayAccess$^{Ind,Now}$(o,X*) \loand \linebreak
-\ablock \ablock \ablock \ablock \quad \loneg MayAccess$^{Dir,Now}$(X*,o)$]$\linebreak
-\ablock \ablock \ablock \ablock \loor \linebreak
-\ablock \ablock \loneg \loexists Y*\loin Obj. $[$MayAccess$^{Ind,Eve}$(X*,Y*) \loand\linebreak
-\ablock \ablock \ablock \ablock \ablock MayAccess$^{Dir,Eve}$(Y*,o')$]$\linebreak
-\ablock \ablock \ablock \ablock  $\}$\linebreak
-\ablock \loimplies\linebreak
-\ablock \loneg MayAffect$^{Eve}$(o,o')\
-\linebreak
-\hrule
-\end{logic}
-We are at the finishing line. The Ind,Now predicates can be expanded using GPEC to obtain direct capabilities. Ind,Eve can be expanded using EPC2. Dir,Eve can be expanded using EEC.
-\subsubsection{A note on execution and protection}
+\subsubsection{Protection is about object-path denial}
 \begin{logic}
@@
 M,\losigma\ \loturns\ MayAffect$^{Eve}$(o,o') \loimplies 
-MayAccess$^{Ind,Eve}$(o,o')
+MayAccess$^{Ind,Eve}$(o,o')\linebreak
+\\
+M,\losigma\ \loturns\ \loneg MayAccess$^{Ind,Eve}$(o,o') \loimplies 
+\loneg MayAffect$^{Eve}$(o,o')\linebreak
+\\
+M,\losigma\ \loturns\ MayAccess$^{Dir,Eve}$(o,o') \loimplies 
+MayAccess$^{Ind,Eve}$(o,o')\linebreak
+\\
+M,\losigma\ \loturns\ \loneg MayAccess$^{Ind,Eve}$(o,o') \loimplies 
+\loneg MayAccess$^{Dir,Eve}$(o,o') \loand\ \loneg MayAffect$^{Eve}$(o,o')\linebreak
 \end{logic}
+We make the important result that protection of an object's state is path denial to the object, where path is a necessary condition for an object to affect another object.
+An object o having the capability of o' is \textbf{NOT} a necessary condition for o to modify the state of o'.
+However, the existence of \textit{some} object X* (that can refer to o, or other objects) having the capability of o', is a necessary condition for o to modify o'.
+See \textsc{nec1} and \textsc{nec2} below which illustrates this.
+This is precisely why attenuating objects are so powerful and useful in OCap patterns.
+We can always deny o the capability of o', but still allow o to modify o' in some protected way.
+To say it in another way, denial of paths is much stronger, as it implies protection \textit{and} denial of capability, but denial of capability does \textit{not} imply protection and denial of paths.\\
+
 \begin{logic}
@@
 \end{logic}
-\textsc{nec1} is a stronger version of \textsc{npc1} and says that in order for an object o to modify the state of o', there must exist an object X* that has a direct path to o', and that X* \textit{can} traverse such a path.
-This is derived trivially from our \textsc{pfa} assumption in the beginning.
+\textsc{nec1} says that in order for an object o to modify the state of o', there must exist an object X* that has a direct path to o', and that X* \textit{can} traverse such a path.
+This is derived trivially from our \textsc{pfa} assumption.
 Consequently, in our contraposition result, we can deny \textit{all} objects in our system capability of o' to enforce the protection of o' from o.
@@
 \end{logic}
-We now examine the second half of the necessary condition in \textsc{nec2}:\\
-$[$\loexists X*\loin{Object}, X*\loneq o. MayAccess$^{Dir,Eve}$(X*,o')$]$
-\\ which means that there exists object X* that is \textit{not} object o, that must have the capability of o'.
-But this does not say anything about o and X*.
-We need to connect o to X*, because \textsc{npc1} says MayAffect$^{Eve}$(o,o') implies that o must have some indirect path to o'.
-
+With this clarification, we then reason about the relationship between protection of state (MayAffect) and paths (MayAccess$^{Ind,*}$).
+The question we ask is what is the relation between state protection in eventual outcomes, and present path configurations?
+We build this relationship from the basics by progressively finding stronger conditions of MayAffect.
+We begin with the first condition, which says that for an object to affect another, an \textit{eventual} path has to exist.
+This can be derived from the definitions of our predicates.
 \begin{logic}
 \hrule\null
-\textsc{\normalsize *Necessary Execution Condition 3 (NEC3)}\\
-M,\losigma\ \loturns\ MayAffect$^{Eve}$(o,o')\linebreak
-	\ablock \loimplies\linebreak
-	\ablock MayAccess$^{Dir,Eve}$(o,o') \loor \linebreak
-	\ablock $[$\loexists X*\loin Obj, X*\loneq o. MayAccess$^{Dir,Eve}$(o,X*) \loand\linebreak
-	\ablock \ablock MayAccess$^{Dir,Eve}$(X*,o')$]$
-	\linebreak\\
+\textsc{\normalsize *Necessary Path Condition 1 (NPC1)}\\
+M,\losigma\ \loturns\ MayAffect$^{Eve}$(o,o') \loimplies\linebreak
+	\ablock MayAccess$^{Ind,Eve}$(o,o')
+\linebreak \\
 \textit{By contraposition,}\linebreak
-M,\losigma\ \loturns\ \loneg MayAccess$^{Dir,Eve}$(o,o') \loand \linebreak
-	\ablock \loneg$[$\loexists X*\loin Obj, X*\loneq o. MayAccess$^{Dir,Eve}$(o,X*) \loand\linebreak
-	\ablock \ablock MayAccess$^{Dir,Eve}$(X*,o')$]$\linebreak
-	\ablock \loimplies\linebreak
+M,\losigma\ \loturns\ \loneg MayAccess$^{Ind,Eve}$(o,o') \loimplies\linebreak
 	\ablock \loneg MayAffect$^{Eve}$(o,o')
@@
 \end{logic}
-\textsc{nec3} is not very useful because we do not have any knowledge of how the system's path or capability configuration should look like at \losigma\ for us to enable protection of o' in all possible states arising from \losigma.
-We now incorporate our results of \textsc{eec} to expand MayAccess$^{Dir,Eve}$ that will give us the present configuration of paths and capabilities.\\
+Here, we immediately see a first defensive outcome of the OCap model. Having no eventual paths from object o to o' guarantees that the state of object o' cannot be modified by object o.\\
+
+So far our results serve as a good base to enforce \loneg MayAffect$^{Eve}(o,o')$ but we require \textit{stronger} necessary conditions for MayAffect$^{Eve}(o,o')$, because we need to understand the present path configurations that can deny o from modifying o'.
+Luckily, we have already built the necessary ingredients in our formalizations of OCap rules in \ref{ocaprules}, where have painstakingly derive the relationship between present path-capability configurations and eventual paths.
+Since \textsc{npc1} tells us that the eventual paths is a necessary condition for object state modification, and \textsc{epc1} tells us the the relationship between eventual paths and present path-capability configurations, we combine the two to present:
 
@@
 \hrule\null
-\textsc{\normalsize *Necessary Execution Condition 4 (NEC4)}\\
-M,\losigma\ \loturns\ MayAffect$^{Eve}$(o,o') \loimplies\linebreak
-\ablock \loexists X*\loin Obj.$[($MayAccess$^{Dir,Now}$(o,Y*) \loor \linebreak
-\ablock \ablock \ablock \ablock MayAccess$^{Dir,Now}$(Y*,o)$)$\linebreak
+\textsc{\normalsize *Necessary Path Condition 2 (NPC2)}\\
+M,\losigma\ \loturns\ MayAffect$^{Eve}$(o,o')\linebreak
+\ablock \loimplies\linebreak
+\ablock \loexists X*\loin Obj.$[($MayAccess$^{Ind,Now}$(o,X*) \loor \linebreak
+\ablock \ablock \ablock \ablock MayAccess$^{Dir,Now}$(X*,o)$)$\linebreak
 \ablock \ablock \ablock \ablock \loand \linebreak
-\ablock \ablock \ablock \ablock MayAccess$^{Dir,Eve}$(Y*,o')$]$\linebreak
-\ablock \loor \linebreak
-\ablock $\{$\loexists X*\loin Obj, X*\loneq o.\linebreak
-\ablock \quad \loexists X$^1$\loin Obj.$[($MayAccess$^{Dir,Now}$(o,X$^1$) \loor \linebreak
-\ablock \ablock \ablock \ablock MayAccess$^{Dir,Now}$(X$^1$,o)$)$\linebreak
-\ablock \ablock \ablock \ablock \loand \linebreak
-\ablock \ablock \ablock \ablock MayAccess$^{Dir,Eve}$(X$^1$,X*)$]$\linebreak
-\ablock \quad \loand \linebreak
-\ablock \quad \loexists X$^2$\loin Obj.$[($MayAccess$^{Dir,Now}$(X*,X$^2$) \loor \linebreak
-\ablock \ablock \ablock \ablock MayAccess$^{Dir,Now}$(X$^2$,X*)$)$\linebreak
-\ablock \ablock \ablock \ablock \loand \linebreak
-\ablock \ablock \ablock \ablock MayAccess$^{Dir,Eve}$(X$^2$,o')$]$\linebreak
-\ablock $\}$
-	\linebreak\\
+\ablock \ablock \ablock \ablock MayAccess$^{Ind,Eve}$(X*,o')$]$
+\linebreak \\
 \textit{By contraposition,}\linebreak
-...\linebreak
-\\
+M,\losigma\ \loturns\ 
+\loforall X*\loin Obj. $[$\loneg MayAccess$^{Ind,Now}$(o,X*) \loand \linebreak
+\ablock \ablock \ablock \ablock \loneg MayAccess$^{Dir,Now}$(X*,o)$]$\linebreak
+\ablock \ablock \ablock \ablock \loor \linebreak
+\ablock \ablock \ablock \ablock \loneg MayAccess$^{Ind,Eve}$(X*,o')$]$\linebreak
+\ablock \loimplies \linebreak
+	\ablock\loneg MayAffect$^{Eve}(o,o')$\linebreak
 \hrule
 \end{logic}
+, where we can always use \textsc{gpec} to expand indirect paths into direct capabilities till we reach the terminating origin o and recursively use the condition to expand the condition till we reach the terminating o'.\\
 
-From \textsc{nec4} we see the power of OCap patterns hence lies in providing concrete examples of a system of cooperation that allows the existence of paths between objects for cooperation while still dictating the degree of control of one object can have over another.
+We now see the power of OCap patterns hence lies in providing concrete examples of a system of cooperation that allows the existence of paths between objects for cooperation while still dictating the degree of control of one object can have over another.
 The logics we have developed so far illuminate the power of attenuating objects X* that can enable protection.
@@
 Consequently, this allows an object which has the capability of any one node in the tree to navigate up to the root node (Document), and consequently navigate to all other nodes in the tree.
-
-\begin{logic}[NodeP1]
-\hrule\null
-\textsc{\normalsize *Policy 1: Necessary Condition}\\
-\loforall n,o, RN\losubseteq ReNode.\linebreak
-\ablock Dom(RN,n) \loand\ MayAffect$^{Eve}$(o,n)\linebreak \ablock\loimplies \linebreak
-\ablock \loexists rn\loin RN. MayAccess$^{Dir,Now}$(o,rn)\linebreak
-\ablock \ablock \:\:\:\: \loand\ MayAccess$^{Dir,Now}$(rn,n)
-\end{logic}
-\hrule\null
-This policy states that if an Object o may affect the state of a Node n, and that n is dominated(protected) by a set of ReNodes RN, then it implies that a path exists from o to some rn in RN, and that rn has strong access to n.
-From Policy 1, using contraposition, we derive the relation:
-
-\begin{logic}[NodeP1a]
-\loforall n,o, RN\losubseteq ReNode.\linebreak
-\ablock [ \loforall rn\loin RN. \loneg MayAccess$^{Ind,Eve}$(o,rn) \loor \linebreak
-\ablock \loforall rn\loin RN. \loneg MayAccess$^{Dir,Eve}$(rn,n) ]\linebreak
-\ablock \loimplies
-\loneg MayAffect$^{Eve}$(o,n) \loor\ \loneg Dom(RN,n)
-\end{logic}
-
-Note that without prescribing any rules to ReNode, we can deny object o the ability to affect node n by enforcing that no path exists from o to any rn in the set RN that dominates n, which by implication is another way of saying isolation from o to n, since:
-\begin{logic}[NodeP1b]
-\loforall n,o. Dom(RN,n) \loand \linebreak \loforall rn\loin RN. \loneg MayAccess$^{Ind,Eve}$(o,rn) \loimplies \linebreak
-\ablock \loneg MayAccess$^{Ind,Eve}$(o,n).
-\end{logic}
-
-Another implication is that we can also deny object o the ability to affect node n by enforcing \loforall rn\loin RN.
-\loneg MayAccess$^{Dir,Eve}$(rn,n).
-This is however, \textit{not} enforceable.
-This is because by definition a ReNode rn holds a field containing the capability of the node n it is meant to protect when the ReNode is constructed, and field that holds the capability of the node is private and constant, that cannot be subsequently changed or removed.
-Consequently, there must exist at least one ReNode rn in the dominating RN set that holds the capability of n:
-\begin{logic}[NodeP1c]
-\loexists rn \loin RN. Dom(RN,n) \loand MayAccess$^{Dir,Eve}$(rn,n)\linebreak
-\ablock \loimplies \linebreak
-\ablock \loneg( \loforall rn\loin RN. \loneg MayAccess$^{Dir,Eve}$(rn,n) )
-\end{logic}
-
-\textsc{Policy 2: Domination of Nodes}
-\begin{logic}[NodeP2]
-\end{logic}
-
-\subsection{Pattern 2: Caretaker}
-\subsection{Pattern 3: Membrane}
-
-\textbf{Policy 1}:
-
 %----------------------------------------------------------------------------------------

### 2017-08-17 article.tex
new condition

@@
 Looking from the left to right direction, since the path from Y* to o' exists, then there must exist an object X* with a capability of o' so that the path from Y* to o' is well-established.
-Moving from the right to left direction, \textsc{gpe} says that if an object X* has a capability then there exist an object Y* with a path to o', and this can be proven from \textsc{pec} or by definition when (X*==Y*).\\
-The contraposition result of \textsc{gpe} says that we can prevent all paths to o' through one of these conditions:
-\begin{itemize}
-\item Deny all direct paths leading into o' 
-\end{itemize}
+Moving from the right to left direction, \textsc{gpe} says that if an object X* has a capability then there exist an object Y* with a path to o', and this can be proven from \textsc{pec} or by definition when (X*==Y*).
+The contraposition result of \textsc{gpe} says that we can prevent all paths to o' through one of these conditions through denying all direct paths leading into o'.\\
 
 However, \textsc{gpe} only postulates the existence of some object with some capability iff there exists some object with some path, and says nothing about how the objects are connected.
-From the connectivity begets connectivity rule, \textsc{pec}, and \textsc{gpe}, we construct global path-execution and execution-execution relationships over the entire system:
+From the connectivity begets connectivity rule, \textsc{pec}, and \textsc{gpe}, we construct global path-execution relationships over the entire system:
 \begin{logic}
@@
 \textsc{gpec} introduces an intermediate object X* between o and o' that makes explicit the missing connection in \textsc{gpe}.\\
-The contraposition result of \textsc{gpec} says that we can prevent all paths from o to o' through one of these conditions:
 
@@
 If not, the capability of X* must be introduced to o.
-For X* to introduce itself, X* must have the capability of o in \losigma', and the only way we can guarantee is to have X* possess the capability of o in \losigma.
+For X* to introduce itself, X* must have the capability of o in \losigma', and the only way we can guarantee this is to have X* possess the capability of o in \losigma.
 Well, what if the capability of X* is introduced by some \textit{other} object \~X*?
@@
 
-\subsubsection{Protection is about object-path denial}
+\subsubsection{Protection is about local object-path denial and global capability denial}
 \begin{logic}
@@
 \ablock \loexists X*\loin Obj.$[($MayAccess$^{Ind,Now}$(o,X*) \loor \linebreak
-\ablock \ablock \ablock \ablock MayAccess$^{Dir,Now}$(X*,o)$)$\linebreak
+\ablock \ablock \ablock \quad MayAccess$^{Dir,Now}$(X*,o)$)$\linebreak
 \ablock \ablock \ablock \ablock \loand \linebreak
-\ablock \ablock \ablock \ablock MayAccess$^{Ind,Eve}$(X*,o')$]$
+\ablock \ablock \ablock \quad MayAccess$^{Ind,Eve}$(X*,o')$]$
 \linebreak \\
@@
 \end{logic}
-, where we can always use \textsc{gpec} to expand indirect paths into direct capabilities till we reach the terminating origin o and recursively use the condition to expand the condition till we reach the terminating o'.\\
+, where we can always use \textsc{gpec} to expand indirect paths into direct capabilities till we reach the terminating origin o and recursively use the condition to expand the condition till we reach the terminating o'.
+\textsc{npc2} gives us the present path configurations that will allow an eventual path from o to o' as a necessary condition for o to modify o'.But if objects are path-isolated, how do we allow them to cooperate?
+
+\textsc{nec2} and \textsc{npc2} show the necessary capability and path conditions for protection.
+Is there a relation that can capture the essence of both?
+Notice how, we have said in \textsc{nec2} that there must exist some object in the system (that is not necssarily o) that has the capability of o', while in \textsc{npc2}, we said that o must have a path to o'.
+Surely it must mean then that the object X* in \textsc{nec2} that has the capability of o' refers to the last object on the path from o to o' mentioned in \textsc{npc2}.
+We now bring these results together, by making use of \textsc{gpec} to make explicit the last object in the eventual path holding the capability of o' in \textsc{npc2}:\\
+\begin{logic}
+\hrule\null
+\textsc{\normalsize *Necessary Execution-Path Condition* (NEPC*)}\\
+M,\losigma\ \loturns\ MayAffect$^{Eve}$(o,o')\linebreak
+\ablock \loimplies\linebreak
+\ablock \loexists X*\loin Obj.$[($MayAccess$^{Ind,Now}$(o,X*) \loor \linebreak
+\ablock \ablock \ablock \ablock MayAccess$^{Dir,Now}$(X*,o)$)$\linebreak
+\ablock \ablock \ablock \ablock \loand \linebreak
+\ablock \loexists Y*\loin{Obj}. $[$MayAccess$^{Ind,Eve}$(X*, Y*) \loand\linebreak
+\ablock \ablock \ablock \ablock MayAccess$^{Dir,Eve}$(Y*,o')$]$
+\linebreak \\
+\textit{By contraposition,}\linebreak
+M,\losigma\ \loturns\ 
+\loforall X*\loin Obj. $[$\loneg MayAccess$^{Ind,Now}$(o,X*) \loand \linebreak
+\ablock \ablock \ablock \ablock \loneg MayAccess$^{Dir,Now}$(X*,o)$]$\linebreak
+\ablock \ablock \ablock \ablock \loor \linebreak
+\ablock \quad\loforall Y*\loin{Obj}. $[$MayAccess$^{Ind,Eve}$(X*, Y*) \loor\linebreak
+\ablock \ablock \ablock \ablock MayAccess$^{Dir,Eve}$(Y*,o')$]$
+\ablock \loimplies \linebreak
+	\ablock\loneg MayAffect$^{Eve}(o,o')$\linebreak
+\hrule
+\end{logic}
+\textsc{nepc*} is nice, because we can see the necessary path condition and the condition that an object must hold the capability of o'.
+But what does it mean to modify an object?
+No matter how long the path from o to o' is, surely to modify o' requires the last object on our path to \textit{be} the 'final messenger' that will call the object's behaviour to modify it---which means MayAffect$^{Eve}$(Y*,o') \textit{must} hold.
+We express this by extending \textsc{nepc1*} to \textsc{nepc2*}.
+\begin{minipage}{\linewidth}
+\begin{logic}
+
+\hrule\null
+\textsc{\normalsize *Necessary Execution-Path Condition 2* (NEPC2*)}\\
+M,\losigma\ \loturns\ MayAffect$^{Eve}$(o,o')\linebreak
+\ablock \loimplies\linebreak
+\ablock \loexists X*\loin Obj.$[($MayAccess$^{Ind,Now}$(o,X*) \loor \linebreak
+\ablock \ablock \ablock \quad MayAccess$^{Dir,Now}$(X*,o)$)$\linebreak
+\ablock \ablock \ablock \ablock \loand \linebreak
+\ablock \loexists Y*\loin{Obj}. $[$MayAccess$^{Ind,Eve}$(X*, Y*) \loand\linebreak
+\ablock \ablock \ablock \quad MayAccess$^{Dir,Eve}$(Y*,o') \loand\linebreak
+\ablock \ablock \ablock \quad MayAffect$^{Eve}$(Y*,o')$]$
+\linebreak \\
+\textit{By contraposition,}\linebreak
+M,\losigma\ \loturns\ 
+\loforall X*\loin Obj. $[$\loneg MayAccess$^{Ind,Now}$(o,X*) \loand \linebreak
+\ablock \ablock \ablock \ablock \loneg MayAccess$^{Dir,Now}$(X*,o)$]$\linebreak
+\ablock \ablock \ablock \ablock \loor \linebreak
+\ablock \quad\loforall Y*\loin{Obj}. $[$MayAccess$^{Ind,Eve}$(X*, Y*) \loor\linebreak
+\ablock \ablock \ablock \ablock MayAccess$^{Dir,Eve}$(Y*,o') \loor\linebreak
+\ablock \ablock \ablock \ablock MayAffect$^{Eve}$(Y*,o')$]$
+\ablock \loimplies \linebreak
+	\ablock\loneg MayAffect$^{Eve}(o,o')$\linebreak
+\hrule\null
+\end{logic}
+\end{minipage}
+\linebreak\\
+
+\textsc{nepc2*} has the intuitive meaning that in order to say object o may modify object o' in some eventual state, \textit{all} these three conditions must hold:
+\begin{itemize}
+\item there must be some present path configuration such that an eventual path from o to o' can arise
+\item there must be an object Y* that sits as the last object along the eventual path to o', that has the capability of o'
+\item object Y* must be able to modify o'
+\end{itemize}
+
+The logics we have developed show that security of an object can be enforced in an OCap model by:
+\begin{itemize}
+\item being careful about our present path configurations so that we can ensure eventual path-isolation between the object and untrusted object
+\item specifying security policies on how \textit{all} objects with the direct capability of our trusted object may modify the object
+\end{itemize}
+Therein lies also the difference between OCap and non-OCap models, because non-OCap models that allow forging of capabilities or a global ambient authoriy \textit{cannot} enforce security through path isolation, which is a distinctive feature of OCap models.
+How do we then expect objects with the direct capability of o' may modify o' through some security policy?
+The nice thing is we already have a predicate MayAffect$^{Eve}$(*,*) that can accomodate these security policies.
+Notice how in \textsc{NEPC*}, object Y* sits as the last object in the path from o to o' and has the capability of o'.
+This gives us a clue as to how we can enable protection of o' without denying paths and writing security policies that allow
+
+\subsection{What about Cooperation?}
+At this juncture, it is good to know we can enforce protection through denying possible path configurations, but in practice this is still not useful.
+Cooperation between objects require paths to exist between objects so it is rare or almost impossible to have two objects o and o' that cooperate while \loneg MayAccess$^{Ind,Eve}$(o,o') holds.
+The quintessential question in OCap systems is whether we can enforce protection such as controlling whether o may affect o' while \textit{still} allowing paths to exist between them.
+This is possible, because of Rule 3 that says objects may have private encapsulation of state and behaviour.
+To begin, let us assume a simple object o' that has \textit{only} private behaviours.
+In such a scenario, since all fields in our system are also assumed to be private, then we know:
+\begin{logic} \loturns\ \loforall X*\loin Obj, X*\loneq o'. \loneg MayAffect(X*,o') \end{logic}
+Notice, how protection of o' can occur independent of paths and capability, due to object encapsulation. This means that objects can specify security policies that it chooses to enforce.
+
+\subsubsection{How can security policies be enforced by objects?}
+In the previous example we have demonstrated in OCap systems that objects can, at the very least enforce protection policies on \textit{itself}.
+But can we have objects enforcing security policies on \textit{other} objects?
+Let us examine this with another example.\\
+
+If we have some starting configuration in a system where there is only one object o that has the capability of o', and no other objects have the capability of o', then we know any eventual path to o' has to include and end with the sub-path from o to o', \textit{unless} o leaks the capability of o' to some other object X*.
+Notice how the sub-path o to o' is also a direct capability from o to o'.
+Object o, therefore, has the power to dictate the policies that can affect o' through itself, and state more precise stronger necessary conditions for the MayAffect$^{Eve}${(o,o') predicate, conditions other than path existence which we have argued.
+To put it in another way, we can now allow a path to exist from o to o', \textit{and} still enforce \loneg MayAffect$^{Eve}$(o,o'), because the negation of these stronger security policy conditions results in weaker antecedents for protection consequence.\\
 

### 2017-08-17 article.tex
cooperation

@@
 , where we can always use \textsc{gpec} to expand indirect paths into direct capabilities till we reach the terminating origin o and recursively use the condition to expand the condition till we reach the terminating o'.
-\textsc{npc2} gives us the present path configurations that will allow an eventual path from o to o' as a necessary condition for o to modify o'.But if objects are path-isolated, how do we allow them to cooperate?
+\textsc{npc2} gives us the present path configurations that will allow an eventual path from o to o' as a necessary condition for o to modify o'.
+But if objects are path-isolated, how do we allow them to cooperate?
 
@@
 \end{itemize}
-Therein lies also the difference between OCap and non-OCap models, because non-OCap models that allow forging of capabilities or a global ambient authoriy \textit{cannot} enforce security through path isolation, which is a distinctive feature of OCap models.
-How do we then expect objects with the direct capability of o' may modify o' through some security policy?
-The nice thing is we already have a predicate MayAffect$^{Eve}$(*,*) that can accomodate these security policies.
-Notice how in \textsc{NEPC*}, object Y* sits as the last object in the path from o to o' and has the capability of o'.
-This gives us a clue as to how we can enable protection of o' without denying paths and writing security policies that allow
+Therein lies also the difference between OCap and non-OCap models, because non-OCap models that allow forging of capabilities or a global ambient authoriy \textit{cannot} enforce security through path isolation, which is a distinctive feature of OCap models.\\
+
+The similarity is that both models allow us to specify policies \textit{within} objects as to whether an object's behaviour can be called (although non OCap systems can do so 'externally' through a global ambient authority).
+'Internal' object protection can be done through using defensive programming techniques like encapsulation and checking of conditions for behaviours to succeed.
+In a way these forms of protection can be a form of 'stack-based' access control from the called object's perspective, because they only happen \textit{after} the calling object already has already called the object.\\
+
+How do we then these internal object security policies in our logical framework?
+The nice thing is we already have a predicate MayAffect$^{Eve}$(*,*) that can accomodate this.
+Notice how in \textsc{NEPC2*}, object Y* sits as the last object in the path from o to o' \textit{and} has the capability of o'.
+This gives us a clue as to how we can enable protection of o' without denying paths, by \textit{specifying whether Y* can affect o' explicitly in security policies} that are independent of paths.
 
 \subsection{What about Cooperation?}
-At this juncture, it is good to know we can enforce protection through denying possible path configurations, but in practice this is still not useful.
+It is good to know we can enforce protection through denying possible path configurations, but in practice this is still not useful.
 Cooperation between objects require paths to exist between objects so it is rare or almost impossible to have two objects o and o' that cooperate while \loneg MayAccess$^{Ind,Eve}$(o,o') holds.
@@
 
-If we have some starting configuration in a system where there is only one object o that has the capability of o', and no other objects have the capability of o', then we know any eventual path to o' has to include and end with the sub-path from o to o', \textit{unless} o leaks the capability of o' to some other object X*.
+If we have some starting configuration in a system where there is only one object o that has the capability of o', and no other objects have the capability of o', then we know any eventual path to o' has to include and end with the sub-path from o to o', \textit{unless} o leaks the capability of o' to some other object.
 Notice how the sub-path o to o' is also a direct capability from o to o'.
-Object o, therefore, has the power to dictate the policies that can affect o' through itself, and state more precise stronger necessary conditions for the MayAffect$^{Eve}${(o,o') predicate, conditions other than path existence which we have argued.
-To put it in another way, we can now allow a path to exist from o to o', \textit{and} still enforce \loneg MayAffect$^{Eve}$(o,o'), because the negation of these stronger security policy conditions results in weaker antecedents for protection consequence.\\
+Object o, therefore, has the power to dictate the policies that can affect o' through itself (since it holds the direct capability of o'), and state more precise stronger necessary conditions for the MayAffect$^{Eve}${(o,o') predicate.
+To put it in another way, we can now allow a path to exist from o to o', \textit{and} still enforce \loneg MayAffect$^{Eve}$(o,o'), because the negation of these stronger security policy conditions results in weaker antecedents for \loneg MayAffect$^{Eve}$(o,o').
+This can be done simply, be ensuring that object o has the capability of o' stored in some private field but has no behaviours that will make use and call this capability.\\
+
+A key result from this analysis shows the importance of attenuating objects, objects that sit as the last line of defence in the eventual paths to our protected objects.
+In the OCap pattern examples we will go through in the next section, protection of an object is done through some attenuating object that the trusted object knows will never leak its capability directly (the trusted object can ensure this if it creates those attenuating objects), and that protection is enforced through some security policies within those attenuating objects.\\\
+
+If the protected object can guarantee through the system of configuration that those attenuating objects are the \textit{only} objects in the entire system that holds its capability in \textit{all} eventual states, then the protected object can dictate the conditions and degree of control on how it can be used by untrusted objects in all eventual states through those attenuating objects.
+This ability of the protected object to dictate control across time and conditions enables cooperation, where conditions on how the protected object will cooperate is written in those attenuating objects.
+\\
 
 We now see the power of OCap patterns hence lies in providing concrete examples of a system of cooperation that allows the existence of paths between objects for cooperation while still dictating the degree of control of one object can have over another.
-The logics we have developed so far illuminate the power of attenuating objects X* that can enable protection.
-Indeed attenuating objects feature prominently in the literature of OCap patterns which we shall see in the next section.
 
@@
 Consequently, this allows an object which has the capability of any one node in the tree to navigate up to the root node (Document), and consequently navigate to all other nodes in the tree.
+
 %----------------------------------------------------------------------------------------

### 2017-08-17 article.tex
some RNode

@@
 M,\losigma\ \loturns\ MayAffect$^{Eve}$(x,e) \loiff \linebreak
-\ablock \loexists \losigma \loin Arising(M,\losigma).\linebreak
+\ablock \loexists \losigma' \loin Arising(M,\losigma).\linebreak
 \ablock \loexists \loconj{m},\loconj{a},\losigma'. x.\loconj{m}(\loconj{a}) \loleadsto\ \losigma' \loand \loexec{e}{\losigma} \loneq \loexec{e}{\losigma'}
@@
 \ablock \ablock \ablock \quad\: MayAccess$^{Dir,Now}$(X*,o)$]$\linebreak
+\ablock \ablock \ablock \ablock\loor \linebreak
+\ablock \ablock \ablock \quad\: MayCreate$^{Dir,Now}$(o,X*)$]$\linebreak
 \ablock \ablock \ablock \quad \loand \linebreak
@@
 \textsc{npc2} gives us the present path configurations that will allow an eventual path from o to o' as a necessary condition for o to modify o'.
-But if objects are path-isolated, how do we allow them to cooperate?
+\textsc{npc2} tells us that we can enforce protection of o' from o by denying all eventual paths from o to o'.
+But if those objects are path-isolated, how do we allow them to cooperate?
+
+\subsection{What about Cooperation?}
+It is good to know we can enforce protection through some present path configurations that can guarantee eventual path-isolation, but in practice this is still not useful.
+Cooperation between objects require paths to exist between objects so it is rare or almost impossible to have two objects o and o' that cooperate while \loneg MayAccess$^{Ind,Eve}$(o,o') holds.
+The quintessential question in OCap systems is whether we can enforce protection such as controlling whether o may affect o' while \textit{still} allowing paths to exist between them.
+This is possible, because of Rule 3 that says objects may have private encapsulation of state and behaviour.
+To begin, let us assume a simple object o' that has \textit{only} private behaviours.
+In such a scenario, since all fields in our system are also assumed to be private, then we know:
+\begin{logic} \loturns\ \loforall X*\loin Obj, X*\loneq o'. \loneg MayAffect$^{Eve}$(X*,o') \end{logic}
+
+Notice, how protection of o' can occur independent of paths and capability, due to object encapsulation.
+This means that objects can specify security policies that it chooses to enforce \textit{internally}.
+Can we further develop our necessary conditions to show this?\\
 
@@
 No matter how long the path from o to o' is, surely to modify o' requires the last object on our path to \textit{be} the 'final messenger' that will call the object's behaviour to modify it---which means MayAffect$^{Eve}$(Y*,o') \textit{must} hold.
-We express this by extending \textsc{nepc1*} to \textsc{nepc2*}.
+We express this by extending \textsc{nepc1*} to get \textsc{nepc2*}:\linebreak\\
 \begin{minipage}{\linewidth}
 \begin{logic}
-
 \hrule\null
@@
 
-The similarity is that both models allow us to specify policies \textit{within} objects as to whether an object's behaviour can be called (although non OCap systems can do so 'externally' through a global ambient authority).
+The similarity is that both models allow us to specify policies \textit{within} objects as to whether an object's behaviour can be called (non-OCap systems can do so 'externally' through a global ambient authority).
 'Internal' object protection can be done through using defensive programming techniques like encapsulation and checking of conditions for behaviours to succeed.
-In a way these forms of protection can be a form of 'stack-based' access control from the called object's perspective, because they only happen \textit{after} the calling object already has already called the object.\\
+In a way these forms of protection can be a form of 'stack-based' access control from the called object's perspective, because they only happen \textit{after} the calling object has already called the object.\\
 
-How do we then these internal object security policies in our logical framework?
+How do we then express these internal object security policies in our logical framework?
 The nice thing is we already have a predicate MayAffect$^{Eve}$(*,*) that can accomodate this.
@@
 
-\subsection{What about Cooperation?}
-It is good to know we can enforce protection through denying possible path configurations, but in practice this is still not useful.
-Cooperation between objects require paths to exist between objects so it is rare or almost impossible to have two objects o and o' that cooperate while \loneg MayAccess$^{Ind,Eve}$(o,o') holds.
-The quintessential question in OCap systems is whether we can enforce protection such as controlling whether o may affect o' while \textit{still} allowing paths to exist between them.
-This is possible, because of Rule 3 that says objects may have private encapsulation of state and behaviour.
-To begin, let us assume a simple object o' that has \textit{only} private behaviours.
-In such a scenario, since all fields in our system are also assumed to be private, then we know:
-\begin{logic} \loturns\ \loforall X*\loin Obj, X*\loneq o'. \loneg MayAffect(X*,o') \end{logic}
-Notice, how protection of o' can occur independent of paths and capability, due to object encapsulation. This means that objects can specify security policies that it chooses to enforce.
-
 \subsubsection{How can security policies be enforced by objects?}
-In the previous example we have demonstrated in OCap systems that objects can, at the very least enforce protection policies on \textit{itself}.
+In the previous example of an object with all fields and behaviours private, we have demonstrated in OCap systems that objects can, at the very least enforce protection policies on \textit{itself}.
 But can we have objects enforcing security policies on \textit{other} objects?
@@
 
-If we have some starting configuration in a system where there is only one object o that has the capability of o', and no other objects have the capability of o', then we know any eventual path to o' has to include and end with the sub-path from o to o', \textit{unless} o leaks the capability of o' to some other object.
-Notice how the sub-path o to o' is also a direct capability from o to o'.
-Object o, therefore, has the power to dictate the policies that can affect o' through itself (since it holds the direct capability of o'), and state more precise stronger necessary conditions for the MayAffect$^{Eve}${(o,o') predicate.
-To put it in another way, we can now allow a path to exist from o to o', \textit{and} still enforce \loneg MayAffect$^{Eve}$(o,o'), because the negation of these stronger security policy conditions results in weaker antecedents for \loneg MayAffect$^{Eve}$(o,o').
-This can be done simply, be ensuring that object o has the capability of o' stored in some private field but has no behaviours that will make use and call this capability.\\
-
+If we have some starting configuration in a system where there is some object X* that belongs to a set of objects belonging to the same class called AttenObj that has the capability of o', and no other objects have the capability of o', then we know any eventual path to o' has to include and end with the sub-path from X* to o', \textit{unless} X* leaks the capability of o' to some other object.
+If object X* has no way of leaking the capability of object o', then it has the power to dictate the policies in all eventual states that can affect o' through itself (since it holds the direct capability of o'), and state more precise stronger necessary conditions for the MayAffect$^{Eve}${(X*,o') predicate.
+To put it in another way, we can now allow a path to exist from X* to o', \textit{and} still enforce \loneg MayAffect$^{Eve}$(X*,o'), because the negation of these stronger security policy conditions results in weaker antecedents that imply \loneg MayAffect$^{Eve}$(X*,o').
+\begin{itemize}
+\item X* has no behaviour that will make use of and call the capability of o':
+\begin{logic} M,\losigma\ \loturns\ \loforall X*\loin AttenObj. MayAffect$^{Eve}$(X*,o')\linebreak
+\ablock \loimplies $\bot$\end{logic}
+\item X* has only one behaviour that will modify o' if it holds a boolean field \texttt{access} that is evaluated to true:
+\begin{logic} M,\losigma\ \loturns\ \loforall X*\loin AttenObj. MayAffect$^{Eve}$(X*,o')\linebreak
+\ablock \loimplies\linebreak 
+\ablock \loexists \losigma' \loin Arising(M,\losigma).\linebreak
+\ablock \ablock \loexec{x.\texttt{allow}}{\losigma'} = \loexec{\texttt{true}}{\losigma'} \end{logic}
+\end{itemize}
 A key result from this analysis shows the importance of attenuating objects, objects that sit as the last line of defence in the eventual paths to our protected objects.
@@
 
+\begin{logic} 
+\hrule\null
+\textsc{\normalsize *ReNode to Node Condition*}\\
+M,\losigma\ \loturns\ \loforall rn\loin ReNode. MayAffect$^{Eve}$(rn,n)\linebreak
+\ablock \loimplies \linebreak
+\ablock \loexists \losigma' \loin Arising(M,\losigma).\linebreak
+\ablock \ablock \loexec{rn.\texttt{node}}{\losigma'} = \loexec{e}{\losigma'} \loand\ \{\texttt{e is n}\}\linebreak
+\hrule
+\end{logic}
+
+\begin{logic} 
+\hrule\null
+\textsc{\normalsize *ReNode Parenthood Condition*}\\
+M,\losigma\ \loturns\ \loforall rn,rn.parent \loin ReNode. MayAccess$^{Dir,Eve}$(rn,rn.parent)<---Or do we need a MayCreate predicate?\linebreak
+\ablock \loimplies \linebreak
+\ablock \loexists \losigma' \loin Arising(M,\losigma).\linebreak
+\ablock \loand\ \loexec{rn.\texttt{depth}}{\losigma'} = \loexec{e}{\losigma'} \loand\ e > 0\linebreak
+\ablock \loand\ \loexec{rn.parent \texttt{depth}}{\losigma'} = \loexec{e'}{\losigma'} \loand\ e' = e - 1\linebreak
+\hrule
+\end{logic}
 %----------------------------------------------------------------------------------------

### 2017-08-18 article.tex
some RNode
RN

--- 2017-08-17 article.tex
@@
 \item \textbf{Initial Conditions or Endowment:} o has an existing path to o' in \losigma, therefore X* refers to o'
-\item \textbf{Parenthood:} o will create o' in some arising \losigma', therefore X* refers to o
+\item \textbf{Parenthood:} since o can create o' in some arising \losigma', then o can also create o' in \losigma, X* refers to o'
 \item \textbf{Introduction:} o will only obtain the path to o' in some arising \losigma', therefore X* refers to an object that is not the same object as o (X*\loneq o). 
@@
 \end{itemize}
-Note that object x can refer to the same object as o.
+Note that object x can refer to the same object as o'.
 There is one final critical result from \textsc{epc1}.
@@
 \end{itemize}
-Note how, these four configurations are present state configurations at \losigma, but allows us to reason about whether a potential path from o to o' can exist in all arising states \losigma' from \losigma.\\
+Note how, these four configurations are direct capability configurations at \losigma, but allows us to reason about whether a potential path from o to o' can exist in some arising state \losigma' from \losigma.
+This is expressed more succintly below in \textsc{epc3***}
 
-So far we have developed results for paths from o to o', but what about a direct path from o to o'?
+\begin{logic}
+\hrule\null
+\textsc{\normalsize *Eventual Path Connectivity 3*** (EPC3***)}\\
+M,\losigma\ \loturns\ MayAccess$^{Ind,Eve}$(o,o')\linebreak
+\ablock \loimplies\linebreak
+\ablock \loexists X*, x$^{1...n}$\loin Obj.$\{[$MayAccess$^{Ind,Now}$(o,o)\linebreak
+\ablock \ablock \ablock \ablock \ablock \loand\linebreak
+\ablock \ablock \ablock \ablock \ablock MayAccess$^{Dir,Now}$(o,x$^1$)\linebreak
+\ablock \ablock \ablock \ablock \ablock $\vdots$\linebreak
+\ablock \ablock \ablock \ablock \ablock MayAccess$^{Dir,Now}$(x$^{n-1}$,x$^n$)\linebreak
+\ablock \ablock \ablock \ablock \ablock \loand\linebreak
+\ablock \ablock \ablock \ablock \ablock MayAccess$^{Dir,Now}$(x$^n$,X*)$]$\linebreak
+\ablock \ablock \ablock \ablock \quad\: \loor \linebreak
+\ablock \ablock \ablock \ablock \quad\: MayAccess$^{Dir,Now}$(X*,o)$\}$\linebreak
+\ablock \ablock \ablock \ablock \loand \linebreak
+\ablock \ablock \ablock \quad MayAccess$^{Ind,Eve}$(X*,o')\linebreak\\
+\ablock \textit{where if o exists in \losigma,} \linebreak
+\ablock \loturns\ MayAccess$^{Ind,Now}$(o,o) is true\linebreak\linebreak\
+\ablock \textit{where if o' exists in some \losigma',}\linebreak
+\ablock \loturns\ MayAccess$^{Ind,Eve}$(o',o') is true\linebreak
+\hrule
+\hrule
+\end{logic}
+
+So far we have developed results for paths from o to o', but what about a direct path from o to o' (what are the conditions such that o will have the direct capability of o' eventually)?
 Luckily, we only need construct \textsc{eec} from \textsc{epc1}, the only difference being we now require the path from o to X* in \textsc{epc1} to be traversable.
@@
 \end{logic}
-The power of \textsc{eec} is similar to \textsc{epc}\ in that it allows an arbitrary recursive expansion of the MayAccess$^{Dir,Eve}$(X*,o') till the terminating MayAccess$^{Dir,Eve}$(o',o') that is always true.\\
+The power of \textsc{eec} is similar to \textsc{epc1}\ in that it allows an arbitrary recursive expansion of the MayAccess$^{Dir,Eve}$(X*,o') till the terminating MayAccess$^{Dir,Eve}$(o',o').\\
 
-We can always use an automated theorem prover to prove whether a path can arise between two objects in a present configuration system.
 We now turn our attention formalizing Ocap state changes rules in the system.\\
@@
 
-\subsubsection{Protection is about local object-path denial and global capability denial}
+\subsubsection{Protection is path-denial}
 \begin{logic}
@@
 However, the existence of \textit{some} object X* (that can refer to o, or other objects) having the capability of o', is a necessary condition for o to modify o'.
-See \textsc{nec1} and \textsc{nec2} below which illustrates this.
+See \textsc{ncc1} and \textsc{ncc2} below which illustrates this.
 This is precisely why attenuating objects are so powerful and useful in OCap patterns.
@@
 \hrule\null
-\textsc{\normalsize *Necessary Execution Condition 1 (NEC1)}\\
+\textsc{\normalsize *Necessary Capability Condition 1 (NCC1)}\\
 M,\losigma\ \loturns\ MayAffect$^{Eve}$(o,o') \loimplies\linebreak
@@
 \textit{By contraposition and quantifier equivalence,}\linebreak
-M,\losigma\ \loturns\ $[$\loforall X*\loin{Obj}. \loneg MayAccess$^{Dir,Eve}$(X*,o')$]$\loimplies\linebreak
+M,\losigma\ \loturns\ \loforall X*\loin{Obj}. \loneg MayAccess$^{Dir,Eve}$(X*,o')\loimplies\linebreak
 	\ablock \loneg MayAffect$^{Eve}$(o,o') \linebreak
@@
 \end{logic}
-\textsc{nec1} says that in order for an object o to modify the state of o', there must exist an object X* that has a direct path to o', and that X* \textit{can} traverse such a path.
+\textsc{ncc1} says that in order for an object o to modify the state of o', there must exist an object X* that has a direct path to o', and that X* \textit{can} traverse such a path.
 This is derived trivially from our \textsc{pfa} assumption.
@@
 
-In \textsc{nec1}, if X* refers to the same object as o, then we have a straightforward configuration where object o has the capability of o' and can therefore affect o'.
+In \textsc{ncc1}, if X* refers to the same object as o, then we have a straightforward configuration where object o has the capability of o' and can therefore affect o'.
 However, very crucially, o having the capability of o' is \textit{not} a sufficient condition---denying o the capability of o' does \textit{not} deny o the ability to affect o', since X* can refer to an object that is \textit{not} o.
-\textsc{nec2} below is an expansion of \textsc{nec1} to illustrate this crucial point.
+\textsc{ncc2} below is an expansion of \textsc{ncc1} to illustrate this crucial point.
 \begin{logic}
 \hrule\null
-\textsc{\normalsize *Necessary Execution Condition 2 (NEC2)}\\
+\textsc{\normalsize *Necessary Capability Condition 2 (NCC2)}\\
 M,\losigma\ \loturns\ MayAffect$^{Eve}$(o,o') \loimplies\linebreak
@@
 
-So far our results serve as a good base to enforce \loneg MayAffect$^{Eve}(o,o')$ but we require \textit{stronger} necessary conditions for MayAffect$^{Eve}(o,o')$, because we need to understand the present path configurations that can deny o from modifying o'.
-Luckily, we have already built the necessary ingredients in our formalizations of OCap rules in \ref{ocaprules}, where have painstakingly derive the relationship between present path-capability configurations and eventual paths.
-Since \textsc{npc1} tells us that the eventual paths is a necessary condition for object state modification, and \textsc{epc1} tells us the the relationship between eventual paths and present path-capability configurations, we combine the two to present:
+So far our results serve as a good base to enforce \loneg MayAffect$^{Eve}$(o,o') but we require \textit{stronger} necessary conditions for MayAffect$^{Eve}$(o,o'), because we need to understand the present path configurations that can deny o from modifying o'.
+Luckily, we have already built the necessary ingredients in our formalizations of OCap rules in \ref{ocaprules}, where have painstakingly derive the relationship between present path configurations and eventual paths.
+Since \textsc{npc1} tells us that the eventual paths is a necessary condition for object state modification, and \textsc{epc1} tells us the the relationship between eventual paths and present path configurations, we combine the two to present:
 
@@
 \end{logic}
-, where we can always use \textsc{gpec} to expand indirect paths into direct capabilities till we reach the terminating origin o and recursively use the condition to expand the condition till we reach the terminating o'.
-\textsc{npc2} gives us the present path configurations that will allow an eventual path from o to o' as a necessary condition for o to modify o'.But if objects are path-isolated, how do we allow them to cooperate?
-
-\textsc{nec2} and \textsc{npc2} show the necessary capability and path conditions for protection.
-Is there a relation that can capture the essence of both?
-Notice how, we have said in \textsc{nec2} that there must exist some object in the system (that is not necssarily o) that has the capability of o', while in \textsc{npc2}, we said that o must have a path to o'.
-Surely it must mean then that the object X* in \textsc{nec2} that has the capability of o' refers to the last object on the path from o to o' mentioned in \textsc{npc2}.
-We now bring these results together, by making use of \textsc{gpec} to make explicit the last object in the eventual path holding the capability of o' in \textsc{npc2}:\\
+, where \textsc{npc2} is similar to \textsc{epc2}, where we can always use \textsc{gpec} to expand indirect paths into direct capabilities till we reach the terminating origin o and recursively use \textsc{epc2} to expand MayAccess$^{Ind,Eve}$(X*,o') till we reach the terminating MayAccess$^{Ind,Eve}$(o',o').
+The final result through such as expansion is similar to \textsc{epc3***}.
+We use \textsc{epc3***} to make explicit the relationship between a configuration of direct capabilities between objects at \losigma, and the ability to modify state at some arising \losigma', which we call \textsc{ncc3***}:
 \begin{logic}
 \hrule\null
-\textsc{\normalsize *Necessary Execution-Path Condition* (NEPC*)}\\
-M,\losigma\ \loturns\ MayAffect$^{Eve}$(o,o')\linebreak
+\textsc{\normalsize *Necessary Capability Condition 3*** (NCC3***)}\\
+M,\losigma\ \loturns\ MayAffect$^{Ind,Eve}$(o,o')\linebreak
 \ablock \loimplies\linebreak
-\ablock \loexists X*\loin Obj.$[($MayAccess$^{Ind,Now}$(o,X*) \loor \linebreak
-\ablock \ablock \ablock \ablock MayAccess$^{Dir,Now}$(X*,o)$)$\linebreak
+\ablock \loexists X*, x$^{1...n}$\loin Obj.$\{[$MayAccess$^{Ind,Now}$(o,o)\linebreak
+\ablock \ablock \ablock \ablock \ablock \loand\linebreak
+\ablock \ablock \ablock \ablock \ablock MayAccess$^{Dir,Now}$(o,x$^1$)\linebreak
+\ablock \ablock \ablock \ablock \ablock $\vdots$\linebreak
+\ablock \ablock \ablock \ablock \ablock MayAccess$^{Dir,Now}$(x$^{n-1}$,x$^n$)\linebreak
+\ablock \ablock \ablock \ablock \ablock \loand\linebreak
+\ablock \ablock \ablock \ablock \ablock MayAccess$^{Dir,Now}$(x$^n$,X*)$]$\linebreak
+\ablock \ablock \ablock \ablock \quad\: \loor \linebreak
+\ablock \ablock \ablock \ablock \quad\: MayAccess$^{Dir,Now}$(X*,o)$\}$\linebreak
 \ablock \ablock \ablock \ablock \loand \linebreak
-\ablock \loexists Y*\loin{Obj}. $[$MayAccess$^{Ind,Eve}$(X*, Y*) \loand\linebreak
-\ablock \ablock \ablock \ablock MayAccess$^{Dir,Eve}$(Y*,o')$]$
-\linebreak \\
-\textit{By contraposition,}\linebreak
-M,\losigma\ \loturns\ 
-\loforall X*\loin Obj. $[$\loneg MayAccess$^{Ind,Now}$(o,X*) \loand \linebreak
-\ablock \ablock \ablock \ablock \loneg MayAccess$^{Dir,Now}$(X*,o)$]$\linebreak
-\ablock \ablock \ablock \ablock \loor \linebreak
-\ablock \quad\loforall Y*\loin{Obj}. $[$MayAccess$^{Ind,Eve}$(X*, Y*) \loor\linebreak
-\ablock \ablock \ablock \ablock MayAccess$^{Dir,Eve}$(Y*,o')$]$
-\ablock \loimplies \linebreak
-	\ablock\loneg MayAffect$^{Eve}(o,o')$\linebreak
+\ablock \ablock \ablock \quad MayAccess$^{Ind,Eve}$(X*,o')\linebreak\\
+
+\ablock \textit{where if o exists in \losigma,} \linebreak
+\ablock \loturns\ MayAccess$^{Ind,Now}$(o,o) is true\linebreak\linebreak\
+\ablock \textit{where if o' exists in some \losigma',}\linebreak
+\ablock \loturns\ MayAccess$^{Ind,Eve}$(o',o') is true\linebreak
 \hrule
 \end{logic}
-\textsc{nepc*} is nice, because we can see the necessary path condition and the condition that an object must hold the capability of o'.
-But what does it mean to modify an object?
-No matter how long the path from o to o' is, surely to modify o' requires the last object on our path to \textit{be} the 'final messenger' that will call the object's behaviour to modify it---which means MayAffect$^{Eve}$(Y*,o') \textit{must} hold.
-We express this by extending \textsc{nepc1*} to \textsc{nepc2*}.
+
+\textsc{ncc3***} gives us the present \textit{capability} configurations that will allow an eventual path from o to o', which is a necessary condition for o to modify o', and tells us that we can enforce protection of o' from o by denying all eventual paths from o to o'.
+But if those objects must \textit{always} and at all states be path-isolated, how do we allow them to cooperate?
+
+\subsection{What about Cooperation?}
+
+It is good to know we can enforce protection through some present path configurations that can guarantee eventual path-isolation, but in practice this is still not entirely useful.
+The quintessential question in OCap systems is whether we can enforce protection such as controlling whether o may affect o', on \textit{some condition}.
+In our framework, this just means that cooperation between objects require paths to exist based on some condition.
+This is made possible, because of Rule 3 that says objects may have private encapsulation of state and behaviour, which enables an object to deny paths leading in and leading out of it.
+Can our framework incorporate this?\\
+
+How do we then express these internal object security policies in our logical framework?
+
+It is worth revisiting our general specifications for the MayAccess$^{Dir,Eve}$ predicate, which is extremely \textit{weak}, because they have no special conditions:
+\begin{logic}[MayAccess$^{Dir,Eve}$]
+M,\losigma\ \loturns\ MayAccess$^{Dir,Eve}$(x,e) \loiff \linebreak
+\ablock \loexists \losigma' \loin Arising(M,\losigma).\linebreak
+\ablock \loexists f. \loexec{x.f}{\losigma'} = \loexec{e}{\losigma'}
+\loor \linebreak
+\ablock (\losigma'(this) = \loexec{x}{\losigma'} \loand\ \loexists y. \losigma'(y) = \loexec{e}{\losigma'})\linebreak
+\ablock \loand\ $\{\}$\texttt{ // No conditions}
+\end{logic}
+
+The reason why we have not used any special conditions is because this also allows us to generalize any kind of object which we do not know how it behaves.
+The most dangerous kind of objects are those that exposes all their behaviours to be public, and can expose all capabilities that is possesses through those behaviours.
+In that end, our general specifications are meant to assume a worst-case scenario, and hence it is no surprise that we insist on path-isolation to guarantee protection.\\
+
+But, what if we have code knowledge of some set of objects in our system so we know how they behave?\\
+
+Let us assume we know a simple object o' that has \textit{only} private behaviours, and belongs to class called IsoObj that we have code knowledge of.
+In such a scenario, since all fields in our system are also assumed to be private, then we know:
+\begin{logic}[MayAccess$^{Dir,Eve}$(X*, o']
+M,\losigma\ \loturns\ \loforall\ X*\loin Obj, o'\loin IsoObj. MayAccess$^{Dir,Eve}$(X*,o')\linebreak
+\ablock \loiff \linebreak
+\ablock \loexists \losigma' \loin Arising(M,\losigma).\linebreak
+\ablock \loexists f. \loexec{X*.f}{\losigma'} = \loexec{o'}{\losigma'}
+\loor \linebreak
+\ablock (\losigma'(this) = \loexec{X*}{\losigma'} \loand\ \loexists y. \losigma'(y) = \loexec{o'}{\losigma'})\linebreak
+\ablock \loand\ $\bot$. $\{$\texttt{//Condition: Always false}$\}$
+\end{logic}
+Even though in theory any object may possess the capability of o', as long as we know there is no way for anybody to \textit{exercise} the capability of o' in under any condition (because o' has restricted all incoming messaages as it has no behaviours), in terms of security concerns, it \textit{is equivalent} to o' being an isolated object with all objects not having any direct paths to o' in all eventual states.
+That is, o' might very well be an isolated object \textit{despite} every object having the capability of o':
+\begin{itemize}
+\item \begin{logic} M,\losigma\ \loturns\ \loforall X*\loin Obj, o'\loin IsoObj. MayAccess$^{Dir,Eve}$(X*,o')\linebreak 
+\ablock \loiff $\bot$ \end{logic}
+\end{itemize}
+Furthermore, since our results in \textsc{npc1} tells us that path is a necessary condition to modify an object, and \textsc{gpe} says denying \textit{all} capabilities in the system to o' is equivalent to denying all paths to o' in the system, then we know:
+\begin{logic} M,\losigma\ \loturns\ \loforall X*\loin Obj. \loneg MayAccess$^{Dir,Eve}$(X*,o')\linebreak
+\ablock \ablock \loimplies \linebreak
+\ablock \ablock \loneg MayAccess$^{Ind,Eve}$(X*,o')\linebreak
+\ablock \ablock \loimplies \linebreak
+\ablock \ablock \loneg MayAffect$^{Eve}$(X*,o')\linebreak
+\end{logic}
+
+Since \textit{all} objects do not have the path of o', then we know o' is protected.
+To put it in another way, we can now determine paths into o',  based on some condition \textit{written} in o'.
+This means that objects can specify security policies on the incoming messages it receives and enforce protection on \textit{itself} \textit{internally}.
+But, can objects specify security policies \textit{externally} on the outgoing messages to the capabilities that is possesses?
+Of course this is possible.
+Let us see how we further develop this in our framework with another example.\\
+
+We have some starting configuration in a system where there is some object Y* that belongs to a class called AttenObj which we have code knowledge of, and that Y* has the capability of o' that belongs to a class called ProtectObj.
+Note that we do not need code knowledge of ProtectObj to enforce a policy of how Y* might want to interact with o'.
+Since Y* has the capabilitiy of o', we say Y* has the power to enforce the conditions on when it will exercise it.
+To put it in another way, we can now allow an outgoing path to exist from Y* to o', based on some condition \textit{written in Y*}:
+
+\begin{itemize}
+\item If Y* has no behaviour that will make use of and call the capability of o' in any outgoing messages:
+\begin{logic}
+M,\losigma\ \loturns\ \loforall\ Y*\loin AttenObj, o'\loin ProtectObj. MayAccess$^{Dir,Eve}$(Y*,o')\linebreak
+\ablock \loiff \linebreak
+\ablock \loexists \losigma' \loin Arising(M,\losigma).\linebreak
+\ablock \loexists f. \loexec{Y*.f}{\losigma'} = \loexec{o'}{\losigma'}
+\loor \linebreak
+\ablock (\losigma'(this) = \loexec{Y*}{\losigma'} \loand\ \loexists y. \losigma'(y) = \loexec{o'}{\losigma'})\linebreak
+\ablock \loand\ $\bot$. \texttt{//Condition: Always false}
+\end{logic}
+\item Y* has only one behaviour that will use the capability of o' if it holds a boolean field \texttt{access} that is evaluated to true at the time of the access call:
+\begin{logic}
+M,\losigma\ \loturns\ \loforall\ Y*\loin AttenObj, o'\loin ProtectObj. MayAccess$^{Dir,Eve}$(Y*,o')\linebreak
+\ablock \loiff \linebreak
+\ablock \loexists \losigma' \loin Arising(M,\losigma').\linebreak
+\ablock \loexists f. \loexec{Y*.f}{\losigma'} = \loexec{o'}{\losigma'}
+\loor \linebreak
+\ablock (\losigma(this) = \loexec{Y*}{\losigma'} \loand\ \loexists y. \losigma'(y) = \loexec{o'}{\losigma'})\linebreak
+\ablock \loand\ \loexec{x.\texttt{allow}}{\losigma'} = \loexec{\texttt{true}}{\losigma'}.
+\ablock \texttt{//Condition: only when boolean\\ \quad variable allow is true}
+\end{logic}
+\end{itemize}
+Note the difference in this example where we \textit{cannot} say that Y* may not modify o', because direct capability from Y* to o' is \textit{not} a sufficient condition, and Y* does not represent all objects in our system (Y*\loin AttenObj, AttenObj \loin Obj).
+\textit{Y* may still find some \textit{other} path to o' to modify o'}.
+\begin{logic} M,\losigma\ \loturns\ \loforall Y*\loin AttenObj. \loneg MayAccess$^{Dir,Eve}$(Y*,o')\linebreak
+\ablock \ablock \lonimplies \linebreak
+\ablock \ablock \loneg MayAccess$^{Ind,Eve}$(Y*,o')\linebreak
+\ablock \ablock \lonimplies \linebreak
+\ablock \ablock \loneg MayAffect$^{Eve}$(Y*,o')\linebreak
+\end{logic}
+
+We can now make use of our result of Domination, to say that if in the system of configuration all paths to o' must go through a set of X*, or conversely, Y* are the only objects that hold the direct capability of o', then we can reason:
+
+\begin{logic} M,\losigma\ \loturns\ \loforall X*\loin Obj, Y*\loin AttenObj, AttenObj \loin Obj. \linebreak
+\ablock \ablock Dom(Y*,o') \loand\ MayAccess$^{Ind,Eve}$(X*,o')\linebreak
+\ablock \ablock \loimplies \linebreak
+\ablock \ablock MayAccess$^{Dir,Eve}$(Y*,o')\linebreak\\
+
+\textit{By contraposition,}\linebreak
+\ablock \ablock \loneg MayAccess$^{Dir,Eve}$(Y*,o')\linebreak
+\ablock \ablock \loimplies \linebreak
+\ablock \ablock \loneg Dom(Y*,o') \loor\ \loneg MayAccess$^{Ind,Eve}$(X*,o')\linebreak\\
+
+\textit{Imposing dominance in antecedent,}\linebreak
+\ablock \ablock \loneg MayAccess$^{Dir,Eve}$(Y*,o') \loand\ Dom(Y*,o')\linebreak
+\ablock \ablock \loimplies \linebreak
+\ablock \ablock \loneg MayAccess$^{Ind,Eve}$(X*,o')\linebreak
+\ablock \ablock \loimplies \linebreak
+\ablock \ablock \loneg MayAccess$^{Ind,Eve}$(Y*,o') \textit{, since Y*\loin X*}\linebreak
+\ablock \ablock \loand\ MayAffect$^{Eve}$(X*,o')\textit{, from \textsc{npc1}}\linebreak
+\ablock \ablock \loimplies \linebreak
+\ablock \ablock \loneg MayAffect$^{Eve}$(Y*,o')\linebreak
+\end{logic}
+This means that if we deny Y* an eventual direct path to o' \textit{and} Y* dominates o', then equivalently, we deny all objects an eventual path to o' and deny all objects, including Y*, the ability to modify o'.\\
+
+We can also show dominance by extending our necessary capability condition \textsc{ncc3***}.
+If we can guarantee that Y* is the only object in the system that holds the direct capability of o', and we deny Y* a direct path to o', then we know the last object in the eventual path to o' is Y* and we can simply expand the last term in \textsc{ncc3***} to state explicitly that Y* has to be the object that connects directly eventually to o':
 \begin{minipage}{\linewidth}
 \begin{logic}
+\hrule\null
+\textsc{\normalsize *Necessary Dominating Condition*** (NDC***)}\\
+If Y* dominates o' then:\\
+M,\losigma\ \loturns\ MayAffect$^{Ind,Eve}$(o,o') \loand\ Dom(Y*,o')\linebreak
+\ablock \loimplies\linebreak
+\ablock \loexists X*, x$^{1...n}$\loin Obj.$\{[$MayAccess$^{Ind,Now}$(o,o)\linebreak
+\ablock \ablock \ablock \ablock \ablock \loand\linebreak
+\ablock \ablock \ablock \ablock \ablock MayAccess$^{Dir,Now}$(o,x$^1$)\linebreak
+\ablock \ablock \ablock \ablock \ablock $\vdots$\linebreak
+\ablock \ablock \ablock \ablock \ablock MayAccess$^{Dir,Now}$(x$^{n-1}$,x$^n$)\linebreak
+\ablock \ablock \ablock \ablock \ablock \loand\linebreak
+\ablock \ablock \ablock \ablock \ablock MayAccess$^{Dir,Now}$(x$^n$,X*)$]$\linebreak
+\ablock \ablock \ablock \ablock \quad\: \loor \linebreak
+\ablock \ablock \ablock \ablock \quad\: MayAccess$^{Dir,Now}$(X*,o)$\}$\linebreak
+\ablock \ablock \ablock \ablock \loand \linebreak
+\ablock \ablock \loexists Y*\loin{AttenObj}.$[$MayAccess$^{Ind,Eve}$(X*, Y*) \loand\linebreak
+\ablock \ablock \ablock \ablock \ablock \quad MayAccess$^{Dir,Eve}$(Y*,o')$]$
 
-\hrule\null
-\textsc{\normalsize *Necessary Execution-Path Condition 2* (NEPC2*)}\\
-M,\losigma\ \loturns\ MayAffect$^{Eve}$(o,o')\linebreak
-\ablock \loimplies\linebreak
-\ablock \loexists X*\loin Obj.$[($MayAccess$^{Ind,Now}$(o,X*) \loor \linebreak
-\ablock \ablock \ablock \quad MayAccess$^{Dir,Now}$(X*,o)$)$\linebreak
-\ablock \ablock \ablock \ablock \loand \linebreak
-\ablock \loexists Y*\loin{Obj}. $[$MayAccess$^{Ind,Eve}$(X*, Y*) \loand\linebreak
-\ablock \ablock \ablock \quad MayAccess$^{Dir,Eve}$(Y*,o') \loand\linebreak
-\ablock \ablock \ablock \quad MayAffect$^{Eve}$(Y*,o')$]$
-\linebreak \\
-\textit{By contraposition,}\linebreak
-M,\losigma\ \loturns\ 
-\loforall X*\loin Obj. $[$\loneg MayAccess$^{Ind,Now}$(o,X*) \loand \linebreak
-\ablock \ablock \ablock \ablock \loneg MayAccess$^{Dir,Now}$(X*,o)$]$\linebreak
-\ablock \ablock \ablock \ablock \loor \linebreak
-\ablock \quad\loforall Y*\loin{Obj}. $[$MayAccess$^{Ind,Eve}$(X*, Y*) \loor\linebreak
-\ablock \ablock \ablock \ablock MayAccess$^{Dir,Eve}$(Y*,o') \loor\linebreak
-\ablock \ablock \ablock \ablock MayAffect$^{Eve}$(Y*,o')$]$
-\ablock \loimplies \linebreak
-	\ablock\loneg MayAffect$^{Eve}(o,o')$\linebreak
-\hrule\null
+\ablock \textit{where if o exists in \losigma,} \linebreak
+\ablock \loturns\ MayAccess$^{Ind,Now}$(o,o) is true\linebreak\linebreak\
+\ablock \textit{where if o' exists in some \losigma',}\linebreak
+\ablock \loturns\ MayAccess$^{Dir,Eve}$(o',o') is true\linebreak
+\hrule
 \end{logic}
-\end{minipage}
-\linebreak\\
-
-\textsc{nepc2*} has the intuitive meaning that in order to say object o may modify object o' in some eventual state, \textit{all} these three conditions must hold:
-\begin{itemize}
-\item there must be some present path configuration such that an eventual path from o to o' can arise
-\item there must be an object Y* that sits as the last object along the eventual path to o', that has the capability of o'
-\item object Y* must be able to modify o'
-\end{itemize}
+\end{minipage}\linebreak\\
+Our analysis shows the importance of attenuating objects---objects that perform their roles as \textit{defensive} objects along the eventual paths to our protected objects.
+If the protected object can guarantee through the system of configuration that those attenuating objects are the \textit{only} objects in the entire system that holds its capability in \textit{all} eventual states, then the protected object can dictate the conditions and degree of control on how it can be used by untrusted objects in all eventual states through those attenuating objects.
+This ability of the protected object to dictate control across time and conditions enables cooperation, where conditions on how the protected object will cooperate is written in those attenuating objects.\\
 
@@
 \item being careful about our present path configurations so that we can ensure eventual path-isolation between the object and untrusted object
-\item specifying security policies on how \textit{all} objects with the direct capability of our trusted object may modify the object
+\item specifying security policies {within} objects that dictate how incoming messages and outgoing messages are handled, so that path-isolation or existence rests on some condition
 \end{itemize}
-Therein lies also the difference between OCap and non-OCap models, because non-OCap models that allow forging of capabilities or a global ambient authoriy \textit{cannot} enforce security through path isolation, which is a distinctive feature of OCap models.
-How do we then expect objects with the direct capability of o' may modify o' through some security policy?
-The nice thing is we already have a predicate MayAffect$^{Eve}$(*,*) that can accomodate these security policies.
-Notice how in \textsc{NEPC*}, object Y* sits as the last object in the path from o to o' and has the capability of o'.
-This gives us a clue as to how we can enable protection of o' without denying paths and writing security policies that allow
+Therein lies also the difference between OCap and non-OCap models, because non-OCap models that allow forging of capabilities or a global ambient authoriy \textit{cannot} enforce security through path isolation, which is a distinctive feature of OCap models.\\
 
-\subsection{What about Cooperation?}
-At this juncture, it is good to know we can enforce protection through denying possible path configurations, but in practice this is still not useful.
-Cooperation between objects require paths to exist between objects so it is rare or almost impossible to have two objects o and o' that cooperate while \loneg MayAccess$^{Ind,Eve}$(o,o') holds.
-The quintessential question in OCap systems is whether we can enforce protection such as controlling whether o may affect o' while \textit{still} allowing paths to exist between them.
-This is possible, because of Rule 3 that says objects may have private encapsulation of state and behaviour.
-To begin, let us assume a simple object o' that has \textit{only} private behaviours.
-In such a scenario, since all fields in our system are also assumed to be private, then we know:
-\begin{logic} \loturns\ \loforall X*\loin Obj, X*\loneq o'. \loneg MayAffect(X*,o') \end{logic}
-Notice, how protection of o' can occur independent of paths and capability, due to object encapsulation. This means that objects can specify security policies that it chooses to enforce.
+The similarity is that both models allow us to specify policies \textit{within} objects as to whether an object's behaviour can be called (note that non-OCap systems can also do so 'externally' through a global ambient authority).
+'Internal' object protection can be done through using defensive programming techniques like encapsulation and checking of conditions for behaviours to succeed.
+In a way these forms of protection can be considered a form of 'stack-based' access control from the called object's perspective, because they only happen \textit{after} the calling object has already called the object.\\
 
-\subsubsection{How can security policies be enforced by objects?}
-In the previous example we have demonstrated in OCap systems that objects can, at the very least enforce protection policies on \textit{itself}.
-But can we have objects enforcing security policies on \textit{other} objects?
-Let us examine this with another example.\\
-
-If we have some starting configuration in a system where there is only one object o that has the capability of o', and no other objects have the capability of o', then we know any eventual path to o' has to include and end with the sub-path from o to o', \textit{unless} o leaks the capability of o' to some other object X*.
-Notice how the sub-path o to o' is also a direct capability from o to o'.
-Object o, therefore, has the power to dictate the policies that can affect o' through itself, and state more precise stronger necessary conditions for the MayAffect$^{Eve}${(o,o') predicate, conditions other than path existence which we have argued.
-To put it in another way, we can now allow a path to exist from o to o', \textit{and} still enforce \loneg MayAffect$^{Eve}$(o,o'), because the negation of these stronger security policy conditions results in weaker antecedents for protection consequence.\\
-
-We now see the power of OCap patterns hence lies in providing concrete examples of a system of cooperation that allows the existence of paths between objects for cooperation while still dictating the degree of control of one object can have over another.
-The logics we have developed so far illuminate the power of attenuating objects X* that can enable protection.
-Indeed attenuating objects feature prominently in the literature of OCap patterns which we shall see in the next section.
+In the OCap pattern examples we will go through in the next section, protection of an object is typically accomplished through some attenuating object that the protected object knows will never leak its capability directly such that the attenuating objects dominate the trusted object (the trusted object can ensure this if it creates those attenuating objects), and that protection is enforced through some security policies \textit{within} those attenuating objects.\\
 
@@
 A node also has a public field \texttt{parent} that will return the capability of its parent node.
-Consequently, this allows an object which has the capability of any one node in the tree to navigate up to the root node (Document), and consequently navigate to all other nodes in the tree.
+Consequently, this allows an object which has the capability of any one node in the tree to navigate up to the root node (Document), and consequently navigate to all other nodes in the tree.\\
+
+\begin{minipage}{\linewidth}
+\begin{logic}
+\hrule\null
+\textsc{\normalsize *Necessary Node Condition***}\\
+If Y* dominates o' then:\\
+M,\losigma\ \loturns\ \loforall o\loin Obj, rn\loin ReNode, n\loin Node\linebreak
+ReNode\loin Obj, Node\loin Obj.\linebreak
+\ablock MayAffect$^{Eve}$(o,n) \loand\ Dom(RN*,n)\linebreak
+\ablock \loimplies\linebreak
+\ablock \loexists X*\loin Obj.$\{[$MayAccess$^{Ind,Eve}$(o,X*)\linebreak
+\ablock \ablock \ablock \ablock \quad\: \loor \linebreak
+\ablock \ablock \ablock \ablock \quad\: MayAccess$^{Dir,Now}$(X*,o)$\}$\linebreak
+\ablock \ablock \ablock \ablock \loand \linebreak
+\ablock \ablock \loexists rn\loin{RN*}.$[$MayAccess$^{Ind,Eve}$(X*, rn) \loand\linebreak
+\ablock \ablock \ablock \ablock \ablock \quad MayAccess$^{Dir,Eve}$(rn,o')$]$
+
+\ablock \textit{where if o exists in \losigma,} \linebreak
+\ablock \loturns\ MayAccess$^{Ind,Now}$(o,o) is true\linebreak\linebreak\
+\ablock \textit{where if o' exists in some \losigma',}\linebreak
+\ablock \loturns\ MayAccess$^{Dir,Eve}$(o',o') is true\linebreak
+\hrule
+
+\hrule
+\end{logic}
+\end{minipage}\linebreak\\
+\begin{logic} 
+\hrule\null
+\textsc{\normalsize *ReNode to Node Condition*}\\
+M,\losigma\ \loturns\ \loforall rn\loin ReNode. MayAccess$^{Dir,Eve}$(rn,n)\linebreak
+\ablock \loimplies \linebreak
+\ablock \loexists \losigma' \loin Arising(M,\losigma).\linebreak
+\ablock \ablock \loexec{rn.\texttt{node}}{\losigma'} = \loexec{e}{\losigma'} \loand\ \{\texttt{e is n}\}\linebreak
+\hrule
+\end{logic}
+
+\begin{logic} 
+\hrule\null
+\textsc{\normalsize *ReNode Parenthood Condition*}\\
+M,\losigma\ \loturns\ \loforall rn,rn.parent \loin ReNode. MayAccess$^{Dir,Eve}$(rn,rn.parent)\linebreak
+\ablock \loimplies \linebreak
+\ablock \loexists \losigma' \loin Arising(M,\losigma).\linebreak
+\ablock \loand\ \loexec{rn.\texttt{depth}}{\losigma'} = \loexec{e}{\losigma'} \loand\ e > 0\linebreak
+\hrule
+\end{logic}
+
+\begin{logic} 
+\hrule\null
+\textsc{\normalsize *ReNode Definition*}\\
+M,\losigma\ \loturns\ \loforall rn,rn' \loin ReNode, n,n' \loin Node.\linebreak
+\ablock \loforall \losigma' \loin Arising(M,\losigma).\linebreak
+\ablock \loexec{rn.\texttt{depth}}{\losigma'} = \loexec{d}{\losigma'} \loand \linebreak
+\ablock \loexec{rn.\texttt{node}}{\losigma'} = \loexec{n}{\losigma'} \loand \linebreak
+\ablock \loexec{rn.\texttt{parent}}{\losigma'} = \loexec{rn'}{\losigma'} \loand\linebreak
+\ablock \loexec{rn'.\texttt{depth}}{\losigma'} = \loexec{d'}{\losigma'} \loand\ d' = d - 1\linebreak
+\ablock \loexec{rn'.\texttt{node}}{\losigma'} = \loexec{n'}{\losigma'} \loand\ n' = \loexec{n.\texttt{parent}}{\losigma'}\linebreak
+\hrule
+\end{logic}
+\begin{logic} 
+\hrule\null
+\textsc{\normalsize *Node Definition*}\\
+M,\losigma\ \loturns\ \loforall n,n.parent \loin Node.\linebreak
+\ablock \loforall \losigma' \loin Arising(M,\losigma).\linebreak
+\ablock \loexec{n.\texttt{parent}}{\losigma'} = \loexec{p}{\losigma'} \loand \linebreak
+\hrule
+\end{logic}
+
 %----------------------------------------------------------------------------------------

### 2017-08-18 article.tex
update

@@
 \ablock \ablock \ablock \quad MayAccess$^{Dir,Now}$(X*,o')$]$
-\linebreak\\
+\\
+\ablock \textit{where if z exists in \losigma,} \linebreak
+\ablock \loturns\ MayAccess$^{Ind,Now}$(z,z) is true \linebreak
+\\
 \textit{By contraposition,} \\
@@
 \ablock \loiff\linebreak
-\ablock \loneg MayAccess$^{Ind,Now}$(o,o')\linebreak
+\ablock \loneg MayAccess$^{Ind,Now}$(o,o')
+
 \hrule
@@
 We now look for a relation between a path configuration at a state \losigma\ ($Now$) and eventual path configurations in states \losigma' arising from \losigma\ ($Eve$).\vfill
-\begin{minipage}[b]{\linewidth}
+\begin{minipage}{\linewidth}
 \begin{logic}
@@
 \ablock \ablock \ablock \quad MayAccess$^{Ind,Eve}$(X*,o')$\}$
-\linebreak \\
+\\
+\ablock \textit{where if z' exists in \losigma,} \linebreak
+\ablock \loturns\ MayAccess$^{Ind,Eve}$(z',z') is true \linebreak
+\\
 \textit{By contraposition,}\linebreak
@@
 \ablock \loimplies\linebreak
-	\ablock\loneg MayAccess$^{Ind,Eve}(o,o')$\linebreak
+	\ablock\loneg MayAccess$^{Ind,Eve}$(o,o')\linebreak
 \hrule
@@
 \ablock \ablock $\}$
-\linebreak \\
+\\
+\ablock \textit{where if z' exists in \losigma,} \linebreak
+\ablock \loturns\ MayAccess$^{Ind,Eve}$(z',z') is true \linebreak
+\\
 \textit{By contraposition,}\linebreak
@@
 \ablock \ablock \ablock \ablock \loand \linebreak
-\ablock \ablock \ablock \quad MayAccess$^{Ind,Eve}$(X*,o')\linebreak\\
-\ablock \textit{where if o exists in \losigma,} \linebreak
-\ablock \loturns\ MayAccess$^{Ind,Now}$(o,o) is true\linebreak\linebreak\
-\ablock \textit{where if o' exists in some \losigma',}\linebreak
-\ablock \loturns\ MayAccess$^{Ind,Eve}$(o',o') is true\linebreak
-\hrule
+\ablock \ablock \ablock \quad MayAccess$^{Ind,Eve}$(X*,o')
+\\
+\ablock \textit{where if z' exists in \losigma,} \linebreak
+\ablock \loturns\ MayAccess$^{Ind,Eve}$(z',z') is true \linebreak
+\\
 \hrule
@@
 \ablock \ablock \ablock \ablock MayAccess$^{Dir,Eve}$(X*,o')$]$
-\linebreak \\
+\\
+\ablock \textit{where if o' exists in some \losigma',}\linebreak
+\ablock \loturns\ MayAccess$^{Dir,Eve}$(o',o') is true\linebreak
+\\
 \textit{By contraposition,}\linebreak
@@
 \ablock \ablock \ablock \ablock \loand \linebreak
-\ablock \ablock \ablock \quad MayAccess$^{Ind,Eve}$(X*,o')\linebreak\\
-
-\ablock \textit{where if o exists in \losigma,} \linebreak
-\ablock \loturns\ MayAccess$^{Ind,Now}$(o,o) is true\linebreak\linebreak\
-\ablock \textit{where if o' exists in some \losigma',}\linebreak
-\ablock \loturns\ MayAccess$^{Ind,Eve}$(o',o') is true\linebreak
+\ablock \ablock \ablock \quad MayAccess$^{Ind,Eve}$(X*,o')
+\linebreak\\
 \hrule
@@
 \ablock \ablock \ablock \ablock \loand \linebreak
-\ablock \ablock \loexists Y*\loin{AttenObj}.$[$MayAccess$^{Ind,Eve}$(X*, Y*) \loand\linebreak
+\ablock \ablock \loexists Y*\loin{AttenObj}.$[$MayAccess$^{Ind,Eve}$(X*,Y*) \loand\linebreak
 \ablock \ablock \ablock \ablock \ablock \quad MayAccess$^{Dir,Eve}$(Y*,o')$]$
-
-\ablock \textit{where if o exists in \losigma,} \linebreak
-\ablock \loturns\ MayAccess$^{Ind,Now}$(o,o) is true\linebreak\linebreak\
-\ablock \textit{where if o' exists in some \losigma',}\linebreak
-\ablock \loturns\ MayAccess$^{Dir,Eve}$(o',o') is true\linebreak
+\linebreak\\
 \hrule
@@
 M,\losigma\ \loturns\ \loforall o\loin Obj, rn\loin ReNode, n\loin Node\linebreak
-ReNode\loin Obj, Node\loin Obj.\linebreak
+ReNode\loin Obj, Node\loin Obj, RN*\losubseteq ReNode.\linebreak
 \ablock MayAffect$^{Eve}$(o,n) \loand\ Dom(RN*,n)\linebreak
 \ablock \loimplies\linebreak
-\ablock \loexists X*\loin Obj.$\{[$MayAccess$^{Ind,Eve}$(o,X*)\linebreak
+\ablock \loexists X*\loin Obj.$[$MayAccess$^{Ind,Eve}$(o,X*)\linebreak
 \ablock \ablock \ablock \ablock \quad\: \loor \linebreak
-\ablock \ablock \ablock \ablock \quad\: MayAccess$^{Dir,Now}$(X*,o)$\}$\linebreak
+\ablock \ablock \ablock \: MayAccess$^{Dir,Now}$(X*,o)$]$\linebreak
 \ablock \ablock \ablock \ablock \loand \linebreak
-\ablock \ablock \loexists rn\loin{RN*}.$[$MayAccess$^{Ind,Eve}$(X*, rn) \loand\linebreak
-\ablock \ablock \ablock \ablock \ablock \quad MayAccess$^{Dir,Eve}$(rn,o')$]$
-
-\ablock \textit{where if o exists in \losigma,} \linebreak
-\ablock \loturns\ MayAccess$^{Ind,Now}$(o,o) is true\linebreak\linebreak\
-\ablock \textit{where if o' exists in some \losigma',}\linebreak
-\ablock \loturns\ MayAccess$^{Dir,Eve}$(o',o') is true\linebreak
-\hrule
+\ablock \ablock \loexists rn\loin{RN*}.$[$MayAccess$^{Ind,Eve}$(X*,rn) \loand\linebreak
+\ablock \ablock \ablock \ablock \quad MayAccess$^{Dir,Eve}$(rn,o')$]$\linebreak\\
 
@@
 \end{minipage}\linebreak\\
+\begin{minipage}{\linewidth}
+\begin{logic}
+\hrule\null
+\textsc{\normalsize *Eventual Path Connectivity 1 (EPC1)}\\
+M,\losigma\ \loturns\ MayAccess$^{Ind,Eve}$(z,z')\linebreak
+\ablock \loimplies\linebreak
+\ablock \loexists X*\loin Obj. $\{[$MayAccess$^{Ind,Now}$(z,X*)\linebreak
+\ablock \ablock \ablock \ablock\loor \linebreak
+\ablock \ablock \ablock \quad\: MayAccess$^{Dir,Now}$(X*,z)$]$\linebreak
+\ablock \ablock \ablock \quad \loand \linebreak
+\ablock \ablock \ablock \quad MayAccess$^{Ind,Eve}$(X*,z')$\}$
+\\
+\ablock \textit{where if z' exists in some \losigma',}\linebreak
+\ablock \loturns\ MayAccess$^{Ind,Eve}$(z',z') is true
+\linebreak \\
+\hrule
+\end{logic}
+\end{minipage}
+\begin{logic}
+\hrule\null
+\textsc{\normalsize *Global Path-Execution Connect. (GPEC)}\\
+M,\losigma\ \loturns\ MayAccess$^{Ind,Now}$(z,z')\linebreak
+\ablock \loiff\linebreak
+\ablock \loexists X*\loin{Obj}. $[$MayAccess$^{Ind,Now}$(z, X*) \loand\linebreak
+\ablock \ablock \ablock \quad MayAccess$^{Dir,Now}$(X*,z')$]$
+\\
+\ablock \textit{where if z exists in \losigma,} \linebreak
+\ablock \loturns\ MayAccess$^{Ind,Now}$(z,z) is true
+\linebreak \\
+\hrule
+\end{logic}
+
 \begin{logic} 

### 2017-08-23 article.tex
update

--- 2017-08-17 article.tex
@@
 \usepackage{microtype} % Slightly tweak font spacing for aesthetics
-
+\usepackage{float}
 \usepackage[english]{babel} % Language hyphenation and typographical rules
@@
 \usepackage{color}
+\usepackage{pdfpages}
 \setlist[itemize]{noitemsep} % Make itemize lists more compact
-
+\usepackage{ latexsym }
 \usepackage{abstract} % Allows abstract customization
@@
 \setlength\parindent{0pt}
-
 %----------------------------------------------------------------------------------------
@@
 %environment formating
-\newenvironment{logic}[1][]
-{\begin{flushleft} \small }
-{\end{flushleft}}
+\newenvironment{logic}
+{\begin{minipage}[c]{\linewidth}  \small \vspace{0.5em}\begin{tabbing}}
+{\end{tabbing}\end{minipage}\vspace{0.5em}}
 %logic shortcuts
@@
 \newcommand{\loin}{$\in$}
+\newcommand{\lonin}{$\centernot\in$}
 \newcommand{\loforall}{$\forall$}
+\newcommand{\locontrad}{$\bot$}
 \newcommand{\loexists}{$\exists$}
@@
 \newcommand{\lonimpliedby}{$\centernot\Longleftarrow$}
-\newcommand{\losigma}{$\upsigma$}
+\newcommand{\losigma}{\text{$\upsigma$}}
 \newcommand{\loturns} {$\vDash$}
@@
 \newcommand{\ablock} {\null\qquad}
-
+\newcommand{\lotiff} {\textit{\textls[-20]{iff}}}
+\newcommand{\hr}{\rule{\linewidth}{0.4pt}}
 %-----------------------
@@
 \begin{document}
+\raggedbottom
 \renewcommand{\ref}[1]{\textit{\autoref{#1}}}
@@
 \item failure to defend against malicious execution of \textit{unknown} code components\end{itemize}
-\noindent These two issue are often closely intertwined in any system of execution that has both trusted and untrusted code components (the second issue is often a result of the first).
+\noindent These two issue are often closely intertwined in any system of execution that has both trusted and untrusted code components (the second issue is often a result of the first).\\
 
@@
 \section{OCap Model}
+\subsection{From Object to Capability}
 The OCap model uses the reference graph of the objects as the access graph, and strictly requires objects to interact with each other only by sending messages on object references\cite{miller2003b}.
+When these references cannot be forged in the system, and when there there are no default globally accessible objects (for an object to be globally accessible, the reference to use that object has to be \textit{explicitly} granted to every object globally), then it becomes necessary that one needs to obtain references to objects in order to call the associated methods of the object.
+Object references therefore embody authority, and we use the word capability to describe the encapsulation of both reference (designation) and authority which are unforgeable.
+
 \subsection{From Capability to Object-Capability}
+Origins of capabilities date back to Dennis and Van Horn\cite{dennis1966}'supervisor as a mechanism of protecting low-level resources like memory segments.
+Early attempts to implement the capability model include the MIT PDP-1 computer and the CAL-TSS system.
+The capability model was then extended by the early operating system designers of HYDRA\cite{wulf1974} into a protected ability to invoke arbitrary services provided by other processes.
+Capabilities were later implemented on computer and operating systems such as CAP\cite{wilkes1979}, KeyKOS\cite{hardy1985}, and EROS\cite{shapiro1999}.
+More recent work includes the Capsicum kernel framework that provides capabilities support for UNIX\cite{watson2010}, and Google's open-source capabilities-based operating system Fuchsia OS\cite{matte2017}.
+The development of capabilities and a thorough examination of the object-capability model with the capability-safe programming language E, can be found in Miller's PhD thesis\cite{miller2006}.
+
 \subsection{OCap Languages}
@@
 Objects interact with each other by sending messages on capabilities.
-An OCap pattern may be visualised as a directed graph---nodes represent objects, and each edge from an object \textit{o} to another \textit{o'} represents \textit{o} holding a capability that allows it to directly access \textit{o'}.
+An OCap pattern may be visualised as a directed graph---nodes represent objects, and each edge from an object o to another o' represents o holding a capability that allows it to directly access o'.
 %----------------------------------------------------------------------------------------
@@
 %----------------------------------------------------------------------------------------
+\newpage
 \section{Formal specifications}
@@
 The heap maps addresses to objects.
-Objects are tuples consisting of the class of the object, and a mapping from field identifiers onto values.
+Objects are tuples consisting of the class of the object, and a mapping from field identifiers onto values.\\
+
  \begin{logic}[Runtime state]
@@
 StackId = \{\textit{this}\} \locup\ VarId \locup\ ParId \\
-object = ClassId \lotimes (FieldId $\rightarrow$ val)
+object = ClassId \lotimes (FieldId $\rightarrow$ val)\\
 \end{logic}
+
 \textbf{Module:}
-\begin{logic}[Module]
-M \loin\ Module = ClassId \locup\ SpecId\linebreak\ablock\qquad\qquad\quad $\rightarrow$ \linebreak
+
+\begin{logic}
+M \loin\ Module = ClassId \locup\ SpecId\\\ablock\qquad\qquad\quad $\rightarrow$ \\
 \ablock \qquad \qquad \quad (ClassDescr \locup\ Specification)
@@
 
-\textbf{Reach and Execution:}\\
-\begin{logic}[Reach and Execution]\end{logic}
+\textbf{Execution:}\\
+\begin{logic}\end{logic}
 
-\textbf{Arising Configurations}\\
-\begin{logic}[Arising]\end{logic}
-
-\textbf{Domination:}\\
-\begin{logic}[Domination]\end{logic}
+\textbf{Arising State Configurations}\\
+\begin{logic}\end{logic}
 
 \subsubsection{MayAccess Definitions}
-We define in total four flavours of MayAccess predicates that describe the relation between two entities in a system (of arity 2 that represent the identifiers of these entities).
-These four flavours represent a combination of space (distance) and time (state):
+We define in total four flavours of \textit{MayAccess} predicates that are inspired by the work in Drossopoulou et al.\cite{drossopoulou2015b}, but our four MayAccess predicates extend and modify the definition in \cite{drossopoulou2015b} to more precisely capture the different configurations of access in terms of mode and time:
 \begin{itemize}
-\item distance: directly ($Dir$) or indirectly ($Ind$)
-\item state: now ($Now$) or eventually ($Eve$)
+\item Mode: directly ($Dir$) or indirectly ($Ind$)
+\item Time: now ($Now$) or eventually ($Eve$)
 \end{itemize}
 and are broad enough to describe both non-OCap and OCap models:
-\begin{logic}[MayAccessDirNow]
-M,\losigma\ \loturns\ MayAccess$^{Dir,Now}$(x,e) \loiff \linebreak
-\ablock \loexists f. \loexec{x.f}{\losigma} = \loexec{e}{\losigma}
-\loor \linebreak
-\ablock (\losigma(this) = \loexec{x}{\losigma} \loand\ \loexists y. \losigma(y) = \loexec{e}{\losigma})
-\end{logic}
-\begin{logic}[MayAccessDirEve]
- M,\losigma\ \loturns\ MayAccess$^{Dir,Eve}$(x,e) \loiff \linebreak
-\ablock \loexists \losigma' \loin Arising(M,\losigma).\linebreak
-\ablock M,\losigma'\ \loturns\ MayAccess$^{Dir,Now}$(x,e) 
-\end{logic}
-\begin{logic}[MayAccessIndNow]
-M,\losigma\ \loturns\ MayAccess$^{Ind,Now}$(x,e) \loiff \linebreak
-\ablock \loexists \loconj{f}. \loexec{x.\loconj{f}}{\losigma} = \loexec{e}{\losigma}
-\loor \linebreak
-\ablock (\losigma(this) = \loexec{x}{\losigma} \loand\ \loexists y. \losigma(y.\loconj{f}) = \loexec{e}{\losigma})
-\end{logic}
-\begin{logic}[MayAccessIndEve]
-M,\losigma\ \loturns\ MayAccess$^{Ind,Eve}$(x,e) \loiff \linebreak
-\ablock \loexists \losigma' \loin Arising(M,\losigma).\linebreak
-\ablock M,\losigma'\ \loturns\ MayAccess$^{Ind,Now}$(x,e) 
+
+\begin{logic}
+\hr\\
+\textsc{\normalsize *Four MayAccess Definitions}\\\vspace{0.5em}
+M,\losigma\ \loturns\ \=MayAccess$^{Dir,Now}$(x,e) \loiff \\
+\>\loexists f. \loexec{x.f}{\losigma} = \loexec{e}{\losigma}
+\loor \\
+\> (\losigma(this) = \loexec{x}{\losigma} \loand\ \loexists y. \losigma(y) = \loexec{e}{\losigma})\\
+\\
+ M,\losigma\ \loturns\ \=MayAccess$^{Dir,Eve}$(x,e) \loiff \\
+\>\loexists \losigma' \loin Arising(M,\losigma).\\
+\>M,\losigma'\ \loturns\ MayAccess$^{Dir,Now}$(x,e)\\
+\\
+M,\losigma\ \loturns\ \=MayAccess$^{Ind,Now}$(x,e) \loiff \\
+\>\loexists \loconj{f}. \loexec{x.\loconj{f}}{\losigma} = \loexec{e}{\losigma}
+\loor \\
+\>(\losigma(this) = \loexec{x}{\losigma} \loand\ \loexists y. \losigma(y.\loconj{f}) = \loexec{e}{\losigma})\\
+\\
+M,\losigma\ \loturns\ \=MayAccess$^{Ind,Eve}$(x,e) \loiff \\
+\>\loexists \losigma' \loin Arising(M,\losigma).\\
+\>M,\losigma'\ \loturns\ MayAccess$^{Ind,Now}$(x,e)\\
+\hr
 \end{logic}
 
-A note on \textit{f} and \textit{\loconj{f}}\ : While f can considered as a field, it can also represent a method that returns a val.
-Similarly \loconj{f} can be considered a series of fields, or methods that return a val, or a combination of both.\\
+where, we say MayAccess$^{Dir,Now}$(x,e) holds \lotiff\ in some current state \losigma, we have a field in x that points to e, or that we have x as a receiver in \losigma\ and there exists some y in the same \losigma\ that points to e.
+We also say MayAccess$^{Ind,Now}$(x,e) holds \lotiff\ in some current state \losigma, we have a \textit{series of fields} in x that leads to e, or that we have x as a receiver in \losigma\ and there exists some y with a series of fields in the same \losigma\ that leads to e.
+Lastly, the 'eventual' definitions of the two predicates are then defined as there existing an arising configuration \losigma' from \losigma, where the 'now' definitions hold at \losigma'.\\
 
-We surmarise the relationships between the four flavours of MayAccess in \ref{MayAccessTable}:
-\begin{table}[hbtp]
+Note that MayAccess$^{*,Now}$ holds imply MayAccess$^{*,Eve}$ holds, since an eventual state \losigma' that arises from \losigma\, can refer to \losigma, and therefore, the 'now' definitions implies and are stronger than the 'eventual' definitions.
+Also since a series of fields \loconj{f}\ can mean a singular field f, MayAccess$^{Dir,*}$ holds imply MayAccess$^{Ind,*}$ holds, and therefore the 'direct' definitions are stronger than the 'indirect' definitions.
+We surmarise the relationships between the four flavours of MayAccess in \ref{MayAccessRelations}.\\
+\begin{table}[htb]
 \small
-\caption{MayAccess$^{*,*}$(x,e) Relations}
+\caption{MayAccess Relations}
 \centering
@@
 & & \lonimpliedby &\\
-\label{MayAccessTable}
+\label{MayAccessRelations}
 \end{tabular*}
@@
 
-Let us assume that both x and e are well-defined.
-Note that without imposing any further assumptions (such as those from the OCap model), we have defined \textit{both} MayAccess$^{Dir,Now}$(x,e) and MayAccess$^{Ind,Now}$(x,e) to mean forms of very weak access---that a directed path exists from x to e, but we do \textit{not} imply that such a path is traverseable by x (it might or might not be traverseable).
-Indeed, these definitions by themselves represent mere \textit{possibilities} of interaction (or possible authorities), they do \textit{not} represent that interaction (or authority) would succeed.
-The difference between MayAccess$^{Dir,Now}$(x,e) and MayAccess$^{Ind,Now}$(x,e) is only the computational distance between x and e on the reference graph, where the latter involves possibly intermediate entities (or objects in an object-oriented model).\\
+Note that without imposing any further assumptions (such as those from the OCap model), we have defined \textit{both} MayAccess$^{Dir,Now}$(x,e) and MayAccess$^{Ind,Now}$(x,e) to mean forms of very weak access---that a directed path exists from x to e, and that we do \textit{not} imply that such a path is \textit{traverseable} by x.
+By a path being traverseable from x to e, we mean that x has a method that upon invocation, can lead to a method invocaton in e, and therefore x can interact with e, and possibly, modify e.
+These MayAccess definitions by themselves represent mere paths, or mere \textit{possibilities} of object interaction---they do \textit{not} imply that interaction would succeed.
+We therefore need another predicate describes method invocation between objects.
 
-What do the definitions mean then for non-OCap and OCap models in an object-oriented world?
-Again, let us assume both o and o' are well-defined, valid object references.
-In non-OCap models, the possible presence of a global ambient authority means the predicates above say \textit{nothing} about whether any interaction between an object o and o' would succeed.
-This is true even if o possesses directly the reference of o', where MayAccess$^{Dir,Now}$(o,o') holds.
-For all we know, we could easily have in a non-OCap language a feature to completely restrict access to o' using a global ambient authority, such that any object in the programming world which possesses references of o' cannot use them, and all paths leading to o' represent merely possibilities of interaction, but from which no authority can possibly arise.\\
-
-Could we say more about OCap systems?
-In OCap systems, there can be no global ambient authority so that an object reference by itself represents both the designation and the authority to use the object.
-This therefore leads us to be able to make a crucial distinction between MayAccess$^{Ind,Now}$(o,o') and MayAccess$^{Dir,Now}$(o,o') in the OCap model:
-\begin{itemize}
-\item MayAccess$^{Ind,Now}$(o,o') means---\textit{only}\\
-\ablock 1. a directed path from o to o' exists\\
-\ablock\ablock (possible authority)
-\item MayAccess$^{Dir,Now}$(o,o') means---\textit{both}\\
-\ablock 1. a directed path from o to o' exists\\
-\ablock\ablock (possible authority) \textit{and}\\
-\ablock 2. o's authority to use o' \textit{will} succeed
-\end{itemize}
-We elaborate the distinction with the following OCap example where there is a particular state \losigma\ of the system where, o$_1$.next points to o$_2$, and o$_2$.next points to o$_3$.
-o$_2$.next is a private method that can only be called internally by o$_2$.
-In this example, M,\losigma ~\loturns ~MayAccess$^{Ind,Now}$(o$_1$,o$_3$) holds, regardless of whether o$_2$.next is traverseable by o$_1$.
-We say here that a path from o$_1$ to o$_3$ exists, but is not traverseable by o$_1$.
-On the other hand, MayAccess$^{Dir,Now}$(o$_1$,o$_3$) does not hold true, because o$_3$ is not reachable from $o_1$ in a single step---o$_2$ sits between them on the reference graph as an intermediate object, and can possibly prevent or allow traversal from o$_1$ to o$_3$ (in this example, o$_2$ prevents such a traversal).\\
-
-What does it mean then for MayAccess$^{Dir,Now}$(*,*) to hold in an OCap model?
-From the same example, MayAccess$^{Dir,Now}$(o$_1$,o$_2$) holds and implies a stronger form of access---it means that a path exists from o$_1$ to o$_2$, and that such a path \textit{is} traverseable.
-This is because, by the definition of the predicate and configuration of the example, the object reference of o$_2$ exists within o$_1$'s state.
-Therefore, o$_2$ is guaranteed therefore to be accessible, and its authority exercisable, by o$_1$, without the interference of any ambient authority.
-Notice how this o$_1$-o$_2$ relationship differs from the o$_1$-o$_3$ relationship, where o$_1$ cannot guarantee that it can exercise the authority of o$_2$.next which points to o$_3$, since o$_2$.next does not exist within o$_1$'s state---it is possible that o$_2$.next is protected by x$_2$ through encapsulation and data-hiding.\\
-
-Within an OCap model, we can now be convinced that the stronger
+\subsubsection{MayAffect Definitions}
+With our access predicates in place, we define a \textit{MayAffect} predicate that describe what it means for x to be able to affect e:\\
 \begin{logic}
-MayAccess$^{Dir,Now}$(o$_1$,o$_2$) \losimeq \linebreak
-\ablock o$_1$ has the capability of o$_2$\end{logic}
-while the weaker MayAccess$^{Ind,Now}$(o$_1$,o$_3$) does not say anything about whether o$_1$ has the capability of o$_3$, but that only a directed path exists.
-It does however, represent a necessary condition for capability.
-\begin{logic}
-MayAccess$^{Ind,Now}$(o$_1$,o$_3$) \losimeq\ \linebreak
-\ablock there is a directed path from o$_1$ to o$_2$\linebreak
-\\
-o$_1$ has the capability of o$_3$ \loimplies \linebreak
-\ablock MayAccess$^{Ind,Now}$(o$_1$,o$_3$)\end{logic}
-\subsubsection{MayAffect Definitions}
-With our access predicates in place, we introduce a set of predicates that describe changes to the state of a system.
-Again, we highlight that these predicates are broad enough to describe both non-OCap and OCap models.
-\begin{logic}
-M,\losigma\ \loturns\ MayAffect$^{Now}$(x,e) \loiff \linebreak
-\ablock \loexists \loconj{m},\loconj{a},\losigma'. x.\loconj{m}(\loconj{a}) \loleadsto\ \losigma' \loand \loexec{e}{\losigma} \loneq \loexec{e}{\losigma'}
-\linebreak \\
-M,\losigma\ \loturns\ MayAffect$^{Eve}$(x,e) \loiff \linebreak
-\ablock \loexists \losigma' \loin Arising(M,\losigma).\linebreak
-\ablock \loexists \loconj{m},\loconj{a},\losigma'. x.\loconj{m}(\loconj{a}) \loleadsto\ \losigma' \loand \loexec{e}{\losigma} \loneq \loexec{e}{\losigma'}
-\linebreak \\
-\textit{If e is an object:} \\
-\loforall e\loin Object. \loexec{e}{\losigma} \loneq \loexec{e}{\losigma'} \loiff \linebreak
-\ablock \loexists f. \loexec{e.f}{\losigma} \loneq \loexec{e.f}{\losigma'}
+\hr\\
+\textsc{\normalsize *MayAffect Definition}\\\vspace{0.5em}
+M,\losigma\ \loturns\ \=MayAffect(x,e) \loiff \\
+\> \losigma(this) = \loexec{x}{\losigma} \loand\ \loexists \losigma'\loin Arising(M,\losigma)\\
+\> \loand\ \losigma'(this) = \loexec{e}{\losigma'}\\
+\hr
 \end{logic}
 
-\begin{table}[hbtp]
-\caption{MayAffect(o,o') Relations in OCap}
-\centering
-\begin{tabular*}{\linewidth}{ccc}\toprule
-\multirow{2}{*}{MayAffect$^{Now}$} & \loimplies & \multirow{2}{*}{MayAffect$^{Eve}$} \\
-& \lonimpliedby &  \\
-\bottomrule
-\end{tabular*}
-\end{table}
+where we say MayAffect(x,e) holds \lotiff\ we have x as a receiver in some state \losigma, and in some arising state \losigma' from \losigma, we have e as a receiver in \losigma'.
+This means that there must exist some method in x that upon invocation, will result in some state \losigma' where e is the receiver in the state \losigma'.
+Other possible definitions for MayAffect and why we have settled for the definition above arguing about method invocation, can be found in the Appendix (\ref{appendix}).\\
 
-\subsection{OCap Security Implications}
+\subsection{Object-Oriented Paradigm}
+
+As our paper focuses on security between trusted and untrusted objects, we shift our use of these predicates to an object-oriented paradigm.
+So, what do our above definitions mean in an object-oriented world?
+
+\begin{logic}
+\hr\\
+\textsc{\normalsize *Object as Receiver Nec. Cond.}\\\vspace{0.5em}
+M,\losigma\ \loturns\ \=\losigma(this) = \loexec{o$_x$}{\losigma}\loimplies \\
+\> \loexists \losigma*,m,\loconj{a}. M,\losigma*, o$_x$.m(\loconj{a}) \loleadsto\ \losigma\\
+\hr
+\end{logic}
+
+We first make clear that for an object to be a receiver in a particular state \losigma, we require a method to exist in the object in a prior state \losigma* that upon invocation can lead to state \losigma.
+This follows from our definition of execution.
+Let us illustrate this with an example where there is a particular state \losigma\ of the system with a configuration of 3 objects o$_1$, o$_2$, and o$_3$.
+In this example, to illustrate object encapsulation, we assume all fields in objects are private.
+o$_1$.next is a private field that holds the object reference of o$_2$, and o$_2$.next is a private field that holds the object reference of o$_3$.
+In this example:
+\begin{itemize}
+\item M,\losigma ~\loturns MayAccess$^{Ind,Now}$(o$_1$,o$_3$) holds true, regardless of whether we know o$_1$ can invoke any methods on o$_3$.
+
+\item If o$_1$ contains no method that can invoke o$_2$, or o$_2$ has no methods that can be invoked externally, then according to our definitions, we say M,\losigma ~\loturns \loneg MayAffect(o$_1$,o$_2$), meaning that o$_1$ cannot affect o$_2$.
+
+\item If o$_1$ contains a method that can invoke o$_2$, and o$_2$ contains some public method that can be invoked externally which is empty and does nothing, but o$_2$ contains no method that can invoke o$_3$, then according to our definitions, we mean M,\losigma ~\loturns\ \loneg MayAffect(o$_2$,o$_3$), since we know o$_2$ does not have any way of invoking o$_3$.
+We also mean M,\losigma ~\loturns\ MayAffect(o$_1$,o$_2$) to hold, because given this configuration, we know o$_1$ has a method that can invoke o$_2$ and, that o$_2$ has a public method that can be invoked, even if it does nothing.
+\end{itemize}
+Note that if we assume all fields are private, for an object o to \textit{eventually modify} an object o', it requires:
+\begin{itemize}
+\item there exist a method in o that can eventually invoke a method in o'
+\item there eventually exist a method in o' that can be invoked externally to modify its own internal state
+\end{itemize}
+Our MayAffect(o,o') predicate however, is not concerned whether or not the state o' is modified ---the fact that a method in o' can be eventually invoked by o is good enough grounds for us to mean that o can affect o'.
+Our justification is that in the object-oriented paradigm, programmers almost always enforce security of a sensitive field in a protected object to be declared private, and define some public method in the object that can modify the concerned private field based on some condition:
+\begin{logic}
+\hr\\
+\textsc{\normalsize *Private Field Definition}\\\vspace{0.5em}
+M,\losigma\ \loturns\ \=PrivField(o,f) \loiff\\
+\>$[$\loexists \losigma'\loin Arising(M,\losigma) \loand\ \loexec{f}{\losigma'}\loneq\loexec{f}{\losigma} \loimplies \\
+\> \loexists m,\loconj{a}. M,\losigma, o.m(\loconj{a}) \loleadsto\ \losigma'$]$\\
+\hr
+\end{logic}
+Hence, the question of whether an untrusted object can modify a private field in a protected object can often be expressed as a question of whether the untrusted object can invoke a method in the protected object.
+Furthermore, given our definitions of PrivField and MayAffect, we can derive a lemma that states that a necessary condition for modification of a private field in o$_x$ is the existence of some arbitrary object o* that can affect o$_x$, where it is also possible that o* can refer to o$_x$.
+Proving the falsity of our MayAffect predicate is therefore sufficient to prove that all private fields in o$_x$ cannot be modified:
+\begin{logic}
+\hr\\
+\textsc{\normalsize *Private Field Modification Lemma}\\\vspace{0.5em}
+M,\losigma\ \loturns\ \loforall \losigma',f. \=$[$\losigma'\loin Arising(M,\losigma) \loand\ \loexec{f}{\losigma'}\loneq\loexec{f}{\losigma}\\
+\>\loand\ PrivField(o,f) \loimplies \\
+\>\loexists o*. M,\losigma\ \loturns\ MayAffect(o*,o)\\
+\hr
+\end{logic}
+
+Therefore, our MayAffect(o,o') requires:
+\begin{itemize}
+\item there exist a method in o$_x$ (\losigma(this)=\loexec{o}{\losigma}) that can eventually invoke a method in o' (\loexists \losigma'\loin Arising(M,\losigma) \loand\ \losigma'(this) = \loexec{o'}{\losigma'})
+\item which by collorary means there eventually exist a method in o' that can be invoked externally \losigma'(this) = \loexec{o'}{\losigma'}
+\end{itemize}
+
+Note, that in some languages like JavaScript, methods can be added to objects dynamically at run-time, such that a method of an object that does not exist in some current state can potentially exist in some eventual later state.
+However, if we are operating in a language that does not allow dynamically adding of methods to objects, or that the protected object o$_y$ is defined such that methods cannot be added dynamically to it, proving there does not exist a public method now in o$_y$ is sufficient to prove that there will not eventually exist a public method in o$_y$:
+
+\begin{logic}
+\hr\\
+\textsc{\normalsize *No dynamic adding of methods to objects.}\\\vspace{0.5em}
+M,\losigma\ \loturns\
+\= \loforall o,\losigma',m,\loconj{a$_\losigma$},\loconj{a$_{\losigma'}$}.\\
+\>[ \losigma'\loin Arising(M,\losigma) \loand\ \losigma'(o.m(\loconj{a$_{\losigma'}$}))\\
+\>\loimplies \losigma(o.m(\loconj{a$_\losigma$})) ]\\
+\hr 
+\end{logic}
+\subsubsection{Domination}
+Here, we introduce the concept of domination that is inspired by the work on ownership types by Clarke et al.\cite{clarke1998}, where they define that the owner o of an object o' dominates o' in the object graph such that for any path p from some arbitrary object o'' to o', either o'' belongs to the set of owners or that p passes through o.
+Similarly, we say that a set S dominates x \lotiff\ for every object X* in the system that has a path to x, X* must either have a path to some arbitrary object y that is a member of the set S, or X* belongs to the set S.
+\begin{logic}
+\hr\\
+\textsc{\normalsize *Domination Definition (DOM)}\\\vspace{0.5em}
+M,\losigma\ \loturns\ \=Dom(S,x) \loiff \\
+\> \loforall X*. $[$\=MayAccess$^{Ind,Now}$(X*,x) \loimplies\\
+\> \>\loexists y\loin S. \=MayAccess$^{Ind,Now}$(X*,y) \loor\ X*\loin S$]$\\
+\hr
+\end{logic}
+
+\subsubsection{Corollaries}
+
+\begin{logic}
+\hr\\
+\textsc{\normalsize *Path-Execution Condition (PEC)}\\\vspace{0.5em}
+M,\losigma\ \loturns\
+\= MayAffect(o,o') \loimplies \\
+\> MayAccess$^{Ind,Eve}$(o,o')\\
+\hr
+\end{logic}
+
+\begin{logic}
+\hr\\
+\textsc{\normalsize *Global Path Chain (GPC)}\\
+M,\losigma\ \loturns\ \=MayAccess$^{Ind,Now}$(o,o')\\
+\>\loiff\\
+\>\loexists X*. $[$\=MayAccess$^{Ind,Now}$(o,X*) \loand\\
+\>\> MayAccess$^{Dir,Now}$(X*,o')$]$\\
+\hr
+\end{logic}
+
+\begin{logic}
+\hr\\
+\textsc{\normalsize *Global Path Chain II (GPC II)}\\
+M,\losigma\ \loturns\ \=MayAccess$^{Ind,Now}$(o,o') \loand\ Dom(S,o')\\
+\>\loimplies\\
+\>\loexists X*\loin{S}. $[$\=MayAccess$^{Ind,Now}$(o,X*) \loand\\
+\>\> MayAccess$^{Dir,Now}$(X*,o')$]$\\
+\hr
+\end{logic}
+
+\begin{logic}
+\hr\\
+\textsc{\normalsize *Global Execution Chain (GEC)}\\
+M,\losigma\ \loturns\ \=MayAffect(o,o')\\
+\> \loiff\\
+\> \loexists X*. $[$\=MayAffect(o,X*) \loand\\
+\>\>MayAffect(X*,o')$]$\\
+\hr
+\end{logic}
+
+\begin{logic}
+\hr\\
+\textsc{\normalsize *Global Execution Chain II (GEC II)}\\
+M,\losigma\ \loturns\ \=MayAffect(o,o') \loand\ Dom(S,o')\\
+\> \loimplies\\
+\> MayAccess$^{Ind,Eve}$(o,o') \loand\\
+\> \loexists x\loin{S}. $[$\=MayAffect(o,x) \loand\\
+\> \>MayAffect(x,o')$]$\\
+\hr
+\end{logic}
+\subsubsection{Lemmas}
+\begin{logic}
+\hr\\
+\textsc{\normalsize *Object Existence Lemma (OEL)}\\
+\textit{if z exists in \losigma,} \\
+\loturns\ M,\losigma\ \loturns\ \=MayAccess$^{Ind,Now}$(z,z) is true \loand\\
+\loturns\ M,\losigma\ \loturns\ \=MayAccess$^{Dir,Now}$(z,z) is true \loand\\
+\loturns\ M,\losigma\ \loturns\ \=MayAffect(z,z) is true\\
+\hr
+\end{logic}
+
+\subsection{OCap Specifications}
 
 \subsubsection{What is protection?}
-In an object-oriented world\footnote{
-To simplify our discussion, we work with the variables \{o,o'\}\loin Object for this entire section.}
-, security concerns between objects are often a question of whether what one object can do to another object in \textit{any} eventual state of a system.
-Because an object encapsulates both internal state and behaviour, strictly speaking, security of an object should govern over both the integrity of the object's fields (internal state) and whether the objects' methods can be called (behaviour).
-Our predicates are broad enough to enable a discussion of both protection of state (MayAffect$^{Eve}$) and behaviour (MayAccess$^{Dir,Eve}$)\footnote{ Our MayAccess$^{Dir,Eve}$ is not weak enough to reason specifically which behaviours can be called.
-This is however not a big issue in \textit{pure} OCap systems, where often giving away the capability of an object typically means allowing \textit{all} public behaviours of the object to be called without distinction.}.
-We emphasize however that protecting either state or behaviour of an object, does \textit{not} necessarily imply the other.
-In fact, a common feature in OCap patterns is being able to protect a sensitive object's behaviours (they cannot be called directly by untrusted objects), but at the some allowing the same untrusted objects to modify the object's state in some controlled way.\\
+In an object-oriented world, security concerns between objects are often a question of whether what one object can do to another object in \textit{any} eventual state of a system.
+Because an object encapsulates both state and behaviour, strictly speaking, security of an object should govern over both the integrity of the object's fields (object state) and whether the objects' methods can be called (object behaviour).
+Our \textit{MayAffect} predicate enables a discussion of protection of behaviour\footnote{ Our MayAffect is too strong to reason specifically which behaviours can be called.
+This is however not a big issue in \textit{pure} OCap systems, where often giving away the capability of an object typically means allowing \textit{all} public behaviours of the object to be called without distinction.}
+and because we assume that protected objects should have fields that are privately encapsulated, we only need to focus our reasoning on object to object method invokations as a security concern.
+With these simplifications, protection for us then becomes solely a matter of whether we can allow or deny an object to invoke a method of another object.
+To help us reason about protection, we first formalize our assumptions and the rules of OCap using our predicates in \ref{ocaprules} which will help guide us in constructing our necessary conditions for MayAffect later in \ref{protection}.
 
-There are however, some flexibility in working with objects, that allows us to simplify our discussion and work with only a broad definition of state protection in terms of our MayAffect$^{Eve}$ predicate, \textit{without} thinking too much about specific fields of the object we want to protect or the protection of behaviour.
-Moving away from our broad definition of whether an object may be affected (we defined it as being able to change at least one field of the object), to more precise spcifications of which particular field(s) of the object may be affected, should be trivial.
-We can in theory also stay and reason within our framework by separating the particular concerned field(s) of the object into separately encapsulated objects.
-We merely have to be careful as to \textit{which} object's state we want to protect.
-Furthermore, in theory one can easily introduce a field within an object that behaves like a 'signal' which would be modified whenever a specified behaviour is called.
-Preventing a particular behaviour to be called by an untrusted object then becomes equivalent to denying the untrusted object the ability to modify the particular state of the signal field of the object.\\
+\subsection{Formal specifications in the OCap Model}\label{ocaprules}
 
-With these simplifications, protection for us then becomes solely a matter of whether we can allow or deny an object to modify the state of another.
-To help us reason about protection, we first formalize our assumptions and the rules of OCap, using our predicates, to help guide us in constructing our necessary conditions in \ref{ocaprules}.
-In the following \ref{protection} below we build the neccessary conditions for MayAffect$^{Eve}$, where MayAffect$^{Eve}$ is placed in the antecedent, and we examine which of the family of MayAccess predicates is placed in the consequent.
+\subsubsection{Passing object references}
+Objects cannot forge capabilities, and only connectivity begets connectivity.\\
+\begin{logic}
+\hr\\
+\textsc{\normalsize *OCap Introduction Lemma (OCI)}\\\vspace{0.5em}
+M,\losigma\ \loturns\
+\= \loneg MayAccess$^{Dir,Now}$(o,o')\\
+\> \loand\ MayAccess$^{Dir,Eve}$(o,o')\\
+\> \loimplies \\
+\> \loexists X*. \=MayAccess$^{Dir,Now}$(X*,o')\\
+\> \> \loand\  MayAffect(X*,o)\\
+\hr
+\end{logic}
+The OCap introduction lemma states that in given a state \losigma, if o does not have the capability o' but o may eventually obtain the capability of o', then there must exist an object X* that has the capability of o' at \losigma, and that X* can affect o, so that X* can introduce the capability of o' to o.\\
 
-\subsection{Formal specifications of OCap Rules}\label{ocaprules}
+Note that this lemma does not hold in non-OCap systems, if object capability or references can be forged.
+That is, even though object o might not have the capability of o' in \losigma, the object o may contain a capability forging method that can forge and return the capability of o', such that after execution of the method to some state \losigma', o would have the capability of o' in \losigma', without involving any other object.
 
-\textbf{Rule 1: Objects can only interact with each other through sending messages on capabilities.}
-We begin our reasoning of protection in our OCap model, by first re-asserting the necessary but not sufficient path condition of capability, and calling it \textsc{pec}.
-By definition of our predicates, an object having a capability of o' implies object o having a path to o', but not vice-versa:
 \begin{logic}
-\hrule\null
-\textsc{\normalsize *Path-Execution Connectivity (PEC)}\\
-M,\losigma\ \loturns\ MayAccess$^{Dir,Now}$(o,o') \loimplies \linebreak
-\ablock \ablock  MayAccess$^{Ind,Now}$(o,o')
-\linebreak \\
-\textit{By contraposition,}\linebreak
-M,\losigma\ \loturns\ \loneg MayAccess$^{Ind,Now}$(o,o')\loimplies \linebreak
-\ablock \ablock \loneg MayAccess$^{Dir,Now}$(o,o')
-\linebreak
-\hrule
+\hr\\
+\textsc{\normalsize *OCap Eventual Path Connect Lemma (EPC)}\\
+M,\losigma\ \loturns\ \=MayAccess$^{Ind,Eve}$(o,o')\\
+\> \loimplies\\
+\> \loexists X*. \=$[ ($\=MayAccess$^{Dir,Now}$(o,X*)\\
+\>\>\> \loor \\
+\>\>\> MayAccess$^{Dir,Now}$(X*,o)$)$\\
+\>\>\loand \\
+\>\>MayAccess$^{Ind,Eve}$(X*,o')$ ]$\\
+\hr
 \end{logic}
-Note also that \textsc{pec} is a direct intepretation from Rule 1, since rule 1 says owning a capability is the only way of sending messages to another object, and therefore implies a path between the objects.\\
 
-\textbf{Rule 2: Objects cannot forge capabilities, and only connectivity begets connectivity.}
-\begin{logic}
-\hrule\null
-\textsc{\normalsize *Global Path Existence (GPE)}\\
-M,\losigma\ \loturns\ $[$\loexists Y*\loin{Obj}. MayAccess$^{Ind,Now}$(Y*,o')$]$\linebreak
-\ablock \loiff\linebreak
-\ablock $[$\loexists X*\loin{Obj}. MayAccess$^{Dir,Now}$(X*,o')$]$
-\linebreak\\
-\textit{By contraposition,}\linebreak
-M,\losigma\ \loturns\ $[$\loforall X*\loin{Obj}. \loneg MayAccess$^{Dir,Now}$(X*,o')$]$\linebreak
-\ablock \loiff\linebreak
-\ablock $[$\loforall Y*\loin{Obj}. \loneg MayAccess$^{Ind,Now}$(Y*,o')$]$\linebreak
-\hrule
-\end{logic}
-In addition to path being a necessary condition for capability, \textsc{gpe} gives us an additional new relation between the two concepts over the entire system.
-\textsc{gpe} says that in a given system of objects, $iff$ there exists an object Y* which has a path to o', then there exists an object X* that has the capability of o'.
-Looking from the left to right direction, since the path from Y* to o' exists, then there must exist an object X* with a capability of o' so that the path from Y* to o' is well-established.
-Moving from the right to left direction, \textsc{gpe} says that if an object X* has a capability then there exist an object Y* with a path to o', and this can be proven from \textsc{pec} or by definition when (X*==Y*).
-The contraposition result of \textsc{gpe} says that we can prevent all paths to o' through one of these conditions through denying all direct paths leading into o'.\\
-
-However, \textsc{gpe} only postulates the existence of some object with some capability iff there exists some object with some path, and says nothing about how the objects are connected.
-From the connectivity begets connectivity rule, \textsc{pec}, and \textsc{gpe}, we construct global path-execution relationships over the entire system:
-\begin{logic}
-\hrule\null
-\textsc{\normalsize *Global Path-Execution Connect. (GPEC)}\\
-M,\losigma\ \loturns\ MayAccess$^{Ind,Now}$(o,o')\linebreak
-\ablock \loiff\linebreak
-\ablock \loexists X*\loin{Obj}. $[$MayAccess$^{Ind,Now}$(o, X*) \loand\linebreak
-\ablock \ablock \ablock \quad MayAccess$^{Dir,Now}$(X*,o')$]$
-\linebreak\\
-\textit{By contraposition,} \\
-M,\losigma\ \loforall X*\loin{Obj}. $[$\loneg MayAccess$^{Ind,Now}$(o, X*) \loor\linebreak
-\ablock \ablock \ablock \quad \loneg MayAccess$^{Dir,Now}$(X*,o')$]$\linebreak
-\ablock \loiff\linebreak
-\ablock \loneg MayAccess$^{Ind,Now}$(o,o')\linebreak
-\hrule
-\end{logic}
-\textsc{gpec} introduces an intermediate object X* between o and o' that makes explicit the missing connection in \textsc{gpe}.\\
-
-The contraposition result of \textsc{gpec} says that for there to be no paths from o to o' these configurations must hold:
-\begin{itemize}\setlength\itemsep{0.5em}
-\item if there is some object x that o has a path to in \losigma, then x cannot have a path to o' in \losigma.
-\item if there is some object x that has a path to o' in \losigma, then o cannot have a path to x in \losigma.
-\end{itemize}
-
-The power in \textsc{gpec} lies in the recursive predicate MayAccess$^{Ind,Now}$(o, X*) that can eventually be expanded into a set of MayAccess$^{Dir,Now}$ predicates that finishes with terminating predicate MayAccess$^{Ind,Now}$(o, o), where the terminating predicate will always be true, unless the object o does not exist in \losigma.
-We now look for a relation between a path configuration at a state \losigma\ ($Now$) and eventual path configurations in states \losigma' arising from \losigma\ ($Eve$).\vfill
-\begin{minipage}[b]{\linewidth}
-\begin{logic}
-\hrule\null
-\textsc{\normalsize *Eventual Path Connectivity 1 (EPC1)}\\
-M,\losigma\ \loturns\ MayAccess$^{Ind,Eve}$(o,o')\linebreak
-\ablock \loimplies\linebreak
-\ablock \loexists X*\loin Obj. $\{[$MayAccess$^{Ind,Now}$(o,X*)\linebreak
-\ablock \ablock \ablock \ablock\loor \linebreak
-\ablock \ablock \ablock \quad\: MayAccess$^{Dir,Now}$(X*,o)$]$\linebreak
-\ablock \ablock \ablock \ablock\loor \linebreak
-\ablock \ablock \ablock \quad\: MayCreate$^{Dir,Now}$(o,X*)$]$\linebreak
-\ablock \ablock \ablock \quad \loand \linebreak
-\ablock \ablock \ablock \quad MayAccess$^{Ind,Eve}$(X*,o')$\}$
-\linebreak \\
-\textit{By contraposition,}\linebreak
-M,\losigma\ \loturns\ \linebreak
-\ablock \loforall X*\loin Obj. $\{[$\loneg MayAccess$^{Ind,Now}$(o,X*)\linebreak
-\ablock \ablock \ablock \ablock \quad \loand \linebreak
-\ablock \ablock \ablock \quad\: \loneg MayAccess$^{Dir,Now}$(X*,o)$]$\linebreak
-\ablock \ablock \ablock \quad \loor \linebreak
-\ablock \ablock \ablock \quad \loneg MayAccess$^{Ind,Eve}$(X*,o')$\}$\linebreak
-\ablock \loimplies\linebreak
-	\ablock\loneg MayAccess$^{Ind,Eve}(o,o')$\linebreak
-\hrule
-\end{logic}
-\end{minipage}
-\textsc{epc1} has the meaning that if an object o has an eventual path to o' in some arising state \losigma', then there must exist an object X* that has an eventual path to o', and there must exist a way for o to have a path to X* in \losigma'.
-o either must already have a path to X* in \losigma, or that it will eventually receive the capability of X* in \losigma'.
+\textsc{epc} has the meaning that if an object o has an eventual path to o' in some arising state \losigma', then there must exist an object X* that has an eventual path to o', and there must exist a way for o to have a path to X* in \losigma'.
+Object o either must already have a path to X* in \losigma, or that it must be able to receive the capability of X* through introduction by X* in \losigma.
 \textsc{epc1} is actually a formal representation of connectivity begets connectivity across time:
@@
 \begin{itemize}
-\item \textbf{Initial Conditions or Endowment:} o has an existing path to o' in \losigma, therefore X* refers to o'
-\item \textbf{Parenthood:} o will create o' in some arising \losigma', therefore X* refers to o
-\item \textbf{Introduction:} o will only obtain the path to o' in some arising \losigma', therefore X* refers to an object that is not the same object as o (X*\loneq o). 
+\item \textbf{Initial Conditions or Endowment:} o has the capability of o' in \losigma, therefore X* refers to o'
+\item \textbf{Parenthood:} if o can create o' in some arising \losigma', then o can also create o' in \losigma, therefore X* refers to o'
+\item \textbf{Introduction:} o will only obtain the path to o' in some arising \losigma' through another object introducing o a path to o', therefore X* refers to an object that is not the same object as o (X*\loneq o).
 \end{itemize}
 
-Note that for o to have an eventual path to o', we require only o to have an eventual path to X* since we have stated that X* will have an eventual path to o'.
-If o already has a path to X* in \losigma\ then we know o can reach X* in \losigma'.
+Note that for o to have an eventual path to o', we only require o to have a a direct capability to some arbitrary object X* or that X* has a direct capability of o so that X* can introduce itself to o.
+This is because we have stated that X* will have an eventual path to o'.
+If o already has the capability of X* in \losigma\ then we know o can reach X* in \losigma' by definition.
 If not, the capability of X* must be introduced to o.
-For X* to introduce itself, X* must have the capability of o in \losigma', and the only way we can guarantee this is to have X* possess the capability of o in \losigma.
+For X* to introduce itself, X* must have the capability of o in \losigma.\\
+
 Well, what if the capability of X* is introduced by some \textit{other} object \~X*?
-Note that in such a case, \~X* will have the capability of X*, and will therefore also be able to eventually have a path to o'.
+Note that in such a case, \~X* must have the capability of object X* (for \~X* to even introduce X* to o in the first place), and therefore \~X* will also be able to eventually have a path to o'.
 Also \~X* must also be able to introduce itself to o.
@@
 
-The contraposition result of \textsc{epc1} says that for there to be no eventual paths from o to o' these configurations must hold:
-\begin{itemize}\setlength\itemsep{0.5em}
-\item if there is some object x that o has a path to in \losigma, then x cannot have an eventual path to o'
-\item if there is some object x that possesses the capability of o in \losigma, then x cannot have an eventual path to o'
-\end{itemize}
-Note that object x can refer to the same object as o.
-There is one final critical result from \textsc{epc1}.
+There is one final critical result from \textsc{epc}.
 Notice how, there is a 'recursive' MayAccess$^{Ind,Eve}$(X*,o') in our condition for MayAccess$^{Ind,Eve}$(o,o').
 This allows us to recursively expand the condition to incorporate \textit{all} X* intermediate objects in the path leading to o'.
-This recursive expansion gives us both MayAccess$^{Dir,Now}$ and MayAccess$^{Ind,Now}$ predicates, where the terminating MayAccess$^{Ind,Eve}$(o',o') always returns true if o' exists.
-This result allows us to define MayAccess$^{Ind,Eve}$(X*,o') in a configuration of paths completely based in \losigma.\\
+Repeated recursive expansions will eventually give us conditions that are only defined in MayAccess$^{Dir,Now}$ predicates, where the terminating MayAccess$^{Ind,Eve}$(o',o') can be determined to be true or false based on whether o' exists, according to \textsc{oel}.
+This result allows us to define an eventual path between two objects based purely on a present configuration of objects on the reference graph in \losigma.\\
 
-What about capabilities?
-Using \textsc{gpec}, we can always expand MayAccess$^{Ind, Now}$(o,X*) into a set of MayAccess$^{Dir, Now}$ predicates that terminate with MayAccess$^{Ind, Now}$(o,o) which is always true if o exists.
-The final result we will get is a configuration based purely on MayAccess$^{Dir, Now}$ predicates.
-We expand \textsc{epc1} into \textsc{epc2} to illustrate:
+In the example, we make concrete \textsc{epc} using only 3 objects o, x, o'.
+Let us assume o, x, and o' always exists, such that the terminating recursive predicates would return true.
+This example then illustrates an eventual path from o to o' can only exist if one of these present 6 configurations in \losigma\ holds:
+\begin{itemize}\setlength\itemsep{0.5em}
+\item If X*=x, then o has the capability of x, and x has an eventual path to o'---\subitem by x having the direct capability of o'
+\item If X*=x, then o has the capability of x, and x has an eventual path to o'---\subitem by o' having the direct capability of x
+\item If X*=x, then x has the capability of o, and x has an eventual path to o'---\subitem by x having the direct capability of o'
+\item If X*=x, then x has the capability of o, and x has an eventual path to o'---\subitem by o' having the direct capability of x
+\item If X*=o', then o has the capability of o'
+\item If X*=o', then o' has the capability of o
+\end{itemize}
+Note how, these 6 configurations are direct capability configurations at \losigma, but allows us to reason about whether a potential path from o to o' can exist in some arising state \losigma' from \losigma.\\
+
+Lastly, from the OCI lemma, we know that for an object o to introduce a capability to another object o', we require that o can affect o'. Therefore, \textsc{epc} can be modified:
 
 \begin{logic}
-\hrule\null
-\textsc{\normalsize *Eventual Path Connectivity 2 (EPC2)}\\
-M,\losigma\ \loturns\ MayAccess$^{Ind,Eve}$(o,o')\linebreak
-\ablock \loimplies\linebreak
-\ablock \loexists X*\loin Obj.$\{$\linebreak
-\ablock \ablock $[$\null
-\textcolor{red}{\loexists X$^1$\loin{Obj}. MayAccess$^{Ind,Now}$(o, X$^1$) \loand\linebreak
-\ablock \ablock \ablock \ablock \quad MayAccess$^{Dir,Now}$(X$^1$,X*)\linebreak
-}
-\ablock \ablock \quad \loor \linebreak
-\ablock \ablock MayAccess$^{Dir,Now}$(X*,o)$ ]$\linebreak
-\ablock \ablock \loand \linebreak
-\textcolor{blue}{
-\ablock \ablock \loexists Y*\loin Obj.$\{[$MayAccess$^{Ind,Now}$(X*,Y*)\linebreak
-\ablock \ablock \ablock \ablock \quad \quad \loor \linebreak
-\ablock \ablock \ablock \ablock \quad MayAccess$^{Dir,Now}$(Y*,X*)$]$\linebreak
-\ablock \ablock \ablock \ablock \quad \loand \linebreak
-\ablock \ablock \ablock \ablock \quad MayAccess$^{Ind,Eve}$(Y*,o')$\}$
-}\linebreak
-\ablock \ablock $\}$
-\linebreak \\
-\textit{By contraposition,}\linebreak
-M,\losigma\ \loturns\ \linebreak
-\ablock \loforall X*\loin Obj.$\{$\linebreak
-\ablock \quad $[$\null
-\textcolor{red}{\loforall X$^1$\loin{Obj}. \loneg MayAccess$^{Ind,Now}$(o,X$^1$) \loor\linebreak
-\ablock \ablock \ablock \ablock \loneg MayAccess$^{Dir,Now}$(X$^1$,X*)\linebreak
-}
-\ablock \ablock \quad \loand \linebreak
-\ablock \quad \loneg MayAccess$^{Dir,Now}$(X*,o)$ ]$\linebreak
-\ablock \ablock \loor \linebreak
-\textcolor{blue}{
-\ablock \loforall Y*\loin Obj. $\{[$\loneg MayAccess$^{Ind,Now}$(X*,Y*)\linebreak
-\ablock \ablock \ablock \ablock \quad \loand \linebreak
-\ablock \ablock \ablock \quad\: \loneg MayAccess$^{Dir,Now}$(Y*,X*)$]$\linebreak
-\ablock \ablock \ablock \quad \loor \linebreak
-\ablock \ablock \ablock \quad \loneg MayAccess$^{Ind,Eve}$(Y*,o')$\}$
-}\linebreak
-\ablock \loimplies\linebreak
-	\ablock\loneg MayAccess$^{Ind,Eve}(o,o')$\linebreak
-\hrule
+\hr\\
+\textsc{\normalsize *OCap EPC Lemma II (EPC II)}\\
+M,\losigma\ \loturns\ \=MayAccess$^{Ind,Eve}$(o,o')\\
+\> \loimplies\\
+\> \loexists X*. \=$[ ($\=MayAccess$^{Dir,Now}$(o,X*)\\
+\>\>\> \loor \\
+\>\>\> MayAffect(X*,o)$)$\\
+\>\>\loand \\
+\>\>MayAccess$^{Ind,Eve}$(X*,o')$ ]$\\
+\hr
 \end{logic}
 
-To give a concrete example, let us assume a simple system with only three objects, o, x, o'.
-\begin{logic}
-\hrule\null
-\textsc{\normalsize *Eventual Path Example}\\
-M,\losigma\ \loturns\ MayAccess$^{Ind,Eve}$(o,o')\linebreak
-\ablock \loimplies\linebreak
-\ablock \ablock $[$\null
-\textcolor{red}{MayAccess$^{Ind,Now}$(o,o) \loand\linebreak
-\ablock \ablock \quad MayAccess$^{Dir,Now}$(o,x)\linebreak
-}
-\ablock \ablock \quad \loor \linebreak
-\ablock \ablock MayAccess$^{Dir,Now}$(x,o)$ ]$\linebreak
-\ablock \ablock \loand \linebreak
-\textcolor{blue}{
-\ablock \ablock $\{[$
-\textcolor{red}{MayAccess$^{Ind,Now}$(x,x) \loand\linebreak
-\ablock \ablock \quad MayAccess$^{Dir,Now}$(x,o')\linebreak}
-\ablock \ablock \quad \quad \loor \linebreak
-\ablock \ablock \quad MayAccess$^{Dir,Now}$(o',x)$]$\linebreak
-\ablock \ablock \quad \loand \linebreak
-\ablock \ablock \quad MayAccess$^{Ind,Eve}$(o',o')$\}$
-}\linebreak
-\ablock \ablock $\}$
-\linebreak \\
-\hrule
-\end{logic}
-In the example, we make concrete \textsc{epc2} using X* = x and Y* = o', where the expansion terminates at the three predicates, MayAccess$^{Ind,Now}$(o,o), MayAccess$^{Ind,Now}$(x,x) and MayAccess$^{Ind,Eve}$(o',o').
-Let us assume o, x, and o' always exists, such that these predicates would return true.
-This example then illustrates an eventual path from o to o' can only exist if one of these four conditions in \losigma\ hold:
-\begin{itemize}\setlength\itemsep{0.5em}
-\item o has the capability of x, and x has the capability of o'
-\item o has the capability of x, and o' has the capability of x
-\item x has the capability of o, and x has the capability of o'
-\item o has the capability of x, and x has the capability of o'
-\end{itemize}
-Note how, these four configurations are present state configurations at \losigma, but allows us to reason about whether a potential path from o to o' can exist in all arising states \losigma' from \losigma.\\
+\subsection{Protection and Cooperation}
 
-So far we have developed results for paths from o to o', but what about a direct path from o to o'?
-Luckily, we only need construct \textsc{eec} from \textsc{epc1}, the only difference being we now require the path from o to X* in \textsc{epc1} to be traversable.
+It is good to know we can enforce protection through some present path configurations that can guarantee eventual path-isolation, but in practice this is still not entirely useful.
+The quintessential question in OCap systems is whether we can enforce protection such as controlling whether o may affect o', on \textit{some condition}.
+In our framework, this just means that cooperation between objects first require paths to exist, but \textit{how} objects cooperate are capture in the object method definitions and availability.\\
+
+So, let us now assume we have code knowledge of some objects in our system so we know how they behave. If we have a simple object o' that has \textit{only} private methods, then we know:
 
 \begin{logic}
-\hrule\null
-\textsc{\normalsize *Eventual Execution Connectivity (EEC)}\\
-M,\losigma\ \loturns\ MayAccess$^{Dir,Eve}$(o,o')\linebreak
-\ablock \loimplies\linebreak
-\ablock \loexists X*\loin Obj.$[($MayAccess$^{Dir,Now}$(o,X*) \loor \linebreak
-\ablock \ablock \ablock \ablock MayAccess$^{Dir,Now}$(X*,o)$)$\linebreak
-\ablock \ablock \ablock \ablock \loand \linebreak
-\ablock \ablock \ablock \ablock MayAccess$^{Dir,Eve}$(X*,o')$]$
-\linebreak \\
-\textit{By contraposition,}\linebreak
-M,\losigma\ \loturns\ 
-\loforall X*\loin Obj. $\{[$\loneg MayAccess$^{Dir,Now}$(o,X*) \loand \linebreak
-\ablock \ablock \ablock \ablock \quad \loneg MayAccess$^{Dir,Now}$(X*,o)$]$\linebreak
-\ablock \ablock \ablock \ablock \loor \linebreak
-\ablock \ablock \ablock \ablock \loneg $[$MayAccess$^{Dir,Eve}$(X*,o')$]\}$\linebreak
-\ablock \loimplies\linebreak
-	\ablock\loneg MayAccess$^{Dir,Eve}(o,o')$\linebreak
-\hrule
+\hr\\
+M,\losigma\ \loturns\ \loforall X*. \=MayAffect(X*,o') \loimplies \\
+\> \loexec{X*}{\losigma} = \loexec{o'}{\losigma}\\
+\hr
 \end{logic}
-The power of \textsc{eec} is similar to \textsc{epc}\ in that it allows an arbitrary recursive expansion of the MayAccess$^{Dir,Eve}$(X*,o') till the terminating MayAccess$^{Dir,Eve}$(o',o') that is always true.\\
 
-We can always use an automated theorem prover to prove whether a path can arise between two objects in a present configuration system.
-We now turn our attention formalizing Ocap state changes rules in the system.\\
-
-\subsection{Protection}\label{protection}
-With our formalizations of OCap rules in place, we can now reason about enforcing protection.
-In this subsection we focus on finding the necessary conditions for the weaker predicate MayAffect$^{Eve}$ rather than MayAffect$^{Now}$ in the antecedent.
-This is because negation on both sides of the implication, would yield a stronger \loneg MayAffect$^{Eve}$ in the consequent.
-In practical terms, if we are concerned with the protection of o' from o, it is also often not very useful to have a policy where \loneg MayAffect$^{Now}$(o,o') holds but \loneg MayAffect$^{Eve}$(o,o') does not.
-Furthermore, by ensuring \loneg MayAffect$^{Eve}$(o,o') holds, we can also ensure \loneg MayAffect$^{Now}$(o,o') holds since by contraposition:
+Even though in theory any object may possess the capability of o', as long as we know there is no way for anybody but o' to \textit{exercise} the capability of o', then we know that for all objects X* that are not o', X* has no way of affecting of o' Recall, in our definition of MayAffect, the requirement that for an object o to affect o', we require a method to exist in o that can invoke o', and a method in o' that can be invoked externally.
+Consequently, with the knowledge that o' has only private methods, we also know o' has no way of invoking any other objects but itself:
 
 \begin{logic}
-$[$MayAffect$^{Now}$(o,o') \loimplies MayAffect$^{Eve}$(o,o')$]$\linebreak
-\ablock \loimplies\linebreak
-$[$\loneg MayAffect$^{Eve}$(o,o') \loimplies \loneg MayAffect$^{Now}$(o,o')$]$\\
+\hr\\
+M,\losigma\ \loturns\ \loforall X*. \=MayAffect(o',X*) \loimplies \\
+\> \loexec{X*}{\losigma} = \loexec{o'}{\losigma}\\
+\hr
 \end{logic}
 
-Furthermore, in building the necessary conditions for the predicate MayAffect$^{Eve}$, we are also more concerned with finding some structure that contains the configuration of MayAccess$^{*,Now}$, rather than MayAccess$^{*,Eve}$.
-This is because it is much easier to prove a configuration of relations that holds in \textit{one} specific state than think about whether a configuration holds in \textit{all} possible eventual states, which makes the former much easier to implement.\\
+The two outcomes from knowing an object with only private behaviours are:
+\begin{itemize}
+\item \loturns\ $[$M,\losigma\ \loturns\ \loforall X*\loneq o'.\loneg MayAffect(X*,o')$]$
+\item \loturns\ $[$M,\losigma\ \loturns\ \loforall X*\loneq o'.\loneg MayAffect(o',X*)$]$
+\end{itemize}
+This means that methods in objects play a dual important role of what objects can do to others, and what others can do to the object.
 
-\textbf{Rule 3: Objects may have private encapsulation of state and behaviour}\\
-To begin, we make clean in our reasoning that all fields in our objects can only be declared private.
-Consequently, this implies the necessary condition that an object's state can only be modified if one of it's behaviour is called, either by itself or one other object in the system that holds its capability.
-This is also implied by the OCap rule that there is no ambient authority that can interact with any object's behaviour.
-\begin{logic}
-\hrule\null
-\textsc{\normalsize *Private Fields Assumption (PFA)}\\
-M,\losigma\ \loturns\ \loexists X*\loin Obj. MayAffect$^{Now}$(X*,o') \loimplies \linebreak
-	\ablock \ablock \loexists Y*\loin Obj. MayAccess$^{Dir,Now}$(Y*,o')
-\linebreak \\
-\textit{By contraposition,}\linebreak
-M,\losigma\ \loturns\ 	\loforall Y*\loin Obj. \loneg MayAccess$^{Dir,Now}$(Y*,o') \loimplies \linebreak
-	\ablock \ablock \loforall X*\loin Obj. \loneg MayAffect$^{Now}$(X*,o')
-\linebreak
-\hrule
-\end{logic}
+Let us see how we further develop this in our framework with another example.\\
 
-\textsc{pfa} means that in order for an object o' to change, it must be done through some object Y* calling its behaviour (Y* can refer to the same object as X*).
-Equivalently, denying all objects in the system the ability to call an object's behaviour implies that no object can modify the object's state.\\
+We have some starting configuration in a system where there is some object o which we have code knowledge of, and that o has the capability of o'.
+Note that we do not need code knowledge of o' to enforce a policy of how o might want to interact with o'.
+Since o has the capabilitiy of o', we say o has the power to enforce the conditions on when and how it will \textit{exercise} o'.
+To put it in another way, we can now allow an outgoing path to exist from o to o', based on some condition \textit{written in o} on how the path to o' can be \textit{used}.\\
+
+If o has only \textit{one} public method that will use the capability of o' based on the condition that the boolean field \texttt{access} within o is evaluated to true at the time of the method call, then we have:
 
 \begin{logic}
-\hrule\null
-\textsc{\normalsize *Global State Change Existence (GSCE)}\\
-M,\losigma\ \loturns\ $[$\loexists X*\loin{Obj}. MayAffect$^{Now}$(X*,o')$]$\linebreak
-\ablock \loimplies \linebreak
-\ablock $[$\loexists Y*\loin{Obj}. MayAccess$^{Dir,Now}$(Y*,o')$]$\linebreak
-\ablock \loimplies \linebreak
-\ablock $[$\loexists Z*\loin{Obj}. MayAccess$^{Ind,Now}$(Z*,o')$]$
-\linebreak\\
-\hrule
+\hr\\
+M,\losigma\ \loturns\ \=\loforall X*. MayAffect(o,X*) \loimplies \\
+\> \loexec{X*}{\losigma} = \loexec{o'}{\losigma} \loand\\
+\> \loexists \losigma'. $[$\=\losigma'\loin Arising(M, \losigma) \loand\\
+\> \>\loexec{o.\texttt{allow}}{\losigma'} = \loexec{\texttt{true}}{\losigma'}$]$\\
+\\
+M,\losigma\ \loturns\ \=\loforall Y*. MayAffect(Y*,o) \loimplies \\
+\> \loexists \losigma'. $[$\=\losigma'\loin Arising(M, \losigma) \loand\\
+\> \>\loexec{o.\texttt{allow}}{\losigma'} = \loexec{\texttt{true}}{\losigma'}$]$\\
+
+\hr
 \end{logic}
-\textsc{gsce} can be derived from our predicate definitions, and makes explicit the connection betwen state change, paths, and capabilities.
-The first implication states that if an object can change the state of o', then there must exist an object in the system that holds the capability of o'.
-The second implication is that there must exist an object in the system that holds a path to o'.
-We view reasoning about state change as reasoning about protection, which we elaborate in the next subsection.
+The logics we have developed show that security of an object can be enforced in an OCap model by:
 
-\subsubsection{Protection is about local object-path denial and global capability denial}
-\begin{logic}
-M,\losigma\ \loturns\ MayAffect$^{Eve}$(o,o') \lonimplies 
-MayAccess$^{Dir,Eve}$(o,o')\linebreak
-\\
-M,\losigma\ \loturns\ MayAffect$^{Eve}$(o,o') \loimplies 
-MayAccess$^{Ind,Eve}$(o,o')\linebreak
-\\
-M,\losigma\ \loturns\ \loneg MayAccess$^{Ind,Eve}$(o,o') \loimplies 
-\loneg MayAffect$^{Eve}$(o,o')\linebreak
-\\
-M,\losigma\ \loturns\ MayAccess$^{Dir,Eve}$(o,o') \loimplies 
-MayAccess$^{Ind,Eve}$(o,o')\linebreak
-\\
-M,\losigma\ \loturns\ \loneg MayAccess$^{Ind,Eve}$(o,o') \loimplies 
-\loneg MayAccess$^{Dir,Eve}$(o,o') \loand\ \loneg MayAffect$^{Eve}$(o,o')\linebreak
-\end{logic}
-We make the important result that protection of an object's state is path denial to the object, where path is a necessary condition for an object to affect another object.
-An object o having the capability of o' is \textbf{NOT} a necessary condition for o to modify the state of o'.
-However, the existence of \textit{some} object X* (that can refer to o, or other objects) having the capability of o', is a necessary condition for o to modify o'.
-See \textsc{nec1} and \textsc{nec2} below which illustrates this.
-This is precisely why attenuating objects are so powerful and useful in OCap patterns.
-We can always deny o the capability of o', but still allow o to modify o' in some protected way.
-To say it in another way, denial of paths is much stronger, as it implies protection \textit{and} denial of capability, but denial of capability does \textit{not} imply protection and denial of paths.\\
-
-\begin{logic}
-\hrule\null
-\textsc{\normalsize *Necessary Execution Condition 1 (NEC1)}\\
-M,\losigma\ \loturns\ MayAffect$^{Eve}$(o,o') \loimplies\linebreak
-	\ablock \loexists X*\loin{Obj}. MayAccess$^{Dir,Eve}$(X*,o') 
-\linebreak\\
-\textit{By contraposition and quantifier equivalence,}\linebreak
-M,\losigma\ \loturns\ $[$\loforall X*\loin{Obj}. \loneg MayAccess$^{Dir,Eve}$(X*,o')$]$\loimplies\linebreak
-	\ablock \loneg MayAffect$^{Eve}$(o,o') \linebreak
-\hrule
-\end{logic}
-\textsc{nec1} says that in order for an object o to modify the state of o', there must exist an object X* that has a direct path to o', and that X* \textit{can} traverse such a path.
-This is derived trivially from our \textsc{pfa} assumption.
-Consequently, in our contraposition result, we can deny \textit{all} objects in our system capability of o' to enforce the protection of o' from o.
-\textsc{nec1} does not yield a very useful result, as this is just another way of implementing protection of o' from o through denying all paths to o' in the system reference graph.
-To see why denying all objects the capability of o' is equivalent to denying all paths to o', see the contraposition result from \textsc{gpe}.\\
-
-In \textsc{nec1}, if X* refers to the same object as o, then we have a straightforward configuration where object o has the capability of o' and can therefore affect o'.
-However, very crucially, o having the capability of o' is \textit{not} a sufficient condition---denying o the capability of o' does \textit{not} deny o the ability to affect o', since X* can refer to an object that is \textit{not} o.
-\textsc{nec2} below is an expansion of \textsc{nec1} to illustrate this crucial point.
-\begin{logic}
-\hrule\null
-\textsc{\normalsize *Necessary Execution Condition 2 (NEC2)}\\
-M,\losigma\ \loturns\ MayAffect$^{Eve}$(o,o') \loimplies\linebreak
-	\ablock $[$MayAccess$^{Dir,Eve}$(o,o') \loor\linebreak
-	\ablock \loexists X*\loin{Obj}, X*\loneq o. MayAccess$^{Dir,Eve}$(X*,o')$]$
-\linebreak\\
-\textit{By contraposition,}\linebreak
-M,\losigma\ \loturns\ $[$\loneg MayAccess$^{Dir,Eve}$(o,o') \loand\linebreak
-	\ablock \loforall X*\loin{Obj}, X*\loneq o. \loneg MayAccess$^{Dir,Eve}$(X*,o')$]$ \linebreak
-	\ablock \loimplies\loneg MayAffect$^{Eve}$(o,o')\linebreak
-\hrule
-\end{logic}
-With this clarification, we then reason about the relationship between protection of state (MayAffect) and paths (MayAccess$^{Ind,*}$).
-The question we ask is what is the relation between state protection in eventual outcomes, and present path configurations?
-We build this relationship from the basics by progressively finding stronger conditions of MayAffect.
-We begin with the first condition, which says that for an object to affect another, an \textit{eventual} path has to exist.
-This can be derived from the definitions of our predicates.
-\begin{logic}
-\hrule\null
-\textsc{\normalsize *Necessary Path Condition 1 (NPC1)}\\
-M,\losigma\ \loturns\ MayAffect$^{Eve}$(o,o') \loimplies\linebreak
-	\ablock MayAccess$^{Ind,Eve}$(o,o')
-\linebreak \\
-\textit{By contraposition,}\linebreak
-M,\losigma\ \loturns\ \loneg MayAccess$^{Ind,Eve}$(o,o') \loimplies\linebreak
-	\ablock \loneg MayAffect$^{Eve}$(o,o')
-\linebreak
-\hrule
-\end{logic}
-Here, we immediately see a first defensive outcome of the OCap model. Having no eventual paths from object o to o' guarantees that the state of object o' cannot be modified by object o.\\
-
-So far our results serve as a good base to enforce \loneg MayAffect$^{Eve}(o,o')$ but we require \textit{stronger} necessary conditions for MayAffect$^{Eve}(o,o')$, because we need to understand the present path configurations that can deny o from modifying o'.
-Luckily, we have already built the necessary ingredients in our formalizations of OCap rules in \ref{ocaprules}, where have painstakingly derive the relationship between present path-capability configurations and eventual paths.
-Since \textsc{npc1} tells us that the eventual paths is a necessary condition for object state modification, and \textsc{epc1} tells us the the relationship between eventual paths and present path-capability configurations, we combine the two to present:
-
-\begin{logic}
-\hrule\null
-\textsc{\normalsize *Necessary Path Condition 2 (NPC2)}\\
-M,\losigma\ \loturns\ MayAffect$^{Eve}$(o,o')\linebreak
-\ablock \loimplies\linebreak
-\ablock \loexists X*\loin Obj.$[($MayAccess$^{Ind,Now}$(o,X*) \loor \linebreak
-\ablock \ablock \ablock \quad MayAccess$^{Dir,Now}$(X*,o)$)$\linebreak
-\ablock \ablock \ablock \ablock \loand \linebreak
-\ablock \ablock \ablock \quad MayAccess$^{Ind,Eve}$(X*,o')$]$
-\linebreak \\
-\textit{By contraposition,}\linebreak
-M,\losigma\ \loturns\ 
-\loforall X*\loin Obj. $[$\loneg MayAccess$^{Ind,Now}$(o,X*) \loand \linebreak
-\ablock \ablock \ablock \ablock \loneg MayAccess$^{Dir,Now}$(X*,o)$]$\linebreak
-\ablock \ablock \ablock \ablock \loor \linebreak
-\ablock \ablock \ablock \ablock \loneg MayAccess$^{Ind,Eve}$(X*,o')$]$\linebreak
-\ablock \loimplies \linebreak
-	\ablock\loneg MayAffect$^{Eve}(o,o')$\linebreak
-\hrule
-\end{logic}
-, where we can always use \textsc{gpec} to expand indirect paths into direct capabilities till we reach the terminating origin o and recursively use the condition to expand the condition till we reach the terminating o'.
-\textsc{npc2} gives us the present path configurations that will allow an eventual path from o to o' as a necessary condition for o to modify o'.
-\textsc{npc2} tells us that we can enforce protection of o' from o by denying all eventual paths from o to o'.
-But if those objects are path-isolated, how do we allow them to cooperate?
-
-\subsection{What about Cooperation?}
-It is good to know we can enforce protection through some present path configurations that can guarantee eventual path-isolation, but in practice this is still not useful.
-Cooperation between objects require paths to exist between objects so it is rare or almost impossible to have two objects o and o' that cooperate while \loneg MayAccess$^{Ind,Eve}$(o,o') holds.
-The quintessential question in OCap systems is whether we can enforce protection such as controlling whether o may affect o' while \textit{still} allowing paths to exist between them.
-This is possible, because of Rule 3 that says objects may have private encapsulation of state and behaviour.
-To begin, let us assume a simple object o' that has \textit{only} private behaviours.
-In such a scenario, since all fields in our system are also assumed to be private, then we know:
-\begin{logic} \loturns\ \loforall X*\loin Obj, X*\loneq o'. \loneg MayAffect$^{Eve}$(X*,o') \end{logic}
-
-Notice, how protection of o' can occur independent of paths and capability, due to object encapsulation.
-This means that objects can specify security policies that it chooses to enforce \textit{internally}.
-Can we further develop our necessary conditions to show this?\\
-
-\textsc{nec2} and \textsc{npc2} show the necessary capability and path conditions for protection.
-Is there a relation that can capture the essence of both?
-Notice how, we have said in \textsc{nec2} that there must exist some object in the system (that is not necssarily o) that has the capability of o', while in \textsc{npc2}, we said that o must have a path to o'.
-Surely it must mean then that the object X* in \textsc{nec2} that has the capability of o' refers to the last object on the path from o to o' mentioned in \textsc{npc2}.
-We now bring these results together, by making use of \textsc{gpec} to make explicit the last object in the eventual path holding the capability of o' in \textsc{npc2}:\\
-\begin{logic}
-\hrule\null
-\textsc{\normalsize *Necessary Execution-Path Condition* (NEPC*)}\\
-M,\losigma\ \loturns\ MayAffect$^{Eve}$(o,o')\linebreak
-\ablock \loimplies\linebreak
-\ablock \loexists X*\loin Obj.$[($MayAccess$^{Ind,Now}$(o,X*) \loor \linebreak
-\ablock \ablock \ablock \ablock MayAccess$^{Dir,Now}$(X*,o)$)$\linebreak
-\ablock \ablock \ablock \ablock \loand \linebreak
-\ablock \loexists Y*\loin{Obj}. $[$MayAccess$^{Ind,Eve}$(X*, Y*) \loand\linebreak
-\ablock \ablock \ablock \ablock MayAccess$^{Dir,Eve}$(Y*,o')$]$
-\linebreak \\
-\textit{By contraposition,}\linebreak
-M,\losigma\ \loturns\ 
-\loforall X*\loin Obj. $[$\loneg MayAccess$^{Ind,Now}$(o,X*) \loand \linebreak
-\ablock \ablock \ablock \ablock \loneg MayAccess$^{Dir,Now}$(X*,o)$]$\linebreak
-\ablock \ablock \ablock \ablock \loor \linebreak
-\ablock \quad\loforall Y*\loin{Obj}. $[$MayAccess$^{Ind,Eve}$(X*, Y*) \loor\linebreak
-\ablock \ablock \ablock \ablock MayAccess$^{Dir,Eve}$(Y*,o')$]$
-\ablock \loimplies \linebreak
-	\ablock\loneg MayAffect$^{Eve}(o,o')$\linebreak
-\hrule
-\end{logic}
-\textsc{nepc*} is nice, because we can see the necessary path condition and the condition that an object must hold the capability of o'.
-But what does it mean to modify an object?
-No matter how long the path from o to o' is, surely to modify o' requires the last object on our path to \textit{be} the 'final messenger' that will call the object's behaviour to modify it---which means MayAffect$^{Eve}$(Y*,o') \textit{must} hold.
-We express this by extending \textsc{nepc1*} to get \textsc{nepc2*}:\linebreak\\
-\begin{minipage}{\linewidth}
-\begin{logic}
-\hrule\null
-\textsc{\normalsize *Necessary Execution-Path Condition 2* (NEPC2*)}\\
-M,\losigma\ \loturns\ MayAffect$^{Eve}$(o,o')\linebreak
-\ablock \loimplies\linebreak
-\ablock \loexists X*\loin Obj.$[($MayAccess$^{Ind,Now}$(o,X*) \loor \linebreak
-\ablock \ablock \ablock \quad MayAccess$^{Dir,Now}$(X*,o)$)$\linebreak
-\ablock \ablock \ablock \ablock \loand \linebreak
-\ablock \loexists Y*\loin{Obj}. $[$MayAccess$^{Ind,Eve}$(X*, Y*) \loand\linebreak
-\ablock \ablock \ablock \quad MayAccess$^{Dir,Eve}$(Y*,o') \loand\linebreak
-\ablock \ablock \ablock \quad MayAffect$^{Eve}$(Y*,o')$]$
-\linebreak \\
-\textit{By contraposition,}\linebreak
-M,\losigma\ \loturns\ 
-\loforall X*\loin Obj. $[$\loneg MayAccess$^{Ind,Now}$(o,X*) \loand \linebreak
-\ablock \ablock \ablock \ablock \loneg MayAccess$^{Dir,Now}$(X*,o)$]$\linebreak
-\ablock \ablock \ablock \ablock \loor \linebreak
-\ablock \quad\loforall Y*\loin{Obj}. $[$MayAccess$^{Ind,Eve}$(X*, Y*) \loor\linebreak
-\ablock \ablock \ablock \ablock MayAccess$^{Dir,Eve}$(Y*,o') \loor\linebreak
-\ablock \ablock \ablock \ablock MayAffect$^{Eve}$(Y*,o')$]$
-\ablock \loimplies \linebreak
-	\ablock\loneg MayAffect$^{Eve}(o,o')$\linebreak
-\hrule\null
-\end{logic}
-\end{minipage}
-\linebreak\\
-
-\textsc{nepc2*} has the intuitive meaning that in order to say object o may modify object o' in some eventual state, \textit{all} these three conditions must hold:
 \begin{itemize}
-\item there must be some present path configuration such that an eventual path from o to o' can arise
-\item there must be an object Y* that sits as the last object along the eventual path to o', that has the capability of o'
-\item object Y* must be able to modify o'
+\item being careful about our present path configurations so that we can ensure eventual path-isolation between the object and untrusted object
+\item specifying security policies {within} objects that dictate how incoming messages and outgoing messages are handled
 \end{itemize}
 
-The logics we have developed show that security of an object can be enforced in an OCap model by:
+Therein lies also the difference between OCap and non-OCap models, because non-OCap models that allow forging of capabilities or a global ambient authoriy \textit{cannot} enforce security through pure path isolation, which is a distinctive feature of OCap models.\\
+
+'Internal' object protection can be done through using defensive programming techniques like encapsulation and checking of conditions for methods to succeed.
+In a way these forms of protection can be considered a form of 'stack-based' access control from the called object's perspective, because they only happen \textit{after} the calling object has already called the object.
+The Ethereum blockchain platform uses prominently such a form of 'stack-based' access control.\\
+
+In the OCap pattern examples we will go through in the next section, protection of an object is typically accomplished through some attenuating object that the protected object knows will never leak its capability directly such that the attenuating objects dominate the trusted object (the trusted object can ensure this if it creates those attenuating objects), and that protection is enforced through some security policies \textit{within} those attenuating objects.
+
+\newpage
+\subsection{Pattern 1: The JavaScript DOM Tree}
+We use a JavaScript Document Object Model (DOM) Tree pattern largely inspired by the example in Devriese et al.\cite{devriese2016} where they use a Kripke worlds framework to reason about the pattern but here instead, we will use our prediate logic framework that we developed in the previous section.
+With the DOM, JavaScript can access and change all the elements of an HTML document.
+Typically, a website will have some kind of external third-party advertisement object, where it is given a particular node (we call it an Ad Node) within the DOM Tree, to render in the website.
+Additionally, we have the Document root node which all other nodes descend from, and hence these nodes form a kind of access hierachy, where higher in the tree are more security-sensitive and powerful.
+We therefore need a way to protect the other nodes in the DOM Tree from the external third party advertisement object.
+Can we protect the other nodes convincingly in an OCap language?
+We illutrate this by first explaining why giving away a direct capability to a node is dangerous.
+We next construct an attenuating restricted node object called ReNode with some configuration that is meant to address the vulnerabilities of the node it is meant to protect.
+This ReNode can prevent access to the immediate parent of the node it is protecting, or allow access to a specified number of levels in the tree.
+Once configured, it can then be safely given away to an external third-party, if we can prove convincingly that the other nodes in the DOM tree are protected.
+We then use our predicate logic framework to prove why, in an OCap system, we can be assured that giving away a ReNode is safe.\\
+
+We begin by stating the variables we will be using, where Object is the most general object in our system which supersets Node and ReNode.
+We only have code knowledge of Node and ReNode, and the question is if giving away a ReNode of a specific configuration to any arbitrary unknown object that we do not have code knowledge of, is safe.
 \begin{itemize}
-\item being careful about our present path configurations so that we can ensure eventual path-isolation between the object and untrusted object
-\item specifying security policies on how \textit{all} objects with the direct capability of our trusted object may modify the object
-\end{itemize}
-Therein lies also the difference between OCap and non-OCap models, because non-OCap models that allow forging of capabilities or a global ambient authoriy \textit{cannot} enforce security through path isolation, which is a distinctive feature of OCap models.\\
-
-The similarity is that both models allow us to specify policies \textit{within} objects as to whether an object's behaviour can be called (non-OCap systems can do so 'externally' through a global ambient authority).
-'Internal' object protection can be done through using defensive programming techniques like encapsulation and checking of conditions for behaviours to succeed.
-In a way these forms of protection can be a form of 'stack-based' access control from the called object's perspective, because they only happen \textit{after} the calling object has already called the object.\\
-
-How do we then express these internal object security policies in our logical framework?
-The nice thing is we already have a predicate MayAffect$^{Eve}$(*,*) that can accomodate this.
-Notice how in \textsc{NEPC2*}, object Y* sits as the last object in the path from o to o' \textit{and} has the capability of o'.
-This gives us a clue as to how we can enable protection of o' without denying paths, by \textit{specifying whether Y* can affect o' explicitly in security policies} that are independent of paths.
-
-\subsubsection{How can security policies be enforced by objects?}
-In the previous example of an object with all fields and behaviours private, we have demonstrated in OCap systems that objects can, at the very least enforce protection policies on \textit{itself}.
-But can we have objects enforcing security policies on \textit{other} objects?
-Let us examine this with another example.\\
-
-If we have some starting configuration in a system where there is some object X* that belongs to a set of objects belonging to the same class called AttenObj that has the capability of o', and no other objects have the capability of o', then we know any eventual path to o' has to include and end with the sub-path from X* to o', \textit{unless} X* leaks the capability of o' to some other object.
-If object X* has no way of leaking the capability of object o', then it has the power to dictate the policies in all eventual states that can affect o' through itself (since it holds the direct capability of o'), and state more precise stronger necessary conditions for the MayAffect$^{Eve}${(X*,o') predicate.
-To put it in another way, we can now allow a path to exist from X* to o', \textit{and} still enforce \loneg MayAffect$^{Eve}$(X*,o'), because the negation of these stronger security policy conditions results in weaker antecedents that imply \loneg MayAffect$^{Eve}$(X*,o').
-\begin{itemize}
-\item X* has no behaviour that will make use of and call the capability of o':
-\begin{logic} M,\losigma\ \loturns\ \loforall X*\loin AttenObj. MayAffect$^{Eve}$(X*,o')\linebreak
-\ablock \loimplies $\bot$\end{logic}
-\item X* has only one behaviour that will modify o' if it holds a boolean field \texttt{access} that is evaluated to true:
-\begin{logic} M,\losigma\ \loturns\ \loforall X*\loin AttenObj. MayAffect$^{Eve}$(X*,o')\linebreak
-\ablock \loimplies\linebreak 
-\ablock \loexists \losigma' \loin Arising(M,\losigma).\linebreak
-\ablock \ablock \loexec{x.\texttt{allow}}{\losigma'} = \loexec{\texttt{true}}{\losigma'} \end{logic}
-\end{itemize}
-A key result from this analysis shows the importance of attenuating objects, objects that sit as the last line of defence in the eventual paths to our protected objects.
-In the OCap pattern examples we will go through in the next section, protection of an object is done through some attenuating object that the trusted object knows will never leak its capability directly (the trusted object can ensure this if it creates those attenuating objects), and that protection is enforced through some security policies within those attenuating objects.\\\
-
-If the protected object can guarantee through the system of configuration that those attenuating objects are the \textit{only} objects in the entire system that holds its capability in \textit{all} eventual states, then the protected object can dictate the conditions and degree of control on how it can be used by untrusted objects in all eventual states through those attenuating objects.
-This ability of the protected object to dictate control across time and conditions enables cooperation, where conditions on how the protected object will cooperate is written in those attenuating objects.
-\\
-
-We now see the power of OCap patterns hence lies in providing concrete examples of a system of cooperation that allows the existence of paths between objects for cooperation while still dictating the degree of control of one object can have over another.
-
-\subsection{Pattern 1: The JavaScript DOM Tree}
-We use a JavaScript DOM Tree OCap pattern largely inspired by the example in Devriese et al.\cite{devriese2016} where they use a Kripke worlds framework to reason about the pattern.
-We define the following variables throughout our pattern:
-\begin{itemize}
-\item o,o' \loin Object
-\item Node,ReNode \losubseteq Object
+\item o,o' \loin UntrustedObject
 \item n,n' \loin Node
@@
 
-\begin{logic}[NodeVul]
-\hrule\null
+\begin{logic}
+\hr\\
 \textsc{\normalsize *Node Vulnerability}\\
-\loforall o, n. MayAccess$^{Dir,Now}$(o,n)\linebreak \ablock\ablock \loimplies \linebreak
-\ablock \ablock MayAffect$^{Now}$(o,n) \loand \linebreak
-\ablock \ablock \loforall n'. MayAccess$^{Dir,Now}$(o,n')
+M,\losigma\ \loturns\ \loforall o,n,n'. \=MayAccess$^{Dir,Now}$(o,n)\\
+\> \loimplies \\
+\> MayAffect$^{Now}$(o,n) \loand \\
+\> MayAccess$^{Dir,Now}$(o,n')\\
+\hr
 \end{logic}
-The vulnerability of a node lies in the fact that it contains a public method \texttt{setProperty(key,value)} that will modify an internal mapping data structure.
-A node also has a public field \texttt{parent} that will return the capability of its parent node.
-Consequently, this allows an object which has the capability of any one node in the tree to navigate up to the root node (Document), and consequently navigate to all other nodes in the tree.
+The vulnerability of a node lies in the fact that it contains one public method \texttt{setProperty(key,value)} that can be called by anyone holding its capability which will modify the properties of the node (by creating or modifying a key-value pair in a map within the node).
+A node also has a public field \texttt{parent} that will return the capability of its parent node in the DOM Tree.
+Consequently, this allows an object which has the direct capability of any one node in the tree to navigate up to the root node (Document), and consequently navigate to all other nodes in the tree and modify them.\\
 
+The question is therefore how do we prove that by giving away an attenuating object 'ReNode' that restricts access to a Node, that the recipient of the ReNode is adhering to the security policies of the system.
+For example, if an unknown object has the capability of a ReNode that allows the unknown object to traverse 2 depth levels upwards in the hierachy of the DOM Tree to modify a node at that level, how can we be sure that the unknown object has no way to traverse more than just 2 levels in the Tree and modify nodes above that level?
+We illustrate this OCap pattern using an example, following which we show the specifications and prove protection of nodes using our predicate framework.
+
+\newpage
 \begin{logic}
-\hrule\null
-\textsc{\normalsize *ReNode to Node Condition*}\\
-M,\losigma\ \loturns\ \loforall rn\loin ReNode. MayAffect$^{Eve}$(rn,n)\linebreak
-\ablock \loimplies \linebreak
-\ablock \loexists \losigma' \loin Arising(M,\losigma).\linebreak
-\ablock \ablock \loexec{rn.\texttt{node}}{\losigma'} = \loexec{e}{\losigma'} \loand\ \{\texttt{e is n}\}\linebreak
-\hrule
+\hr\\
+\textsc{\normalsize *Obj->Node Weakest Condition*}\\
+M,\losigma\ \loturns\ \=\loforall o,rn,n. MayAffect(o,n)\loimplies\\
+\> MayAccess$^{Ind,Eve}$(o,n) \loand\\
+\> \loexists o'. $[$\=MayAffect(o,o') \loand\\
+\> \>MayAffect(o',n)$]$\\
+\hr
 \end{logic}
@@
 \begin{logic}
-\hrule\null
-\textsc{\normalsize *ReNode Parenthood Condition*}\\
-M,\losigma\ \loturns\ \loforall rn,rn.parent \loin ReNode. MayAccess$^{Dir,Eve}$(rn,rn.parent)<---Or do we need a MayCreate predicate?\linebreak
-\ablock \loimplies \linebreak
-\ablock \loexists \losigma' \loin Arising(M,\losigma).\linebreak
-\ablock \loand\ \loexec{rn.\texttt{depth}}{\losigma'} = \loexec{e}{\losigma'} \loand\ e > 0\linebreak
-\ablock \loand\ \loexec{rn.parent \texttt{depth}}{\losigma'} = \loexec{e'}{\losigma'} \loand\ e' = e - 1\linebreak
-\hrule
+\hr\\
+\textsc{\normalsize *Obj->Node with Dom Condition*}\\
+M,\losigma\ \loturns\ \=\loforall o,rn,n. MayAffect(o,n) \loand\ Dom(RN,n)\loimplies\\
+\> MayAccess$^{Ind,Eve}$(o,n) \loand\\
+\> \loexists rn\loin{RN}. $[$\=MayAffect(o,rn) \loand\\
+\> \>MayAffect(rn,n)$]$\\
+\hr
 \end{logic}
+
+\begin{logic} 
+\hr\\
+\textsc{\normalsize *ReNode->Node Condition*}\\
+M,\losigma\ \loturns\ \=\loforall rn,n. MayAffect(rn,n)\loimplies \\
+\>\loexists \losigma' \loin Arising(M,\losigma). \=\loexec{rn.\texttt{node}}{\losigma'} = \loexec{n}{\losigma'}\\
+\hr
+\end{logic}
+
+\begin{logic} 
+\hr\\
+\textsc{\normalsize *ReNode->ReNode Condition*}\\
+M,\losigma\ \loturns\ \loforall rn,rn',n,n'. MayAffect(rn,rn')\\
+\loimplies \loexists \losigma' \loin Arising(M,\losigma).\\
+// \textit{if d>0 then we can create new parent renode} \\
+$[$\=\loexec{rn.\texttt{depth}}{\losigma'} = \loexec{d}{\losigma'} \loand\ $($\loexec{d}{\losigma'}>0$)$ \loand\\
+\>\loexec{rn.\texttt{parent}}{\losigma'} = \loexec{rn'}{\losigma'} \loand\\
+\>\loexec{rn'.\texttt{depth}}{\losigma'} = \loexec{d'}{\losigma'} \loand\ $[($\loexec{d'}{\losigma'}=$($\loexec{d}{\losigma'}-1 $)]$ \loand\\
+\>\loexec{rn'.\texttt{node}}{\losigma'} = \loexec{rn.\texttt{node.parent}}{\losigma'}$]$\\
+\loor\\
+// \textit{create new child node} \\
+$[$\=\loexists i\loin$\mathbb{R}_{\geq 0}$ \loexec{rn.\texttt{child(i)}}{\losigma'} = \loexec{rn'}{\losigma'} \loand\\ 
+\>\loexec{rn'.\texttt{parent}}{\losigma'} = \loexec{rn}{\losigma'} \loand\\
+\>\loexec{rn'.\texttt{depth}}{\losigma'} = \loexec{d'}{\losigma'}  \loand\ $[($\loexec{d'}{\losigma'}=$($\loexec{d}{\losigma'}+1 $)]$ \loand\\\>\loexec{rn'.\texttt{node}}{\losigma'} = \loexec{n.\texttt{child(i)}}{\losigma'}$]$\\
+\hr
+\end{logic}
+\begin{logic} 
+\hr\\
+\textsc{\normalsize *ReNode->Untrusted Condition*}\\
+M,\losigma\ \loturns\ \=\loforall rn,o. MayAffect(rn,o)\loimplies \\
+\> o\loin ReNode \loor\ o\loin Node\\
+\hr
+\end{logic}
+
+\begin{logic} 
+\hr\\
+\textsc{\normalsize *ReNode Definition*}\\
+M,\losigma\ \loturns\ \= \loforall rn,rn',rn'',n, \loforall \losigma'\loin Arising(M,\losigma).\\
+\>\loexec{rn.\texttt{depth}}{\losigma'} = \loexec{d}{\losigma'} \loand \\
+\>\loexec{rn.\texttt{node}}{\losigma'} = \loexec{n}{\losigma'} \loand \\
+\>$[$\=\loexec{rn.\texttt{parent}}{\losigma'} = \loexec{p}{\losigma'} \loand \\
+\>\>$($\loexec{p}{\losigma'}=\loexec{rn'}{\losigma}$)$ \loor\ $($\loexec{p}{\losigma'}=\texttt{null}$)]$\\
+\>\loand \\
+\>$[$\loforall i\loin$\mathbb{R}_{\geq 0}$. \loexec{rn.\texttt{child(i)}}{\losigma'} = \loexec{c}{\losigma'} \loand \\
+\>\>$($\loexec{c}{\losigma'}=\loexec{rn''}{\losigma'}$)$ \loor\ $($\loexec{c}{\losigma'}=\texttt{null}$)]$ \\
+\hr
+\end{logic}
+
+\begin{logic} 
+\hr\\
+\textsc{\normalsize *Node Definition*}\\
+M,\losigma\ \loturns\ \= \loforall n,n',n'',\loforall \losigma'\loin Arising(M,\losigma).\\
+\>$[$\=\loexec{n.\texttt{parent}}{\losigma'} = \loexec{p}{\losigma'} \loand \\
+\>\>$($\loexec{p}{\losigma'}=\loexec{n'}{\losigma'}$)$ \loor\ $($\loexec{p}{\losigma'}=\texttt{null}$)$ $]$\\
+\>\loand \\
+\>$[$\=\loforall i\loin$\mathbb{R}_{\geq 0}$. \loexec{n.\texttt{child(i)}}{\losigma'} = \loexec{c}{\losigma'} \loand \\
+\>\>$($\loexec{c}{\losigma'}=\loexec{n''}{\losigma'}$)$ \loor\ $($\loexec{c}{\losigma'}=\texttt{null}$)$ $]$\\
+\hr
+\end{logic}
+\newpage
+\noindent%
+\begin{minipage}{\textwidth}
+\centering
+\includegraphics[width=1.05\textwidth]{figures/DOM.pdf}
+ \captionof{figure}{JavaScript DOM Tree OCap Pattern}
+  \label{DOM example}
+\end{minipage}
+\newpage\null\newpage
+\section{Appendix}\label{appendix}
 %----------------------------------------------------------------------------------------

### 2017-09-07 shupeng_msc_report.tex
update

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Journal Article
% LaTeX Template
% Version 1.4 (15/5/16)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% Original author:
% Frits Wenneker (http://www.howtotex.com) with extensive modifications by
% Vel (vel@LaTeXTemplates.com)
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------
\newcommand{\reporttitle}{Risk and Trust in Open Systems}
\newcommand{\reportauthor}{Shu Peng Loh}
\newcommand{\supervisor}{Sophia Drossopoulou}
\newcommand{\degreetype}{Computing Science}
\documentclass[a4paper,11pt,twoside]{article}
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
\usepackage[hang,flushmargin]{footmisc} 
\definecolor{light-gray}{gray}{0.95}
\usepackage{listings}
\lstset{
    breaklines=true,
    columns=flexible,
    backgroundcolor=\color{light-gray},
    basicstyle=\sffamily\small,
    escapechar=\&
}
\usepackage[sc]{mathpazo} % Use the Palatino font
\usepackage[T1]{fontenc} % Use 8-bit encoding that has 256 glyphs
\linespread{1.15} % Line spacing - Palatino needs more space between lines
\usepackage{microtype} % Slightly tweak font spacing for aesthetics
\usepackage[english]{babel} % Language hyphenation and typographical rules
\usepackage{multirow}
\usepackage{multicol}
\usepackage[left=2.5cm,right=2.5cm,top=3cm,bottom=3cm, hmarginratio=1:1,headheight=14pt,columnsep=15pt]{geometry} % Document margins
\usepackage[hang, normal,labelfont=bf,up,textfont=sc,up]{caption} % Custom captions under/above floats in tables or figures
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{enumitem} % Customized lists
\usepackage{color}
\usepackage{pdfpages}
\setlist[itemize]{noitemsep} % Make itemize lists more compact
\usepackage{ latexsym }
\usepackage{abstract} % Allows abstract customization
\renewcommand{\abstractnamefont}{\normalfont\bfseries} % Set the "Abstract" text to bold
\renewcommand{\abstracttextfont}{\normalfont\small} % Set the abstract itself to small italic text

\usepackage{fancyhdr} % Headers and footers
\fancyhead{} % Blank out the default header
\fancyfoot{} % Blank out the default footer
%\fancyhead[C]{Running title $\bullet$ May 2016 $\bullet$ Vol. XXI, No. 1} % Custom header text
\fancyfoot[RO,LE]{\thepage} % Custom footer text

\usepackage{titling} % Customizing the title section
\usepackage{titlesec}
\titleformat{name=\section}{\flushleft\bf\scshape\Large}{\thesection}{.5em}{}
\titleformat{name=\subsection}{\flushleft\bf\scshape\large}{\thesubsection}{.5em}{}
\titleformat{name=\subsubsection}{\flushleft\bf\scshape\normalsize}{\thesubsubsection}{.5em}{}
\usepackage[export]{adjustbox}
\usepackage{hyperref} % For hyperlinks in the PDF
\usepackage{cleveref}
\crefname{paragraph}{\S}{\S\S}
\crefformat{section}{\S#2#1#3}
\crefformat{subsection}{\S#2#1#3}
\crefformat{subsubsection}{\S#2#1#3}
\usepackage{amsfonts} % For more logic symbols
\usepackage{amssymb} % For more logic symbols
\usepackage[fleqn]{amsmath} % left aligned math
\usepackage{centernot} % to use not implies
\usepackage{array}
\usepackage{upgreek}
\setlength\parindent{0pt}
%----------------------------------------------------------------------------------------
%	Custom Logic Environment and  Bindings
%----------------------------------------------------------------------------------------

%environment formating
\newenvironment{logic}
{\begin{minipage}[c]{\linewidth}  \sffamily \mdseries \begin{tabbing}}
{\end{tabbing}\end{minipage}\vspace{0.3em}}
%logic shortcuts
\newcommand*\rot{\rotatebox{90}}
\newcommand{\loin}{$\in$}
\newcommand{\lonin}{$\centernot\in$}
\newcommand{\loforall}{$\forall$}
\newcommand{\locontrad}{$\bot$}
\newcommand{\loexists}{$\exists$}
\newcommand{\loand}{$\land$}
\newcommand{\loor} {$\lor$}
\newcommand{\loneq} {$\neq$}
\newcommand{\losimeq} {$\simeq$}
\newcommand{\losubseteq}{$\subseteq$}
\newcommand{\loimplies}{$\implies$}
\newcommand{\lonimplies}{$\centernot\implies$}
\newcommand{\loimpliedby}{$\Longleftarrow$}
\newcommand{\lonimpliedby}{$\centernot\Longleftarrow$}
\newcommand{\losigma}{\text{$\upsigma$}}
\newcommand{\loturns} {$\vDash$}
\newcommand{\lochi}{$\upchi$}
\newcommand{\lophi}{$\upvarphi$}
\newcommand{\lorarrow}{$\rightarrow$}
\newcommand{\lonturns} {$\nvDash$}
\newcommand{\loiff} {$\iff$}
\newcommand{\loleadsto} {$\leadsto$}
\newcommand{\lomapsto} {$\mapsto$}
\newcommand{\locup} {$\cup$}
\newcommand{\lobigcup} {$\bigcup$}
\newcommand{\lotimes} {$\times$}
\newcommand{\loneg}{$\boldsymbol \neg$}
\newcommand{\loexec}[2] {$\lfloor$#1$\rfloor _{\text{#2}}$}
\newcommand{\loconj}[1] {$\bar{\text{#1}}$}
\newcommand{\ablock} {\null\qquad}
\newcommand{\lotiff} {\textit{\textls[-20]{iff}}}
\newcommand{\hr}{\rule{\linewidth}{0.4pt}}
\newcommand{\lonat}[1][] {$\mathbb{N}_{#1}$}
\newcommand{\losup}[1]{$^{\text{#1}}$}
\newcommand{\logeq}{$\geq$\ }

\setcounter{secnumdepth}{4}

\titleformat{\paragraph}
{\normalfont\normalsize\bfseries}{\theparagraph}{1em}{}
\titlespacing*{\paragraph}
{0pt}{3.25ex plus 1ex minus .2ex}{1.5ex plus .2ex}

%-----------------------
%----------------------------------------------------------------------------------------
%	TITLE SECTION
%----------------------------------------------------------------------------------------
\setlength{\droptitle}{-4\baselineskip} % Move the title up

\pretitle{\begin{center}\Huge\bfseries} % Article title formatting
\posttitle{\end{center}} % Article title closing formatting
\title{Risk and Trust in Open Systems:\\ \Large Towards formalising Permission and Authority, and specifying policies for Object Capability Patterns} % Article title
\author{%
\textsc{Shu-Peng Loh} \\[1ex] % Your name
\normalsize Imperial College London \\ % Your institution
}
\date{} % Leave empty to omit a date

\makeatletter
\DeclareRobustCommand{\emp}{%
  \@nomath\em \if b\expandafter\@car\f@series\@nil
  \normalfont \else \sffamily \bfseries \fi}
\makeatother
\renewcommand{\maketitlehookd}{%
\begin{abstract}
\noindent  We revisit the concepts of \textit{permission} and \textit{authority} in the works of Miller\cite{miller2006} and \linebreak Drossopoulou et al. \cite{drossopoulou2016}, and we contribute to the literature by proposing new formal definitions of permission and authority (MayAccess and MayCall predicates in our paper). We show how our definitions can be used to reason about object-to-object interactions, and in particular, cooperation between objects and protection of objects from unknown code. Another contribution we made is that we propose and show how the concept of domination over objects can be used to specify and reason about safe or vulnerable cooperation, which is novel but inspired by the work of Clarke et al. on ownership types \cite{clarke1998}. Our paper then show how using a mix of our proposed formalisations of permission and authority, and Hoare logics, we can formally specify well-established Object-Capability (OCap) patterns in the form of \textit{OCap policies} \cite{drossopoulou2013,drossopoulou2014,drossopoulou2015,drossopoulou2015b}.  In particular, we reason about protecting the properties of nodes in a DOM Tree, inspired by the works of Maffeis et al.\cite{maffeis2010} and Devriese et al.\cite{devriese2016}, but we do so within the capability-safe language Pony \cite{clebsch2015}. We believe our methodology and specifications of OCap patterns are less complex than other methodologies in the literature\cite{devriese2016,swasey2017}, which will allow a programmer to reason convincingly how the use of attenuating objects in capability-safe languages can help to facilitate cooperation betweeen known and unknown code, while still preventing vulnerabilities by preserving desired properties in an open system. Lastly, we end our paper with some preliminary insights on thinking about security and risk in a non-OCap system by using the Ethereum\cite{wood2014} blockchain as a case study.\end{abstract}
}
\begin{document}
\let\sf\textsf
\input{titlepage}
\raggedbottom
\thispagestyle{empty}
\tableofcontents
\thispagestyle{empty}

\maketitle
\setcounter{page}{1}
\pagestyle{fancy} % All pages have headers and footers

%----------------------------------------------------------------------------------------
%	ARTICLE CONTENTS
%----------------------------------------------------------------------------------------

\section{Introduction}
The power of distributed modern computing lies in facilitating cooperation between multiple agents, but it comes with risk as an agent is vulnerable to \textit{unexpected} outcomes\footnote{Representing in general any outcome arising from a piece of code execution that has deviated from an agent's original intention or objective independent from code.}. This might generally arise from two issues: 
\begin{itemize}
\item oversight or misconception of the outcomes of executing a piece of \textit{known} code
\item failure to defend against malicious execution of \textit{unknown} code components\end{itemize}
\noindent These two issue are often closely intertwined in any system of execution that has both trusted and untrusted code components (the second issue is often a result of the first).\\

It is therefore not surprising that \textit{thinking} about security while building a complex system of cooperation that involves the interplay between trusted(known) and untrusted(unknown) code components is extremely challenging to any programmer. To some programmers, security, is often treated as a \textit{separate} mental burden to be considered separately from the functionality of a system. In recent years the Object-Capability (OCap) model has received increasing attention as a compelling approach to building robust, distributed systems that promote what Miller\cite{miller2006} calls \textit{cooperation without vulnerability}. The OCap model attemps to address these two issues by alleviating security as a separate concern from the mind of the programmer. This is done through leveraging the object-oriented programming paradigm and imposing certain prohibitions or restrictions. Within the OCap literature, there are two central concepts that describe how objects interact: permission and authority. The OCap literature also features prominently how using an OCap system would accomplish cooperation without vulnerability through Object-Capability Patterns (OCap patterns). OCap patterns are important, because they demonstrate how objects can cooperate with unknown code while preserving desirable properties in a system.\\
\subsection{Our contributions}
While there have been considerable work done on OCap patterns in programming languages \cite{miller2003b,miller2006,murray2010}, less work has been done on formally specifying them. The only works we know so far are \cite{drossopoulou2015b,devriese2016,swasey2017}, which are all very recent in the literature. Our paper builds on the works of \cite{drossopoulou2015b}, because we claim that the methodology of using OCap policies to specify OCap patterns to be intuitive and straightforward. While they may seem less sophisticated than the methodologies in other works, we believe they are powerful enough to reason convincingly about the interactions between trusted code and code of unknown provenance. To formally specify OCap patterns in the style of OCap policies, our paper contributes to the literature by proposing formal predicates for permission and authority that describe access dynamics between objects. Miller first proposed the concepts of permission and authority in his PhD thesis\cite{miller2006}, following which the formal definitions of these concepts were proposed and refined in the work of Drossopoulou et al. \cite{drossopoulou2016}. Our paper defines permission and authority in a formal way that is different from the aforementioned works, but we generally adopt the same informal meanings (that having permission means the right to invoke an object's behaviours, and authority to mean the ability to cause effects). More precisely, our paper defines \textit{permission} as having the reference of another object (right to invoke that object's behaviours), either directly or indirectly---through multiple object reference(s)---and either currently or eventually---in some current or future state of the system; and defines \textit{authority} to mean being \textit{able} to invoke an object's behaviours; i.e. we distinguish between having the \textit{right to do} something (permission) and \textit{being able to do} something (authority)\footnote{An anology would be having the right to perform an impossible task does not imply the task becomes possible.}. Furthermore, our paper implicitly treats invoking behaviours of objects as 'effects' in the system. We claim that these insights are novel and our motivation for defining them can be found in \cref{sec:permauth}.\\

To put our formal definitions of permission and authority in practice and demonstrate their usefulness, we further propose the novel use of the concept of domination over objects that is inspired by the works of \cite{clarke1998} on ownership types. Our domination predicate describes a special set of objects that dominate a particular object---all permissions to use the particular object have to involve the permission of a member of the dominating set, and all authorities have to involve the authority of a member of the dominating set.\\

With our formal definitions of permission, authority, and domination, our paper then argues how we can use them to reason and distinguish concepts of \textit{isolation}, \textit{cooperation}, \textit{vulnerability} and \textit{protection}. We claim that our framework allows one to understand and reason about these concepts in a highly straightforward and intuitive way without treating security as a separate burden from the functionality of the system, which we hope would encourage programmers to seek the use of the object-capability model to build robust and secure systems.\\

To show this in practice, our paper provides an illustration of three OCap patterns found in \cref{sec:figdom,sec:figcaretaker,sec:figmem}, but we focus only on providing the OCap policy specifications for the DOM Tree pattern in \cref{sec:dompolicies}. We choose the DOM Tree pattern because we are convinced of the importance of reasoning about unknown code using the object-capability framework in the web security space, where guest/unknown code (such as third party scripted advertising or widgets) are pervasive in the web today (details can be found in the work by Maffeis et al.\cite{maffeis2010}). We show how our policies can specify the preservation of properties of nodes, in the presence of unknown code execution. Lastly, in \cref{sec:ethereum}, we offer some preliminary insights on the security properties of smart contracts on the Ethereum blockchain system that has experienced explosive growth in recent years, and we recognise that further work has to be done in this exciting new field.

\subsection{Overview of Paper}
In our paper, we first give an overview background history on the OCap model in \cref{sec:ocapmodel}. Readers familiar with the object-capability model and its languages are encouraged to skim through this section. In \cref{sec:specs}, we seek to formally specify the definitions of permission and authority and use our specifications to describe and reason how objects interact with each other. Our formal specifications are broad enough to reason in general about how two objects interact in our small specification language. That said, there are benefits to using these concepts to reason within an OCap system, which we demonstrate in \cref{sec:ocapspecs}, because the restrictions placed by OCap languages or what we call capability-safe languages allow us to enforce protection of objects more easily. We further solidify how our methodology can be used to reason convincgly in an OCap system by applying it on a DOM Tree pattern within an OCap framework, in \cref{sec:ocappatterns}.
%----------------------------------------------------------------------------------------
%   OCap Model
%----------------------------------------------------------------------------------------
\section{Object-Capability Model}\label{sec:ocapmodel}
\subsection{From Object to Capability}
The OCap model uses the reference graph of the objects as the access graph, and strictly requires objects to interact with each other only by sending messages on object references\cite{miller2003b}. When these references cannot be forged in the system, and when there there are no default globally accessible objects (for an object to be globally accessible, the reference to use that object has to be \textit{explicitly} granted to every object globally), then it becomes necessary that for an object to call the methods of another object, an object needs to explicitly obtain the reference of the object it is calling.  Object references therefore embody authority, and we use the word capability to describe the encapsulation of both reference (designation) and authority which are unforgeable.

\subsection{From Capability to Object-Capability}
Origins of capabilities date back to Dennis and Van Horn\cite{dennis1966}'supervisor as a mechanism of protecting low-level resources like memory segments. Early attempts to implement the capability model include the MIT PDP-1 computer and the CAL-TSS system. The capability model was then extended by the early operating system designers of HYDRA\cite{wulf1974} into a protected ability to invoke arbitrary services provided by other processes. Capabilities were later implemented on computer and operating systems such as CAP\cite{wilkes1979}, KeyKOS\cite{hardy1985}, and EROS\cite{shapiro1999}. The development of capabilities and a detailed examination of the object-capability model with the capability-safe programming language E, can be found in Miller's PhD thesis\cite{miller2006}. Capability-safe programming languages that allow building a system based on the object-capability model are:\begin{itemize}
\item Joe-E (inspired by Java)
\item Emily (inspired by OCaml)
\item Caja (inspired by JavaScript)
\item E
\item Pony
\end{itemize}
\subsubsection{Difference between capability and object-capability}
In his seminal paper \textit{Protection}\cite{lampson1974}, Lampson defined protection as a general term for all the mechanisms that control the access of a program to other things in the system. Specifically, Lampson introduced the idea of an Access Control Matrix, an abstract, formal security model that describes the rights of each subject with respect to every object in the system, at a given point in time. The matrix contains one row per subject and one column per object, and each cell entry (for a particular subject-object pair) indicates the access mode that the subject is permitted to exercise on the object. Each column represents an "access control list" (ACL) for the object; and each row represents an "access profile" for the subject. Since a capability model associates each subject with a list of capabilities, it would appear that the model can be represented by a row-based view of the matrix.\\

However, in the technical report \textit{Capability Myths Demolished}\cite{miller2003}, Miller et al. identified that a capabilities-as-rows interpretation assumes an ambient authority security property---subjects are not required to indicate a specific authority in order to exercise it---and in total identified seven security properties that capture the distinction between ACL, capabilities-as-rows, capabilities-as-keys, and object-capabilities. In the same report, Miller et al. argue that the "true" capability model \textit{is} the object-capability model, because all known major capability systems take the object-based approach and capability-based systems are explicitly characterised as "object-based". The object-based model of computation can be seen as a dynamic reference graph of objects, where there is no distinction between subjects and objects. The object-capability model uses the reference graph as the access graph, requiring that objects can interact only by sending messages on references. An object's state therefore consists of both data and the capabilities---unforgeable object references that allow its holder to send messages to the object it references by invoking the capability.

\subsection{Capability-safe Language Restrictions}
In this section, we briefly mention three critical restrictions required in a capability-safe language: memory safety, no global mutable variables (objects can only interact with each other on the reference graph), and the taming of reflection.
\subsubsection{No Capability Forging/Memory-safety}
The object-capability model requires a capability to be \textit{unforgeable}. Without memory safety in the programming language, a reference to a particular object in a memory can potentially be forged externally and anywhere. For example in C++, because the language allows memory manipulation, designation is not opaque and hence forgeable. Let us assume a pointer ptr that holds the designation of a newly created int object \texttt{\small int* ptr = new int()}. The int object can thus be accessed  through the use of ptr. However C++ allows the programmer access to the memory and we can obtain the a string of bits that represent the memory address of the int object. Let us assume this memory address is represented by the bits 0x12345678. We say Designation is forgeable and accessible by bits because any object can perform the following operation:\\
\ablock\texttt{\small int* ptr2 = reinterpret\_cast<int*>(0x12345678)}\\
to obtain a new pointer variable to the same int object. On the other hand, in java, object designations are completely opaque/indivisible--even with knowledge of the bits that constitute a pointer (obtained by Object.toString()); there is no operation for obtaining the corresponding pointer from this bit information and therefore designation cannot be forged.
\subsubsection{No Global Mutable Variables}
This restriction is necessary because if there are global mutable variables, then objects can now communicate without needing the object capabilities of each other. Global mutable variables can behave like a message carrier, such that two objects can both communicate by monitoring and modifying the state of the global variable. The capability-safe language Pony, which we use to show the code of OCap patterns, do not allow global variables.
\subsubsection{No 'Untamed' Reflection}
Reflection is the ability of a piece of code to "reflect" on the structure of other code in the same system or itself. More concretely in object-oriented systems, an object that can perform untamed reflection is about to examine and modify the code of other objects or itself. Take for example in Java, reflection allows code to perform operations that would otherwise be illegal in non-reflective code, such as accessing private fields and methods. Let us assume a simple Java program where object A possesses the reference capability of object B, and object B possesses the reference capability of object C. Object B stores the reference capability of object C in a private field and does not have any public method to divulge this data. With reflection, we cannot guarantee that object A will never obtain the capability of object C - this is because with the capability of object B, object A can reflect on object B's code and get all information stored in object B - including the private field that holds the reference capability of object C. In order to ensure capability-safety, the reflection feature in a language has to be 'tamed' or restricted. Joe-E\cite{mettler2010}, a security oriented subset of Java, has such a feature.
\subsection{Related Works}

Drossopoulou et al. \cite{drossopoulou2015b} 
propose that reasoning about object-capability patterns require extensions in the logics that talk about programs, and those extensions are not about what currently is in the heap, but about what a program might do in the future.
They use predicates modelling trust and risk, and used those predicates to specify a capability-based version of Miller's escrow exchange example\cite{miller2000}. Drossopoulou and Noble\cite{drossopoulou2013} first re-wrote the escrow exchange example in Joe-E and E, and in the process of writing the code, they argue that code written in existing capability safe languages is too focused on the low-level mechanism rather than higher-level security policies. Furthermore, the code required to describe such security policies tend to be interwoven and mixed with the parts implementing the functionality, resulting in these policies being \textit{implicit}. Thus, they propose  a specification language where policies are made explicit, to define policies required in the escrow exchange example\cite{drossopoulou2014, drossopoulou2015, drossopoulou2015b}. More recently, Drossopoulou et al.\cite{drossopoulou2016}'s work further introduced formalisations with regards to the distinction between permission and authority in the object-capability literature.\\

Maffeis et al.\cite{maffeis2010}'s formal system defines authority in a topological manner where objects are represented as nodes in a graph, and a path between two nodes implies that the source node can access the destination node. Their work brings object-capability to web security, which is also a motivation for our paper's work on specifying the DOM Tree OCap pattern. They formalized concepts of capability safety and authority safety, and proved that capability safety implies authority safety, which in turn implies resource isolation. They demonstrated these guarantees in a Caja-based subset of JavaScript.\\

Devriese et al.\cite{devriese2016} present a formalisation of reasoning about object capabilities that is based on a Kripke logical relations model, in a language with higher-order state. Using a simple core calculus for JavaScript, they allow bounding the behaviour of a program fragment based on the capabilities it has access to, and use their model to verify and reason about several simple capability patterns, including the DOM Tree pattern. \\

Swasey et al.\cite{swasey2017} present a logics for OCP called OCPL, a formal system for compositionally specifying and verifying the security guarantees provided by OCPs, with a language that has higher-order functions, state, and concurrency. OCPL allows modular reasoning about OCP implementations and user code that depends on them, and to specify a general property on user code that ensures such code can be safely shared with untrusted code without having its internal invariants violated.

%----------------------------------------------------------------------------------------
%   Formal Specifications
%----------------------------------------------------------------------------------------
\newpage
\section{Formal specifications}\label{sec:specs}
In this section we describe a small object-oriented language in \cref{sec:speclang} and define predicates that allow us to specify object-capability policies in \cref{sec:permauth}. For our language, we use the definitions of modules, classes, runtime state, interpretation, execution and arising configurations from the specification language in the appendix of \cite{drossopoulou2015b}. To express more precisely how an object can interact with another object, we use the ideas of permission and authority in Miller's\cite{miller2006} and Drossopoulou et al.'s\cite{drossopoulou2016} works. In their papers and our paper, we generally try to capture the informal meaning of:
\begin{itemize}
\item permission: access right of an object to invoke a behaviour of an object
\item authority: ability to cause effects
\end{itemize}
In the approach of our paper however, we contribute to the literature by taking a novel approach in formally defining both permission and authority, though we argue they generally capture the same informal meanings. For permission, we have four different modes permissions specified by four MayAccess predicates. Given a module M and state \losigma,:
\begin{tabbing}
MayAccess$^{Dir,Now}$(o,o$'$): \=o has an object reference of o$'$\\
MayAccess$^{Ind,Now}$(o,o$'$): o contains a reference that points to an object that has a reference\\
\>to another object---thus forming a series of object reference(s)---\\ \> where the last object in the series has an object reference of o$'$\\
MayAccess$^{Dir,Eve}$(o,o$'$): there will be some eventual state \losigma$'$ arising from \losigma\,\\ \> where M,\losigma$'$\ \loturns\ MayAccess$^{Dir,Now}$(o,o$'$) holds\\
MayAccess$^{Ind,Eve}$(o,o$'$): there will be some eventual state \losigma$'$ arising from \losigma\,\\ \> where M,\losigma$'$\ \loturns\ MayAccess$^{Ind,Now}$(o,o$'$) holds\\
\end{tabbing}
For authority, we propose a new MayCall predicate that deviates from the MayAffect predicate proposed in Drossopoulou et al.\cite{drossopoulou2016}. While their MayAffect predicate captures the idea of a field of an object being modified, our MayCall predicate is much weaker in the sense that mean the ability to invoke a method of an object \textit{being}. Given a module M and state \losigma,:
\begin{tabbing}
MayCall(o,o$'$): \=there exist a method in o that can call a method in o$'$,\\ \> such that there will be some eventual state \losigma$'$ arising from \losigma,\\ \> where the receiver in the stack has transitioned from o in \losigma, to o$'$ in \losigma$'$\\
\end{tabbing}

We further elaborate these predicates in \cref{MayAccess} and \cref{MayCall}.
\subsection{Specification Language}\label{sec:speclang}
In our small specification language, we briefly describe the defintions of module, class, runtime state, interpretation, execution, arising configurations, and assertion.
\textbf{Module:}\\
A module maps class identifiers to class descriptions:\\
\begin{logic}
M \loin\ Module = ClassId \locup\ SpecId\\\ablock\qquad\qquad\quad $\rightarrow$ \\
\ablock \qquad \qquad \quad (ClassDescr \locup\ Specification)
\end{logic}\\

\textbf{Class:}\\
Class definitions describe how a class contributes to the runtime behaviour of a program for which we use its field and method declarations. We define method bodies as consisting a sequences of statements; these statements can be field read or internal field assignments (only allowed if the object is \texttt{this}), conditionals, and method calls. We also define all classes to be private which mean only objects belonging to class C may construct objects of class C.\\

\begin{logic}
ClassDescr\quad\=:\=:= [\textbf{private}] \textbf{class} \=ClassId\\
\>\>\>\textbf{implements} SpecId*\\
\>\>\{ \=(\textbf{field} FieldId)*\\
\>\>\> (methBody)*\\
\>\>\> (FunDescr)*\\
\>\>\> (PredDescr)* \}\\
methBody\>::= \textbf{method} m (ParId*)\\
\>\>\{ Stmts; \textbf{return} Arg \}\\
Stmts\>::= Stmt | Stmt; Stmts\\
Stmt\>::= \textbf{var} VarId := Rhs\\
\>\>$|$ VarId := Rhs\\
\>\>$|$ \textbf{this}.FieldId := Rhs\\
\>\>$|$ \textbf{if} Arg then Stmts \textbf{else} Stmts\\
\>\>$|$ \textbf{skip}\\
Rhs\>::= Arg.MethId( Arg* ) | Arg\\
\>\>$|$ \textbf{new} ClassId( Arg* )\\
Arg\>::= ParId | VarId | \textbf{this}\\
\>\>$|$ \textbf{this}.FieldId\\
\end{logic}\\
Note that we work with all fields being private in our model - this restricts each object to being able to modify only its own fields.\\

\textbf{Runtime state:}\\
\losigma\ consists of a stack frame \lophi, and a heap \lochi. A stack frame is a mapping from receiver (this) to its address, and from the local variables (VarId) and parameters (ParId) to their values. Values are integers, the booleans true or false, addresses, or null. The heap maps addresses to objects. Objects are tuples consisting of the class of the object, and a mapping from field identifiers onto values.\\

 \begin{logic}
\losigma\ \loin\ state \quad\= = frame \lotimes\ heap \\
\lophi \loin\ frame \>= StackId $\rightarrow$ val \\
\lochi \loin\ heap \>= addr $\rightarrow$ object \\
v \loin\ val \>= \{\texttt{null}, \texttt{true}, \texttt{false}\} \locup\ \textit{addr} \locup\ $\mathbb{N}$ \\
StackId \>= \{\textit{this}\} \locup\ VarId \locup\ ParId \\
object \>= ClassId \lotimes (FieldId $\rightarrow$ val)\\
$\iota$,$\iota$',.. \>\loin\ \textit{addr}\\
\end{logic}

\textbf{Interpretation:}\\
For a state \losigma = (\lophi,\lochi), we define the partial function:\\

\begin{logic}
\quad \loexec{\_}{\_}: state $\times$ Path $\rightarrow$ Value\\
\end{logic}

as follows:\\
\begin{logic}
\quad\=\loexec{\textbf{null}}{\losigma}\quad \= = \textbf{null}\\
\>\loexec{\textbf{true}}{\losigma} \>= \textbf{true}\\
\>\loexec{\textbf{false}}{\losigma} \>= \textbf{false}\\
\>\loexec{x}{\losigma} \>= \lophi(x) for x \loin StackId,\\
\>\>undefined, otherwise.\\
\> \loexec{p.f}{\losigma} \>= \lochi(\loexec{p}{\losigma})(f)\quad  \=if \loexec{p}{\losigma} is defined,\\
\>\>\>and f is a field of \loexec{this}{\losigma}\\
\>\>undefined, otherwise.\\
where,\\
\> p \loin path \>::= x $|$ p.f\\
We also define the lookup of the class of an
expression:\\
\> Class(e)$_\losigma$ \> = (\losigma$\downarrow_2$\loexec{e}{\losigma})$\downarrow_1$ if \loexec{e}{\losigma} defined\\
\>\>undefined, otherwise.\\
\end{logic}\\

\textbf{Execution:}\\
Execution uses module M, and maps a runtime state \losigma\ and statements stmts onto a new state \losigma$'$. We keep our model simple by not giving execution rules for outcomes like null-pointer-exception or stuck execution. Execution of statements and expressions has the following shape:\\

\begin{logic}
\quad \loleadsto \quad : Module $\times$ state $\times$ Stmts $\rightarrow$ state\\
\quad \loleadsto \quad : Module $\times$ state $\times$ Rhs $\rightarrow$ heap $\times$ val\\
\end{logic}

\textbf{Arising Configurations:}\\
We define Arising(M, \losigma) as the set of runtime configurations which may be reached during execution of some initial context (\losigma$_0$, stmts$_0$). The mappings of Init and Arising are as follows:\\

\begin{logic}
Init \ablock \=: Module $\rightarrow$ P(State $\times$ Stmt)\\
Arising \>: Module $\rightarrow$ P(State $\times$ Stmts)\\
\end{logic}

We say a context is initial if its heap contains only objects of class Object, and initial configuration should be kept as minimal as possible, with a heap that has only one object and executing a method call on a newly created object with another newly created object as an argument:\\

\begin{logic}
Init\=$\{$(M,\losigma) = (\losigma$_0$, \textbf{new} c.m(\textbf{new} c')) $|$ c,c' \loin dom(M)\\ \> where \losigma$_0$ = (($\iota$,\textbf{null}),\lochi$_0$),\\
\>and \lochi$_0$($\iota$) = (Object, $\emptyset$)$\}$\\
Arising(M,\losigma) = $\bigcup_{\text{(\losigma,stmts)\loin Init(M)}}$ Reach(M,\losigma,stmts)\\
\end{logic}

We also say a configuration is reachable from another configuration, if the former may be required for the evaluation of the latter after any number of steps. Reach takes the following shape:\\

\begin{logic}
Reach : Module $\times$ state $\times$ Stmts\\
\null\qquad $\rightarrow$ P(Stmts $\times$ state)\\
\end{logic}

The set Reach(M,\losigma,stmts) collects all configurations reachable during execution of \losigma,stmts. Note that the function Reach(M,\losigma,stmts) is defined even when execution diverges so it may be an infinite set. We therefore can give meaning to capability policies without requiring termination.\\

\textbf{Assertion}\\
The syntax of assertion can be found in the works of \cite{drossopoulou2015b}, where we have placed the relevant section in Appendix \cref{app:assertion} of our paper , and is more or less as expected. Validity of assertion has the shape:\\

\begin{logic}
M,\losigma\ \loturns\ A\\
\end{logic}

and holds in context of a module \sf{M} and runtime configuration \sf{\losigma}, where more details can be found in the mentioned appendix section above.\\

We say that a class \sf{C} satisfies an assertion \sf{A} in the context of a module \sf{M}, if in all runtime configuration \losigma$'$ arising from execution of any module \sf{M$'$} linked with module \sf{M}, all objects of class \sf{C} satisfies \sf{A}:\\

\begin{logic}
M \loturns\ C:A \loiff\ \loforall M$'$,\losigma\loin Arising(M$*$M$'$,\losigma). [ M$*$M$'$,\losigma\ \loturns\ x:C \loimplies A[x/this] ]\\
\end{logic}

We define linking of modules, M$*$M$'$ to be the union of their respective mappings, provided that the domain of the two modules are disjoint with respect to classes.We need module linking to allow us to reason about policies in the execution of unknown code, where module \sf{M$'$} can be unknown.  More details on module linking can be found in Appendix \cref{app:modulelinking}.
\subsection{Permission and Authority}\label{sec:permauth}
The concepts of permission and authority are central to reasoning about object-to-object interactions, and are especially important for reasoning about OCap patterns. Miller first proposed the concepts of permission and authority in his PhD thesis\cite{miller2006}, following which the formal definitions of these concepts were proposed and refined in the work of Drossopoulou et al. \cite{drossopoulou2016}. Our paper defines permission and authority in a formal way that is different from the aforementioned works, but we generally adopt the same informal meanings (that having permission means the right to invoke an object's behaviours, and authority to mean the ability to cause effects).

\subsubsection{Permission - MayAccess Definition}\label{MayAccess}
We define in total four flavours of MayAccess predicates that are inspired by the work in Drossopoulou et al.\cite{drossopoulou2015b}, but our four MayAccess predicates extend and modify the definition in \cite{drossopoulou2015b} to more precisely capture the different configurations of access in terms of mode and time: 
\begin{itemize}
\item Mode: directly ($Dir$) or indirectly ($Ind$)
\item Time: now ($Now$) or eventually ($Eve$)
\end{itemize}
and are general enough to describe generic object-oriented systems:

\begin{logic}
\hr\\
\emp{Definition---[MayAccess]}\\

M,\losigma\ \loturns\ \=MayAccess$^{Dir,Now}$(x,e) \loiff \\
\>\loexists f. \loexec{x.f}{\losigma} = \loexec{e}{\losigma}
\loor \\
\> \loexec{\texttt{this}}{\losigma} = \loexec{x}{\losigma} \loand\ \loexists y. \loexec{y}{\losigma} = \loexec{e}{\losigma})\\
\\
 M,\losigma\ \loturns\ \=MayAccess$^{Dir,Eve}$(x,e) \loiff \\
\>\loexists \losigma$'$ \loin Arising(M,\losigma).\\
\>M,\losigma$'$\ \loturns\ MayAccess$^{Dir,Now}$(x,e)\\
\\
M,\losigma\ \loturns\ \=MayAccess$^{Ind,Now}$(x,e) \loiff \\
\>\loexists \loconj{f}. \loexec{x.\loconj{f}}{\losigma} = \loexec{e}{\losigma}
\loor \\
\> \loexec{\texttt{this}}{\losigma} = \loexec{x}{\losigma} \loand\ \loexists y. \loexec{y.\loconj{f}}{\losigma} = \loexec{e}{\losigma})\\
\\
M,\losigma\ \loturns\ \=MayAccess$^{Ind,Eve}$(x,e) \loiff \\
\>\loexists \losigma$'$ \loin Arising(M,\losigma).\\
\>M,\losigma$'$\ \loturns\ MayAccess$^{Ind,Now}$(x,e)\\
\hr
\end{logic}

where, we say MayAccess$^{Dir,Now}$(x,e) holds \lotiff\ in some current state \losigma, we have a field in x that points to e, or that we have x as a receiver in \losigma\ and there exists some y in the same \losigma\ that points to e. We also say MayAccess$^{Ind,Now}$(x,e) holds \lotiff\ in some current state \losigma, we have a \textit{series of fields} from x that leads to e, or that we have x as a receiver in \losigma\ and there exists some y with a series of fields in the same \losigma\ that leads to e. Lastly, the 'eventual' definitions of the two predicates are then defined as there existing an arising configuration \losigma$'$ from \losigma, where the 'now' definitions hold at \losigma$'$.\\

Note that MayAccess$^{\_,Now}$ holds imply MayAccess$^{\_,Eve}$ holds, since an eventual state \losigma$'$ that arises from \losigma\, can refer to \losigma, and therefore, the 'now' definitions implies and are stronger than the 'eventual' definitions. Also since a series of fields \loconj{f}\ can mean a singular field f, MayAccess$^{Dir,\_}$ holds imply MayAccess$^{Ind,\_}$ holds, and therefore the 'direct' definitions are stronger than the 'indirect' definitions. \\
\begin{table}[htb]
\caption{MayAccess Relations}\label{tab:accessrelations}
\centering
\begin{tabular*}{0.5\linewidth}{c|ccc}\toprule
& \bf Now & & \bf Eventually\\
\hline
\multirow{5}{*}{\rot{\bf Indirect \enspace Direct \:}} & \multirow{2}{*}{MayAccess$^{Dir,Now}$} & \loimplies & \multirow{2}{*}{MayAccess$^{Dir, Eve}$} \\
& & \lonimpliedby &  \\
& \rot{\loimpliedby} \rot{\lonimplies}& &\rot{\loimpliedby} \rot{\lonimplies} \\
& \multirow{2}{*}{MayAccess$^{Ind,Now}$} & \loimplies & \multirow{2}{*}{MayAccess$^{Ind, Eve}$} \\
& & \lonimpliedby &\\
\end{tabular*}
\end{table}\\
We surmarise the relationships between the four flavours of MayAccess in \cref{tab:accessrelations}. Note that in an object-to-object paradigm, all four flavours of MayAccess involve object capabilities. MayAccess$^{Dir,\_}$(o,o$'$) says object o has the direct reference of o$'$, while MayAccess$^{Ind,\_}$ says object o has a path of reference(s) that lead to o$'$ (that is, o has the reference of some object x$_1$, and object x$_1$ has the reference of object x$_2$ $\dots$ and object x$_n$ has the reference of object o$'$). The 'now' and 'eventual' flavours simply state that the "direct" and "indirect" predicates either they hold at \losigma\ (Now) or they can hold at some arising eventual state \losigma' from \losigma\ (eventual).

\subsubsection{Authority - MayCall Definition}\label{MayCall}
With our access predicates in place, we define a MayCall predicate that describe what it means for x to be able to call e. Note that our definition of MayCall is different from the MayAffect predicate in \cite{drossopoulou2016}. Our MayCall predicate captures the meaning of an object o having authority over another object o$'$ without o necessarily modifying the fields of o$'$. In that sense, our MayCall predicate is weaker than the MayAffect predicate in \cite{drossopoulou2016}:\\
\begin{logic}
\hr\\
\emp{Definition---[MayCall]}\\
M,\losigma\ \loturns\ \=MayCall(x,e) \loiff \\
\> \loexec{\texttt{this}}{\losigma} = \loexec{x}{\losigma} \loand\ \loexists \losigma$'$\loin Arising(M,\losigma). \loand\ \loexec{\texttt{this}}{\losigma$'$} = \loexec{e}{\losigma$'$}\\
\hr
\end{logic}

where we say MayCall(x,e) holds \lotiff\ we have x as a receiver in some state \losigma, and in some arising state \losigma$'$ from \losigma, we have e as a receiver in \losigma$'$. This means that there must exist some method in x that upon invocation, will result in some state \losigma$'$ where e is the receiver in the state \losigma$'$.\\

\begin{logic}
\hr\\
\emp{Definition---\emp{[MayAffect from Drossopoulou et al.\cite{drossopoulou2016}]}}\\
M,\losigma\ \loturns\ \=MayAffect(x,e) if there exists a method \sf{m}, argument(s) \loconj{a}, state \losigma', such that:\\
\ablock M,\losigma\ \=e.m(\loconj{a}) \loleadsto\ \losigma', and\\
\>\loexec{e}{M,\losigma} \loneq\ \loexec{e$'$}{M,\losigma$'$}\\
\hr
\end{logic}

\textbf{Motivation for MayCall:} The motivation for defining authority as MayCall in our paper is that we would like to describe a \textit{chain} of authority between objects, without needing objects in such a chain to necessarily modify the fields of other objects in the chain. In other words, when object A wants to use object B to modify a field in C but can only do so through object B, we want to express that we can deny A from modifying C either by denying A from calling B, or deny B from calling C. Note that object A does not need to be able to modify the fields of B to eventually modify the state of C. Hence, the MayAffect predicate used in \cite{drossopoulou2016} is too strong to decompose a long chain of authority into smaller object-to-object authorities.\\

\begin{logic}
\hr\\
\emp{Lemma---[Object as receiver requires method invocation]}\\
M,\losigma\ \loturns\ \=\losigma(this) = \loexec{o}{\losigma}\loimplies \\
\> \loexists \losigma$'$,m,\loconj{a}. [ M,\losigma$'$, o.m(\loconj{a}) \loleadsto\ \losigma\ ]\\
\hr
\end{logic}

We make clear that for an object to be a receiver in a particular state \losigma, we require a method to exist in the object in a prior state \losigma$'$ that upon invocation can lead to state \losigma. This follows from our definition of execution.\\

Our MayCall(o,o$'$) predicate however, is not concerned whether or not the state o$'$ is modified ---the fact that a method in o$'$ can be eventually invoked by o is good enough grounds for us to mean that o has authority of o$'$. Our justification is that in the object-oriented paradigm, programmers almost always enforce security of a sensitive field in a protected object to be declared private, and define some public method in the object that can modify the concerned private field based on some condition. Hence, the question of whether an \textit{untrusted} object can modify a private field in an object \sf{o$'$} can be expressed as a question of whether the untrusted object can invoke a method of \sf{o$'$}. Given our definitions of MayCall(o,o$'$), it follows that a necessary condition for modification of a private field in o$'$ is the invocation of a method of o$'$. Proving the falsity of our MayCall(o,o$'$) predicate is therefore sufficient to prove that for a field \sf{f} in o$'$, MayAffect(o,o$'$.f) does not hold.\\
\begin{logic}
\hr\\
\emp{Lemma---[Field Modification Requires Authority]}\\
M,\losigma\ \loturns\ \loforall \losigma$'$,f. [\=\losigma$'$\loin Arising(M,\losigma) \loand\ \loexec{o$'$.f}{\losigma$'$}\loneq\loexec{o$'$.f}{\losigma}\loimplies \\
\>\loexists o. M,\losigma\ \loturns\ MayCall(o,o$'$) ]\\
\hr
\end{logic}

\subsection{Domination}\label{sec:domination}
Here, we introduce the concept of domination that is inspired by the work on ownership types by Clarke et al.\cite{clarke1998}, where they define that the owner o of an object o$'$ dominates o$'$ in the object graph such that for any path p from some arbitrary object o$'$' to o$'$, either o$'$' belongs to the set of owners or that p passes through o.\\

We say that a set S dominates an object x \lotiff\ for every object y in the system that has a path to x, y must have a path to some arbitrary object o that is a member of the set S. The power of having such a predicate will be demonstrated in \cref{sec:ocappatterns}, where we show how attenuating objects that is meant to protect an object o can ensure certain properties of members in a dominating set over s under any unknown code execution.\\

\begin{logic}
\hr\\
\emp{Definition---[Domination]}\\
M,\losigma\ \loturns\ \=Dom(S,x) \loiff \\
\> \loforall y.\loconj{f},n. [ \= x\loneq y \loand\ \loexec{y.f$_1$...f$_n$}{\losigma} = \loexec{x}{\losigma} \loimplies\\
\>\> \loexists k,o. \loexec{y.f$_1$...f$_k$}{\losigma} = \loexec{o}{\losigma} \loand\ o\loin S ]\\
\hr
\end{logic}
The usefulness of domination will be made more clear in the next section when we argue about cooperation without vulnerability.
\section{Reasoning in an OCap System}\label{sec:ocapspecs}

In this section, we formally describe the ideas of \textit{isolation} and \textit{cooperation}, and how \textit{cooperation} can either be vulnerable or protected, within the OCap system. We use the word \textit{capability} (of an object), interchangeably with reference (of an object), to represent that both object reference and the right to use the object reference are indistinguishable in OCap; i.e., there is no global ambient authority to prevent an object from using another object reference in general. Given our definition of \textit{permission}, it follows that o has the permission of o$'$ if o has the capability of o'. However, if we were to use \textit{our} formal definition of \textit{authority} (causing effects where effects are method invocations), we cannot say that an object reference embodies authority - we elaborate more in \cref{sec:isolation and cooperation}. We use the word \textit{path} interchangeably with indirect capability or permission, to emphasise that an object o need not necessarily require the direct capability of o$'$ to invoke the behaviours of o$'$---invocation of o$'$ can be done by o through a network of paths consisting of other object capabilities.\\

In \cref{sec:epc} we first explain that objects require permission (direct capabilities or a chain of capabilities) to communicate, i.e. permission forms the necessary condition for authority. We also propose a lemma that states that under an OCap system, we can reason about eventual indirect permission, from a current permission configuration of an initial closed system\footnote{We require a closed system because unknown object or code might introduce capabilities of objects in the system that the unknown object or code might already have.}. That is, given a system where we know the entire system's present reference graph, OCap rules confine the possible reference graphs that can be born eventually from the present reference graph. Next in \cref{sec:isolation and cooperation}, we reason about \textit{isolation} and \textit{cooperation} using our formal definition of permission. Following which in \cref{sec:vulnerability and protection}, we expand the concept of \textit{cooperation} to include cooperation that is vulnerable, and cooperation that is protected, i.e. \textit{vulnerability} and \textit{protection}. \textit{Protection} and \textit{vulnerability} are built on our formal definitions of authority and domination.

\subsection{Eventual permission from present permission configuration}\label{sec:epc}
Here, we propose a lemma \textsc{epc} that states how eventual permission can be born from a constellation of direct permissions within an OCap system. We use paths interchangeably with direct and indirect permission, to emphasise that an object o need not necessarily have the direct capability of o$'$ to invoke the behaviours of o$'$, invocation can be done through a network of paths consisting of permissions.

\begin{logic}
\hr\\
\emp{Lemma---[OCap Eventual Paths from Current Paths (EPC)]}\\
M,\losigma\ \loturns\ \=MayAccess$^{Ind,Eve}$(o,o$'$)\\
\> \loimplies\\
\> \loexists x. \=$[ ($\=MayAccess$^{Dir,Now}$(o,x)\\
\>\>\> \loor \\
\>\>\> MayAccess$^{Dir,Now}$(x,o)$)$\\
\>\>\loand \\
\>\>MayAccess$^{Ind,Eve}$(x,o$'$)$ ]$\\
\hr
\end{logic}

\textsc{Lemma epc} has the meaning that if an object o has an eventual path to o$'$ in some arising state \losigma$'$, then there must exist an object x that has an eventual path to o$'$, and there must exist a way for o to have a path to x in \losigma$'$. Object o either must already have a path to x in \losigma, or that it must be able to receive the capability of x through introduction by x in \losigma. \textsc{epc} is actually a formal representation of the well-known OCap idiom of \textit{connectivity begets connectivity}:

\begin{itemize}
\item \textbf{Initial Conditions or Endowment:} o has the capability of o$'$ in \losigma\ due to initiate conditions or endowment in the system, therefore x refers to o$'$
\item \textbf{Parenthood:} if o can create o$'$ in some arising \losigma$'$, then o can also create o$'$ in \losigma, therefore x refers to o$'$
\item \textbf{Introduction:} o will only obtain the path to o$'$ in some arising \losigma$'$ through another object introducing o a path to o$'$, therefore x refers to some object that is not o (x\loneq o). 
\end{itemize}

Note that for o to have an eventual path to o$'$, we only require now that o have a direct capability to some arbitrary object x, or that x has a direct capability of o so that x can introduce itself to o. If o already has the capability of x in \losigma\ then we know o can reach x in \losigma$'$ by definition. If not, the capability of x must be introduced to o. For x to introduce itself, x must have the capability of o in \losigma.\\

Well, what if the capability of x is introduced by some \textit{other} object x'? Note that in such a case, x' must have the capability of object x (for x' to even introduce x to o in the first place), and therefore x' can also have an eventual path to o$'$. Also x' must also be able to introduce itself to o. With these observations, we note that there is hence no logical difference between x' and x in our formal description and x' might simply be referred to as x.\\

There is one final critical result from \textsc{Lemma epc}.
Notice how, there is a 'recursive'\linebreak MayAccess$^{Ind,Eve}$(x,o$'$) in our condition for MayAccess$^{Ind,Eve}$(o,o$'$).
This allows us to recursively expand the condition to incorporate \textit{all} x intermediate objects in the path leading to o$'$. Repeated recursive expansions will eventually give us conditions that are only defined in MayAccess$^{Dir,Now}$ predicates, where the terminating MayAccess$^{Ind,Eve}$(o$'$,o$'$) can be determined to be true or false based on whether o$'$ exists in some arising state. This result allows us to define an eventual path between two objects based purely on a present configuration of objects on the reference graph in \losigma.\\

In the example, we show an example using \textsc{Lemma epc} with 3 objects o, y, o$'$. Let us assume o, y, and o$'$ always exists, such that the terminating recursive predicates would return true. This example then illustrates an eventual path from o to o$'$ can only exist if one of these present 6 configurations in \losigma\ holds (where x refers to the intermediate object in \textsc{epc}):
\begin{tabbing}
1) o has the capability of y, and \=y has an eventual path to o$'$---\\
\>by y having the direct capability of o$'$\\
2) o has the capability of y, and y has an eventual path to o$'$---\\
\>by o$'$ having the direct capability of y\\
3) y has the capability of o, and y has an eventual path to o$'$---\\
\>by y having the direct capability of o$'$\\
4) y has the capability of o, and y has an eventual path to o$'$---\\
\>by o$'$ having the direct capability of y\\
5) o has the direct capability of o$'$\\
6) o$'$ has the direct capability of o\\
\end{tabbing}
Note how, these 6 configurations are direct capability configurations at \losigma, but allows us to reason about whether a potential path from o to o$'$ can exist in some arising state \losigma$'$ from \losigma.\\

\subsection{Isolation and Cooperation}\label{sec:isolation and cooperation}
In this section, we reason about isolation and cooperation using our formal definition of permission and authority. What exactly is object isolation when we say that object o$'$ is isolated from object o, and what exactly is cooperation?\\

We say isolation is simply that there is no eventual directed path from o to o$'$:\\

\begin{logic}
\emp{Isolation of o$'$ from o:}\\
\loturns\ \loneg MayAccess$^{Ind,Eve}$(o,o$'$)\\
\end{logic}

Given \textsc{lemma epc}, we can enforce isolation of o$'$ from o given the present path configuration of a system at \losigma, and when the closed system is eventually opened such that an unknown object has permission and authority of \textit{only} object o in some arising state \losigma', we know that the unknown code will not have permission of o$'$, since o is isolated from o$'$.\\

When we have isolation, we know from contraposition of the corollary from our definitions of permission and authority:\\

\begin{logic}
\hr\\
\emp{Corollary---[Permission(Indirect,Eventual) is necessary for Authority]}\\
M,\losigma\ \loturns\
\= MayCall(o,o$'$) \loimplies \\
\> MayAccess$^{Ind,Eve}$(o,o$'$)\\
\hr
\end{logic}

Therefore,\\

\begin{logic}
\emp{Isolation implies no authority}\\
\loturns\ \loneg MayAccess$^{Ind,Eve}$(o,o$'$) \loimplies \loneg Maycall(o,o$'$)\\
\end{logic}

It is good to know we can guarantee the internal state and behaviour of an object o$'$ in the presence of unknown code on object o through isolation, but in practice this is not entirely useful. The quintessential question in OCap systems is whether we can allow cooperation without vulnerability. What then is cooperation?\\

We say cooperation between two objects o and o$'$ is simply that o has permission of o$'$:\\

\begin{logic}
\emp{Cooperation between o and o$'$ (o $\rightarrow$ o$'$):}\\
\loturns\ MayAccess$^{Ind,Eve}$(o,o$'$)\\
\end{logic}

\begin{logic}
\emp{Cooperation is necessary but not sufficient for authority:}\\
\loturns\ MayAccess$^{Ind,Eve}$(o,o$'$) \lonimplies\ MayCall(o,o$'$)\\
\end{logic}

We provide some analogies as to why cooperation does not imply authority. Object o can have the reference(capability) of o$'$, and hence there is cooperation, but o contains no methods that can invoke(use) the capability of o$'$, and thus o has the capability of o$'$ but not the authority of o$'$. Another analogy, from the other viewpoint of o$'$, would be that o$'$ has no available methods that can be invoked by o ---this can happen if o has no methods, or o only contains private methods or methods that are restricted for objects inherited from the same class as o of which o$'$ is not.\\

The problem with cooperation is when we know o$'$ has exposed methods (or public methods), and when object o is of unknown provenance. Since we assume object o is unknown, then we must assume that given the permission of o$'$, o will contain some method that will try to invoke the behaviours of o$'$. This then leads us to our next section on vulnerability and protection in the context of cooperation.

\subsection{Vulnerability and Protection}\label{sec:vulnerability and protection}
\begin{itemize}\item We assume that object o$'$ will always have some exposed public method for the rest of this section, and is the object of interest that we want to protect.\end{itemize}
For the concepts of vulnerability and protection, our arguments are based primarily on whether the source object is trusted (known) or untrusted (unknown).\\

While we can always protect o$'$ through isolation from o (o will have no paths to o$'$), what is probably more useful is whether cooperation between o$'$ and o, is possible in a \textit{protected way}. What does that mean? Notice that if we \textit{know} or \textit{trust} the code of o, then \textbf{cooperation can actually be safe}, even if o contains methods that can eventually invoke behaviours of o$'$.\\

\begin{logic}
\hr\\
\emp{Lemma---[Known-Safe cooperation]}\\
o:KnownObj \loand\ o$'$:ExposedObj \loand\ MayAccess$^{Ind,Eve}$(o,o$'$)\\
\hr
\end{logic}

Then regardless of whether MayCall(o,o$'$) holds or not, since we know or trust the code of o, we know that o will not do something that will lead to undesirable method invocations of o$'$.\\

However, if we have another object o* that is of unknown provenance such that:\\

\begin{logic}
\hr\\
\emp{Lemma---[Unknown-Vulnerable cooperation]}\\
o*:UnknownObj \loand\ o$'$:ExposedObj \loand\ MayAccess$^{Ind,Eve}$(o*,o$'$)\\
\hr
\end{logic}

then we must always be prepared for the worst in terms of what o* will do, and cooperation becomes vulnerable if o* can invoke behaviours of o such that MayCall(o*,o$'$) holds.\\

So, this then leads us to the most important question, how do we enable safe cooperation in the face of unknown code?\\

\subsubsection{Attenuating objects as trusted messengers}\label{sec:atten}
Notice how we already have a result about a Known-Safe cooperation. In this section, we show how can make use of it, and four corollaries from our formal specifications, including the concept of domination that we have proposed.\\

\begin{logic}
\hr\\
\emp{Corollary---[Indirect permission is made of direct permissions]}\\
M,\losigma\ \loturns\ [ \=MayAccess$^{Ind,Now}$(o,o$'$)\\
\>\loiff\\
\>\loexists x. [ \=MayAccess$^{Ind,Now}$(o,x) \loand\\
\>\> MayAccess$^{Dir,Now}$(x,o$'$) ] ]\\
\hr
\end{logic}

\begin{logic}
\hr\\
\emp{Corollary---[Authority can be a chain of smaller authorities]}\\
M,\losigma\ \loturns\ \=MayCall(o,o$'$)\\
\> \loimplies\\
\> \loexists x. $[$\=MayCall(o,x) \loand\ MayCall(x,o$'$)$]$\\
\hr
\end{logic}

The above two corollaries tell us that we do not need to give a direct permission or capability of o$'$ to the unknown object o* for cooperation to occur between the two objects. Cooperation between o* and o$'$ is possible with \textit{intermediate} objects.

\begin{logic}
\hr\\
\emp{Corollary---[Domination of Permission]}\\
M,\losigma\ \loturns\ \=MayAccess$^{Ind,Now}$(o,o$'$) \loand\ Dom(S,o$'$)\\
\>\loimplies\\
\>\loexists x\loin{S}. $[$\=MayAccess$^{Ind,Now}$(o,x) \loand\ MayAccess$^{Ind,Now}$(x,o$'$)$]$\\
\hr
\end{logic}

\begin{logic}
\hr\\
\emp{Corollary---[Domination of Authority]}\\
M,\losigma\ \loturns\ \=MayCall(o,o$'$) \loand\ Dom(S,o$'$)\\
\> \loimplies\\
\> \loexists x\loin{S}. $[$\=MayCall(o,x) \loand\ MayCall(x,o$'$)$]$\\
\hr
\end{logic}

The above two corollaries on domination tell us that if we know a set S that dominates our object of interest o$'$, then an unknown object o* \textit{must} use a member of this dominating set in the path of permissions to reach o$'$ and in the chain of authority to invoke o$'$.\\

With these four corollaries, and from knowing that cooperation between trusted code is safe, we have:\\

\begin{logic}
\hr\\
\emp{Lemma---[Unknown-Protected cooperation]}\\
o*:UnknownObj \loand\ o$'$:ExposedObj \loand\ MayAccess$^{Ind,Eve}$(o*,o$'$)\\
\loand\ Dom(S,o$'$) \loand\ \loforall s. [ s\loin S \loimplies s:KnownObj ]\\
\hr
\end{logic}

where objects s are known as attenuating objects in the OCap literature. To see why Unknown-Protected cooperation is safe, see how it follows that:

\begin{logic}
\hr\\
\emp{Lemma---[Unknown-Protected cooperation implies Known-Safe cooperation]}\\
o*:UnknownObj \loand\ o$'$:ExposedObj \loand\ MayAccess$^{Ind,Eve}$(o*,o$'$)\\
\loand\ Dom(S,o$'$) \loand\ \loforall s. [ s\loin S \loimplies s:KnownObj ]\\
\loimplies\
\loexists s$'$:KnownObj. [ MayAccess$^{Ind,Eve}$(o*,s$'$) \loand\ MayAccess$^{Ind,Eve}$(s$'$,o$'$) ]\\
\hr
\end{logic}

This says that if we have a configuration where object o$'$ is dominated by known objects, then the permission to use o$'$ by o* must involve at least a known object. This shows that we can use attenuating objects to mitigate vulnerability between unknown code and known code in OCap systems.\\

Therefore, this shows that we can always protect some object of interest by:

\begin{itemize}
\item being careful about our present path configurations so that we can ensure eventual path-isolation between the untrusted object and the object of interest
\item ensuring that trusted objects dominate the object of interest and attenuates the permission and authority of the untrusted object \end{itemize}

Therein lies also the difference between OCap and non-OCap models, because non-OCap models that allow forging of capabilities or a global ambient authority \textit{cannot} enforce security through pure path isolation, which is a distinctive feature of OCap models.\\

Protection strategies for attenuating objects can be implemented using defensive programming techniques like encapsulation and checking of conditions of method arguments. At the extreme ends, attenuating objects can take the shapes of either deciding to forward messages from the unknown object o* in an intact way, thereby allowing MayCall(o*,o$'$) to hold; or if some condition fails, the attenuating object can choose not to forward the message at all such that MayCall(o*,o$'$) does not hold. Somewhere in the middle, attenuating objects can also choose to modify the message from o* such that it becomes safe to forward, where MayCall(o*,o$'$) holds but since it has been attenuated by the attenuating object, we would know it is safe. What and how the attenuating objects decide to do with messages from unknown objects before forwarding them can be best expressed as OCap policies for the attenuating object.\\

In the OCap pattern examples we will go through in the next section, protection of an object is accomplished through some attenuating object that will never leak the capability of the object of interest. A safe configuration is also expressed as attenuating objects dominating the object of interest in such a way that unknown code is forced to use them to cooperate.

\newpage
\section{OCap Patterns}\label{sec:ocappatterns}

An OCap pattern is a a set of objects connected to each other by capabilities in such a way that some effects can only happen under certain conditions. Objects interact with each other by sending messages on capabilities. An OCap pattern may be visualised as a directed graph---nodes represent objects, and each edge from an object o to another o$'$ represents o holding a capability that allows it to directly access o$'$. Indeed, in \cref{sec:figdom}, \cref{sec:figcaretaker}, and \cref{sec:figmem}, we represent such diagrams for three important OCap patterns in the literature. These three OCap patterns are: the DOM Tree pattern, the Caretaker pattern, and the Membrane pattern. In addition to the visual diagrams of these patterns, we also show the code of these three patterns in the capability-safe language Pony\cite{clebsch2015}. To make use of our formal specifications, we focus exclusively on the DOM Tree pattern, where we show how we can use our formal specifications to write policies that can guarantee the preservation of properties of specific nodes in the pattern.

\subsection{The DOM Tree Pattern}

According to the World Wide Web Consortium (W3C)\footnote{https://www.w3.org/DOM/}, the Document Object Model (DOM) is a platform- and language-neutral interface that allow programs and scripts to dynamically access and update the content, structure and style of documents. A DOM Tree describes a model where objects in a system are arranged according to some hierarchical configuration in the structure of a tree. In this pattern, permission and authority of nodes can be reasoned according to whether one object can traverse up or down the tree to access or call other nodes in the tree. This type of reasoning is useful when we need to protect the integrity of more sensitive objects higher up in the hierarchy of the tree, while at the same time, we need to allow untrusted external objects the ability to affect less-sensitive objects lower in the hierarchy of the tree. Our contribution in our paper is that we implement the pattern and give the code for the pattern classes \sf{Nodes} and \sf{ReNodes} in the capability-safe language Pony. Another contribution of ours is that provide the formal specifications of such a pattern using our methodology, which then allows us to assert conditions when properties of a Node in the tree are modified. \\

We first illustrate the vulnerability of a \sf{Node} object in the Tree in \cref{domnode}. Subsequently, we show how we can create an attenuating object \sf{ReNode} that protects access to a \sf{Node}, in \cref{domrenode}. In \cref{domexample}, we show the code of a simulation where we have a JavaScript web document (simulated in Pony) where we need to give some arbitrary 3rd party advertisement company external access to a particular node called \sf{AdNode} in the document (so that the website is able to run advertisements and make money). In the simulation, we show how using an attenuating restricted node \sf{RAdNode} with a configuration of depth = 0, is able to protect all other nodes higher in the hierarchy of the DOM Tree, while \textit{still} allowing the 3rd-party advertiser the ability to modify properties of the \sf{AdNode}. We then show a graphical illustration of \sf{AdNode} and \sf{RAdNode} simulation \cref{fig:figdom}. Following which, we write the object-capability policies of the DOM Tree pattern using our methodology, in \cref{sec:dompolicies}. Lastly, we show how we can begin to reason the protection of a node using those policies.

\subsubsection{Vulnerability of Node in Tree}\label{domnode}
We call objects in the hierarchical tree to be a \sf{Node}, and by default, nodes are vulnerable because they have unprotected methods that can access other nodes in the tree. In particular, the methods \sf{parent()} and \sf{getChild(id:String)} return the object references of other nodes in the tree.
\begin{lstlisting}
class Node
    let name: String
    let _parent: (Node ref|None)
    let _children: collections.Map[String val, Node ref] = _children.create()
    let _props: collections.Map[String val, String val] = _props.create()

    new ref create(name':String, parent': Node ref)=>
        name = name'; _parent = parent'
    new ref createRoot(name':String)=>
        name = name'; _parent = None 
    fun ref getChild(id:String val): Node ref?=>
        try _children(id) else error end
    fun ref parent(): Node ref?=>
        match _parent
        |let x: Node ref => x 
        else error end
    fun box getProp(id:String val):String val ?=>
        try _props(id) else error end
    fun ref setProp(id:String val,prop:String val) =>
        _props(id)=prop
    fun ref addChild(id:String val):Node?=>
        _children(id) = Node.create((name+"-child-"+id.string()), this)
        _children(id)
    fun ref delChild(id:String val) ?=>
        try _children.remove(id) else error end
\end{lstlisting}

Internal state modification of a node is possible because the class \sf{Node} contains a public method \sf{setProp(key,value)} that can be called by anyone holding a node's capability. The method will modify the internal state properties of the node (by creating or modifying a key-value pair in a map within the node). Consequently, this means that having the direct capability of a node implies having the authority to modify the node's properties. Also, since having a direct capability of a node can eventually translate to having the direct capability and authority over all other nodes in the tree, this means that an untrusted object \sf{o} that obtains access to a node \sf{n} can eventually obtain authority over \textit{all} other nodes \sf{n'} in the same tree:\\

\begin{logic}
\loforall o:Object, \loforall n,n':Node. [ \=MayAccess$^{Dir,Now}$(o,n) \loand\ MayAccess$^{Ind,Now}$(n,n')\\
\> \loimplies\\
\> MayCall(o,n') ]\\
\end{logic}

In the OCap literature, letting an object \sf{o} have access to an object \sf{o$'$} (\sf{o} having the reference of \sf{o$'$}) \textit{usually} implies we want to let object \sf{o} use or call methods on \sf{o$'$}. More precisely, in such a configuration, we are typically prepared to let o have the ability to modify the state of o$'$. However, there are situations where object \sf{o$'$} has not just the ability to modify its own internal state, but in addition, the ability to call methods  and modify the state of other objects \sf{o$''$}. This then becomes problematic, because while we are typically prepared to let \sf{o} modify the state of \sf{o$'$}, we typically are \textbf{not} prepared to give \sf{o} the authority to modify the state of all other objects \sf{o$''$} that are reachable from \sf{o$'$}.\\

Nodes are problematic precisely because of the same reasons described in the general case above, because while we are typically prepared to give some unknown and untrusted object the authority to modify the state of a particular node \sf{n}, we are \textbf{not} prepared to give the untrusted object the authority to modify all other nodes reachable from \sf{n}. Notice the implication MayCall(o,n$'$) where n' refers to all other nodes in the tree, such that the untrusted object can now call all other nodes in the tree, which is an extremely undesirable outcome in terms of security. Therefore, we need to find ways where we can give to an untrusted object \sf{o} access and authority to a node \sf{n} so that \sf{o} can modify or call methods on node \sf{n}, \textit{but} at the same time, we do \textit{not} compromise the properties or state of other nodes in the tree.

\subsubsection{Protection using Restricted Node}\label{domrenode}

In the OCap literature, the use of attenuating objects help to mediate authority of a protected object, and help to remedy the security problem we have just described above. In the DOM Tree pattern, we make use of an attenuating restricted node class \sf{ReNode} whose objects serve to attenuate the authority of objects belonging to class \sf{Node}. A restricted node \sf{rn} of class \sf{ReNode} attenuates the capability of a node of class \sf{Node} in the following ways:

\begin{itemize}
\item A restricted node \sf{rn} of class \sf{ReNode} contains a field which points to a node \sf{n} of class \sf{Node} that \sf{rn} is meant to attenuate, and contains a depth field \sf{d} of type \sf{integer}. 
\item  It protects \sf{n} such that with an untrusted object having only the capability of the restricted node \sf{rn}, \sf{rn} \textbf{only allows} the properties of \sf{n}, all of \sf{n}'s descended child nodes in the tree, and all of \sf{n}'s ancestor parent nodes up to depth \sf{d} in the tree, to be modified.
\end{itemize}

The restricted node accomplishes this primarily by specifying the conditions on which method calls are allowed to be \textit{forwarded} to \sf{n}, and specifying how the results of those method calls to \sf{n} should be \textit{processed}. Specifically, the sensitive \sf{parent()} method in \sf{n} that returns the direct capability of the parent node of \sf{n}, is now protected by the restricted node \sf{rn}. By calling \sf{parent()} method on \sf{rn}, \sf{rn} will only \sf{forward} the parent method call to \sf{n} \textbf{only} when the condition that \sf{rn}'s depth level stored in the integer variable \sf{d} must have a value of at least more than 0. Even if \sf{rn} has a depth level more than 0, the restricted node does not return the direct capability of the parent node of \sf{n} to the method caller on \sf{rn}, but instead \textit{processes} the result to the method caller on \sf{rn}, by only returning a new restricted node object \sf{rn'} with a field that points to the parent node of \sf{n}, where \sf{rn'} inherits the depth level of \sf{rn} decremented by 1. Lastly, the restricted node does not have any methods that return the direct capability of the node that it protects and all the node's children. Method calls done by the restricted node on the underlying node that return the capabilities of the child nodes are always processed by the restricted node such that the restricted node only returns newly created restricted nodes of those child nodes configured with a depth level incremented by 1.

\begin{lstlisting}
class ReNode&\footnote{In the Pony language, a field is declared private with a leading underscore "\_", so we know that the capability of the protected node stored in a restricted node cannot be read directly by external objects using the restricted node, and can only do so through some method. Also, a field is declared using "let" to restrict the field from being modified. Therefore, we know that the node that a restricted node protects cannot change.}&
    let \_node: Node ref
    let _d: U32
    new ref create(node':Node ref, d':U32)=>
        _node = node'; _d = d'
    fun ref getChild(id:String val):ReNode ref?=>
        try ReNode.create(_node.getChild(id), _d+1) else error end
    fun ref parent():ReNode ref?=>
        if \_d > 0 then ReNode.create(\_node.parent(), \_d-1) else error end
    fun box getProp(id:String val):String val ?=>
        try _node.getProp(id) else error end
    fun ref setProp(id:String val,prop:String val)=>
        _node.setProp(id, prop)
    fun ref addChild(id:String val):ReNode ref?=>
        try ReNode.create(_node.addChild(id), _d+1) else error end
    fun ref delChild(id:String val)?=>
        try _node.delChild(id) else error end
\end{lstlisting}

\subsubsection{DOM Tree Example}\label{domexample}
We will now demonstrate how \sf{ReNode} attenuates authority over \sf{Node}. We assume we only have two nodes in the tree, where the root node \sf{n$'$} has a child node \sf{n}. In this example, we want to allow an untrusted object \sf{o} the authority to modify the properties of \sf{n}, but not the root node \sf{n$'$}. To illustrate this, we use a DOM Tree and call the root note \sf{n$'$} a \sf{document} node, and the child node \sf{n} an \sf{adnode}.  Typically, a website that wants to make money from having advertisements will have a particular node (\sf{adnode}) within the DOM Tree, so that a dynamic advertisement object can be rendered properly in the document. Suppose that the website owner is called Alice and she is also required to give an external party (advertisement company) the authority to modify the properties of \sf{adnode} in her HTML document. Obviously, Alice will not want the external party to modify the properties of the other nodes in her document---she would want to only allow the external party the authority to modify the \sf{adnode} and its descended nodes. Alice therefore needs a way to protect the other nodes in the DOM Tree from the external party, but still allow the untrusted external party the authority to modify \sf{adnode}.\\
\begin{lstlisting}
actor Main
    let env: Env
    let main: Main tag
    new create(env':Env)=>
        env = env'
        main = this
        let document: Node = Node.createRoot("Document")
        let access: Bool = initWebPage(document, {(radnode: ReNode ref)=>
                try radnode.parent().setProp("title", "Bob website") end 
                }ref)
        env.out.print(access.string())

fun ref initWebPage(document: Node, ad_lambda:{(ReNode)}):Bool val=>
    try
        document.setProp("title","Alice website")
        let adnode = document.addChild("ad_div")
        let radnode = ReNode(adnode, 0)
        ad_lambda(radnode)
        if document.getProp("title") is "Alice website" then return true
            else return false end
    else return false end
\end{lstlisting}

In the DOM Tree -adnode- code example above, we simulate initialising a webpage with a root \sf{document} node that we set the property of a map-key "title" having a map-value of "Alice website". We then create a node for an advertisement called \sf{adnode}. To allow the external party the authority to modify the properties of \sf{adnode}, we create a restricted node object \sf{radnode} with a depth level of 0 that points to \sf{adnode}. In the simulation, we execute a method \sf{ad\_lambda} that attempts to call code on \sf{radnode}, where in our simulation, \sf{ad\_lambda} attempts to modify the root \sf{Document} node's properties by attempting to change the value of the root node's map-key "title", to "Bob's website". This would fail, because we have constructed \sf{radnode} with a depth level of 0, so we know that \sf{radnode} will not be able to traverse up to the root node of the tree. Therefore, we know the title of the root\sf{document} node remains intact as "Alice website" after execution of \sf{ad\_lambda}.
\begin{minipage}{\textwidth}
\subsubsection{DOM Tree Pattern Illustration}\label{sec:figdom}
Here, we illustrate in more detail the example of the DOM Tree pattern in a visual diagram.\\
\centering
\includegraphics[width=1.05\textwidth]{figures/DOM.pdf}
 \captionof{figure}{DOM Tree Example}\label{fig:figdom}
\end{minipage}
\clearpage

\subsubsection{DOM Tree Policies} \label{sec:dompolicies}
Here, we provide the policies for the DOM Tree, using our methodology and Hoare tuples. The use of ":" indicates the variables of the LHS of ":" belonging to a primitive type or class stated in the RHS of ":". Before we dive into the policies, we define the lookup of the k-th parent of a restricted node as follows:\\

\begin{logic}
Definition ---[ReNode.parent]\\
Given rn,rn$'$:ReNode, k:\lonat, then [ rn.parent\losup{k} = rn$'$ \loimplies \=(k $>$ 1 \loand\ rn.parent.parent\losup{k-1} = rn$'$) \loor\\
\>(k = 1 \loand\ rn.parent = rn$'$) \loor\\
\>(k = 0 \loand\ rn = rn$'$) ] \\
\end{logic}
\\
This definition says that given a superscript \losup{k} placed on a \sf{parent}, it means that the result on the RHS of such a look up is the restricted node's parent on the LHS being called \sf{k} multiple times; if \sf{rn'} refers to the same restricted node as \sf{rn}, then k = 0.
\paragraph{Policies for calling and modifying nodes}
The following two policies describe the condition under which an Object or ReNode may call methods on a node.\\

\begin{logic}
\emp{Policy ---[Object calls Node]}\\
\loforall o:Object, \loforall n:Node. [ \=MayCall(o,n) \loand\ Dom(S,n) \loand\ \loforall s. [ s\loin S \loimplies s:ReNode ]\\
\>\loimplies\\
\>\loexists rn:ReNode. [ rn\loin S \loand\ MayCall(o,rn) \loand\ MayCall(rn,n) ] ]\\
\end{logic}
\\
\begin{logic}
\emp{Policy ---[ReNode calls Node (Necessary Condition)]}\\
\loforall rn:ReNode, \loforall n:Node. [ \=MayCall(rn,n) \loimplies\
\loexists rn$'$:ReNode. [ MayCall(rn,rn$'$) \loand\ rn$'$.node = n ] ]\\
\end{logic}
\\
\begin{logic}
\emp{Policy ---[ReNode calls Node (Sufficient Condition)]}\\
\loforall rn:ReNode, \loforall n:Node. [ \=rn.node = n \loimplies\
MayCall(rn,n) ]\\
\end{logic}

Policy [Object calls Node] says that if we have an object \sf{o} that can call the methods of node \sf{n}, and we have a set \sf{S} that dominates n such that all members in \sf{S} are of class \sf{ReNode}, then it implies that object \sf{o} must be able to call methods of a restricted node \sf{rn} out of \sf{S} and that \sf{rn} must able to call methods of \sf{n}. This just means that the chain of authority from \sf{o} to \sf{n} must involve \sf{rn} in the chain.\\
Policy [ReNode calls Node] says something about the last restricted node \sf{rn$'$} at the end of the chain of authority to n: that a restricted node \sf{rn} can call \sf{n} by calling another restricted node \sf{rn$'$} that points to \sf{n}. That is, a restricted node \sf{rn} that points to a node \sf{n} (\sf{rn.node = n)} \\
\hr
\paragraph{Policy for restricted node calling other restricted nodes}
\begin{logic}
\emp{Policy ---[ReNode calls ReNode (Necessary Condition)]}\\
\loforall rn,rn$'$:ReNode. [ \=MayCall(rn,rn$'$)\\
\>\loimplies\\
\>\loexists k,j:\lonat. [ \=(rn.parent\losup{k} = rn$'$.parent\losup{j}) \loand\ rn.depth \logeq k ] ]\\
\end{logic}\\

Policy [ReNode calls ReNode] specifies a necessary condition and a sufficient condition for which a restricted node \sf{rn} may call another restricted node \sf{rn$'$}. This policy is subtle and says that a restricted node rn is allowed to navigate to all the descendants from all its accessible ancestors. $k$ is a natural number that indexes all the accessible ancestors of rn $k$ levels up the tree from rn, while we use another natural number $j$ that describes all the descendant nodes $j$ levels down the tree from a particular accessible ancestor. If rn$'$ is a direct descendant from rn, then k = 0. If \sf{rn$'$} is a direct ancestor of rn, then j = 0. An ancestor $k$ levels up the tree is only accessible if rn has sufficient depth access (rn \logeq k); there are no restrictions on accessing the descendants $j$ levels down the tree from a particular accessible ancestor.\\

\hr
\paragraph{Policy for reasoning about new paths from ReNodes}
\begin{logic}
\emp{Policy ---[Domination No Leaked Paths from ReNode]}\\
\loforall n:Node, RND$_{old}$. \=[ Dom(RND$_{old}$,n) \loand\ \loforall s. [ s\loin RND$_{old}$\loimplies s:ReNode ]\\
\>\{ \texttt{\textit{code}} \}\\
\loexists RND$_{new}$. [ \=Dom(RND$_{new}$,n) \loand\ \loforall s$'$\loin RND$_{new}$. [ \=s$'$:ReNode \loand\\
\>\loexists s$''$\loin RND$_{old}$, \loexists k,j:\lonat. [ \=(s$'$.parent\losup{k} = s$''$.parent\losup{j})  \loand\ (s$''$.depth \logeq j) \loand\\
\>\>(s$'$.depth = s$''$.depth - j + k) ] ] ] ]\\
\end{logic}

This policy is a very important security policy of the DOM Tree pattern.
The policy confines all possible paths to a node \sf{n} that might arise from running \textit{any} code, if all members in the initial set that dominates n belong to class \sf{ReNode}.
For all restricted nodes in the new dominating set, they must have been descended from some restricted node (\sf{rn$''$.parent\losup{j}}), where (\sf{rn$''$.parent\losup{j}}) is accessible from a restricted node rn$''$ in the old dominating set (rn$''$.depth \logeq j). The policy says that an accessible ancestor rn$''$ from a restricted node rn$'$ in the old dominating set is indexed by a depth level j so that the accesible ancestor would inherit a depth level j higher than rn$''$'s depth level---(rn$''$.parent\losup{j}.depth = rn$''$.depth - j). For a newly created restricted node \sf{rn$'$} in the new dominating set that is descended k levels down from one of these accessible ancestors, it would thus inherit a depth of (rn$''$.depth - j + k). In other words every restricted node in the new dominating set must fall into one of these categories:
\begin{itemize}
\item a newly created restricted node descended k levels down directly from a restricted node in the old dominating set---\textbf{(j = 0, k > 0)}
\item a newly created restricted node descended k levels down from some restricted node that is an accessible ancestor j levels up of a restricted node from the old dominating set---\mbox{\textbf{(j > 0, k > 0)}}.
\item a new created restricted node that is a direct ancestor from a restricted node in the old dominating set---\textbf{(j > 0, k = 0)}
\item a restricted node in the new dominating set is also a member of the old dominating set---\textbf{(j = 0, k = 0)}
\end{itemize}
The argument for the policy is as follows: given a set of initial restricted nodes, we know by definition of the callable methods of a restricted node such that it can only create, store and return the capabilities of newly created restricted nodes, or return a \sf{null} value, or return a \sf{String} value, in \textbf{all} of its methods. The policies for these methods can be found in \cref{renodemethods}. The execution policies in 
\cref{renodemethods} imply that all newly created restricted nodes from existing restricted nodes must obey certain policies (such as inheriting an incremented depth from its creator restricted node if its a child, or inheriting a decremented depth if it is a parent). Therefore, all restricted nodes in the new dominating set, if created during the unknown code execution, must be connected to a creator restricted node in the old dominating set through some path formed from the policies in \cref{renodemethods}, or is already a member of the old dominating set, no matter what kind of code execution. 
\paragraph{Policies for methods of ReNode}\label{renodemethods}
The six methods in ReNode are described in the policies as follows:\\

\begin{logic}
\emp{Policy ---[Execution of ReNode.parent]}\\
\loforall rn:ReNode, \loexists i:String. \loforall s. [ \=(rn.depth > 0)\\
\>\{\texttt{s = rn.parent()}\} \\
\>(s:ReNode) \loand\ (s = rn.parent) \loand\ (s.child(i) = rn) \loand\\
\>(s.node = rn.node.parent) \loand\\
\>(s.depth = rn.depth - 1)\\
\>\loor\\
\>s = \texttt{null} ]\\
\end{logic}\\
This policy says that the pre-condition creating a parent restricted node requires that the creator restricted node must have a depth level of more than 0. The post-conditions are that the newly created parent restricted node would be pointing towards a node that is the parent node of the node that its creator is pointing towards, and the newly created parent restricted node inherit the depth level of its creator restricted node decremented by 1.\\

\begin{logic}
\emp{Policy ---[Execution of ReNode.setProp]}\\
\loforall rn:ReNode, \loforall i,j:String. [ \textit{true} \{\texttt{rn.setProp(i,j)}\} rn.node.prop(i) = j ]\\
\end{logic}

\begin{logic}
\emp{Policy ---[Execution of ReNode.getProp]}\\
\loforall rn:ReNode, \loforall i,j:String. [ \textit{true} \{\texttt{j = rn.getProp(i)}\} j = rn.node.prop(i) ]\\
\end{logic}

\begin{logic}
\emp{Policy ---[Execution of ReNode.addChild]}\\
\loforall rn,rn$'$:ReNode, \loforall i:String. [ \=\textit{true}\\
\>\{\texttt{rn$'$ = rn.addChild(i)}\}\\
\>(rn$'$ = rn.child(i)) \loand\ (rn$'$.parent = rn) \loand\\
\>(rn$'$.node = rn.node.child(i)) \loand\\
\>(rn$'$.depth = rn.depth + 1) ]\\
\end{logic}

\begin{logic}
\emp{Policy ---[Execution of ReNode.getChild]}\\
\loforall rn:ReNode, \loforall i:String, \loforall s. [ \=\textit{true}\\
\>\{\texttt{s = rn.getChild(i)}\}\\
\>(s:ReNode) \loand\ (s = rn.child(i)) \loand\ (s.parent = rn) \loand\\
\>(s.node = rn.node.child(i)) \loand\\
\>(s.depth = rn.depth + 1) ]\\
\>\loor\\
\> s = \texttt{null} ]\\
\end{logic}

\begin{logic}
\emp{Policy ---[Execution of ReNode.delChild]}\\
\loforall rn:ReNode, \loforall i:String. [ \textit{true} \{\texttt{rn.delChild(i)}\} \=rn.child(i) = \texttt{null} \loand\\
\> rn.node.child(i) = \texttt{null} ]\\
\end{logic}

\hr
\paragraph{Policy for Node Property Modification}\label{policy_node_mod}

This policy uses the Lemma [Field Modification requires Authority] and says that if a property of a node \sf{n} is modified after the execution of some code by object \sf{o}, then it implies \sf{o} must be able to invoke a method on \sf{n}.\\

\begin{logic}
\emp{Policy ---[Node Property Modification]}\\
\loforall n:Node, \loforall o:Object. [ [ n.prop(i) = j \{ \texttt{o.\textit{code}} \} n.prop(i) \loneq\ j ] \loimplies\ MayCall(o,n) ]
\end{logic}

\hr
\paragraph{Policy for ReNode Constructor}\label{renodeconstructor}
The constructor for \sf{ReNode} is private to the module and class \sf{ReNode}. This means that an unknown object cannot directly use the constructor of \sf{ReNode} even if it holds the capability of an object belonging to class ReNode, unless the unknown object is in the same module as ReNode.\\

\begin{logic}
\emp{Policy ---[Constructor of new ReNode]}\\
\loforall rn:ReNode, \loforall n:Node, \loforall k:\lonat. [ \textit{true} \{\texttt{rn = ReNode(n,k)}\} \=rn.node = n \loand\\
\>rn.depth = k \loand\\
\>rn.parent = \texttt{null} ]\\
\end{logic}

\hr
\paragraph{Policies on invariants and immutable properties}\label{policy_immutable}
The following policies describe the properties that hold in any state of execution. Namely, a node that a restricted node points to will never change [Immutability of ReNode.node], its ancestor(s) will never change [Immutability of ReNode.parent\losup{k}], its depth will never change [Immutability of ReNode depth], and the depth of a restricted node will never be less than 0 [Invariant of minimum ReNode depth].\\

\begin{logic}
\emp{Policy ---[Immutability of ReNode.node]}\\
\loforall rn:ReNode, \loforall n:Node. [ rn.node = n \{\texttt{code}\} rn.node = n ]\\
\end{logic}

\begin{logic}
\emp{Policy ---[Immutability of ReNode.parent\losup{k}]}\\
\loforall rn,rn$'$:ReNode, \loforall k:\lonat. [ rn.parent\losup{k} = rn$'$ \{\texttt{code}\} rn.parent\losup{k} = rn$'$ ]\\
\end{logic}

\begin{logic}
\emp{Policy ---[Immutability of ReNode depth]}\\
\loforall rn:ReNode, \loforall k:\lonat. [ rn.depth = k \{\texttt{code}\} rn.depth = k ]\\
\end{logic}

\begin{logic}
\emp{Policy ---[Invariant of minimum ReNode depth]}\\
\loforall rn:ReNode. [ \textit{true} \{\texttt{code}\} rn.depth \logeq 0 ]\\
\end{logic}

\hr\\

\subsection{Reasoning about the DOM Tree with unknown code using our policies}
In this subsection, we will now discuss how the policies we have developed in \cref{sec:dompolicies} allow us to reason about  the use of restricted nodes of class \sf{ReNode} to attenuate authority of nodes of class \sf{Node} is a DOM tree. In particular, we can argue that authority is attenuated in the presence of the execution of unknown code. We also believe that the argument we have put forth using our logics is relatively simple, and natural to a programmer, compared with the corresponding works by Devriese et al.\cite{devriese2016} that uses Kripke world logics and Swasey et al.\cite{swasey2017} that builds on the Iris framework for concurrent separation logic. Furthermore, our framework allows us to reason about unknown code without needing to deploy more sophisticated tools of more advanced logic frameworks. In the code below, we create a DOM tree with the following structure:\\
\begin{minipage}{\textwidth}
\centering
\includegraphics[width=0.80\textwidth]{figures/assertion.pdf}
 \captionof{figure}{DOM Tree Example With Unknown Code}
  \label{DOM example}
\end{minipage}
\begin{lstlisting}[numbers = left]
node0 = Node.createRoot("node0")
node0.setProp("key", "true")

node1 = node0.addChild("node1")
node1.setProp("key", "true")

node2a = node1.addChild("node2a")
node2a.setProp("key", "true")

node2b = node1.addChild("node2b")
node2b.setProp("key", "true")

reNode2b = ReNode(node2b, 1) //reNode2b given depth level of 1

// we have a mystery object o, which we do not trust
// o has a mystery method that takes an object reference as an argument
// we give o the capability of node2b, and o attempts to execute a mystery method on renode2b:

o.mystery0(renode2b)

&\textbf{Assertion 1:}& node0.getProp("key") = "true"
&&
\end{lstlisting} \captionof{figure}{Code for DOM Tree Pattern with unknown provenance}
\vspace{0.5em}
It is unsafe to hand the capability of \sf{node2b} directly to an untrusted agent, because it is possible the properties and state of the entire DOM tree would be compromised, in particular the root node \sf{node0} whose properties and state we want to protect in this example. Note that an untrusted agent having the direct capability of \sf{node2b} implies being that the untrusted agent is able to obtain the direct capability of \sf{node1}, \sf{node2a}, and \sf{node0}, because \sf{node2b} has methods that freely gives those capabilities to any caller.\\

 How then, can we allow an untrusted agent whose object and code we do not know, to safely modify the properties of \sf{node2b},  and some other nodes (\sf{node2a} and \sf{node1}), whose properties are not critical and we do not mind being modified, but \textbf{not} allow the untrusted agent to modify the properties of \sf{node0} which are critical and which we want to protect?\\

We will accomplish this by creating an attenuating restricted node object called \sf{reNode2b}, which points to \sf{node2b} and has a depth level of \sf{1} (\sf{reNode2b.node = node2b, reNode2b.depth = 1}). On \texttt{line 19}, we pass \sf{reNode2b} to an object \sf{o}, of unknown provenance. While we do not know exactly how \sf{o} might make use of \sf{reNode2b}, to represent unknown code execution on \sf{reNode2b} by \sf{o}, we can assume that \sf{o} would have a function called \sf{mystery} that executes some unknown code on \sf{reNode2b}. Using our policies, we can comfortably argue that the properties of \sf{node0} have been preserved after the execution of the unknown \sf{mystery} function by \sf{o}. However, because we have configured \sf{reNode2b} to allow the unknown object to potentially modify the properties of \sf{node2a}, \sf{node2b} and \sf{node1}, we do not know if the properties of \sf{node2a}, \sf{node2b} and \sf{node1} have been preserved.\\

We argue that \textbf{Assertion 1} at \texttt{line 21} holds, such that after execution of a piece of unknown code, \sf{node0.getProp("key") = "true"}.
We begin by stating several facts at \texttt{line 18}, before the execution of \sf{mystery}; we know that:
\begin{itemize}
\item \textbf{F1}: \sf{node2b.getProp("key") = true}
\item \textbf{F2}: \sf{reNode2b.depth = 1}
\item \textbf{F3}: \sf{reNode2b.node = node2b}
\item \textbf{F4}: \sf{reNode2b.parent.node = node1}
\item \textbf{F5}: \sf{reNode2b.parent.child("node2a").node = node2a}
\item \textbf{F6}: \sf{reNode2b.parent.parent.node = node0}
\end{itemize}

Then at \texttt{line 21}, we consider the effects of the \sf{mystery} function. We argue that:
\begin{itemize}
\item \textbf{(A)}: \textbf{F1} is preserved, unless \textbf{\sf{MayCall(o,node0)}} holds at line 21. This is because to modify the value of "key" in \sf{node2b}, \textbf{Policy [Node Property Modification]} at \cref{policy_node_mod} tells us that a method of \sf{node0} has to be invoked by \sf{o}.
\end{itemize}
Therefore, to show that \textbf{Assertion 1} at \texttt{line 21} holds, it is sufficient to show that at \texttt{line 19}, the assertion MayCall(o,node0) does not hold. We will show this by contradiction.\\

Let us assume:
\begin{itemize}\item \textbf{(B)}: \sf{MayCall(o,node0)}\end{itemize}

During execution of the method call \sf{mystery}, the receiver is \sf{o}, and the method argument is \sf{reNode2}. Moreover, \sf{node0} was created before \sf{o}. Therefore, within this frame, it holds that : 
\begin{itemize}
\item \textbf{(C)}: \sf{Dom(\{reNode2b\}, node0)}
\item \textbf{(D)}: \sf{reNode2b:ReNode}
\end{itemize}

Applying \textbf{Policy [Object calls Node]} on \textbf{(C)} and \textbf{(D)}, we obtain:
\begin{itemize}
\item \textbf{(E)}:
\begin{logic}
MayCall(o,node0) \loimplies\ MayCall(o,reNode2b) \loand\ MayCall(reNode2b,node0)
\end{logic}
\end{itemize}
From \textbf{(D)} and \textbf{(E)} and applying \textbf{Policy [ReNode calls Node]}, we obtain that there exists a rn:ReNode such that:
\begin{itemize}\item \textbf{(F)}: \sf{MayCall(renode2b,node0)} \loimplies \sf{MayCall(renode2b,rn) \loand\ rn$'$.node = node0}\end{itemize}
From \textbf{(D)} and by applying \textbf{Policy [ReNode calls ReNode]} on \textbf{(F)}, we know that there exists a k,j:\lonat\ such that:
\begin{itemize}\item \textbf{(G)}:
\begin{logic}
MayCall(renode2b,rn) \loimplies\ reNode2b.parent\losup{k} = rn.parent\losup{j} \loand\ reNode2b.depth \logeq k
\end{logic}
\end{itemize}
From \textbf{(G)}, and \textbf{(F2)} \sf{which says reNode2b.depth = 1}, we obtain that:
\begin{itemize}\item \textbf{(H)} k = 0 \loor\ k = 1
\item Using proof by cases,\\
\begin{logic}
\textbf{1\losup{st} case} k = 0, \=then for all rn$'$ such that rn$'$.parent\losup{j} = reNode2b,\\
\>we have that rn$'$ are descendants of reNode2b,\\
\>and therefore from \textbf{(F3) and (F6)}, we also know:\\
\>\textbf{rn$'$.node \loneq\ node0}
\\
\textbf{2\losup{st} case} k = 1, \=then for all rn$'$ such that rn$'$.parent\losup{j} = reNode2b.parent,\\
\>we have that rn$'$ are descendants of reNode2b.parent,\\
\>and therefore from \textbf{(F4) and (F6)}, we also know:\\
\>\textbf{rn$'$.node \loneq\ node0}
\end{logic}
\item Therefore, \textbf\sf{{rn$'$.node \loneq\ node0}} from proof by cases.
\end{itemize}
We thus obtain a contradiction from \textbf{(F)}, and we know that assumption \textbf{(B)} must be false:
\begin{itemize} \item \textbf{\sf{\loneg MayCall(o,node0)}} \end{itemize}
***Consequently, from \textbf{(A)}, we know the properties of \sf{node0} is preserved.\\

Note that in the above steps we used without mentioning explicitly that the parent and depth fields of all restricted nodes belonging to \sf{ReNode} are immutable. Particularly, our \textbf{Policies on invariants and immutable properties} in \cref{policy_immutable} tell us that the following must hold in all states of exection: \textbf{(F2-F6)}, and the minimum depth of all restricted nodes must be greater than 0.
\clearpage
\subsection{Caretaker and Membrane Patterns}

The Caretaker pattern originally appeared in Redell's 1974 work {\cite{redell1974}, but features prominently along with its more advanced Membrane pattern in the works of \cite{miller2006,murray2010,swasey2017}. Unlike the previous DOM Tree pattern, protected objects in the Caretaker and Membrane pattern are now no longer necessarily ordered in some hierarchical access structure. Consequently, the attenuating 'caretaker' and 'membrane' objects do not try to maintain the integrity of some hierarchy structure, as compared to the attenuating ReNode object in the DOM Tree, where it maintains hierarchy of the Tree by adjusting the depth of ReNodes that it can return (parent ReNodes have decremented depths while child ReNodes have incremented depths).\\

The Caretaker merely fulfils the simple task of 1) masking the protected object's direct capability, and 2) only forwarding messages to the protected object according to some condition in an associated lock object. Note that a simple caretaker object forwards messages without analysing the contents of the messages, i.e. messages can contain capabilities. The Membrane pattern on the other hand, shares some similarities from both the DOM Tree pattern and Caretaker pattern. The Membrane pattern, is identical to the Caretaker pattern in the sense that it is meant to mask the protected object's direct capability and forward messages to the protected object, but with the exception that it does not forward messages that contain capabilities without first 'wrapping' those capabilities within new membranes. The membrane attenuating object can create more membranes (or deep attenuating objects) that is similar to the DOM Tree pattern where a restricted node creates more restricted nodes when accessing different parts of the tree.\\

We show the visual illustration of the differences between the Caretaker and Membrane patterns in \cref{fig:figcaretaker} and \cref{fig:figmem}. In the Caretaker pattern, while the status of the lock object \sf{Carol-CT-Lock} for the caretaker object \sf{Carol-CT} is disabled (\sf{Carol-CT.status = false}), object \sf{Bob} can pass his own capability to \sf{Carol} using \sf{Carol-CT}, and consequently, with \sf{Bob}'s capability, \sf{Carol} can pass the capability of \sf{Diane}, which \sf{Carol} initially holds, to \sf{Bob}. \textit{After} the \sf{Lock} object is enabled (\sf{Carol-CT.status = true}), \sf{Carol-CT} no longer forwards messages to \sf{Carol}. However, because \sf{Bob} has transferred his own capability to \sf{Carol}, and \sf{Carol} has transferred the capability of \sf{Diane} to \sf{Bob} during \sf{Carol-CT.status = false}, even after \sf{Carol-CT.status = true}, Carol can continue to communicate with \sf{Bob}, and \sf{Bob} can continue to communicate with \sf{Diane}, because these objects can store the capabilities they have received for future use. For the Membrane example however, for all messages that contain capabilities, those messages will be attenuated before being forwarded. When \sf{Bob} tries to send his own capability to \sf{Carol} through the membrane \sf{Carol-M}, \sf{Carol-M} will wrap the capability of \sf{Bob} within the message in a new membrane \sf{Bob-M} and lock \sf{Bob-M-Lock}, and forward the new membrane  \sf{Bob-M} to \sf{Carol}, and when \sf{Carol} tries to send the capability of \sf{Diane} to \sf{Bob} using \sf{Bob-M}, a similar membrane \sf{Diane-M} and lock object is created by \sf{Bob-M}. When \sf{Alice} decides to disable \sf{Carol-M} by setting \sf{Carol-M-Lock = true}, the disabling action can propagate to all membranes descended from \sf{Carol-M}, thereby preventing communication from \sf{Carol} to \sf{Bob} and \sf{Bob} to \sf{Diane}.

\clearpage
\begin{minipage}{\textwidth}
\subsubsection{Caretaker Pattern Illustration}\label{sec:figcaretaker}
\centering
\includegraphics[width=1.05\textwidth]{figures/Caretaker.pdf}
 \captionof{figure}{Caretaker Pattern}
  \label{fig:figcaretaker}
\end{minipage}

\begin{minipage}{\textwidth}
\subsubsection{Membrane Pattern Illustration}\label{sec:figmem}
\centering
\includegraphics[width=1.05\textwidth]{figures/Membrane.pdf}
 \captionof{figure}{Membrane Pattern}
   \label{fig:figmem}
\end{minipage}
\clearpage

\section{Case Study of Ethereum / Solidity as a Non-OCap Model}\label{sec:ethereum}

\subsection{Motivation}
Recent widespread adoption of distributed ledger technology (blockchain) has created multiple decentralised, distributed computational platforms where millions of dollars are transacted over codified constructs called smart contracts. For example, as of 7 September 2017, Ethereum\cite{wood2014} is approximately a US\$30 billion blockchain platform with an in-built Turing-complete programming language that can be used to create and deploy such contracts.

\subsection{Solidity - Language for smart contracts}
Solidity\footnote{https://solidity.readthedocs.io/en/develop/} is a high-level language with a syntax similar to that of JavaScript and is designed to write smart contracts on the Ethereum Virtual Machine (EVM), the Turing-complete 256bit runtime environment of the Ethereum blockchain\footnote{Details on the Ethereum blockchain platform in our paper are based on the Etheruem white-paper: https://github.com/ethereum/wiki/wiki/White-Paper\\}.

\subsection{Objects in Solidity}
On the Ethereum blockchain, there are generally two types of accounts: "externally-owned accounts"(external accounts)  and "contract accounts"(contracts). We define these entities as objects because they encapsulate both state and behaviour. In terms of state, both types of accounts have an Ether balance (the currency of Ethereum) associated with them, while contract accounts differ from external accounts by having additional persistent contract storage that can store data such as integers and strings. In terms of behaviour, both types of accounts come with a predefined set of methods to transfer Ether, but contract accounts can have additional specified methods. The methods of external accounts can only be called by a person who has authenticated himself as the owner of the external account using a private key, while methods of contract accounts that are deployed on the blockchain can be called based on any conditions defined in the contract.
Hence, external accounts are controlled by public-private key cryptography, while contract accounts are controlled by their contract code. An external account has no code but can send messages to other external accounts or contract accounts by creating or signing a transaction. For contract accounts, everytime the contract account receives a message, its code activates, allowing it to read and write to its internal storage and sending other messages to other external accounts or contracts. Transactions can be triggered from both types of accounts, though contracts can only trigger transactions in response to other transactions that they have received. Therefore, all transactions in Ethereum can only originate from external accounts.

\subsubsection{Object References/Addresses} In Ethereum, the designations of all external account and contract objects are made of up a 20 bytes value. Hence, if we treat smart contracts and Ethereum accounts as objects, then these object references are non-opaque---because anybody can specify an address using a 20 bytes string value. That is, anybody on Ethereum only needs the bit information of a particular address of an object to start interacting with the object. Unlike in Java where knowing the bit information of an object address does not imply knowing the designation/reference of an object, in Ethereum, the bit information of an address \textit{is} the object reference.

\subsection{Object Protection in Solidity}
Because all smart contract/account object references in Solidity are available to everybody, the concept of isolating objects do not exist in Ethereum. Consequently, the concept of protecting objects through eventual path-isolation does not exist, because:\\

\begin{logic}
\ablock\loforall o:ExtAcct, \loforall o$'$:Object. [ MayAccess$^{Dir,Now}$(o,o$'$) ]\\
\end{logic}

where we define o as objects representing external accounts and o$'$ as objects representing external accounts or contracts in Solidity. In other words, on the Solidity platform, every external account has a path to every object on the blockchain. Therefore, in Ethereum, private-public key cryptography is used to determine whether one has the authority to use an external account, while for contract accounts, we can only use a form of stack-based access control that allow whether a method called is allowed to succeed or fail---the protected contract object o$'$ can only enforce policies in a state \losigma\ \textit{after} it has been called by another object since:\\

\begin{logic}
\ablock \loforall o:ExtAcct, \loforall o$'$:Object. [ MayCall(o,o$'$) ]\\
\end{logic}

\subsubsection{Protection through Message Sender Authentication}
In Ethereum, protection of an external account is done through private key authentication and protection of a smart contract is done through specifying the conditions on whether a particular method call would be successful. For external accounts, once a person has provided the correct private key, the person is allowed to call any of the predefined methods in an external account that facilitate ether transactions. For contracts, things become more complex because the code of all contracts are transparent and publicly available to everyone. What then, are the kind of protection mechanisms that a smart contract can deploy? As addresses of smart contracts are transparent and globally available, contracts \textit{cannot} prevent anybody from sending them messages, but they can dictate \textit{how} they respond to these messages.\\

All this means is that from the receiver object's viewpoint, there must be a way for the receiver to know \textit{who} or what is the sender address of the message. The Ethereum platform facilitates sender authentication, such then when an external account or contract account calls a method on a contract account o$'$ by sending a message to o$'$, o$'$ as the receiver, can always see the address of the message sender, and can trust the address of the message sender to be authentic\footnote{Note that we are less concerned with how the architecture of the Ethereum platform guarantees the authenticity of who the message sender is, but more with higher-level protection strategies that smart contracts can employ knowing such a guarantee exists.}.\\

In other words, when Object A sends a message to Object B, Object A \textit{cannot} mask or hide its own address or pretend to be another Object in the message to B, and B can always guarantee that the designation of the sender is the true sender of the message. If the sender of the message is an address of an external account, we know that a person has authenticated using a private-key to be the external account's owner and fired off the transaction from the external account. If a sender of a message is a contract address, then we know the message is triggered from the contract (but that the origin of such a transaction has to be an external account).\\

From the receiver viewpoint, Solidity allows a method call to succeed or fail based on function modifiers. Function modifiers represent conditions written by the programmer that can be attached to methods that determine whether calls on that method would succeed or fail. They provide more flexibility than the private-public access modifiers in most object-oriented languages. They differ from access modifiers in the sense that access modifiers restrict an object's method from being called by a caller, i.e. the method \textit{cannot be called}, while function modifiers in Solidity, they can be considered as \textit{part of the method execution}. In a smart contract, one can use the following code in Solidity to determine who exactly is the message sender:
\begin{itemize}
\item \texttt{Msg.sender} gives the designation of the message sender
\item \texttt{Tx.origin} allows stack introspection and gives the designation of the first sender(originator)
\end{itemize}

A common strategy used in smart contracts to protect methods meant only for the owner of the contract is to first store the 20-byte address of the deployer of the smart contract in an arbitrary variable \texttt{owner} during the constructor of the contract:\\

\begin{lstlisting}
//Contract is called TestContract
TestContract {
	address owner;
	function TestContract{ //constructor
		owner = msg.sender;
		}
}
\end{lstlisting}

With the owner of the contract stored as a persistent state variable, we can now write a function modifier called \texttt{ownerOnly()} in TestContract:
\begin{lstlisting}
modifier ownerOnly(){ require(msg.sender == owner);_;}
\end{lstlisting}

Attaching this owner-only modifier on a method \texttt{restrictedMeth()}, we have:

\begin{lstlisting}
//Contract is called TestContract
TestContract {
	address owner;
	modifier ownerOnly(){ require(msg.sender == owner);_;}
	function TestContract{ //constructor
		owner = msg.sender;
		}
	function restrictedMeth() ownerOnly { //does something restricted to owner}
}
\end{lstlisting}

where we now know that while any object in Solidity can call \texttt{restrictedMeth()} in TestContract, the execution of the method will fail if the object caller's address is not the same as the owner of the contract.
\subsection{Further Work}
We note that our \sf{MayCall} predicate is inadequate to describe the protection mechanisms of smart contracts in Ethereum. This is because, on the Ethereum platform, the protection mechanism in a smart contract essentially happens \textit{during} the execution of a method in that object, i.e., that the receiver in the state of execution must have \textit{already} transitioned into that object, \textit{before} that object can protect itself. Through some condition specified in the method call (e.g. checking the identity of the message sender or other conditions), the method call will fail during execution, where the Ethereum virtual machine will, up till the 'point of failure', revert all state changes during the method execution, except the payment of the underlying 'gas' fees to execute the contract on the blockchain. Therefore, the platform on Ethereum requires stronger formal definitions of authority that does not merely mean a transition of the receiver from an object to another object---the stronger definition of authority needs to be able to describe the \textit{successful completion} of a method call.

\section{Conclusion}
To formally specify OCap patterns in the style of OCap policies, our paper contributes to the literature by proposing new formal definitions for permission and authority that describe access dynamics between objects. Using our formal definition, we have reasoned in a novel way about isolation, cooperation, vulnerability, and protection of objects in the face of trusted and untrusted code. Our paper has also provided an illustration of three OCap patterns and more specifically, the OCap policy specifications using our proposed methodology for the DOM Tree pattern.  Using our OCap policies in the DOM Tree pattern, we show how we can preserve properties of a system that interacts with code of unknown provenance. In the course of coming up with our OCap policies, we have also proposed the novel use of the concept of domination over objects. We have further demonstrated that it is challenging to reason about security properties of modern open systems like the Ethereum blockchain platform that does not adhere to an OCap system, and further work needs to be done in this area given the meteoric rise of capital traded and number of smart contracts on blockchain technology.

%----------------------------------------------------------------------------------------
%	REFERENCE LIST
%----------------------------------------------------------------------------------------
\small
\section{Bibliography}
\bibliographystyle{alpha}
\bibliography{ref}
\clearpage
\section{Appendix}
\subsection{Module Linking}\label{app:modulelinking}
Here we show the definition of module linking found in the appendix of \cite{drossopoulou2015b}.\\

\includegraphics[width=0.5\textwidth,valign=t]{figures/app_modlink1.pdf}\linebreak
\includegraphics[width=0.5\textwidth,valign=t]{figures/app_modlink2.pdf}
\clearpage
\subsection{Assertion}\label{app:assertion}
Here we show the definition of assertion found in the appendix of \cite{drossopoulou2015b}.\\

\includegraphics[width=0.5\textwidth,valign=t]{figures/app_assertion1.pdf}
\includegraphics[width=0.5\textwidth,valign=t]{figures/app_assertion2.pdf}
%----------------------------------------------------------------------------------------

\end{document}

### 2017-09-07 shupeng_msc_report.tex
update

@@
 \usepackage{booktabs}
+\usepackage{pdfpages}
 \usepackage{graphicx}
@@
 on ownership types \cite{clarke1998}.
-Our paper then show how using a mix of our proposed formalisations of permission and authority, and Hoare logics, we can formally specify well-established Object-Capability (OCap) patterns in the form of \textit{OCap policies} \cite{drossopoulou2013,drossopoulou2014,drossopoulou2015,drossopoulou2015b}.
+Our paper then show how using a mix of our proposed formalisations of permission, authority, domination, and Hoare logics, we can formally specify well-established Object-Capability (OCap) patterns in the form of \textit{OCap policies} \cite{drossopoulou2013,drossopoulou2014,drossopoulou2015,drossopoulou2015b}.
 In particular, we reason about protecting the properties of nodes in a DOM Tree, inspired by the works of Maffeis et al.\cite{maffeis2010} and Devriese et al.\cite{devriese2016}, but we do so within the capability-safe language Pony \cite{clebsch2015}.
 We believe our methodology and specifications of OCap patterns are less complex than other methodologies in the literature\cite{devriese2016,swasey2017}, which will allow a programmer to reason convincingly how the use of attenuating objects in capability-safe languages can help to facilitate cooperation betweeen known and unknown code, while still preventing vulnerabilities by preserving desired properties in an open system.
-Lastly, we end our paper with some preliminary insights on thinking about security and risk in a non-OCap system by using the Ethereum\cite{wood2014} blockchain as a case study.\end{abstract}
+Lastly, we end our paper with some preliminary insights on thinking about security and risk in a non-OCap system, using the Ethereum\cite{wood2014} blockchain as a case study.\end{abstract}
 }
@@
 \thispagestyle{empty}
-
+\clearpage
+\thispagestyle{empty}
+\begin{minipage}{\textwidth}
+\begin{center}
+\textit{Acknowledgements}
+\end{center}
+\small
+I am immensely grateful to my supervisor Prof.
+Sophia Drossopoulou, for her dedication and guidance, where many ideas included in this thesis were derived and inspired from.
+I will miss the countless hours we have worked together on this thesis.\\
+
+I am also thankful for the emotional support and encouragement from my friends and family; and the Monetary Authority of Singapore for offering me a scholarship to pursue this MSc degree at Imperial College London.
+\end{minipage}
 \maketitle
@@
 
-With our formal definitions of permission, authority, and domination, our paper then argues how we can use them to reason and distinguish concepts of \textit{isolation}, \textit{cooperation}, \textit{vulnerability} and \textit{protection}.
+With our formal definitions of permission, authority, and domination, our paper then argues in \cref{sec:ocapreasoning} on how we can use them to reason and distinguish concepts of \textit{isolation}, \textit{cooperation}, \textit{vulnerability} and \textit{protection}.
 We claim that our framework allows one to understand and reason about these concepts in a highly straightforward and intuitive way without treating security as a separate burden from the functionality of the system, which we hope would encourage programmers to seek the use of the object-capability model to build robust and secure systems.\\
 
-To show this in practice, our paper provides an illustration of three OCap patterns found in \cref{sec:figdom,sec:figcaretaker,sec:figmem}, but we focus only on providing the OCap policy specifications for the DOM Tree pattern in \cref{sec:dompolicies}.
+To show this in practice, our paper provides an illustration of three OCap patterns found in \cref{sec:figdom,sec:figcaretaker,sec:figmem}, and we focus only on providing the OCap policy specifications for the DOM Tree pattern in \cref{sec:dompolicies}.
 We choose the DOM Tree pattern because we are convinced of the importance of reasoning about unknown code using the object-capability framework in the web security space, where guest/unknown code (such as third party scripted advertising or widgets) are pervasive in the web today (details can be found in the work by Maffeis et al.\cite{maffeis2010}).
-We show how our policies can specify the preservation of properties of nodes, in the presence of unknown code execution.
-Lastly, in \cref{sec:ethereum}, we offer some preliminary insights on the security properties of smart contracts on the Ethereum blockchain system that has experienced explosive growth in recent years, and we recognise that further work has to be done in this exciting new field.
+We show how using our OCap policies, we can specify the preservation of properties of nodes, in the presence of unknown code execution, in \cref{sec:treereasoning}.
+Lastly, in \cref{sec:ethereum}, we offer some preliminary insights on the security properties of smart contracts on the Ethereum blockchain platform that has experienced explosive growth in recent years, and we recognise that further work has to be done in the field of smart contracts on the blockchain technology.
 
@@
 Readers familiar with the object-capability model and its languages are encouraged to skim through this section.
-In \cref{sec:specs}, we seek to formally specify the definitions of permission and authority and use our specifications to describe and reason how objects interact with each other.
+In \cref{sec:specs}, we seek to formally specify the definitions of permission, authority and domination, and we use our specifications to describe and reason how objects interact with each other.
 Our formal specifications are broad enough to reason in general about how two objects interact in our small specification language.
-That said, there are benefits to using these concepts to reason within an OCap system, which we demonstrate in \cref{sec:ocapspecs}, because the restrictions placed by OCap languages or what we call capability-safe languages allow us to enforce protection of objects more easily.
-We further solidify how our methodology can be used to reason convincgly in an OCap system by applying it on a DOM Tree pattern within an OCap framework, in \cref{sec:ocappatterns}.
+That said, there are benefits to using these concepts to reason within an OCap system, which we demonstrate in \cref{sec:ocapreasoning}, because the restrictions placed by OCap languages or what we call capability-safe languages allow us to enforce protection of objects more easily.
+We further solidify how our methodology can be used to reason convincingly in an OCap system by applying it on a DOM Tree pattern within an OCap framework, in \cref{sec:ocappatterns}.
+We end off with some preliminary insights on Ethereum in \cref{sec:ethereum}.
 %----------------------------------------------------------------------------------------
@@
 When these references cannot be forged in the system, and when there there are no default globally accessible objects (for an object to be globally accessible, the reference to use that object has to be \textit{explicitly} granted to every object globally), then it becomes necessary that for an object to call the methods of another object, an object needs to explicitly obtain the reference of the object it is calling.
-Object references therefore embody authority, and we use the word capability to describe the encapsulation of both reference (designation) and authority which are unforgeable.
 
@@
 Global mutable variables can behave like a message carrier, such that two objects can both communicate by monitoring and modifying the state of the global variable.
-The capability-safe language Pony, which we use to show the code of OCap patterns, do not allow global variables.
+The capability-safe language Pony\cite{clebsch2015}, which we use to show the code of OCap patterns, do not allow global variables.
 \subsubsection{No 'Untamed' Reflection}
@@
 For authority, we propose a new MayCall predicate that deviates from the MayAffect predicate proposed in Drossopoulou et al.\cite{drossopoulou2016}.
-While their MayAffect predicate captures the idea of a field of an object being modified, our MayCall predicate is much weaker in the sense that mean the ability to invoke a method of an object \textit{being}.
+While their MayAffect predicate captures the idea of a field of an object being modified, our MayCall predicate is much weaker ---we mean the ability to invoke a method of an object.
+Therefore implicitly, we consider method invocation as an effect in the system.
 Given a module M and state \losigma,:
@@
 where, we say MayAccess$^{Dir,Now}$(x,e) holds \lotiff\ in some current state \losigma, we have a field in x that points to e, or that we have x as a receiver in \losigma\ and there exists some y in the same \losigma\ that points to e.
-We also say MayAccess$^{Ind,Now}$(x,e) holds \lotiff\ in some current state \losigma, we have a \textit{series of fields} from x that leads to e, or that we have x as a receiver in \losigma\ and there exists some y with a series of fields in the same \losigma\ that leads to e.
+We also say MayAccess$^{Ind,Now}$(x,e) holds \lotiff\ in some current state \losigma, we have a \textit{series of fields} from x that leads to e, or that we have x as a receiver in \losigma\ and there exists some y that leads to e.
 Lastly, the 'eventual' definitions of the two predicates are then defined as there existing an arising configuration \losigma$'$ from \losigma, where the 'now' definitions hold at \losigma$'$.\\
@@
 MayAccess$^{Dir,\_}$(o,o$'$) says object o has the direct reference of o$'$, while MayAccess$^{Ind,\_}$ says object o has a path of reference(s) that lead to o$'$ (that is, o has the reference of some object x$_1$, and object x$_1$ has the reference of object x$_2$ $\dots$ and object x$_n$ has the reference of object o$'$).
-The 'now' and 'eventual' flavours simply state that the "direct" and "indirect" predicates either they hold at \losigma\ (Now) or they can hold at some arising eventual state \losigma' from \losigma\ (eventual).
+The 'now' and 'eventual' flavours simply state that the "direct" and "indirect" predicates either they hold at \losigma\ (now) or they can hold at some arising eventual state \losigma' from \losigma\ (eventual).
 
 \subsubsection{Authority - MayCall Definition}\label{MayCall}
-With our access predicates in place, we define a MayCall predicate that describe what it means for x to be able to call e.
+With our MayAccess predicates in place, we define a MayCall predicate that describe what it means for x to be able to call e.
 Note that our definition of MayCall is different from the MayAffect predicate in \cite{drossopoulou2016}.
@@
 \subsection{Domination}\label{sec:domination}
-Here, we introduce the concept of domination that is inspired by the work on ownership types by Clarke et al.\cite{clarke1998}, where they define that the owner o of an object o$'$ dominates o$'$ in the object graph such that for any path p from some arbitrary object o$'$' to o$'$, either o$'$' belongs to the set of owners or that p passes through o.\\
+Here, we introduce the concept of domination that is inspired by the work on ownership types by Clarke et al.\cite{clarke1998}, where they define that the owner o of an object o$'$, dominates o$'$ in the object graph such that for any path p from some arbitrary object o$''$ to o$'$, either o$''$ belongs to the set of owners or that p passes through o.\\
 
 We say that a set S dominates an object x \lotiff\ for every object y in the system that has a path to x, y must have a path to some arbitrary object o that is a member of the set S.
-The power of having such a predicate will be demonstrated in \cref{sec:ocappatterns}, where we show how attenuating objects that is meant to protect an object o can ensure certain properties of members in a dominating set over s under any unknown code execution.\\
 
@@
 \end{logic}
-The usefulness of domination will be made more clear in the next section when we argue about cooperation without vulnerability.
-\section{Reasoning in an OCap System}\label{sec:ocapspecs}
+The usefulness of domination will be made more clear in the next section where we argue about cooperation without vulnerability.
+The power of having such a predicate will also be demonstrated in \cref{sec:ocappatterns}, where we show how attenuating objects that is meant to protect an object o can ensure certain properties of the system under unknown code execution.\\
+\section{Reasoning in an OCap System}\label{sec:ocapreasoning}
 
 In this section, we formally describe the ideas of \textit{isolation} and \textit{cooperation}, and how \textit{cooperation} can either be vulnerable or protected, within the OCap system.
-We use the word \textit{capability} (of an object), interchangeably with reference (of an object), to represent that both object reference and the right to use the object reference are indistinguishable in OCap; i.e., there is no global ambient authority to prevent an object from using another object reference in general.
-Given our definition of \textit{permission}, it follows that o has the permission of o$'$ if o has the capability of o'.
+We use the word \textit{capability} (of an object), interchangeably with reference (of an object), to represent that both object reference and the right to use the object reference are indistinguishable in OCap; i.e., there is no global ambient authority to prevent an object from using another object reference.
+Given our 4 formal definitions of \textit{permission}, it follows that o has different kinds of permission of o$'$ depending on whether o has a direct or indirect capability of o$'$, and whether we are arguing about a current or eventual state of the system.
 However, if we were to use \textit{our} formal definition of \textit{authority} (causing effects where effects are method invocations), we cannot say that an object reference embodies authority - we elaborate more in \cref{sec:isolation and cooperation}.
@@
 
-In \cref{sec:epc} we first explain that objects require permission (direct capabilities or a chain of capabilities) to communicate, i.e.
+In \cref{sec:epc} we propose a lemma that states that under an OCap system, we can reason about eventual indirect permission, from a current permission configuration of an initial closed system\footnote{We require a closed system because unknown object or code might introduce capabilities of objects of the system that the unknown object or code might already have.}.
+That is, given a system where we know the entire system's present reference graph, OCap rules confine the possible reference graphs that can be born eventually from the present reference graph.
+Next in \cref{sec:isolation and cooperation}, we first explain that objects require permission (direct capabilities or a chain of capabilities) to communicate, i.e.
 permission forms the necessary condition for authority.
-We also propose a lemma that states that under an OCap system, we can reason about eventual indirect permission, from a current permission configuration of an initial closed system\footnote{We require a closed system because unknown object or code might introduce capabilities of objects in the system that the unknown object or code might already have.}.
-That is, given a system where we know the entire system's present reference graph, OCap rules confine the possible reference graphs that can be born eventually from the present reference graph.
-Next in \cref{sec:isolation and cooperation}, we reason about \textit{isolation} and \textit{cooperation} using our formal definition of permission.
+We also reason about \textit{isolation} and \textit{cooperation} using our formal definition of permission.
 Following which in \cref{sec:vulnerability and protection}, we expand the concept of \textit{cooperation} to include cooperation that is vulnerable, and cooperation that is protected, i.e.
@@
 Here, we propose a lemma \textsc{epc} that states how eventual permission can be born from a constellation of direct permissions within an OCap system.
-We use paths interchangeably with direct and indirect permission, to emphasise that an object o need not necessarily have the direct capability of o$'$ to invoke the behaviours of o$'$, invocation can be done through a network of paths consisting of permissions.
 
@@
 
-In the example, we show an example using \textsc{Lemma epc} with 3 objects o, y, o$'$.
+We show an example below using \textsc{Lemma epc} with 3 objects o, y, o$'$.
 Let us assume o, y, and o$'$ always exists, such that the terminating recursive predicates would return true.
@@
 
-We say isolation is simply that there is no eventual directed path from o to o$'$:\\
+We say isolation of o$'$ from o is simply that there is no eventual directed path from o to o$'$:\\
 
@@
 
-When we have isolation, we know from contraposition of the corollary from our definitions of permission and authority:\\
+We know from contraposition of the corollary from our definitions of permission and authority:\\
 
@@
 
-Protection strategies for attenuating objects can be implemented using defensive programming techniques like encapsulation and checking of conditions of method arguments.
+Protection strategies for attenuating objects can be implemented using defensive programming techniques like encapsulation and checking of conditions on method arguments.
 At the extreme ends, attenuating objects can take the shapes of either deciding to forward messages from the unknown object o* in an intact way, thereby allowing MayCall(o*,o$'$) to hold; or if some condition fails, the attenuating object can choose not to forward the message at all such that MayCall(o*,o$'$) does not hold.
-Somewhere in the middle, attenuating objects can also choose to modify the message from o* such that it becomes safe to forward, where MayCall(o*,o$'$) holds but since it has been attenuated by the attenuating object, we would know it is safe.
+Somewhere in the middle, attenuating objects can also choose to modify the message from o* such that it becomes safe to forward, where MayCall(o*,o$'$) holds but since it has been attenuated by the attenuating object, we would know that the message from the attenuating object would be safe.
 What and how the attenuating objects decide to do with messages from unknown objects before forwarding them can be best expressed as OCap policies for the attenuating object.\\
@@
 In the OCap pattern examples we will go through in the next section, protection of an object is accomplished through some attenuating object that will never leak the capability of the object of interest.
-A safe configuration is also expressed as attenuating objects dominating the object of interest in such a way that unknown code is forced to use them to cooperate.
+A safe configuration is also expressed as attenuating objects dominating the object of interest in such a way that unknown code is forced to use them when cooperating with objects of interest in the system.
 
@@
 These three OCap patterns are: the DOM Tree pattern, the Caretaker pattern, and the Membrane pattern.
-In addition to the visual diagrams of these patterns, we also show the code of these three patterns in the capability-safe language Pony\cite{clebsch2015}.
+In addition to the visual diagrams of these patterns, we also show the code of these three patterns in the capability-safe language Pony\cite{clebsch2015}, in the appendix \cref{sec:code_DOM,sec:code_Caretaker,sec:code_Membrane}.
 To make use of our formal specifications, we focus exclusively on the DOM Tree pattern, where we show how we can use our formal specifications to write policies that can guarantee the preservation of properties of specific nodes in the pattern.
@@
 This type of reasoning is useful when we need to protect the integrity of more sensitive objects higher up in the hierarchy of the tree, while at the same time, we need to allow untrusted external objects the ability to affect less-sensitive objects lower in the hierarchy of the tree.
-Our contribution in our paper is that we implement the pattern and give the code for the pattern classes \sf{Nodes} and \sf{ReNodes} in the capability-safe language Pony.
-Another contribution of ours is that provide the formal specifications of such a pattern using our methodology, which then allows us to assert conditions when properties of a Node in the tree are modified.
+Reasoning about unknown code using the object-capability framework in the web security space is important and relevant, where guest/unknown code (such as third party scripted advertising or widgets) are pervasive in the web today.
+Details can be found in the work by Maffeis et al.\cite{maffeis2010}.
+Our contribution is that we implement the pattern and give the code for the pattern classes \sf{Nodes} and \sf{ReNodes} in the capability-safe language Pony.
+Another contribution is that we provide the formal specifications of such a pattern using our methodology, which then allows us to assert conditions when properties of a Node in the tree are modified.
 \\
@@
 Subsequently, we show how we can create an attenuating object \sf{ReNode} that protects access to a \sf{Node}, in \cref{domrenode}.
-In \cref{domexample}, we show the code of a simulation where we have a JavaScript web document (simulated in Pony) where we need to give some arbitrary 3rd party advertisement company external access to a particular node called \sf{AdNode} in the document (so that the website is able to run advertisements and make money).
-In the simulation, we show how using an attenuating restricted node \sf{RAdNode} with a configuration of depth = 0, is able to protect all other nodes higher in the hierarchy of the DOM Tree, while \textit{still} allowing the 3rd-party advertiser the ability to modify properties of the \sf{AdNode}.
-We then show a graphical illustration of \sf{AdNode} and \sf{RAdNode} simulation \cref{fig:figdom}.
+In \cref{domexample}, we show the code of a simulation where we have a web document (simulated in Pony) where we need to give some arbitrary 3rd party advertisement company external access to a particular node called \sf{adnode} in the document (so that the website is able to run advertisements and make money).
+In the simulation, we show how using an attenuating restricted node \sf{radnode} with a configuration of depth = 0, is able to protect all other nodes higher in the hierarchy of the DOM Tree, while \textit{still} allowing the 3rd-party advertiser the ability to modify properties of the \sf{adnode}.
+We then show a graphical illustration of this example in \cref{sec:figdom}.
 Following which, we write the object-capability policies of the DOM Tree pattern using our methodology, in \cref{sec:dompolicies}.
-Lastly, we show how we can begin to reason the protection of a node using those policies.
+Lastly, in \cref{sec:treereasoning}, we show how we can begin to reason the protection of a node using those policies.
 
@@
 
-In the OCap literature, letting an object \sf{o} have access to an object \sf{o$'$} (\sf{o} having the reference of \sf{o$'$}) \textit{usually} implies we want to let object \sf{o} use or call methods on \sf{o$'$}.
+In the OCap literature, letting an object \sf{o} have permission of an object \sf{o$'$} (\sf{o} having the reference of \sf{o$'$}) \textit{usually} implies we want to let object \sf{o} use or call methods on \sf{o$'$}.
 More precisely, in such a configuration, we are typically prepared to let o have the ability to modify the state of o$'$.
@@
 In the DOM Tree pattern, we make use of an attenuating restricted node class \sf{ReNode} whose objects serve to attenuate the authority of objects belonging to class \sf{Node}.
-A restricted node \sf{rn} of class \sf{ReNode} attenuates the capability of a node of class \sf{Node} in the following ways:
+A restricted node \sf{rn} of class \sf{ReNode} attenuates authority in the following ways:
 
@@
 By calling \sf{parent()} method on \sf{rn}, \sf{rn} will only \sf{forward} the parent method call to \sf{n} \textbf{only} when the condition that \sf{rn}'s depth level stored in the integer variable \sf{d} must have a value of at least more than 0.
-Even if \sf{rn} has a depth level more than 0, the restricted node does not return the direct capability of the parent node of \sf{n} to the method caller on \sf{rn}, but instead \textit{processes} the result to the method caller on \sf{rn}, by only returning a new restricted node object \sf{rn'} with a field that points to the parent node of \sf{n}, where \sf{rn'} inherits the depth level of \sf{rn} decremented by 1.
+Even if \sf{rn} has a depth level more than 0, the restricted node does not return the direct capability of the parent node of \sf{n} to the method caller on \sf{rn}, but instead \textit{processes} the result to the method caller on \sf{rn}, by only returning a new restricted node object \sf{rn$'$} with a field that points to the parent node of \sf{n}, where \sf{rn$'$} inherits the depth level of \sf{rn} decremented by 1.
 Lastly, the restricted node does not have any methods that return the direct capability of the node that it protects and all the node's children.
 Method calls done by the restricted node on the underlying node that return the capabilities of the child nodes are always processed by the restricted node such that the restricted node only returns newly created restricted nodes of those child nodes configured with a depth level incremented by 1.
+This ensures that a child restricted node will never have more authority than its parent node.
 
@@
 In this example, we want to allow an untrusted object \sf{o} the authority to modify the properties of \sf{n}, but not the root node \sf{n$'$}.
-To illustrate this, we use a DOM Tree and call the root note \sf{n$'$} a \sf{document} node, and the child node \sf{n} an \sf{adnode}.
+To illustrate this, we call the root note \sf{n$'$} a \sf{document} node, and the child node \sf{n} an \sf{adnode}.
 Typically, a website that wants to make money from having advertisements will have a particular node (\sf{adnode}) within the DOM Tree, so that a dynamic advertisement object can be rendered properly in the document.
@@
 
-In the DOM Tree -adnode- code example above, we simulate initialising a webpage with a root \sf{document} node that we set the property of a map-key "title" having a map-value of "Alice website".
+In the DOM Tree example above, we simulate initialising a webpage with a root \sf{document} node that we set the property of a map-key "title" having a map-value of "Alice website".
 We then create a node for an advertisement called \sf{adnode}.
@@
 \subsubsection{DOM Tree Pattern Illustration}\label{sec:figdom}
-Here, we illustrate in more detail the example of the DOM Tree pattern in a visual diagram.\\
+\raggedright \small Here, we illustrate in more detail the example of the DOM Tree pattern in a visual diagram. Code can be found in found in \cref{sec:code_DOM}\\
 \centering
@@
 
-\subsection{Reasoning about the DOM Tree with unknown code using our policies}
+\subsection{Reasoning about the DOM Tree with unknown code using our policies}\label{sec:treereasoning}
 In this subsection, we will now discuss how the policies we have developed in \cref{sec:dompolicies} allow us to reason about  the use of restricted nodes of class \sf{ReNode} to attenuate authority of nodes of class \sf{Node} is a DOM tree.
@@
 Furthermore, our framework allows us to reason about unknown code without needing to deploy more sophisticated tools of more advanced logic frameworks.
-In the code below, we create a DOM tree with the following structure:\\
+In the code below, we create another DOM Tree example with the following structure:\\
 \begin{minipage}{\textwidth}
@@
 \includegraphics[width=0.5\textwidth,valign=t]{figures/app_assertion2.pdf}
+
+\subsection{Pony Code - DOM Tree}\label{sec:code_DOM}
+\begin{minipage}{\textwidth}
+\includegraphics[width=\textwidth,valign=t,page=1]{figures/code_DOM.pdf}
+\end{minipage}
+\begin{minipage}{\textwidth}
+\includegraphics[width=\textwidth,valign=t,page=2]{figures/code_DOM.pdf}
+\end{minipage}
+\subsection{Pony Code - Caretaker}\label{sec:code_Caretaker}
+\begin{minipage}{\textwidth}
+\includegraphics[width=\textwidth,valign=t,page=1]{figures/code_Caretaker.pdf}
+\end{minipage}
+\begin{minipage}{\textwidth}
+\includegraphics[width=\textwidth,valign=t,page=2]{figures/code_Caretaker.pdf}
+\end{minipage}
+\begin{minipage}{\textwidth}
+\includegraphics[width=\textwidth,valign=t,page=3]{figures/code_Caretaker.pdf}
+\end{minipage}
+\subsection{Pony Code - Membrane}\label{sec:code_Membrane}
+\begin{minipage}{\textwidth}
+\includegraphics[width=\textwidth,valign=t,page=1]{figures/code_Membrane.pdf}
+\end{minipage}
+\begin{minipage}{\textwidth}
+\includegraphics[width=\textwidth,valign=t,page=2]{figures/code_Membrane.pdf}
+\end{minipage}
+\begin{minipage}{\textwidth}
+\includegraphics[width=\textwidth,valign=t,page=3]{figures/code_Membrane.pdf}
+\end{minipage}
+
 %----------------------------------------------------------------------------------------

### 2017-09-08 shupeng_msc_report.tex
updated code

@@
 \noindent  We revisit the concepts of \textit{permission} and \textit{authority} in the works of Miller\cite{miller2006} and \linebreak Drossopoulou et al.
-\cite{drossopoulou2016}, and we contribute to the literature by proposing new formal definitions of permission and authority (MayAccess and MayCall predicates in our paper).
-We show how our definitions can be used to reason about object-to-object interactions, and in particular, cooperation between objects and protection of objects from unknown code.
-Another contribution we made is that we propose and show how the concept of domination over objects can be used to specify and reason about safe or vulnerable cooperation, which is novel but inspired by the work of Clarke et al.
+\cite{drossopoulou2016}, and we contribute by proposing new formal definitions of permission and authority (MayAccess and MayCall predicates in our paper).
+Another contribution we made is that we propose and show how the concept of domination over objects can be used to reason about OCap policies, which is novel but inspired by the work of Clarke et al.
 on ownership types \cite{clarke1998}.
-Our paper then show how using a mix of our proposed formalisations of permission, authority, domination, and Hoare logics, we can formally specify well-established Object-Capability (OCap) patterns in the form of \textit{OCap policies} \cite{drossopoulou2013,drossopoulou2014,drossopoulou2015,drossopoulou2015b}.
-In particular, we reason about protecting the properties of nodes in a DOM Tree, inspired by the works of Maffeis et al.\cite{maffeis2010} and Devriese et al.\cite{devriese2016}, but we do so within the capability-safe language Pony \cite{clebsch2015}.
+We show how our definitions above can be used to reason about object-to-object interactions, and in particular, cooperation between objects and protection of objects from unknown code.
+Our paper then shows how the use of our concepts of permission, authority, and domination within extended Hoare logics can formally specify well-established Object-Capability (OCap) patterns, namely the DOM Tree pattern, Caretaker pattern and Membrane pattern, in the form of \textit{OCap policies} \cite{drossopoulou2013,drossopoulou2014,drossopoulou2015,drossopoulou2015b}.
+In particular, we show how we can reason about protecting the properties of nodes in the DOM Tree pattern, which is inspired by the works of Maffeis et al.\cite{maffeis2010} and Devriese et al.\cite{devriese2016}, but we do so within our methodology and the capability-safe language Pony \cite{clebsch2015}.
 We believe our methodology and specifications of OCap patterns are less complex than other methodologies in the literature\cite{devriese2016,swasey2017}, which will allow a programmer to reason convincingly how the use of attenuating objects in capability-safe languages can help to facilitate cooperation betweeen known and unknown code, while still preventing vulnerabilities by preserving desired properties in an open system.
@@
 In recent years the Object-Capability (OCap) model has received increasing attention as a compelling approach to building robust, distributed systems that promote what Miller\cite{miller2006} calls \textit{cooperation without vulnerability}.
-The OCap model attemps to address these two issues by alleviating security as a separate concern from the mind of the programmer.
+The OCap model addresses these two issues by alleviating security as a separate concern from the mind of the programmer.
 This is done through leveraging the object-oriented programming paradigm and imposing certain prohibitions or restrictions.
@@
 \subsection{Our contributions}
-While there have been considerable work done on OCap patterns in programming languages \cite{miller2003b,miller2006,murray2010}, less work has been done on formally specifying them.
+While there have been considerable work on OCap patterns in programming languages \cite{miller2003b,miller2006,murray2010,drossopoulou2014b}, less work has been done on formally specifying them.
 The only works we know so far are \cite{drossopoulou2015b,devriese2016,swasey2017}, which are all very recent in the literature.
-Our paper builds on the works of \cite{drossopoulou2015b}, because we claim that the methodology of using OCap policies to specify OCap patterns to be intuitive and straightforward.
-While they may seem less sophisticated than the methodologies in other works, we believe they are powerful enough to reason convincingly about the interactions between trusted code and code of unknown provenance.
-To formally specify OCap patterns in the style of OCap policies, our paper contributes to the literature by proposing formal predicates for permission and authority that describe access dynamics between objects.
+Our paper builds on the works of \cite{drossopoulou2015b}, but we propose a different methodology for specifying OCap patterns which we claim is intuitive and straightforward.
+While ours may seem less sophisticated than the methodologies in other works, we believe that ours are powerful enough to reason convincingly about the interactions between trusted code and code of unknown provenance.\\
+
+To formally specify OCap patterns in the style of OCap policies, our paper contributes to the literature by proposing formal predicates for permission and authority that describe dynamic access between objects.
 Miller first proposed the concepts of permission and authority in his PhD thesis\cite{miller2006}, following which the formal definitions of these concepts were proposed and refined in the work of Drossopoulou et al.
 \cite{drossopoulou2016}.
-Our paper defines permission and authority in a formal way that is different from the aforementioned works, but we generally adopt the same informal meanings (that having permission means the right to invoke an object's behaviours, and authority to mean the ability to cause effects).
+Our paper defines permission and authority in a formal way that is different from the aforementioned works, but we generally adopt similar intuitions (that having permission means the right to invoke an object's behaviours, and authority to mean the ability to cause effects).
 More precisely, our paper defines \textit{permission} as having the reference of another object (right to invoke that object's behaviours), either directly or indirectly---through multiple object reference(s)---and either currently or eventually---in some current or future state of the system; and defines \textit{authority} to mean being \textit{able} to invoke an object's behaviours; i.e.
-we distinguish between having the \textit{right to do} something (permission) and \textit{being able to do} something (authority)\footnote{An anology would be having the right to perform an impossible task does not imply the task becomes possible.}.
-Furthermore, our paper implicitly treats invoking behaviours of objects as 'effects' in the system.
+we distinguish between having the \textit{right to do} something (permission) and \textit{being able to do} something (authority)\footnote{An analogy would be having the right to perform an impossible task does not imply the task becomes possible.}.
+Our paper also implicitly treats invoking behaviours of objects as 'effects' in the system.
 We claim that these insights are novel and our motivation for defining them can be found in \cref{sec:permauth}.\\
 
-To put our formal definitions of permission and authority in practice and demonstrate their usefulness, we further propose the novel use of the concept of domination over objects that is inspired by the works of \cite{clarke1998} on ownership types.
+We further propose the novel use of the concept of domination over objects that is inspired by the works of \cite{clarke1998} on ownership types.
 Our domination predicate describes a special set of objects that dominate a particular object---all permissions to use the particular object have to involve the permission of a member of the dominating set, and all authorities have to involve the authority of a member of the dominating set.\\
 
-With our formal definitions of permission, authority, and domination, our paper then argues in \cref{sec:ocapreasoning} on how we can use them to reason and distinguish concepts of \textit{isolation}, \textit{cooperation}, \textit{vulnerability} and \textit{protection}.
+With our formal definitions of permission, authority, and domination, our paper then argues in \cref{sec:ocapreasoning} on how we can use them to reason and distinguish concepts of \textit{isolation}, \textit{cooperation}, \textit{vulnerability}, \textit{safety} and \textit{protection}.
 We claim that our framework allows one to understand and reason about these concepts in a highly straightforward and intuitive way without treating security as a separate burden from the functionality of the system, which we hope would encourage programmers to seek the use of the object-capability model to build robust and secure systems.\\
 
-To show this in practice, our paper provides an illustration of three OCap patterns found in \cref{sec:figdom,sec:figcaretaker,sec:figmem}, and we focus only on providing the OCap policy specifications for the DOM Tree pattern in \cref{sec:dompolicies}.
-We choose the DOM Tree pattern because we are convinced of the importance of reasoning about unknown code using the object-capability framework in the web security space, where guest/unknown code (such as third party scripted advertising or widgets) are pervasive in the web today (details can be found in the work by Maffeis et al.\cite{maffeis2010}).
-We show how using our OCap policies, we can specify the preservation of properties of nodes, in the presence of unknown code execution, in \cref{sec:treereasoning}.
+To put our formal definitions of permission, authority, and domination in practice and demonstrate their usefulness, 
+our paper provides an illustration of three OCap patterns, DOM Tree pattern, Caretaker pattern, and Membrane pattern, which can be found respectively in \cref{sec:figdom,sec:figcaretaker,sec:figmem}.
+Our paper also provides the corresponding OCap policies based on our formal definitions for these three patterns.
+In \cref{sec:treereasoning}, we choose the Dom Tree pattern as an example to demonstrate how the DOM Tree policies help us to reason comfortably about the preservation of properties of nodes in the presence of unknown code execution.
+A motivation for using the DOM Tree pattern is how guest/unknown code (such as third party scripted advertising or widgets) are becoming more pervasive in the web today (where details on reasoning about them within the object-capability framework can be found in the work by Maffeis et al.\cite{maffeis2010}).
 Lastly, in \cref{sec:ethereum}, we offer some preliminary insights on the security properties of smart contracts on the Ethereum blockchain platform that has experienced explosive growth in recent years, and we recognise that further work has to be done in the field of smart contracts on the blockchain technology.
@@
 \subsection{Overview of Paper}
-In our paper, we first give an overview background history on the OCap model in \cref{sec:ocapmodel}.
+We first give an overview background history on the OCap model in \cref{sec:ocapmodel}.
 Readers familiar with the object-capability model and its languages are encouraged to skim through this section.
-In \cref{sec:specs}, we seek to formally specify the definitions of permission, authority and domination, and we use our specifications to describe and reason how objects interact with each other.
+In \cref{sec:specs}, we formally specify the definitions of permission, authority and domination, and we use our specifications to describe and reason how objects interact with each other.
 Our formal specifications are broad enough to reason in general about how two objects interact in our small specification language.
@@
 We further solidify how our methodology can be used to reason convincingly in an OCap system by applying it on a DOM Tree pattern within an OCap framework, in \cref{sec:ocappatterns}.
-We end off with some preliminary insights on Ethereum in \cref{sec:ethereum}.
+We end the paper with some preliminary insights on Ethereum in \cref{sec:ethereum}.
 %----------------------------------------------------------------------------------------
@@
 For our language, we use the definitions of modules, classes, runtime state, interpretation, execution and arising configurations from the specification language in the appendix of \cite{drossopoulou2015b}.
-To express more precisely how an object can interact with another object, we use the ideas of permission and authority in Miller's\cite{miller2006} and Drossopoulou et al.'s\cite{drossopoulou2016} works.
-In their papers and our paper, we generally try to capture the informal meaning of:
+To express more precisely how an object can interact with another object, we give formal definitions to the ideas of permission and authority which were studied but not formally defined in \cite{miller2006}.
+They were later studied in several ways, e.g.
+in the work by Maffeis et al.\cite{maffeis2010}.
+Our formal definition is based on but not identical to that from  the work by Drossopoulou et al.'s\cite{drossopoulou2016}.
+In their papers and our paper, we generally try to capture the intuition of:
 \begin{itemize}
@@
 \end{itemize}
-In the approach of our paper however, we contribute to the literature by taking a novel approach in formally defining both permission and authority, though we argue they generally capture the same informal meanings.
+In this work however, we contribute to the literature by taking a novel approach in formally defining both permission and authority.
 For permission, we have four different modes permissions specified by four MayAccess predicates.
@@
 \begin{tabbing}
-MayCall(o,o$'$): \=there exist a method in o that can call a method in o$'$,\\ \> such that there will be some eventual state \losigma$'$ arising from \losigma,\\ \> where the receiver in the stack has transitioned from o in \losigma, to o$'$ in \losigma$'$\\
+MayCall(o,o$'$): \=there exists a method in o that can call a method in o$'$,\\ \> such that there will be some eventual state \losigma$'$ arising from \losigma,\\ \> where the receiver in the stack has transitioned from o in \losigma, to o$'$ in \losigma$'$\\
 \end{tabbing}
@@
 \subsection{Specification Language}\label{sec:speclang}
-In our small specification language, we briefly describe the defintions of module, class, runtime state, interpretation, execution, arising configurations, and assertion.
+In our small specification language, we briefly describe the definitions of module, class, runtime state, interpretation, execution, arising configurations, and assertion.
+All definitions in this section are repeated from the appendix of \cite{drossopoulou2015b} for the reader's convenience.
+Some parts from \cite{drossopoulou2015b} have also been included in the appendix of this paper.
+\\
+
 \textbf{Module:}\\
@@
 Class definitions describe how a class contributes to the runtime behaviour of a program for which we use its field and method declarations.
-We define method bodies as consisting a sequences of statements; these statements can be field read or internal field assignments (only allowed if the object is \texttt{this}), conditionals, and method calls.
-We also define all classes to be private which mean only objects belonging to class C may construct objects of class C.\\
+We define method bodies as consisting a sequences of statements; these statements can be field read or internal field assignments (only allowed if the object is \texttt{this}), conditionals, and method calls.\\
 
@@
 \begin{logic}
+\textit{Runtime State definition}\\
 \losigma\ \loin\ state \quad\= = frame \lotimes\ heap \\
@@
 
+Further details can be found in appendix \cref{app:execution}.\\
+
 \textbf{Arising Configurations:}\\
@@
 We therefore can give meaning to capability policies without requiring termination.\\
+The definition of Reach can be found in appendix \cref{app:reach}.\\
 
@@
 
-We say that a class \sf{C} satisfies an assertion \sf{A} in the context of a module \sf{M}, if in all runtime configuration \losigma$'$ arising from execution of any module \sf{M$'$} linked with module \sf{M}, all objects of class \sf{C} satisfies \sf{A}:\\
+We say that a class \sf{C} satisfies an assertion \sf{A} in the context of a module \sf{M}, if in all runtime configuration \losigma$'$ arising from execution of any module \sf{M$'$} linked with module \sf{M}, all objects of class \sf{C} satisfy \sf{A}:\\
 
@@
 \>\loexists f. \loexec{x.f}{\losigma} = \loexec{e}{\losigma}
-\loor \\
-\> \loexec{\texttt{this}}{\losigma} = \loexec{x}{\losigma} \loand\ \loexists y. \loexec{y}{\losigma} = \loexec{e}{\losigma})\\
+\loor\
+\loexec{\texttt{this}}{\losigma} = \loexec{x}{\losigma} \loand\ \loexists y. \loexec{y}{\losigma} = \loexec{e}{\losigma}\\
 \\
  M,\losigma\ \loturns\ \=MayAccess$^{Dir,Eve}$(x,e) \loiff \\
-\>\loexists \losigma$'$ \loin Arising(M,\losigma).\\
-\>M,\losigma$'$\ \loturns\ MayAccess$^{Dir,Now}$(x,e)\\
+\>\loexists \losigma$'$ \loin Arising(M,\losigma).
+M,\losigma$'$\ \loturns\ MayAccess$^{Dir,Now}$(x,e)\\
 \\
@@
 \>\loexists \loconj{f}. \loexec{x.\loconj{f}}{\losigma} = \loexec{e}{\losigma}
-\loor \\
-\> \loexec{\texttt{this}}{\losigma} = \loexec{x}{\losigma} \loand\ \loexists y. \loexec{y.\loconj{f}}{\losigma} = \loexec{e}{\losigma})\\
+\loor\
+\loexec{\texttt{this}}{\losigma} = \loexec{x}{\losigma} \loand\ \loexists y. \loexec{y.\loconj{f}}{\losigma} = \loexec{e}{\losigma}\\
 \\
 M,\losigma\ \loturns\ \=MayAccess$^{Ind,Eve}$(x,e) \loiff \\
-\>\loexists \losigma$'$ \loin Arising(M,\losigma).\\
-\>M,\losigma$'$\ \loturns\ MayAccess$^{Ind,Now}$(x,e)\\
+\>\loexists \losigma$'$ \loin Arising(M,\losigma). M,\losigma$'$\ \loturns\ MayAccess$^{Ind,Now}$(x,e)\\
 \hr
@@
 \begin{table}[htb]
-\caption{MayAccess Relations}\label{tab:accessrelations}
+\caption{MayAccess Relations - Observations}\label{tab:accessrelations}
 \centering
@@
 M,\losigma\ \loturns\ \=MayAffect(x,e) if there exists a method \sf{m}, argument(s) \loconj{a}, state \losigma', such that:\\
-\ablock M,\losigma\ \=e.m(\loconj{a}) \loleadsto\ \losigma', and\\
-\>\loexec{e}{M,\losigma} \loneq\ \loexec{e$'$}{M,\losigma$'$}\\
+\ablock M,\losigma\ \=e.m(\loconj{a}) \loleadsto\ \losigma', and \loexec{e}{M,\losigma} \loneq\ \loexec{e$'$}{M,\losigma$'$}\\
 \hr
@@
 
-\begin{logic}
-\hr\\
-\emp{Lemma---[Object as receiver requires method invocation]}\\
-M,\losigma\ \loturns\ \=\losigma(this) = \loexec{o}{\losigma}\loimplies \\
-\> \loexists \losigma$'$,m,\loconj{a}. [ M,\losigma$'$, o.m(\loconj{a}) \loleadsto\ \losigma\ ]\\
-\hr
-\end{logic}
-
-We make clear that for an object to be a receiver in a particular state \losigma, we require a method to exist in the object in a prior state \losigma$'$ that upon invocation can lead to state \losigma.
-This follows from our definition of execution.\\
-
-Our MayCall(o,o$'$) predicate however, is not concerned whether or not the state o$'$ is modified ---the fact that a method in o$'$ can be eventually invoked by o is good enough grounds for us to mean that o has authority of o$'$.
+Our MayCall(o,o$'$) predicate is not concerned whether or not the state of o$'$ is modified ---the fact that a method in o$'$ can be eventually invoked by o is good enough grounds for us to say that o has authority of o$'$.
 Our justification is that in the object-oriented paradigm, programmers almost always enforce security of a sensitive field in a protected object to be declared private, and define some public method in the object that can modify the concerned private field based on some condition.
@@
 \emp{Lemma---[Field Modification Requires Authority]}\\
-M,\losigma\ \loturns\ \loforall \losigma$'$,f. [\=\losigma$'$\loin Arising(M,\losigma) \loand\ \loexec{o$'$.f}{\losigma$'$}\loneq\loexec{o$'$.f}{\losigma}\loimplies \\
-\>\loexists o. M,\losigma\ \loturns\ MayCall(o,o$'$) ]\\
+\loforall \losigma$'$,f. [\=\losigma$'$\loin Arising(M,\losigma) \loand\ \loexec{o$'$.f}{\losigma$'$}\loneq\loexec{o$'$.f}{\losigma}\loimplies \loexists o. M,\losigma\ \loturns\ MayCall(o,o$'$) ]\\
 \hr
@@
 In this section, we formally describe the ideas of \textit{isolation} and \textit{cooperation}, and how \textit{cooperation} can either be vulnerable or protected, within the OCap system.
-We use the word \textit{capability} (of an object), interchangeably with reference (of an object), to represent that both object reference and the right to use the object reference are indistinguishable in OCap; i.e., there is no global ambient authority to prevent an object from using another object reference.
+We use the word \textit{capability} (of an object), interchangeably with reference (of an object), to represent that object reference and the right to use the object reference are indistinguishable in OCap; i.e., there is no global ambient authority to prevent an object from using another object reference.
 Given our 4 formal definitions of \textit{permission}, it follows that o has different kinds of permission of o$'$ depending on whether o has a direct or indirect capability of o$'$, and whether we are arguing about a current or eventual state of the system.
-However, if we were to use \textit{our} formal definition of \textit{authority} (causing effects where effects are method invocations), we cannot say that an object reference embodies authority - we elaborate more in \cref{sec:isolation and cooperation}.
+However, if we were to use \textit{our} formal definition of \textit{authority} (causing effects where effects are method invocations), we would not be able to say that an object reference embodies authority - we elaborate more in \cref{sec:isolation and cooperation}.
 We use the word \textit{path} interchangeably with indirect capability or permission, to emphasise that an object o need not necessarily require the direct capability of o$'$ to invoke the behaviours of o$'$---invocation of o$'$ can be done by o through a network of paths consisting of other object capabilities.\\
@@
 \subsection{Eventual permission from present permission configuration}\label{sec:epc}
-Here, we propose a lemma \textsc{epc} that states how eventual permission can be born from a constellation of direct permissions within an OCap system.
+Here, we propose a lemma that states how eventual permission can be born from a constellation of direct permissions within an OCap system.
 
@@
 
-\textsc{Lemma epc} has the meaning that if an object o has an eventual path to o$'$ in some arising state \losigma$'$, then there must exist an object x that has an eventual path to o$'$, and there must exist a way for o to have a path to x in \losigma$'$.
+Lemma \textsc{epc} guarantees that if an object o has an eventual path to o$'$ in some arising state \losigma$'$, then there must exist an object x that has an eventual path to o$'$, and there must exist a way for o to have a path to x in \losigma$'$.
 Object o either must already have a path to x in \losigma, or that it must be able to receive the capability of x through introduction by x in \losigma.
@@
 In this section, we reason about isolation and cooperation using our formal definition of permission and authority.
-What exactly is object isolation when we say that object o$'$ is isolated from object o, and what exactly is cooperation?\\
-
+Before we define isolation and cooperation, we first propose a lemma on the relationship between permission and authority:\\
+
+\begin{logic}
+\hr\\
+\emp{Lemma---[Permission(Indirect,Eventual) is necessary for Authority]}\\
+M,\losigma\ \loturns\ MayCall(o,o$'$) \loimplies\ MayAccess$^{Ind,Eve}$(o,o$'$)\\
+\hr
+\end{logic}
+
+What exactly is object isolation when we say that object o$'$ is isolated from object o, and what exactly is cooperation?
 We say isolation of o$'$ from o is simply that there is no eventual directed path from o to o$'$:\\
@@
 \begin{logic}
-\emp{Isolation of o$'$ from o:}\\
-\loturns\ \loneg MayAccess$^{Ind,Eve}$(o,o$'$)\\
-\end{logic}
-
-Given \textsc{lemma epc}, we can enforce isolation of o$'$ from o given the present path configuration of a system at \losigma, and when the closed system is eventually opened such that an unknown object has permission and authority of \textit{only} object o in some arising state \losigma', we know that the unknown code will not have permission of o$'$, since o is isolated from o$'$.\\
-
-We know from contraposition of the corollary from our definitions of permission and authority:\\
-
-\begin{logic}
-\hr\\
-\emp{Corollary---[Permission(Indirect,Eventual) is necessary for Authority]}\\
-M,\losigma\ \loturns\
-\= MayCall(o,o$'$) \loimplies \\
-\> MayAccess$^{Ind,Eve}$(o,o$'$)\\
-\hr
-\end{logic}
-
-Therefore,\\
-
-\begin{logic}
+\hr\\
+\emp{Definition---[Isolation of o$'$ from o]}\\
+M,\losigma\ \loturns\ Isolated(o,o$'$) \loiff\ \loneg MayAccess$^{Ind,Eve}$(o,o$'$)\\
+\hr
+\end{logic}
+
+Given lemma \textsc{epc}, we can enforce isolation of o$'$ from o given the present path configuration of a system at \losigma, and when the closed system is eventually opened such that an unknown object has permission and authority of \textit{only} object o in some arising state \losigma', we know that the unknown code will not have permission of o$'$, since o is isolated from o$'$.\\
+
+From the contraposition of the lemma [Permission is necessary for Authority], we have the corollary:\\
+
+\begin{logic}
+\hr\\
 \emp{Isolation implies no authority}\\
-\loturns\ \loneg MayAccess$^{Ind,Eve}$(o,o$'$) \loimplies \loneg Maycall(o,o$'$)\\
-\end{logic}
-
-It is good to know we can guarantee the internal state and behaviour of an object o$'$ in the presence of unknown code on object o through isolation, but in practice this is not entirely useful.
+M,\losigma\ \loturns\ \loneg MayAccess$^{Ind,Eve}$(o,o$'$) \loimplies\ M,\losigma\ \loturns\ \loneg Maycall(o,o$'$)\\
+\hr
+\end{logic}
+
+Therefore, it is good to know we can guarantee the internal state and behaviour of an object o$'$ in the presence of unknown code on object o through isolation, but in practice this is not entirely useful.
 The quintessential question in OCap systems is whether we can allow cooperation without vulnerability.
@@
 
-We say cooperation between two objects o and o$'$ is simply that o has permission of o$'$:\\
-
-\begin{logic}
-\emp{Cooperation between o and o$'$ (o $\rightarrow$ o$'$):}\\
-\loturns\ MayAccess$^{Ind,Eve}$(o,o$'$)\\
-\end{logic}
-
-\begin{logic}
-\emp{Cooperation is necessary but not sufficient for authority:}\\
+We say \textit{directed} cooperation from o to o$'$ is simply that o has permission of o$'$:\\
+
+\begin{logic}
+\hr\\
+\emp{Definition---[Cooperation (o $\rightarrow$ o$'$)]:}\\
+M,\losigma\ \loturns\ Cooperation(o,o$'$) \loiff\ M,\losigma\ \loturns\ MayAccess$^{Ind,Eve}$(o,o$'$)\\
+\hr
+\end{logic}
+
+Furthermore, we make the observation between cooperation and authority that:\\
+\begin{logic}
+\hr\\
+\emp{Observation---[Cooperation is necessary but not sufficient for authority]}\\
 \loturns\ MayAccess$^{Ind,Eve}$(o,o$'$) \lonimplies\ MayCall(o,o$'$)\\
+\hr
 \end{logic}
@@
 What does that mean?
-Notice that if we \textit{know} or \textit{trust} the code of o, then \textbf{cooperation can actually be safe}, even if o contains methods that can eventually invoke behaviours of o$'$.\\
-
-\begin{logic}
-\hr\\
-\emp{Lemma---[Known-Safe cooperation]}\\
-o:KnownObj \loand\ o$'$:ExposedObj \loand\ MayAccess$^{Ind,Eve}$(o,o$'$)\\
-\hr
-\end{logic}
-
-Then regardless of whether MayCall(o,o$'$) holds or not, since we know or trust the code of o, we know that o will not do something that will lead to undesirable method invocations of o$'$.\\
-
-However, if we have another object o* that is of unknown provenance such that:\\
-
-\begin{logic}
-\hr\\
-\emp{Lemma---[Unknown-Vulnerable cooperation]}\\
-o*:UnknownObj \loand\ o$'$:ExposedObj \loand\ MayAccess$^{Ind,Eve}$(o*,o$'$)\\
+Notice that if we \textit{know} or \textit{trust} the code of o, then \textbf{cooperation can actually be safe}, even if o contains methods that can eventually invoke behaviours of o$'$.
+Suppose we have a class KnownObj which we have knowledge of and can trust:\\
+
+\begin{logic}
+\hr\\
+\emp{Definition---[Safe cooperation]}\\
+M,\losigma\ \loturns\ SafeCooperation(o,o$'$) \loiff\ o:KnownObj \loand\ Cooperation(o,o$'$)\\
+\hr
+\end{logic}
+
+Then regardless of whether MayCall(o,o$'$) holds or not, since we know or trust the code of o, we know that o will not do something that will lead to undesirable effects on o$'$.\\
+
+However, if we have another object o* that is of unknown provenance where o* belongs to a class UnknownObj which we do not have code knowledge of, then:\\
+
+\begin{logic}
+\hr\\
+\emp{Definition---[Vulnerable cooperation]}\\
+M,\losigma\ \loturns\ VulnerableCooperation(o*,o$'$) \loiff\ o*:UnknownObj \loand\ Cooperation(o*,o$'$)\\
 \hr
@@
 \>\loiff\\
-\>\loexists x. [ \=MayAccess$^{Ind,Now}$(o,x) \loand\\
-\>\> MayAccess$^{Dir,Now}$(x,o$'$) ] ]\\
+\>\loexists x. [ \=MayAccess$^{Ind,Now}$(o,x) \loand\ MayAccess$^{Dir,Now}$(x,o$'$) ] ]\\
 \hr
@@
 \emp{Corollary---[Authority can be a chain of smaller authorities]}\\
-M,\losigma\ \loturns\ \=MayCall(o,o$'$)\\
-\> \loimplies\\
-\> \loexists x. $[$\=MayCall(o,x) \loand\ MayCall(x,o$'$)$]$\\
+M,\losigma\ \loturns\ \=MayCall(o,o$'$) \loimplies\ \loexists x. $[$\=MayCall(o,x) \loand\ MayCall(x,o$'$)$]$\\
 \hr
@@
 \hr\\
-\emp{Lemma---[Unknown-Protected cooperation]}\\
-o*:UnknownObj \loand\ o$'$:ExposedObj \loand\ MayAccess$^{Ind,Eve}$(o*,o$'$)\\
-\loand\ Dom(S,o$'$) \loand\ \loforall s. [ s\loin S \loimplies s:KnownObj ]\\
+\emp{Definition---[Protected cooperation]}\\
+M,\losigma\ \loturns\ ProtectedCooperation(o*,o$'$) \loiff\ \=o*:UnknownObj \loand\ Cooperation(o*,o$'$) \loand\ \\
+\>\loand\ Dom(S,o$'$) \loand\ \loforall s. [ s\loin S \loimplies s:KnownObj ]\\
 \hr
@@
 \hr\\
-\emp{Lemma---[Unknown-Protected cooperation implies Known-Safe cooperation]}\\
-o*:UnknownObj \loand\ o$'$:ExposedObj \loand\ MayAccess$^{Ind,Eve}$(o*,o$'$)\\
-\loand\ Dom(S,o$'$) \loand\ \loforall s. [ s\loin S \loimplies s:KnownObj ]\\
-\loimplies\
-\loexists s$'$:KnownObj. [ MayAccess$^{Ind,Eve}$(o*,s$'$) \loand\ MayAccess$^{Ind,Eve}$(s$'$,o$'$) ]\\
-\hr
-\end{logic}
-
-This says that if we have a configuration where object o$'$ is dominated by known objects, then the permission to use o$'$ by o* must involve at least a known object.
+\emp{Lemma---[Protected cooperation implies Safe cooperation]}\\
+M,\losigma\ \loturns\ \=ProtectedCooperation(o*,o$'$) \loand\ Dom(S,o$'$) \loand\ \loforall s. [ s\loin S \loimplies s:KnownObj ]\\
+\>\loimplies\\
+\> \loexists s$'$:KnownObj. [ VulnerableCooperation(o*,s$'$) \loand\ SafeCooperation(s$'$,o$'$) ]\\
+\hr
+\end{logic}
+
+This says that if we have a configuration where object o$'$ is dominated by known objects, then the permission to use o$'$ by o* must involve at least a known object s$'$.
 This shows that we can use attenuating objects to mitigate vulnerability between unknown code and known code in OCap systems.\\
@@
 
-Therein lies also the difference between OCap and non-OCap models, because non-OCap models that allow forging of capabilities or a global ambient authority \textit{cannot} enforce security through pure path isolation, which is a distinctive feature of OCap models.\\
+Therein lies also the difference between OCap and non-OCap models.
+Non-OCap models may allow forging of capabilities or a global ambient authority, and thus \textit{cannot} enforce security through pure path isolation, which is a distinctive feature of OCap models.\\
 
 Protection strategies for attenuating objects can be implemented using defensive programming techniques like encapsulation and checking of conditions on method arguments.
-At the extreme ends, attenuating objects can take the shapes of either deciding to forward messages from the unknown object o* in an intact way, thereby allowing MayCall(o*,o$'$) to hold; or if some condition fails, the attenuating object can choose not to forward the message at all such that MayCall(o*,o$'$) does not hold.
-Somewhere in the middle, attenuating objects can also choose to modify the message from o* such that it becomes safe to forward, where MayCall(o*,o$'$) holds but since it has been attenuated by the attenuating object, we would know that the message from the attenuating object would be safe.
+At the extreme end, attenuating objects can take the shape of either deciding to forward messages from the unknown object o* in a totally intact way, thereby allowing MayCall(o*,o$'$) to hold; or if some condition fails, the attenuating object can choose not to forward the message at all such that MayCall(o*,o$'$) does not hold.
+Somewhere in the middle, attenuating objects can also choose to modify the message from o* such that it becomes safe to forward, where MayCall(o*,o$'$) holds, but since it has been attenuated by the trusted attenuating object, we would know that the message from the attenuating object would be safe.
 What and how the attenuating objects decide to do with messages from unknown objects before forwarding them can be best expressed as OCap policies for the attenuating object.\\
@@
 \begin{logic}
-Definition ---[ReNode.parent]\\
-Given rn,rn$'$:ReNode, k:\lonat, then [ rn.parent\losup{k} = rn$'$ \loimplies \=(k $>$ 1 \loand\ rn.parent.parent\losup{k-1} = rn$'$) \loor\\
-\>(k = 1 \loand\ rn.parent = rn$'$) \loor\\
+\hr\\
+\textbf{Definition ---[ReNode.parent]}\\
+Given rn,rn$'$:ReNode, k:\lonat, then [ rn.parent\losup{k} = rn$'$ \loiff\ \=(k $\geq$ 1 \loand\ rn.parent.parent\losup{k-1} = rn$'$) \loor\\
 \>(k = 0 \loand\ rn = rn$'$) ] \\
+\hr
 \end{logic}
 \\
-This definition says that given a superscript \losup{k} placed on a \sf{parent}, it means that the result on the RHS of such a look up is the restricted node's parent on the LHS being called \sf{k} multiple times; if \sf{rn'} refers to the same restricted node as \sf{rn}, then k = 0.
+This definition says that given a superscript \losup{k} placed on a \sf{parent}, the result on the RHS of such a look up is the restricted node's parent on the LHS called \sf{k} times; if \sf{rn'} refers to the same restricted node as \sf{rn}, then k = 0.
 \paragraph{Policies for calling and modifying nodes}
@@
 \begin{logic}
+\hr\\
 \emp{Policy ---[Object calls Node]}\\
@@
 MayCall(rn,n) ]\\
+\hr
 \end{logic}
@@
 That is, a restricted node \sf{rn} that points to a node \sf{n} (\sf{rn.node = n)} \\
-\hr
+
 \paragraph{Policy for restricted node calling other restricted nodes}
 \begin{logic}
+\hr\\
 \emp{Policy ---[ReNode calls ReNode (Necessary Condition)]}\\
@@
 
-Policy [ReNode calls ReNode] specifies a necessary condition and a sufficient condition for which a restricted node \sf{rn} may call another restricted node \sf{rn$'$}.
+\begin{logic}
+\emp{Policy ---[ReNode calls ReNode (Sufficient Condition)]}\\
+\loforall k,j:\lonat, \loforall rn,rn$'$:ReNode. [ \=(rn.parent\losup{k} = rn$'$.parent\losup{j}) \loand\ rn.depth \logeq k \\
+\>\loimplies\\
+\>MayCall(rn,rn$'$) ]\\
+\hr
+\end{logic}\\
+
+Policy [ReNode calls ReNode] describes a necessary condition and a sufficient condition for which a restricted node \sf{rn} may call another restricted node \sf{rn$'$}.
 This policy is subtle and says that a restricted node rn is allowed to navigate to all the descendants from all its accessible ancestors.
@@
 \begin{logic}
+\hr\\
 \emp{Policy ---[Domination No Leaked Paths from ReNode]}\\
@@
 \>\>(s$'$.depth = s$''$.depth - j + k) ] ] ] ]\\
+\hr
 \end{logic}
@@
 The policy confines all possible paths to a node \sf{n} that might arise from running \textit{any} code, if all members in the initial set that dominates n belong to class \sf{ReNode}.
-For all restricted nodes in the new dominating set, they must have been descended from some restricted node (\sf{rn$''$.parent\losup{j}}), where (\sf{rn$''$.parent\losup{j}}) is accessible from a restricted node rn$''$ in the old dominating set (rn$''$.depth \logeq j).
-The policy says that an accessible ancestor rn$''$ from a restricted node rn$'$ in the old dominating set is indexed by a depth level j so that the accesible ancestor would inherit a depth level j higher than rn$''$'s depth level---(rn$''$.parent\losup{j}.depth = rn$''$.depth - j).
-For a newly created restricted node \sf{rn$'$} in the new dominating set that is descended k levels down from one of these accessible ancestors, it would thus inherit a depth of (rn$''$.depth - j + k).
+For all restricted nodes in the new dominating set, they must have been descended from some restricted node (\sf{s$''$.parent\losup{j}}), where (\sf{s$''$.parent\losup{j}}) is accessible from a restricted node s$''$ in the old dominating set (\sf{s$''$.depth \logeq j}).
+The policy says that an accessible ancestor \sf{s$''$} from a restricted node \sf{s$'$} in the old dominating set is indexed by a depth level j so that the accessible ancestor would inherit a depth level j higher than the depth level of \sf{s$''$}---(\sf{s$''$.parent\losup{j}.depth = s$''$.depth - j}).
+For a newly created restricted node \sf{s$'$} in the new dominating set that is descended k levels down from one of these accessible ancestors, it would thus inherit a depth of (\sf{s$''$.depth - j + k}).
 In other words every restricted node in the new dominating set must fall into one of these categories:
@@
 \end{itemize}
-The argument for the policy is as follows: given a set of initial restricted nodes, we know by definition of the callable methods of a restricted node such that it can only create, store and return the capabilities of newly created restricted nodes, or return a \sf{null} value, or return a \sf{String} value, in \textbf{all} of its methods.
-The policies for these methods can be found in \cref{renodemethods}.
-The execution policies in
+The argument for the policy is as follows: given a set of initial restricted nodes, we know by definition of the callable methods of a restricted node such that it can only create, store and return the capabilities of newly created restricted nodes, or return a \sf{null} value, or return a \sf{String} value, in \textbf{all} of its methods.\\
+
+The policies for these methods can be found in \cref{renodemethods}. The execution policies in 
 \cref{renodemethods} imply that all newly created restricted nodes from existing restricted nodes must obey certain policies (such as inheriting an incremented depth from its creator restricted node if its a child, or inheriting a decremented depth if it is a parent).
@@
 
-\begin{logic}
+\begin{logic}\hr\\
 \emp{Policy ---[Execution of ReNode.parent]}\\
@@
 \hr
-\paragraph{Policy for Node Property Modification}\label{policy_node_mod}
-
-This policy uses the Lemma [Field Modification requires Authority] and says that if a property of a node \sf{n} is modified after the execution of some code by object \sf{o}, then it implies \sf{o} must be able to invoke a method on \sf{n}.\\
-
-\begin{logic}
-\emp{Policy ---[Node Property Modification]}\\
-\loforall n:Node, \loforall o:Object. [ [ n.prop(i) = j \{ \texttt{o.\textit{code}} \} n.prop(i) \loneq\ j ] \loimplies\ MayCall(o,n) ]
-\end{logic}
-
-\hr
 \paragraph{Policy for ReNode Constructor}\label{renodeconstructor}
@@
 \begin{logic}
+\hr\\
 \emp{Policy ---[Constructor of new ReNode]}\\
@@
 \hr
-\paragraph{Policies on invariants and immutable properties}\label{policy_immutable}
+\paragraph{Policies on immutable properties}\label{policy_immutable}
 The following policies describe the properties that hold in any state of execution.
@@
 \begin{logic}
-\emp{Policy ---[Invariant of minimum ReNode depth]}\\
+\emp{Policy ---[Immutability of minimum ReNode depth]}\\
 \loforall rn:ReNode. [ \textit{true} \{\texttt{code}\} rn.depth \logeq 0 ]\\
@@
 It is unsafe to hand the capability of \sf{node2b} directly to an untrusted agent, because it is possible the properties and state of the entire DOM tree would be compromised, in particular the root node \sf{node0} whose properties and state we want to protect in this example.
-Note that an untrusted agent having the direct capability of \sf{node2b} implies being that the untrusted agent is able to obtain the direct capability of \sf{node1}, \sf{node2a}, and \sf{node0}, because \sf{node2b} has methods that freely gives those capabilities to any caller.\\
+Note that an untrusted agent having the direct capability of \sf{node2b} implies that the untrusted agent is able to obtain the direct capability of \sf{node1}, \sf{node2a}, and \sf{node0}, because \sf{node2b} has methods that freely gives those capabilities to any caller.\\
 
@@
 We will accomplish this by creating an attenuating restricted node object called \sf{reNode2b}, which points to \sf{node2b} and has a depth level of \sf{1} (\sf{reNode2b.node = node2b, reNode2b.depth = 1}).
-On \texttt{line 19}, we pass \sf{reNode2b} to an object \sf{o}, of unknown provenance.
-While we do not know exactly how \sf{o} might make use of \sf{reNode2b}, to represent unknown code execution on \sf{reNode2b} by \sf{o}, we can assume that \sf{o} would have a function called \sf{mystery} that executes some unknown code on \sf{reNode2b}.
-Using our policies, we can comfortably argue that the properties of \sf{node0} have been preserved after the execution of the unknown \sf{mystery} function by \sf{o}.
+On \texttt{line 19}, we pass \sf{reNode2b} to an object \sf{o} of unknown provenance.
+While we do not know exactly how \sf{o} might make use of \sf{reNode2b}, we can comfortably argue that the properties of \sf{node0} have been preserved after the execution of an unknown \sf{mystery} function by \sf{o}.
 However, because we have configured \sf{reNode2b} to allow the unknown object to potentially modify the properties of \sf{node2a}, \sf{node2b} and \sf{node1}, we do not know if the properties of \sf{node2a}, \sf{node2b} and \sf{node1} have been preserved.\\
 
-We argue that \textbf{Assertion 1} at \texttt{line 21} holds, such that after execution of a piece of unknown code, \sf{node0.getProp("key") = "true"}.
-We begin by stating several facts at \texttt{line 18}, before the execution of \sf{mystery}; we know that:
+We argue that \textbf{Assertion 1} at \texttt{line 21} holds, i.e.
+after execution of a piece of unknown code, we know that \sf{node0.getProp("key") = "true"}.
+We now sketch the proof of this assertion below.\\
+
+We begin by stating several facts at \texttt{line 18}, before the execution of \sf{mystery}. namely:
 \begin{itemize}
@@
 \begin{itemize}
-\item \textbf{(A)}: \textbf{F1} is preserved, unless \textbf{\sf{MayCall(o,node0)}} holds at line 21.
-This is because to modify the value of "key" in \sf{node2b}, \textbf{Policy [Node Property Modification]} at \cref{policy_node_mod} tells us that a method of \sf{node0} has to be invoked by \sf{o}.
+\item \textbf{(A)}: \textbf{F1} is preserved, unless \textbf{\sf{MayCall(o,node0)}} holds at line 19. This is based on the observation below:\\
+\begin{logic}
+\hr\\
+\emp{Observation ---[Node Property Modification]}\\
+\loforall n:Node, \loforall o:Object. [ [ n.prop(i) = j \{ \texttt{o.\textit{code}} \} n.prop(i) \loneq\ j ] \loimplies\ MayCall(o,n) ]\\
+\hr
+\end{logic}
+
+which follows from our Lemma [Field Modification Requires Authority], which says that if a property of a node \sf{n} is modified after the execution of some code by object \sf{o}, then it implies \sf{o} must be able to invoke a method on \sf{n}.
 \end{itemize}
@@
 
-Let us assume:
+Let us assume during execution of the \sf{mystery} function in \texttt{line 19}; at some point:
 \begin{itemize}\item \textbf{(B)}: \sf{MayCall(o,node0)}\end{itemize}
@@
 \begin{itemize} \item \textbf{\sf{\loneg MayCall(o,node0)}} \end{itemize}
-***Consequently, from \textbf{(A)}, we know the properties of \sf{node0} is preserved.\\
+Consequently, from \textbf{(A)}, we know the properties of \sf{node0} is preserved.\\
 
 Note that in the above steps we used without mentioning explicitly that the parent and depth fields of all restricted nodes belonging to \sf{ReNode} are immutable.
-Particularly, our \textbf{Policies on invariants and immutable properties} in \cref{policy_immutable} tell us that the following must hold in all states of exection: \textbf{(F2-F6)}, and the minimum depth of all restricted nodes must be greater than 0.
+Our \textbf{Policies on immutable properties} in \cref{policy_immutable} tell us that the properties \textbf{(F2-F6)} are preserved throughout execution of the body of the \sf{mystery} function, and the minimum depth of all restricted nodes must be greater than 0.
 \clearpage
@@
 The Membrane pattern on the other hand, shares some similarities from both the DOM Tree pattern and Caretaker pattern.
-The Membrane pattern, is identical to the Caretaker pattern in the sense that it is meant to mask the protected object's direct capability and forward messages to the protected object, but with the exception that it does not forward messages that contain capabilities without first 'wrapping' those capabilities within new membranes.
+The Membrane pattern is identical to the Caretaker pattern in the sense that it is meant to mask the protected object's direct capability and forward messages to the protected object, but with the exception that it does not forward messages that contain capabilities without first 'wrapping' those capabilities within new membranes.
 The membrane attenuating object can create more membranes (or deep attenuating objects) that is similar to the DOM Tree pattern where a restricted node creates more restricted nodes when accessing different parts of the tree.\\
 
-We show the visual illustration of the differences between the Caretaker and Membrane patterns in \cref{fig:figcaretaker} and \cref{fig:figmem}.
+We first show the important policies for the Caretaker pattern in \cref{sec:caretakerpolicies} and a visual illustration of it in \cref{fig:figcaretaker}.
+The policies for the Membrane pattern can be found next in \cref{sec:membranepolicies}, and its visual illustration in \cref{fig:figmem}.
 In the Caretaker pattern, while the status of the lock object \sf{Carol-CT-Lock} for the caretaker object \sf{Carol-CT} is disabled (\sf{Carol-CT.status = false}), object \sf{Bob} can pass his own capability to \sf{Carol} using \sf{Carol-CT}, and consequently, with \sf{Bob}'s capability, \sf{Carol} can pass the capability of \sf{Diane}, which \sf{Carol} initially holds, to \sf{Bob}.
@@
 \clearpage
+\subsubsection{Policies for Caretaker Pattern}\label{sec:caretakerpolicies}
+Here we show the important policies for the Caretaker pattern, where we also say how the Caretaker might leak paths in \textbf{Policy [Possible Leaked Paths from Caretaker]}.
+Path leakage is possible because the caretaker might possibly return the capability of its protected object o$'$ from some method call forwarded to o$'$, when o$'$ has a method that exposes its own capability; i.e.
+the caretaker does not create deep attenuating objects to mask capabilities returned by its protected object, or messages containing capabilities passed by its caller.\\
+
+\begin{logic}
+\emp{Policy ---[Object calls Node]}\\
+\loforall o,o$'$:Object, \loforall n:Node. [ \=MayCall(o,o$'$) \loand\ Dom(S,o$'$) \loand\ \loforall s. [ s\loin S \loimplies s:Caretaker ]\\
+\>\loimplies\\
+\>\loexists ct:Caretaker. [ ct\loin S \loand\ MayCall(o,ct) \loand\ MayCall(ct,o$'$) ] ]\\
+\end{logic}
+
+\begin{logic}
+\emp{Policy ---[Possible Leaked Paths from Caretaker]}\\
+\loneg\ [ \loforall o:Object, S$_{old}$. \=[ Dom(S$_{old}$,o) \loand\ \loforall s. [ s\loin S$_{old}$\loimplies s:Caretaker ]\\
+\>\{ \texttt{\textit{code}} \}\\
+\>\loexists S$_{new}$. [ \=Dom(S$_{new}$,n) \loand\ \loforall s$'$\loin S$_{new}$. [ \=s$'$:Caretaker ] ] ] ]\\
+
+\end{logic}
+
+\begin{logic}
+\emp{Policy ---[Caretaker calls Caretaker.target (Necessary Condition)]}\\
+\loforall c:Caretaker. [ MayCall(c,c.target) \loimplies\ c.lock.status = \texttt{false} ]\\
+\end{logic}
+
+\begin{logic}
+\emp{Policy ---[Execution of Lock.lock()]}\\
+\loforall k:Lock. [ \=\textit{true} \{\texttt{k.lock()}\} k.status = \texttt{true} ]\\
+\end{logic}
+
+\begin{logic}
+\emp{Policy ---[Constructor of new Caretaker]}\\
+\loforall ct:Caretaker, \loforall o:Object, \loforall k:Lock. [ \textit{true} \{\texttt{ct = Caretaker(o,k)}\} \=ct.target = o \loand\ ct.lock = k ]\\
+\end{logic}
+
+\begin{logic}
+\emp{Policy ---[Immutability of Caretaker.lock]}\\
+\loforall c:Caretaker. [ \=c.lock = k \{\texttt{\textit{code}}\} c.lock = k ]\\
+\end{logic}
+
+\begin{logic}
+\emp{Policy ---[Immutability of Caretaker.target]}\\
+\loforall c:Caretaker. [ \=c.target = o \{\texttt{\texttt{\textit{code}}}\} c.target = o ]\\
+\hr
+\end{logic}
+
 \begin{minipage}{\textwidth}
@@
 \end{minipage}
+\clearpage
+\subsubsection{Policies for Membrane Pattern}\label{sec:membranepolicies}
+Here we show the important policies for the Membrane pattern.
+Unlike the Caretaker pattern, a Membrane does not leak capabilities because the membrane creates deep attenuating new membranes to mask capabilities returned by its protected object, and capabilities in messages passed by its caller.\\
+
+Every membrane and lock object has a ghost field \sf{parent}: \begin{itemize}
+\item The ghost field \sf{parent} of a membrane object \sf{m$'$} points to a membrane \sf{m} if the constructor for \sf{m$'$} is called  by \sf{m}, otherwise it points to \sf{null}
+\item The ghost field \sf{parent} of a lock object \sf{k$'$} points to a lock \sf{k} if the constructor for \sf{k$'$} is called by a membrane \sf{m}, where \sf{m.lock = k}, otherwise it points to \sf{null}
+\end{itemize}
+
+\begin{logic}
+\emp{Policy ---[Object calls Object$'$]}\\
+\loforall o,o$'$:Object. [ \=MayCall(o,o$'$) \loand\ Dom(S,o$'$) \loand\ \loforall s. [ s\loin S \loimplies s:Membrane ]\\
+\>\loimplies\\
+\>\loexists m:Membrane. [ m\loin S \loand\ MayCall(o,m) \loand\ MayCall(m,o$'$) ] ]\\
+\end{logic}
+
+\begin{logic}
+\emp{Policy ---[Domination No Leaked Paths from Membrane]}\\
+\loforall o:Object, MEM$_{old}$. \=[ Dom(MEM$_{old}$,o) \loand\ \loforall s. [ s\loin MEM$_{old}$\loimplies s:Membrane ]\\
+\>\{ \texttt{\textit{code}} \}\\
+\>\loexists MEM$_{new}$. [ \=Dom(MEM$_{new}$,n) \loand\ \loforall s$'$\loin MEM$_{new}$. [ \=s$'$:Membrane \loand\\
+\>\>\loexists s$''$\loin MEM$_{old}$, \loexists k:\lonat. [ \=(s$'$.parent\losup{k} = s$''$) ] ] ] ]\\
+\end{logic}
+
+\begin{logic}
+\emp{Policy ---[Membrane calls Membrane.target (Necessary Condition)]}\\
+\loforall m:Membrane. [ MayCall(m,m.target) \loimplies\ m.lock.status = \texttt{false} ]\\
+\end{logic}
+
+\begin{logic}
+\emp{Policy ---[Constructor of new Membrane]}\\
+\loforall m:Membrane, \loforall o:Object, \loforall k:Lock. [ \textit{true} \{\texttt{m = Membrane(o,k)}\} \=m.target = o \loand\ m.lock = k ]\\
+\end{logic}
+
+\begin{logic}
+\emp{Policy ---[Each Membrane in the Tree has an associated Lock in the same Tree]}\\
+\loforall m,m$'$:Membrane, \loforall j:\lonat. [ m.parent\losup{j} = m$'$ \loiff\ m.lock.parent\losup{j} = m$'$.lock]\\
+\end{logic}
+
+\begin{logic}
+\emp{Policy ---[Execution of Lock.lock()]}\\
+\loforall k:Lock. [ \=\textit{true} \{\texttt{k.lock()}\} k.status = \texttt{true} ]\\
+\end{logic}
+
+\begin{logic}
+\emp{Policy ---[Execution of Lock.lockall()]}\\
+\loforall k:Lock. [ \=\textit{true} \{\texttt{k.lockall()}\} \loforall k$'$:Lock, \loforall j$>$0. [ k$'$.parent\losup{j} = k \loimplies\ k$'$.status = \texttt{true} ]\\
+\end{logic}
+
+\begin{logic}
+\emp{Policy ---[Immutability of Membrane.lock]}\\
+\loforall m:Membrane. [ \=m.lock = k \{\texttt{\textit{code}}\} m.lock = k ]\\
+\end{logic}
+
+\begin{logic}
+\emp{Policy ---[Immutability of Membrane.target]}\\
+\loforall m:Membrane. [ \=m.target = o \{\texttt{\texttt{\textit{code}}}\} m.target = o ]\\
+\hr
+\end{logic}
 
@@
 \end{minipage}
-\clearpage
-
 \section{Case Study of Ethereum / Solidity as a Non-OCap Model}\label{sec:ethereum}
@@
 \subsubsection{Object References/Addresses} In Ethereum, the designations of all external account and contract objects are made of up a 20 bytes value.
-Hence, if we treat smart contracts and Ethereum accounts as objects, then these object references are non-opaque---because anybody can specify an address using a 20 bytes string value.
+Hence, if we treat smart contracts and Ethereum accounts as objects, then these object references are not unforgeable---because anybody can specify an address using a 20 bytes string value.
 That is, anybody on Ethereum only needs the bit information of a particular address of an object to start interacting with the object.
@@
 In other words, on the Solidity platform, every external account has a path to every object on the blockchain.
-Therefore, in Ethereum, private-public key cryptography is used to determine whether one has the authority to use an external account, while for contract accounts, we can only use a form of stack-based access control that allow whether a method called is allowed to succeed or fail---the protected contract object o$'$ can only enforce policies in a state \losigma\ \textit{after} it has been called by another object since:\\
+Therefore, in Ethereum, private-public key cryptography is used to determine whether one has the authority to use an external account, while for contract accounts, we can only use a form of stack-based access control that determines whether a method call is allowed to succeed or fail---the protected contract object o$'$ can only enforce policies in a state \losigma\ \textit{after} it has been called by another object since:\\
 
@@
 
+Note that because contracts in Ethereum are restricted by their code in terms of what they can perform, it is not true that all contracts may invoke behaviours of all objects in Ethereum.
 \subsubsection{Protection through Message Sender Authentication}
@@
 For external accounts, once a person has provided the correct private key, the person is allowed to call any of the predefined methods in an external account that facilitate ether transactions.
-For contracts, things become more complex because the code of all contracts are transparent and publicly available to everyone.
+For contracts, things become more complex because the code of all contracts is transparent and publicly available to everyone.
 What then, are the kind of protection mechanisms that a smart contract can deploy?
@@
 All this means is that from the receiver object's viewpoint, there must be a way for the receiver to know \textit{who} or what is the sender address of the message.
-The Ethereum platform facilitates sender authentication, such then when an external account or contract account calls a method on a contract account o$'$ by sending a message to o$'$, o$'$ as the receiver, can always see the address of the message sender, and can trust the address of the message sender to be authentic\footnote{Note that we are less concerned with how the architecture of the Ethereum platform guarantees the authenticity of who the message sender is, but more with higher-level protection strategies that smart contracts can employ knowing such a guarantee exists.}.\\
+The Ethereum platform facilitates sender authentication, such that when an external account or contract account calls a method on a contract account o$'$ by sending a message to o$'$, o$'$ as the receiver, can always see the address of the message sender, and can trust the address of the message sender to be authentic\footnote{Note that we are less concerned with how the architecture of the Ethereum platform guarantees the authenticity of who the message sender is, but more with higher-level protection strategies that smart contracts can employ knowing such a guarantee exists.}.\\
 
@@
 
-From the receiver viewpoint, Solidity allows a method call to succeed or fail based on function modifiers.
+From the receiver's viewpoint, Solidity allows a method call to succeed or fail based on function modifiers.
 Function modifiers represent conditions written by the programmer that can be attached to methods that determine whether calls on that method would succeed or fail.
@@
 where we now know that while any object in Solidity can call \texttt{restrictedMeth()} in TestContract, the execution of the method will fail if the object caller's address is not the same as the owner of the contract.
-\subsection{Further Work}
-We note that our \sf{MayCall} predicate is inadequate to describe the protection mechanisms of smart contracts in Ethereum.
+\section{Further Work}
+In \cref{sec:treereasoning}, we rely heavily on Hoare logics, especially when we were using policies on immutable properties to argue how certain properties of restricted nodes will not change  during code execution.
+Hence, we recognise that more work needs to be done to augment the Hoare triple to prove adherence to capability polices so that the logics can guarantee not only some properties to hold after execution of the code, \textbf{but also that these properties hold during execution of the code}.\\
+
+We further note that our \sf{MayCall} predicate is inadequate to describe the protection mechanisms of smart contracts in Ethereum.
 This is because, on the Ethereum platform, the protection mechanism in a smart contract essentially happens \textit{during} the execution of a method in that object, i.e., that the receiver in the state of execution must have \textit{already} transitioned into that object, \textit{before} that object can protect itself.
@@
 checking the identity of the message sender or other conditions), the method call will fail during execution, where the Ethereum virtual machine will, up till the 'point of failure', revert all state changes during the method execution, except the payment of the underlying 'gas' fees to execute the contract on the blockchain.
-Therefore, the platform on Ethereum requires stronger formal definitions of authority that does not merely mean a transition of the receiver from an object to another object---the stronger definition of authority needs to be able to describe the \textit{successful completion} of a method call.
+Therefore, the platform on Ethereum requires stronger formal definitions of authority that do not merely mean a transition of the receiver from an object to another object---the stronger definition of authority needs to be able to describe the \textit{successful completion} of a method call.
+Further work needs to be done in formal reasoning about risk and trust on blockchain platforms, given the meteoric rise of capital traded and number of smart contracts on blockchain technology.
 
@@
 To formally specify OCap patterns in the style of OCap policies, our paper contributes to the literature by proposing new formal definitions for permission and authority that describe access dynamics between objects.
-Using our formal definition, we have reasoned in a novel way about isolation, cooperation, vulnerability, and protection of objects in the face of trusted and untrusted code.
-Our paper has also provided an illustration of three OCap patterns and more specifically, the OCap policy specifications using our proposed methodology for the DOM Tree pattern.
-Using our OCap policies in the DOM Tree pattern, we show how we can preserve properties of a system that interacts with code of unknown provenance.
-In the course of coming up with our OCap policies, we have also proposed the novel use of the concept of domination over objects.
-We have further demonstrated that it is challenging to reason about security properties of modern open systems like the Ethereum blockchain platform that does not adhere to an OCap system, and further work needs to be done in this area given the meteoric rise of capital traded and number of smart contracts on blockchain technology.
+We have also proposed the novel use of the concept of domination over objects to reason about OCap policies.
+Using our formal definitions of these concepts, we have also reasoned in a novel way about isolation, cooperation, vulnerability, and protection of objects in the face of trusted and untrusted code.
+We have further provided an illustration of three OCap patterns and their policies using our methodology.
+By using the DOM Tree pattern, we have also showed how we can use our OCap policies to comfortably reason about which properties of a system can be preserved when interacting with code of unknown provenance.
+Lastly, we have further demonstrated that it is challenging to reason about security properties of modern open systems like the Ethereum blockchain platform that does not adhere to an OCap system.
 
@@
 \section{Appendix}
-\subsection{Module Linking}\label{app:modulelinking}
+\subsection{Definitions in Appendix of \cite{drossopoulou2015b}}
+\subsubsection{Execution}\label{app:execution}
+Here we show the definition of execution and the operational semantics found in the appendix of \cite{drossopoulou2015b}.\\
+\includegraphics[trim={0 5cm 0 5cm},width=0.5\textwidth]{figures/app_execution.pdf}\linebreak
+\includegraphics[trim={0 0 0 0cm},width=\textwidth]{figures/app_op.pdf}
+\clearpage
+
+\subsubsection{Reach}\label{app:reach}
+Here we show the definition of execution and the operational semantics found in the appendix of \cite{drossopoulou2015b}.\\
+
+\includegraphics[trim={0 0 0 0},width=0.5\textwidth]{figures/app_reach.pdf}\linebreak
+\clearpage
+
+\subsubsection{Module Linking}\label{app:modulelinking}
 Here we show the definition of module linking found in the appendix of \cite{drossopoulou2015b}.\\
-
-\includegraphics[width=0.5\textwidth,valign=t]{figures/app_modlink1.pdf}\linebreak
+\includegraphics[trim={0 5cm 0 5cm},width=0.5\textwidth,valign=t]{figures/app_modlink1.pdf}\linebreak
 \includegraphics[width=0.5\textwidth,valign=t]{figures/app_modlink2.pdf}
 \clearpage
-\subsection{Assertion}\label{app:assertion}
+
+\subsubsection{Assertion}\label{app:assertion}
 Here we show the definition of assertion found in the appendix of \cite{drossopoulou2015b}.\\
-
 \includegraphics[width=0.5\textwidth,valign=t]{figures/app_assertion1.pdf}
@@
 \includegraphics[width=\textwidth,valign=t,page=1]{figures/code_DOM.pdf}
-\end{minipage}
+\end{minipage}\clearpage
 \begin{minipage}{\textwidth}
 \includegraphics[width=\textwidth,valign=t,page=2]{figures/code_DOM.pdf}
-\end{minipage}
+\end{minipage}\clearpage
 \subsection{Pony Code - Caretaker}\label{sec:code_Caretaker}
@@
 \includegraphics[width=\textwidth,valign=t,page=1]{figures/code_Caretaker.pdf}
-\end{minipage}
+\end{minipage}\clearpage
 \begin{minipage}{\textwidth}
 \includegraphics[width=\textwidth,valign=t,page=2]{figures/code_Caretaker.pdf}
-\end{minipage}
+\end{minipage}\clearpage
 \begin{minipage}{\textwidth}
 \includegraphics[width=\textwidth,valign=t,page=3]{figures/code_Caretaker.pdf}
-\end{minipage}
+\end{minipage}\clearpage
 \subsection{Pony Code - Membrane}\label{sec:code_Membrane}
@@
 \includegraphics[width=\textwidth,valign=t,page=1]{figures/code_Membrane.pdf}
-\end{minipage}
+\end{minipage}\clearpage
 \begin{minipage}{\textwidth}
 \includegraphics[width=\textwidth,valign=t,page=2]{figures/code_Membrane.pdf}
-\end{minipage}
+\end{minipage}\clearpage
 \begin{minipage}{\textwidth}
 \includegraphics[width=\textwidth,valign=t,page=3]{figures/code_Membrane.pdf}
-\end{minipage}
+\end{minipage}\clearpage
 

### 2017-09-08 shupeng_msc_report.tex
update

@@
 \begin{minipage}{\textwidth}
-\subsubsection{DOM Tree Pattern Illustration}\label{sec:figdom}
-\raggedright \small Here, we illustrate in more detail the example of the DOM Tree pattern in a visual diagram. Code can be found in found in \cref{sec:code_DOM}\\
-\centering
-\includegraphics[width=1.05\textwidth]{figures/DOM.pdf}
+\subsubsection{DOM Tree Pattern - Illustration}\label{sec:figdom}
+\small Here, we illustrate in more detail an example of the DOM Tree pattern. The code can be found in \cref{sec:code_DOM}.\\
+\includegraphics[width=\textwidth]{figures/DOM.pdf}
  \captionof{figure}{DOM Tree Example}\label{fig:figdom}
@@
 \clearpage
-\subsubsection{Policies for Caretaker Pattern}\label{sec:caretakerpolicies}
+\subsubsection{Caretaker Pattern - Policies}\label{sec:caretakerpolicies}
 Here we show the important policies for the Caretaker pattern, where we also say how the Caretaker might leak paths in \textbf{Policy [Possible Leaked Paths from Caretaker]}.
@@
 \begin{minipage}{\textwidth}
-\subsubsection{Caretaker Pattern Illustration}\label{sec:figcaretaker}
-\centering
+\subsubsection{Caretaker Pattern - Illustration}\label{sec:figcaretaker}
+\small Here, we illustrate in more detail an example of the Caretaker pattern. The code can be found in \cref{sec:code_Caretaker}.\\
 \includegraphics[width=1.05\textwidth]{figures/Caretaker.pdf}
@@
 \clearpage
-\subsubsection{Policies for Membrane Pattern}\label{sec:membranepolicies}
+\subsubsection{Membrane Pattern - Policies}\label{sec:membranepolicies}
 Here we show the important policies for the Membrane pattern.
-Unlike the Caretaker pattern, a Membrane does not leak capabilities because the membrane creates deep attenuating new membranes to mask capabilities returned by its protected object, and capabilities in messages passed by its caller.\\
-
-Every membrane and lock object has a ghost field \sf{parent}: \begin{itemize}
+Unlike the Caretaker pattern, a Membrane does not leak capabilities because the membrane creates deep attenuating new membranes to mask capabilities returned by its protected object, and do so also for capabilities in messages passed by its caller.
+A membrane that creates new membranes also creates the associated lock objects for the new membranes, and passes the capabilities of these new locks to its own associated lock object \sf{k}.
+Therefore when some other object calls the method \sf{lockall()}\\on the lock object \sf{k}, \sf{k} will recursively call the same method on all lock objects that it has.
+
+Therefore, every membrane and lock object has a ghost field \sf{parent}: \begin{itemize}
 \item The ghost field \sf{parent} of a membrane object \sf{m$'$} points to a membrane \sf{m} if the constructor for \sf{m$'$} is called  by \sf{m}, otherwise it points to \sf{null}
@@
 \begin{minipage}{\textwidth}
-\subsubsection{Membrane Pattern Illustration}\label{sec:figmem}
-\centering
+\subsubsection{Membrane Pattern - Illustration}\label{sec:figmem}
+\small Here, we illustrate in more detail an example of the Membrane pattern. The code can be found in \cref{sec:code_Membrane}.\\
 \includegraphics[width=1.05\textwidth]{figures/Membrane.pdf}

### 2017-09-08 shupeng_msc_report.tex
update

@@
 \small
-I am immensely grateful to my supervisor Prof.
-Sophia Drossopoulou, for her dedication and guidance, where many ideas included in this thesis were derived and inspired from.
-I will miss the countless hours we have worked together on this thesis.\\
-
-I am also thankful for the emotional support and encouragement from my friends and family; and the Monetary Authority of Singapore for offering me a scholarship to pursue this MSc degree at Imperial College London.
+I am immensely grateful to my supervisor Prof. Sophia Drossopoulou, for her dedication and guidance, where many ideas included in this thesis were derived and inspired from.\\
 \end{minipage}
@@
 To put our formal definitions of permission, authority, and domination in practice and demonstrate their usefulness, 
-our paper provides an illustration of three OCap patterns, DOM Tree pattern, Caretaker pattern, and Membrane pattern, which can be found respectively in \cref{sec:figdom,sec:figcaretaker,sec:figmem}.
-Our paper also provides the corresponding OCap policies based on our formal definitions for these three patterns.
-In \cref{sec:treereasoning}, we choose the Dom Tree pattern as an example to demonstrate how the DOM Tree policies help us to reason comfortably about the preservation of properties of nodes in the presence of unknown code execution.
-A motivation for using the DOM Tree pattern is how guest/unknown code (such as third party scripted advertising or widgets) are becoming more pervasive in the web today (where details on reasoning about them within the object-capability framework can be found in the work by Maffeis et al.\cite{maffeis2010}).
-Lastly, in \cref{sec:ethereum}, we offer some preliminary insights on the security properties of smart contracts on the Ethereum blockchain platform that has experienced explosive growth in recent years, and we recognise that further work has to be done in the field of smart contracts on the blockchain technology.
+our paper provides an illustration of three well-known OCap patterns, DOM Tree pattern, Caretaker pattern, and Membrane pattern, which can be found respectively in \cref{sec:figdom,sec:figcaretaker,sec:figmem}.
+Our paper also provides the corresponding OCap policies for these three patterns in \cref{sec:dompolicies,sec:caretakerpolicies,sec:membranepolicies}.
+In \cref{sec:treereasoning}, we dive deeper and choose the Dom Tree pattern as an example to demonstrate how OCap policies help us to reason comfortably about the preservation of properties of nodes in the presence of unknown code execution.
+A motivation for focusing on the DOM Tree pattern is how guest/unknown code (such as third party scripted advertising or widgets) are becoming more pervasive in the web today (details on reasoning about them within the object-capability framework can be found in the work by Maffeis et al.\cite{maffeis2010}).
+Lastly, in \cref{sec:ethereum}, we offer some preliminary insights on the security properties of smart contracts on the Ethereum blockchain platform that has experienced explosive growth in recent years.
 
@@
 That said, there are benefits to using these concepts to reason within an OCap system, which we demonstrate in \cref{sec:ocapreasoning}, because the restrictions placed by OCap languages or what we call capability-safe languages allow us to enforce protection of objects more easily.
-We further solidify how our methodology can be used to reason convincingly in an OCap system by applying it on a DOM Tree pattern within an OCap framework, in \cref{sec:ocappatterns}.
+We further demonstrate how our methodology can be used to provide the policy specifications for three well-known OCap patterns in \cref{sec:ocappatterns}.
 We end the paper with some preliminary insights on Ethereum in \cref{sec:ethereum}.
@@
 They were later studied in several ways, e.g.
-in the work by Maffeis et al.\cite{maffeis2010}.
-Our formal definition is based on but not identical to that from  the work by Drossopoulou et al.'s\cite{drossopoulou2016}.
+in the work by Maffeis et al.\cite{maffeis2010} and Drossopoulou et al.'s\cite{drossopoulou2016}.
+Our formal definition is based on but not identical to that from  \cite{drossopoulou2016}.
 In their papers and our paper, we generally try to capture the intuition of:
@@
 \end{tabbing}
-For authority, we propose a new MayCall predicate that deviates from the MayAffect predicate proposed in Drossopoulou et al.\cite{drossopoulou2016}.
+For defining authority, we propose a new MayCall predicate that deviates from the MayAffect predicate proposed in Drossopoulou et al.\cite{drossopoulou2016}.
 While their MayAffect predicate captures the idea of a field of an object being modified, our MayCall predicate is much weaker ---we mean the ability to invoke a method of an object.
@@
 
-We further elaborate these predicates in \cref{MayAccess} and \cref{MayCall}.
+We further elaborate these predicates in \cref{MayAccess} and \cref{MayCall}, after introducing a small object-oriented language in the next section, which we would need to define \cref{MayAccess} and \cref{MayCall}.
+Our small language is identical to that found in the appendix of \cite{drossopoulou2015b}.
 \subsection{Specification Language}\label{sec:speclang}
@@
 
-Note that MayAccess$^{\_,Now}$ holds imply MayAccess$^{\_,Eve}$ holds, since an eventual state \losigma$'$ that arises from \losigma\, can refer to \losigma, and therefore, the 'now' definitions implies and are stronger than the 'eventual' definitions.
-Also since a series of fields \loconj{f}\ can mean a singular field f, MayAccess$^{Dir,\_}$ holds imply MayAccess$^{Ind,\_}$ holds, and therefore the 'direct' definitions are stronger than the 'indirect' definitions.
-\\
 \begin{table}[htb]
@@
 \end{tabular*}
-\end{table}\\
-We surmarise the relationships between the four flavours of MayAccess in \cref{tab:accessrelations}.
+\end{table}
+
+We illustrate the relationships between the four flavours of MayAccess in \cref{tab:accessrelations}.
 Note that in an object-to-object paradigm, all four flavours of MayAccess involve object capabilities.
 MayAccess$^{Dir,\_}$(o,o$'$) says object o has the direct reference of o$'$, while MayAccess$^{Ind,\_}$ says object o has a path of reference(s) that lead to o$'$ (that is, o has the reference of some object x$_1$, and object x$_1$ has the reference of object x$_2$ $\dots$ and object x$_n$ has the reference of object o$'$).
-The 'now' and 'eventual' flavours simply state that the "direct" and "indirect" predicates either they hold at \losigma\ (now) or they can hold at some arising eventual state \losigma' from \losigma\ (eventual).
+The 'now' and 'eventual' flavours simply state that the "direct" and "indirect" predicates either they hold at \losigma\ (now) or they can hold at some arising eventual state \losigma' from \losigma\ (eventual).\\
+
+Note that MayAccess$^{\_,Now}$ holds imply MayAccess$^{\_,Eve}$ holds, since an eventual state \losigma$'$ that arises from \losigma\, can refer to \losigma, and therefore, the 'now' definitions implies and are stronger than the 'eventual' definitions.
+Also since a series of fields \loconj{f}\ can mean a singular field f, MayAccess$^{Dir,\_}$ holds imply MayAccess$^{Ind,\_}$ holds, and therefore the 'direct' definitions are stronger than the 'indirect' definitions.
+\\
+
+Note that MayAccess$^{\_,Eve}$ does \textbf{not} imply MayAccess$^{\_,Now}$ holds.
+An analogy will be object A having the capabilities of object B and C at some state \losigma.
+Object B and C do not have the capabilities of each other at \losigma.
+At some arising state \losigma' from \losigma, it is possible that object A shares the capability of C with B through introduction.
+Then at some state \losigma$''$ arising from \losigma$'$ (and arising from \losigma), we know B has access to C after A's introduction.
+Therefore M,\losigma\ \loturns\ MayAccess$^{\_Eve}$(B,C) holds since we have a possible configuration at \losigma$''$ in which B has access to C, but M,\losigma\ \loturns\ MayAccess$^{\_Now}$(B,C) does \textbf{not} hold since no such possible configuration exists at \losigma.\\
+
+Note also that MayAccess$^{\_,Ind}$ does \textbf{not} imply MayAccess$^{\_,Dir}$ holds.
+An analogy will be a configuration where object A does not have the direct capability of object C, but object A has the direct capability of object B and object B has the direct capability of object C.
+Therefore, Object A has a path to C (through B), but it does \textbf{not} imply A has the direct capability of C.\\
 
@@
 
-\textbf{Motivation for MayCall:} The motivation for defining authority as MayCall in our paper is that we would like to describe a \textit{chain} of authority between objects, without needing objects in such a chain to necessarily modify the fields of other objects in the chain.
+\textbf{Motivation for MayCall:} The motivation for defining authority as MayCall in our paper is that we would like to describe a \textit{chain of authority} between objects, without needing objects in such a chain to necessarily modify the fields of other objects in the chain.
 In other words, when object A wants to use object B to modify a field in C but can only do so through object B, we want to express that we can deny A from modifying C either by denying A from calling B, or deny B from calling C.
@@
 Given our definitions of MayCall(o,o$'$), it follows that a necessary condition for modification of a private field in o$'$ is the invocation of a method of o$'$.
-Proving the falsity of our MayCall(o,o$'$) predicate is therefore sufficient to prove that for a field \sf{f} in o$'$, MayAffect(o,o$'$.f) does not hold.\\
+Proving the falsity of our MayCall(o,o$'$) predicate is therefore sufficient to prove that for a field \sf{f} in o$'$, MayAffect(o,o$'$.f) does not hold:
 \begin{logic}
@@
 \loforall \losigma$'$,f. [\=\losigma$'$\loin Arising(M,\losigma) \loand\ \loexec{o$'$.f}{\losigma$'$}\loneq\loexec{o$'$.f}{\losigma}\loimplies \loexists o. M,\losigma\ \loturns\ MayCall(o,o$'$) ]\\
+\\
+\small\textit{Note that o can possibly refer to o$'$}\\
 \hr
@@
 
-In this section, we formally describe the ideas of \textit{isolation} and \textit{cooperation}, and how \textit{cooperation} can either be vulnerable or protected, within the OCap system.
+In this section, we formally describe the ideas of \textit{isolation} and \textit{cooperation}, and how \textit{cooperation} can either be vulnerable, safe, or protected, within the OCap system.
 We use the word \textit{capability} (of an object), interchangeably with reference (of an object), to represent that object reference and the right to use the object reference are indistinguishable in OCap; i.e., there is no global ambient authority to prevent an object from using another object reference.
-Given our 4 formal definitions of \textit{permission}, it follows that o has different kinds of permission of o$'$ depending on whether o has a direct or indirect capability of o$'$, and whether we are arguing about a current or eventual state of the system.
+Given our four formal definitions of \textit{permission}, it follows that o has different kinds of permission of o$'$ depending on whether o has a direct or indirect capability of o$'$, and whether we are arguing about a current or eventual state of the system.
 However, if we were to use \textit{our} formal definition of \textit{authority} (causing effects where effects are method invocations), we would not be able to say that an object reference embodies authority - we elaborate more in \cref{sec:isolation and cooperation}.
@@
 Let us assume o, y, and o$'$ always exists, such that the terminating recursive predicates would return true.
-This example then illustrates an eventual path from o to o$'$ can only exist if one of these present 6 configurations in \losigma\ holds (where x refers to the intermediate object in \textsc{epc}):
+This example then illustrates an eventual path from o to o$'$ can only exist if one of these present 6 configurations in \losigma\ holds (where y refers to the intermediate object x in \textsc{epc}):
 \begin{tabbing}
@@
 \subsection{Isolation and Cooperation}\label{sec:isolation and cooperation}
-In this section, we reason about isolation and cooperation using our formal definition of permission and authority.
-Before we define isolation and cooperation, we first propose a lemma on the relationship between permission and authority:\\
-
-\begin{logic}
-\hr\\
-\emp{Lemma---[Permission(Indirect,Eventual) is necessary for Authority]}\\
+In this section, we reason about isolation and cooperation using our formal definitions of permission and authority.
+Before we define isolation and cooperation, we first observe the relationship between permission and authority:\\
+
+\begin{logic}
+\hr\\
+\emp{Observation---[Permission(Indirect,Eventual) is necessary for Authority]}\\
 M,\losigma\ \loturns\ MayCall(o,o$'$) \loimplies\ MayAccess$^{Ind,Eve}$(o,o$'$)\\
@@
 \end{logic}
+This observation says that in order for an object o to invoke behaviours of o$'$, it needs to be done through a path of capabilities from o to o$'$.\\
 
@@
 
-Given lemma \textsc{epc}, we can enforce isolation of o$'$ from o given the present path configuration of a system at \losigma, and when the closed system is eventually opened such that an unknown object has permission and authority of \textit{only} object o in some arising state \losigma', we know that the unknown code will not have permission of o$'$, since o is isolated from o$'$.\\
-
-From the contraposition of the lemma [Permission is necessary for Authority], we have the corollary:\\
+Given \textsc{lemma epc}, we can enforce isolation of o$'$ from o given the present path configuration of a closed system at \losigma, and when the closed system is eventually opened such that an unknown object has permission and authority of \textit{only} object o in some arising state \losigma', we know that the unknown code will not have permission of o$'$, since o is isolated from o$'$.\\
+
+From the contraposition of the our observation previously that [Permission is necessary for Authority], we have the corollary:\\
 
@@
 
-Therefore, it is good to know we can guarantee the internal state and behaviour of an object o$'$ in the presence of unknown code on object o through isolation, but in practice this is not entirely useful.
+Therefore, it is good to know that through isolation, we can guarantee the internal state and behaviour of an object o$'$ in the presence of unknown code from object o, but in practice this is not entirely useful.
 The quintessential question in OCap systems is whether we can allow cooperation without vulnerability.
@@
 
-We say \textit{directed} cooperation from o to o$'$ is simply that o has permission of o$'$:\\
-
+We say \textit{directed} cooperation from o to o$'$ is simply that o has indirect permission of o$'$:\\
 \begin{logic}
@@
 \end{logic}
-
 Furthermore, we make the observation between cooperation and authority that:\\
@@
 \emp{Observation---[Cooperation is necessary but not sufficient for authority]}\\
-\loturns\ MayAccess$^{Ind,Eve}$(o,o$'$) \lonimplies\ MayCall(o,o$'$)\\
+\loturns\ Cooperation(o,o$'$) \lonimplies\ MayCall(o,o$'$)\\
 \hr
@@
 Since we assume object o is unknown, then we must assume that given the permission of o$'$, o will contain some method that will try to invoke the behaviours of o$'$.
-This then leads us to our next section on vulnerability and protection in the context of cooperation.
-
-\subsection{Vulnerability and Protection}\label{sec:vulnerability and protection}
+This then leads us to our next section on vulnerability, safety, and protection in the context of cooperation.
+
+\subsection{Vulnerability, Safety, and Protection}\label{sec:vulnerability and protection}
 \begin{itemize}\item We assume that object o$'$ will always have some exposed public method for the rest of this section, and is the object of interest that we want to protect.\end{itemize}
-For the concepts of vulnerability and protection, our arguments are based primarily on whether the source object is trusted (known) or untrusted (unknown).\\
+For the concepts of vulnerability, safety, and protection, our arguments are based primarily on whether the source object is trusted (known) or untrusted (unknown).\\
 
@@
 Notice that if we \textit{know} or \textit{trust} the code of o, then \textbf{cooperation can actually be safe}, even if o contains methods that can eventually invoke behaviours of o$'$.
-Suppose we have a class KnownObj which we have knowledge of and can trust:\\
+Suppose we have a class TrustedObj which we have knowledge of and can trust:\\
 
@@
 \emp{Definition---[Safe cooperation]}\\
-M,\losigma\ \loturns\ SafeCooperation(o,o$'$) \loiff\ o:KnownObj \loand\ Cooperation(o,o$'$)\\
+M,\losigma\ \loturns\ SafeCooperation(o,o$'$) \loiff\ o:TrustedObj \loand\ Cooperation(o,o$'$)\\
 \hr
@@
 
-However, if we have another object o* that is of unknown provenance where o* belongs to a class UnknownObj which we do not have code knowledge of, then:\\
+However, if we have another object o* that is of unknown provenance where o* belongs to a class UntrustedObj which we do not have code knowledge of, then:\\
 
@@
 \emp{Definition---[Vulnerable cooperation]}\\
-M,\losigma\ \loturns\ VulnerableCooperation(o*,o$'$) \loiff\ o*:UnknownObj \loand\ Cooperation(o*,o$'$)\\
+M,\losigma\ \loturns\ VulnerableCooperation(o*,o$'$) \loiff\ o*:UntrustedObj \loand\ Cooperation(o*,o$'$)\\
 \hr
@@
 \subsubsection{Attenuating objects as trusted messengers}\label{sec:atten}
-Notice how we already have a result about a Known-Safe cooperation.
+Notice how we already have a result about a safe form of cooperation.
 In this section, we show how can make use of it, and four corollaries from our formal specifications, including the concept of domination that we have proposed.\\
@@
 \emp{Definition---[Protected cooperation]}\\
-M,\losigma\ \loturns\ ProtectedCooperation(o*,o$'$) \loiff\ \=o*:UnknownObj \loand\ Cooperation(o*,o$'$) \loand\ \\
-\>\loand\ Dom(S,o$'$) \loand\ \loforall s. [ s\loin S \loimplies s:KnownObj ]\\
-\hr
-\end{logic}
-
-where objects s are known as attenuating objects in the OCap literature. To see why Unknown-Protected cooperation is safe, see how it follows that:
-
-\begin{logic}
-\hr\\
-\emp{Lemma---[Protected cooperation implies Safe cooperation]}\\
-M,\losigma\ \loturns\ \=ProtectedCooperation(o*,o$'$) \loand\ Dom(S,o$'$) \loand\ \loforall s. [ s\loin S \loimplies s:KnownObj ]\\
+M,\losigma\ \loturns\ ProtectedCooperation(o*,o$'$) \loiff\ \=VulnerableCooperation(o*,o$'$) \loand\ \\
+\>Dom(S,o$'$) \loand\ \loforall s. [ s\loin S \loimplies s:TrustedObj ]\\
+\hr
+\end{logic}
+
+where objects s are known as attenuating objects in the OCap literature. To see why protected cooperation is safe, see how it follows that:
+
+\begin{logic}
+\hr\\
+\emp{Lemma---[Protected cooperation implies a Safe cooperation]}\\
+M,\losigma\ \loturns\ \=ProtectedCooperation(o*,o$'$)\\
 \>\loimplies\\
-\> \loexists s$'$:KnownObj. [ VulnerableCooperation(o*,s$'$) \loand\ SafeCooperation(s$'$,o$'$) ]\\
-\hr
-\end{logic}
-
-This says that if we have a configuration where object o$'$ is dominated by known objects, then the permission to use o$'$ by o* must involve at least a known object s$'$.
-This shows that we can use attenuating objects to mitigate vulnerability between unknown code and known code in OCap systems.\\
+\> \loexists s$'$:TrustedObj. [ VulnerableCooperation(o*,s$'$) \loand\ SafeCooperation(s$'$,o$'$) ]\\
+\hr
+\end{logic}
+
+This says that if we have a configuration where object o$'$ is dominated by known objects, then the permission to use o$'$ by o* must involve at least a trusted object s$'$.
+Therefore we can allow cooperation between o* and o$'$, and still make it safe by using an attenuating trusted object to \textit{shift the vulnerability} from the object of interest o$'$ to the attenuating trusted object.
+This shows that we can use attenuating objects to implement cooperation without vulnerability (of the object of interest) in OCap systems.\\
 
@@
 \item being careful about our present path configurations so that we can ensure eventual path-isolation between the untrusted object and the object of interest
-\item ensuring that trusted objects dominate the object of interest and attenuates the permission and authority of the untrusted object \end{itemize}
+\item ensuring that trusted objects dominate the object of interest and attenuates the authority of the untrusted object \end{itemize}
 
@@
 At the extreme end, attenuating objects can take the shape of either deciding to forward messages from the unknown object o* in a totally intact way, thereby allowing MayCall(o*,o$'$) to hold; or if some condition fails, the attenuating object can choose not to forward the message at all such that MayCall(o*,o$'$) does not hold.
-Somewhere in the middle, attenuating objects can also choose to modify the message from o* such that it becomes safe to forward, where MayCall(o*,o$'$) holds, but since it has been attenuated by the trusted attenuating object, we would know that the message from the attenuating object would be safe.
+Somewhere in the middle, attenuating objects can also choose to modify the message from o* such that it becomes safe to forward (usually by creating more attenuating objects), where MayCall(o*,o$'$) holds, but since it has been attenuated by the trusted attenuating object, we would know that the message from the attenuating object would be safe.
 What and how the attenuating objects decide to do with messages from unknown objects before forwarding them can be best expressed as OCap policies for the attenuating object.\\
 
-In the OCap pattern examples we will go through in the next section, protection of an object is accomplished through some attenuating object that will never leak the capability of the object of interest.
+In the OCap pattern examples we will go through in the next section, protection of an object is accomplished through some attenuating object.
 A safe configuration is also expressed as attenuating objects dominating the object of interest in such a way that unknown code is forced to use them when cooperating with objects of interest in the system.
@@
 An OCap pattern may be visualised as a directed graph---nodes represent objects, and each edge from an object o to another o$'$ represents o holding a capability that allows it to directly access o$'$.
-Indeed, in \cref{sec:figdom}, \cref{sec:figcaretaker}, and \cref{sec:figmem}, we represent such diagrams for three important OCap patterns in the literature.
-These three OCap patterns are: the DOM Tree pattern, the Caretaker pattern, and the Membrane pattern.
-In addition to the visual diagrams of these patterns, we also show the code of these three patterns in the capability-safe language Pony\cite{clebsch2015}, in the appendix \cref{sec:code_DOM,sec:code_Caretaker,sec:code_Membrane}.
-To make use of our formal specifications, we focus exclusively on the DOM Tree pattern, where we show how we can use our formal specifications to write policies that can guarantee the preservation of properties of specific nodes in the pattern.
-
-\subsection{The DOM Tree Pattern}
+Indeed, in \cref{sec:figdom}, \cref{sec:figcaretaker}, and \cref{sec:figmem}, we represent such diagrams for three important OCap patterns in the literature: the DOM Tree pattern, the Caretaker pattern, and the Membrane pattern.
+In addition, we also show the policies of these patterns in \cref{sec:dompolicies,sec:caretakerpolicies,sec:membranepolicies}.
+The code of these three patterns in the capability-safe language Pony\cite{clebsch2015} can be found in the appendix \cref{sec:code_DOM,sec:code_Caretaker,sec:code_Membrane}.
+To show the usefulness of these policies, we have also reasoned in more detail about the DOM Tree pattern in \cref{sec:dom,sec:treereasoning}, where in \cref{sec:treereasoning}, we show how OCap policies can be used to formally reason about the preservation of properties in a system.
+\subsection{The DOM Tree Pattern}\label{sec:dom}
 
@@
 Details can be found in the work by Maffeis et al.\cite{maffeis2010}.
-Our contribution is that we implement the pattern and give the code for the pattern classes \sf{Nodes} and \sf{ReNodes} in the capability-safe language Pony.
-Another contribution is that we provide the formal specifications of such a pattern using our methodology, which then allows us to assert conditions when properties of a Node in the tree are modified.
-\\
+Our key contribution is that we provide the formal specifications of such a pattern using OCap policies, which will then allows us to assert conditions when properties of a Node in the tree are modified.
+We also implement an example of the pattern and give the code for the pattern classes \sf{Nodes} and \sf{ReNodes} in the capability-safe language Pony.\\
 
@@
 Following which, we write the object-capability policies of the DOM Tree pattern using our methodology, in \cref{sec:dompolicies}.
-Lastly, in \cref{sec:treereasoning}, we show how we can begin to reason the protection of a node using those policies.
+Lastly, in \cref{sec:treereasoning}, we show how we can begin to reason about properties of the DOM Tree in the face of unknown code using our proposed OCap policies.
 
@@
 \clearpage
-\subsubsection{Caretaker Pattern - Policies}\label{sec:caretakerpolicies}
-Here we show the important policies for the Caretaker pattern, where we also say how the Caretaker might leak paths in \textbf{Policy [Possible Leaked Paths from Caretaker]}.
-Path leakage is possible because the caretaker might possibly return the capability of its protected object o$'$ from some method call forwarded to o$'$, when o$'$ has a method that exposes its own capability; i.e.
-the caretaker does not create deep attenuating objects to mask capabilities returned by its protected object, or messages containing capabilities passed by its caller.\\
-
-\begin{logic}
-\emp{Policy ---[Object calls Node]}\\
-\loforall o,o$'$:Object, \loforall n:Node. [ \=MayCall(o,o$'$) \loand\ Dom(S,o$'$) \loand\ \loforall s. [ s\loin S \loimplies s:Caretaker ]\\
-\>\loimplies\\
-\>\loexists ct:Caretaker. [ ct\loin S \loand\ MayCall(o,ct) \loand\ MayCall(ct,o$'$) ] ]\\
-\end{logic}
-
-\begin{logic}
-\emp{Policy ---[Possible Leaked Paths from Caretaker]}\\
-\loneg\ [ \loforall o:Object, S$_{old}$. \=[ Dom(S$_{old}$,o) \loand\ \loforall s. [ s\loin S$_{old}$\loimplies s:Caretaker ]\\
-\>\{ \texttt{\textit{code}} \}\\
-\>\loexists S$_{new}$. [ \=Dom(S$_{new}$,n) \loand\ \loforall s$'$\loin S$_{new}$. [ \=s$'$:Caretaker ] ] ] ]\\
-
-\end{logic}
-
-\begin{logic}
-\emp{Policy ---[Caretaker calls Caretaker.target (Necessary Condition)]}\\
-\loforall c:Caretaker. [ MayCall(c,c.target) \loimplies\ c.lock.status = \texttt{false} ]\\
-\end{logic}
-
-\begin{logic}
-\emp{Policy ---[Execution of Lock.lock()]}\\
-\loforall k:Lock. [ \=\textit{true} \{\texttt{k.lock()}\} k.status = \texttt{true} ]\\
-\end{logic}
-
-\begin{logic}
-\emp{Policy ---[Constructor of new Caretaker]}\\
-\loforall ct:Caretaker, \loforall o:Object, \loforall k:Lock. [ \textit{true} \{\texttt{ct = Caretaker(o,k)}\} \=ct.target = o \loand\ ct.lock = k ]\\
-\end{logic}
-
-\begin{logic}
-\emp{Policy ---[Immutability of Caretaker.lock]}\\
-\loforall c:Caretaker. [ \=c.lock = k \{\texttt{\textit{code}}\} c.lock = k ]\\
-\end{logic}
-
-\begin{logic}
-\emp{Policy ---[Immutability of Caretaker.target]}\\
-\loforall c:Caretaker. [ \=c.target = o \{\texttt{\texttt{\textit{code}}}\} c.target = o ]\\
-\hr
-\end{logic}
-
 \begin{minipage}{\textwidth}
@@
 \clearpage
-\subsubsection{Membrane Pattern - Policies}\label{sec:membranepolicies}
-Here we show the important policies for the Membrane pattern.
-Unlike the Caretaker pattern, a Membrane does not leak capabilities because the membrane creates deep attenuating new membranes to mask capabilities returned by its protected object, and do so also for capabilities in messages passed by its caller.
-A membrane that creates new membranes also creates the associated lock objects for the new membranes, and passes the capabilities of these new locks to its own associated lock object \sf{k}.
-Therefore when some other object calls the method \sf{lockall()}\\on the lock object \sf{k}, \sf{k} will recursively call the same method on all lock objects that it has.
-
-Therefore, every membrane and lock object has a ghost field \sf{parent}: \begin{itemize}
-\item The ghost field \sf{parent} of a membrane object \sf{m$'$} points to a membrane \sf{m} if the constructor for \sf{m$'$} is called  by \sf{m}, otherwise it points to \sf{null}
-\item The ghost field \sf{parent} of a lock object \sf{k$'$} points to a lock \sf{k} if the constructor for \sf{k$'$} is called by a membrane \sf{m}, where \sf{m.lock = k}, otherwise it points to \sf{null}
-\end{itemize}
-
-\begin{logic}
-\emp{Policy ---[Object calls Object$'$]}\\
-\loforall o,o$'$:Object. [ \=MayCall(o,o$'$) \loand\ Dom(S,o$'$) \loand\ \loforall s. [ s\loin S \loimplies s:Membrane ]\\
+\subsubsection{Caretaker Pattern - Policies}\label{sec:caretakerpolicies}
+Here we show the important policies for the Caretaker pattern, where we also say how the Caretaker might leak paths in \textbf{Policy [Possible Leaked Paths from Caretaker]}.
+Path leakage is possible because the caretaker might possibly return the capability of its protected object o$'$ from some method call forwarded to o$'$, when o$'$ has a method that exposes its own capability; i.e.
+the caretaker does not create deep attenuating objects to mask capabilities returned by its protected object, or messages containing capabilities passed by its caller.\\
+
+\begin{logic}
+\emp{Policy ---[Object calls Node]}\\
+\loforall o,o$'$:Object, \loforall n:Node. [ \=MayCall(o,o$'$) \loand\ Dom(S,o$'$) \loand\ \loforall s. [ s\loin S \loimplies s:Caretaker ]\\
 \>\loimplies\\
-\>\loexists m:Membrane. [ m\loin S \loand\ MayCall(o,m) \loand\ MayCall(m,o$'$) ] ]\\
-\end{logic}
-
-\begin{logic}
-\emp{Policy ---[Domination No Leaked Paths from Membrane]}\\
-\loforall o:Object, MEM$_{old}$. \=[ Dom(MEM$_{old}$,o) \loand\ \loforall s. [ s\loin MEM$_{old}$\loimplies s:Membrane ]\\
+\>\loexists ct:Caretaker. [ ct\loin S \loand\ MayCall(o,ct) \loand\ MayCall(ct,o$'$) ] ]\\
+\end{logic}
+
+\begin{logic}
+\emp{Policy ---[Possible Leaked Paths from Caretaker]}\\
+\loneg\ [ \loforall o:Object, S$_{old}$. \=[ Dom(S$_{old}$,o) \loand\ \loforall s. [ s\loin S$_{old}$\loimplies s:Caretaker ]\\
 \>\{ \texttt{\textit{code}} \}\\
-\>\loexists MEM$_{new}$. [ \=Dom(MEM$_{new}$,n) \loand\ \loforall s$'$\loin MEM$_{new}$. [ \=s$'$:Membrane \loand\\
-\>\>\loexists s$''$\loin MEM$_{old}$, \loexists k:\lonat. [ \=(s$'$.parent\losup{k} = s$''$) ] ] ] ]\\
-\end{logic}
-
-\begin{logic}
-\emp{Policy ---[Membrane calls Membrane.target (Necessary Condition)]}\\
-\loforall m:Membrane. [ MayCall(m,m.target) \loimplies\ m.lock.status = \texttt{false} ]\\
-\end{logic}
-
-\begin{logic}
-\emp{Policy ---[Constructor of new Membrane]}\\
-\loforall m:Membrane, \loforall o:Object, \loforall k:Lock. [ \textit{true} \{\texttt{m = Membrane(o,k)}\} \=m.target = o \loand\ m.lock = k ]\\
-\end{logic}
-
-\begin{logic}
-\emp{Policy ---[Each Membrane in the Tree has an associated Lock in the same Tree]}\\
-\loforall m,m$'$:Membrane, \loforall j:\lonat. [ m.parent\losup{j} = m$'$ \loiff\ m.lock.parent\losup{j} = m$'$.lock]\\
+\>\loexists S$_{new}$. [ \=Dom(S$_{new}$,n) \loand\ \loforall s$'$\loin S$_{new}$. [ \=s$'$:Caretaker ] ] ] ]\\
+
+\end{logic}
+
+\begin{logic}
+\emp{Policy ---[Caretaker calls Caretaker.target (Necessary Condition)]}\\
+\loforall c:Caretaker. [ MayCall(c,c.target) \loimplies\ c.lock.status = \texttt{false} ]\\
 \end{logic}
@@
 \begin{logic}
-\emp{Policy ---[Execution of Lock.lockall()]}\\
-\loforall k:Lock. [ \=\textit{true} \{\texttt{k.lockall()}\} \loforall k$'$:Lock, \loforall j$>$0. [ k$'$.parent\losup{j} = k \loimplies\ k$'$.status = \texttt{true} ]\\
-\end{logic}
-
-\begin{logic}
-\emp{Policy ---[Immutability of Membrane.lock]}\\
-\loforall m:Membrane. [ \=m.lock = k \{\texttt{\textit{code}}\} m.lock = k ]\\
-\end{logic}
-
-\begin{logic}
-\emp{Policy ---[Immutability of Membrane.target]}\\
-\loforall m:Membrane. [ \=m.target = o \{\texttt{\texttt{\textit{code}}}\} m.target = o ]\\
-\hr
-\end{logic}
-
+\emp{Policy ---[Constructor of new Caretaker]}\\
+\loforall ct:Caretaker, \loforall o:Object, \loforall k:Lock. [ \textit{true} \{\texttt{ct = Caretaker(o,k)}\} \=ct.target = o \loand\ ct.lock = k ]\\
+\end{logic}
+
+\begin{logic}
+\emp{Policy ---[Immutability of Caretaker.lock]}\\
+\loforall c:Caretaker. [ \=c.lock = k \{\texttt{\textit{code}}\} c.lock = k ]\\
+\end{logic}
+
+\begin{logic}
+\emp{Policy ---[Immutability of Caretaker.target]}\\
+\loforall c:Caretaker. [ \=c.target = o \{\texttt{\texttt{\textit{code}}}\} c.target = o ]\\
+\hr
+\end{logic}
 \begin{minipage}{\textwidth}
@@
 \end{minipage}
+\clearpage
+\subsubsection{Membrane Pattern - Policies}\label{sec:membranepolicies}
+Here we show the important policies for the Membrane pattern.
+Unlike the Caretaker pattern, a Membrane does not leak capabilities because the membrane creates deep attenuating new membranes to mask capabilities returned by its protected object, and do so also for capabilities in messages passed by its caller.
+A membrane that creates new membranes also creates the associated lock objects for the new membranes, and passes the capabilities of these new locks to its own associated lock object \sf{k}.
+Therefore when some other object calls the method \sf{lockall()}\\on the lock object \sf{k}, \sf{k} will recursively call the same method on all lock objects that it has.
+
+Therefore, every membrane and lock object has a ghost field \sf{parent}: \begin{itemize}
+\item The ghost field \sf{parent} of a membrane object \sf{m$'$} points to a membrane \sf{m} if the constructor for \sf{m$'$} is called  by \sf{m}, otherwise it points to \sf{null}
+\item The ghost field \sf{parent} of a lock object \sf{k$'$} points to a lock \sf{k} if the constructor for \sf{k$'$} is called by a membrane \sf{m}, where \sf{m.lock = k}, otherwise it points to \sf{null}
+\end{itemize}
+
+\begin{logic}
+\emp{Policy ---[Object calls Object$'$]}\\
+\loforall o,o$'$:Object. [ \=MayCall(o,o$'$) \loand\ Dom(S,o$'$) \loand\ \loforall s. [ s\loin S \loimplies s:Membrane ]\\
+\>\loimplies\\
+\>\loexists m:Membrane. [ m\loin S \loand\ MayCall(o,m) \loand\ MayCall(m,o$'$) ] ]\\
+\end{logic}
+
+\begin{logic}
+\emp{Policy ---[Domination No Leaked Paths from Membrane]}\\
+\loforall o:Object, MEM$_{old}$. \=[ Dom(MEM$_{old}$,o) \loand\ \loforall s. [ s\loin MEM$_{old}$\loimplies s:Membrane ]\\
+\>\{ \texttt{\textit{code}} \}\\
+\>\loexists MEM$_{new}$. [ \=Dom(MEM$_{new}$,n) \loand\ \loforall s$'$\loin MEM$_{new}$. [ \=s$'$:Membrane \loand\\
+\>\>\loexists s$''$\loin MEM$_{old}$, \loexists k:\lonat. [ \=(s$'$.parent\losup{k} = s$''$) ] ] ] ]\\
+\end{logic}
+
+\begin{logic}
+\emp{Policy ---[Membrane calls Membrane.target (Necessary Condition)]}\\
+\loforall m:Membrane. [ MayCall(m,m.target) \loimplies\ m.lock.status = \texttt{false} ]\\
+\end{logic}
+
+\begin{logic}
+\emp{Policy ---[Constructor of new Membrane]}\\
+\loforall m:Membrane, \loforall o:Object, \loforall k:Lock. [ \textit{true} \{\texttt{m = Membrane(o,k)}\} \=m.target = o \loand\ m.lock = k ]\\
+\end{logic}
+
+\begin{logic}
+\emp{Policy ---[Each Membrane in the Tree has an associated Lock in the same Tree]}\\
+\loforall m,m$'$:Membrane, \loforall j:\lonat. [ m.parent\losup{j} = m$'$ \loiff\ m.lock.parent\losup{j} = m$'$.lock]\\
+\end{logic}
+
+\begin{logic}
+\emp{Policy ---[Execution of Lock.lock()]}\\
+\loforall k:Lock. [ \=\textit{true} \{\texttt{k.lock()}\} k.status = \texttt{true} ]\\
+\end{logic}
+
+\begin{logic}
+\emp{Policy ---[Execution of Lock.lockall()]}\\
+\loforall k:Lock. [ \=\textit{true} \{\texttt{k.lockall()}\} \loforall k$'$:Lock, \loforall j$>$0. [ k$'$.parent\losup{j} = k \loimplies\ k$'$.status = \texttt{true} ]\\
+\end{logic}
+
+\begin{logic}
+\emp{Policy ---[Immutability of Membrane.lock]}\\
+\loforall m:Membrane. [ \=m.lock = k \{\texttt{\textit{code}}\} m.lock = k ]\\
+\end{logic}
+
+\begin{logic}
+\emp{Policy ---[Immutability of Membrane.target]}\\
+\loforall m:Membrane. [ \=m.target = o \{\texttt{\texttt{\textit{code}}}\} m.target = o ]\\
+\hr
+\end{logic}
+
 \section{Case Study of Ethereum / Solidity as a Non-OCap Model}\label{sec:ethereum}
@@
 
+This is because even though smart contracts can be deployed without any public methods, all smart contracts have a default \sf{fallback} function that will deal with incoming messages that do not match any of its available methods.
 Note that because contracts in Ethereum are restricted by their code in terms of what they can perform, it is not true that all contracts may invoke behaviours of all objects in Ethereum.
+
 \subsubsection{Protection through Message Sender Authentication}
@@
 \section{Further Work}
-In \cref{sec:treereasoning}, we rely heavily on Hoare logics, especially when we were using policies on immutable properties to argue how certain properties of restricted nodes will not change  during code execution.
+In \cref{sec:treereasoning}, we have relied heavily on Hoare logics, especially when we were using policies on immutable properties to argue how certain properties of restricted nodes will not change \textit{during} unknown code execution.
 Hence, we recognise that more work needs to be done to augment the Hoare triple to prove adherence to capability polices so that the logics can guarantee not only some properties to hold after execution of the code, \textbf{but also that these properties hold during execution of the code}.\\
@@
 Through some condition specified in the method call (e.g.
-checking the identity of the message sender or other conditions), the method call will fail during execution, where the Ethereum virtual machine will, up till the 'point of failure', revert all state changes during the method execution, except the payment of the underlying 'gas' fees to execute the contract on the blockchain.
-Therefore, the platform on Ethereum requires stronger formal definitions of authority that do not merely mean a transition of the receiver from an object to another object---the stronger definition of authority needs to be able to describe the \textit{successful completion} of a method call.
-Further work needs to be done in formal reasoning about risk and trust on blockchain platforms, given the meteoric rise of capital traded and number of smart contracts on blockchain technology.
-
+checking the identity of the message sender or other conditions), the method call will fail during execution, where the Ethereum virtual machine will, up till the 'point of failure', revert all state changes during the method execution, except the payment of the underlying 'gas' fees to execute the contract on the blockchain.\\
+
+Therefore, the Ethereum system requires stronger formal definitions of authority that do not merely mean a transition of the receiver from an object to another object---a stronger definition of authority needs to be able to describe the \textit{successful completion} of a method call.
+Given the meteoric rise in capital traded and number of smart contracts used on open distributed blockchain platforms such as Ethereum, further work needs to be done in formal reasoning about risk and trust on these kind of open systems.
 \section{Conclusion}
@@
 We have also proposed the novel use of the concept of domination over objects to reason about OCap policies.
-Using our formal definitions of these concepts, we have also reasoned in a novel way about isolation, cooperation, vulnerability, and protection of objects in the face of trusted and untrusted code.
-We have further provided an illustration of three OCap patterns and their policies using our methodology.
-By using the DOM Tree pattern, we have also showed how we can use our OCap policies to comfortably reason about which properties of a system can be preserved when interacting with code of unknown provenance.
-Lastly, we have further demonstrated that it is challenging to reason about security properties of modern open systems like the Ethereum blockchain platform that does not adhere to an OCap system.
+Using our formal definitions of these concepts, we have also reasoned in a novel way about isolation, cooperation, vulnerability, safety, and protection of objects in the face of trusted and untrusted code.
+We have further provided an illustration of three OCap patterns and also their policies using our methodology.
+By using the DOM Tree pattern as an example, we have also showed how we can use our OCap policies to comfortably reason about which properties of a system can be preserved when interacting with code of unknown provenance.
+Lastly, we have further demonstrated that it is challenging to reason about security properties of modern open systems like the Ethereum blockchain platform that does not adhere to an OCap model.
 
@@
 \clearpage
-
+\subsubsection{Assertion}\label{app:assertion}
+Here we show the definition of assertion found in the appendix of \cite{drossopoulou2015b}.\\
+\includegraphics[width=0.5\textwidth,valign=t]{figures/app_assertion1.pdf}
+\includegraphics[width=0.5\textwidth,valign=t]{figures/app_assertion2.pdf}
+\clearpage
 \subsubsection{Module Linking}\label{app:modulelinking}
@@
 \clearpage
-
-\subsubsection{Assertion}\label{app:assertion}
-Here we show the definition of assertion found in the appendix of \cite{drossopoulou2015b}.\\
-\includegraphics[width=0.5\textwidth,valign=t]{figures/app_assertion1.pdf}
-\includegraphics[width=0.5\textwidth,valign=t]{figures/app_assertion2.pdf}
-
 \subsection{Pony Code - DOM Tree}\label{sec:code_DOM}

### 2017-09-08 shupeng_msc_report.tex
final

@@
 
-From the contraposition of the our observation previously that [Permission is necessary for Authority], we have the corollary:\\
-
-\begin{logic}
-\hr\\
-\emp{Isolation implies no authority}\\
+From the contraposition of the observation previously that [Permission is necessary for Authority], we have another observation:\\
+
+\begin{logic}
+\hr\\
+\emp{Observation---[Isolation implies no authority]}\\
 M,\losigma\ \loturns\ \loneg MayAccess$^{Ind,Eve}$(o,o$'$) \loimplies\ M,\losigma\ \loturns\ \loneg Maycall(o,o$'$)\\
@@
 Object o can have the reference(capability) of o$'$, and hence there is cooperation, but o contains no methods that can invoke(use) the capability of o$'$, and thus o has the capability of o$'$ but not the authority of o$'$.
-Another analogy, from the other viewpoint of o$'$, would be that o$'$ has no available methods that can be invoked by o ---this can happen if o has no methods, or o only contains private methods or methods that are restricted for objects inherited from the same class as o of which o$'$ is not.\\
+Another analogy, from the other viewpoint of o$'$, would be that o$'$ has no available methods that can be invoked by o ---this can happen if o$'$ has no methods, or o$'$ only contains private methods or methods that are restricted for objects inherited from the same class as o$'$ of which o is not.\\
 
@@
 
-This says that if we have a configuration where object o$'$ is dominated by known objects, then the permission to use o$'$ by o* must involve at least a trusted object s$'$.
+This says that if we have a configuration where object o$'$ is dominated by trusted objects, then the permission to use o$'$ by o* must involve at least a trusted object s$'$.
 Therefore we can allow cooperation between o* and o$'$, and still make it safe by using an attenuating trusted object to \textit{shift the vulnerability} from the object of interest o$'$ to the attenuating trusted object.
